### 2021-02-07 arxiv/arxiv.Rmd
initial with analysis from blog

---
authors:
  - name: Emi Tanaka
    thanks: emitanaka.org
    department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: emi.tanaka@monash.edu
bibliography: references.bib
biblio-style: unsrt
#preamble: >
output: 
  rticles::arxiv_article:
    keep_tex: true
---

```{r, include = FALSE}
paper <- here::here("paper.Rmd")
meta <- rmarkdown::yaml_front_matter(paper)
```

---
title: "`r meta$title`"
abstract: >
  `r meta$abstract`
---

```{r child=paper}
```

### 2021-02-07 arxiv/arxiv.tex
initial with analysis from blog

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{graphicx}

\title{Capitalized Title Here}

\author{
    Emi Tanaka
    \thanks{emitanaka.org}
   \\
    Department of Econometrics and Business Statistics \\
    Monash University \\
  Clayton, VIC 3800 \\
  \texttt{\href{mailto:emi.tanaka@monash.edu}{\nolinkurl{emi.tanaka@monash.edu}}} \\
  }

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

% Pandoc citation processing

\begin{document}
\maketitle

\def\tightlist{}

\begin{abstract}
An abstract of less than 150 words.
\end{abstract}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Introductory section which may include references in parentheses
({\textbf{???}}), or cite a reference such as ({\textbf{???}}) in the
text.

\hypertarget{section-title-in-sentence-case}{%
\subsection{Section title in sentence
case}\label{section-title-in-sentence-case}}

This section may contain a figure such as Figure \ref{fig:Rlogo}.

\begin{figure}[htbp]

{\centering \includegraphics[width=2in]{/Users/etan0038/Dropbox/projects/paper-review-DoE-pkgs/Rlogo}

}

\caption{The logo of R.}\label{fig:Rlogo}
\end{figure}

\hypertarget{another-section}{%
\subsection{Another section}\label{another-section}}

There will likely be several sections, perhaps including code snippets,
such as:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{x \textless{}{-}}\StringTok{ }\DecValTok{1}\OperatorTok{:}\DecValTok{10}
\KeywordTok{plot}\NormalTok{(x)}
\end{Highlighting}
\end{Shaded}

\includegraphics{arxiv_files/figure-latex/unnamed-chunk-3-1.pdf}

\hypertarget{summary}{%
\subsection{Summary}\label{summary}}

This file is only a basic article template. For full details of
\emph{The R Journal} style and information on how to prepare your
article for submission, see the
\href{https://journal.r-project.org/share/author-guide.pdf}{Instructions
for Authors}.

\hypertarget{about-this-format-and-the-r-journal-requirements}{%
\subsubsection{About this format and the R Journal
requirements}\label{about-this-format-and-the-r-journal-requirements}}

\texttt{rticles::rjournal\_article} will help you build the correct
files requirements:

\begin{itemize}
\tightlist
\item
  A R file will be generated automatically using \texttt{knitr::purl} -
  see \url{https://bookdown.org/yihui/rmarkdown-cookbook/purl.html} for
  more information.
\item
  A tex file will be generated from this Rmd file and correctly included
  in \texttt{RJwapper.tex} as expected to build \texttt{RJwrapper.pdf}.
\item
  All figure files will be kept in the default rmarkdown
  \texttt{*\_files} folder. This happens because
  \texttt{keep\_tex\ =\ TRUE} by default in
  \texttt{rticles::rjournal\_article}
\item
  Only the bib filename is to modifed. An example bib file is included
  in the template (\texttt{RJreferences.bib}) and you will have to name
  your bib file as the tex, R, and pdf files.
\end{itemize}

\bibliographystyle{unsrt}
\bibliography{references.bib}

\end{document}

### 2021-02-07 paper.Rmd
initial with analysis from blog

---
title: Current status and propsects of R-packages for the design of experiments
author:
  # see ?rjournal_article for more information
  - name: Emi Tanaka
    affiliation: Monash University
    address:
    - Monash University
    - Clayton campus, VIC 3800, Australia
    url: http://emitanaka.org/
    orcid: 0000-0002-1455-259X
    email:  emi.tanaka@monash.edu
abstract: >
  An abstract of less than 150 words.
preamble: |
  % Any extra LaTeX you need in the preamble

# per R journal requirement, the bib filename should be the same as the output 
# tex file. Don't forget to rename the bib file and change this example value.
bibliography: paper.bib

output: rticles::rjournal_article
---

```{r setup, cache = FALSE, include = FALSE}
library(tidyverse)
library(rvest)
library(lubridate)
library(cranlogs)
library(glue)
library(scales)
library(colorspace)
library(tidytext)
library(pluralize)
library(kableExtra)
library(igraph)
library(ggraph)

knitr::opts_chunk$set(echo = FALSE, 
                      cache = TRUE,
                      cache.path = "cache/",
                      fig.align='center', 
                      fig.pos='htbp')

myggtheme <- 
  theme(panel.background = element_rect(fill = NA),
        panel.grid = element_line(color = "#f6e5ee"),
        axis.text = element_text(color = "#79003e"),
        axis.line = element_line(color = "#79003e", size = 0.7),
        axis.ticks.length = unit(1.4, "mm"),
        axis.ticks = element_line(color = "#79003e", size = 0.7),
        axis.title = element_text(color = "#79003e", face = "bold"),
        strip.background = element_rect(color = "#79003e",
                                        fill = "#AD0059"),
        strip.text = element_text(color = "white"),
        plot.title.position = "plot",
        plot.title = element_text(color = "#79003e", face = "bold")) 
```

```{r cran-titles}
# Thanks to Dirk Eddelbuettel's answer on SO:
# https://stackoverflow.com/questions/11560865/list-and-description-of-all-packages-in-cran-from-within-r
url <- "http://cran.rstudio.com/web/packages/packages.rds"
db <- readRDS(url(url)) %>% 
  as.data.frame()
```

```{r}
nanalysis <- db %>% 
  filter(str_detect(tolower(Title), "analysis")) %>% 
  nrow()

ndesign <- db %>% 
  filter(str_detect(tolower(Title), "design")) %>% 
  nrow()
```

```{r doe-cran}
dat_DoE <- read_html("https://cran.r-project.org/web/views/ExperimentalDesign.html")
date_download <- Sys.Date()
cran_names <- available.packages(contriburl = "http://cran.rstudio.com/src/contrib") %>% 
  rownames() %>% 
  unique() # it should be unique
doe_pkgs <- dat_DoE %>% 
  html_nodes("li") %>% 
  html_nodes("a") %>% 
  html_text() %>% 
  .[. %in% cran_names] %>% 
  unique()

dat_survey <- read_html("https://cran.r-project.org/web/views/OfficialStatistics.html")
survey_pkgs <- dat_survey %>% 
  html_nodes("li") %>% 
  html_nodes("a") %>% 
  html_text() %>% 
  .[. %in% cran_names] %>% 
  unique()
```

```{r bigram}
stop_words_ext <- c(stop_words$word, "doi")

doe_db <- db %>% 
  filter(Package %in% doe_pkgs) %>% 
  mutate(Description = str_replace_all(Description, "\n", " "),
         Description = str_squish(Description),
         Title = str_replace_all(Title, "\n", " "))

bigram_tab <- function(data, col) {
  data %>% 
    unnest_tokens(word, {{col}}, token = "ngrams", n = 2) %>% 
    separate(word, c("word1", "word2"), sep = " ") %>% 
    mutate(word1 = singularize(word1),
           word2 = singularize(word2)) %>% 
    # don't count the same bigram within the same package
    distinct(Package, word1, word2) %>% 
    filter(!word1 %in% stop_words_ext,
           !word2 %in% stop_words_ext,
           !str_detect(word1, "^[0-9.]+$"),
           !str_detect(word2, "^[0-9.]+$")) %>% 
    count(word1, word2, sort = TRUE)  
}
```

```{r bigram-title}
bigram_tab(doe_db, Title) %>% 
  filter(n > 3) %>% 
  mutate(word = paste(word1, word2)) %>% 
  select(word, n) %>% 
  kbl(caption = "The bigram of the R-package titles as provided in the DESCRIPTION file in CRAN.", 
      col.names = c("Bigram", "Count")) 
```

```{r bigram-desc}
bigram_tab(doe_db, Description) %>% 
  filter(n > 4) %>% 
  mutate(word = paste(word1, word2)) %>% 
  select(word, n) %>% 
  kbl(caption = "The bigram of the R-package descriptions as provided in the DESCRIPTION file in CRAN.", 
      col.names = c("Bigram", "Count"), 
      booktabs = TRUE) 
```

```{r}
doe_imports <- doe_db %>% 
  mutate(Depends = str_replace_all(Depends, "\n", " "),
         Depends = str_replace_all(Depends, fixed("("), " ("),
         Imports = str_replace_all(Imports, "\n", " "),
         Imports = str_replace_all(Imports, fixed("("), " ("),
         imports = str_c(Depends, Imports, sep = ","),
         imports = str_split(imports, ","),
         imports = map(imports, ~{
           str_squish(.x) %>% 
             word() %>% 
             .[.!=""]}
         ),
         imports_doe = map(imports, ~.x[.x %in% doe_pkgs])) %>% 
  select(Package, imports_doe) %>% 
  unnest_longer(imports_doe) %>% 
  filter(!is.na(imports_doe)) %>% 
  rename(from = imports_doe, to = Package) %>% 
  select(from, to)
```

```{r doe-network, eval = FALSE}
graph_from_data_frame(doe_imports) %>% 
  ggraph(layout = 'fr') +
  geom_edge_link(aes(start_cap = label_rect(node1.name),
                     end_cap = label_rect(node2.name)), 
                 arrow = arrow(length = unit(2, 'mm')),
                 color = "#79003e") + 
  geom_node_text(aes(label = name),
                 color = "#79003e") +
  theme(panel.background = element_rect(fill = "#f6e5ee",
                                        color = "#79003e"),
        plot.margin = margin(20, 20, 20, 20))
```

```{r cran-data}
end <- Sys.Date() - 2 # usually 1-2 days are not available yet
start <- end - years(5) + days(2)
dldat <- cran_downloads(doe_pkgs, from = start, to = end)
```

```{r download-hist}
dldat %>% 
  group_by(package) %>% 
  summarise(total = sum(count)) %>%
  ggplot(aes(total)) + 
  geom_histogram(color = "white", fill = "#AD0059") + 
  scale_x_log10(label = comma) + 
  myggtheme + 
  labs(x = glue("Total download counts from {start} to {end}"),
       y = "Number of packages") +
  scale_y_continuous(expand = c(0, 0))
```

```{r top5}
ntop <- 5

top5sum_df <- dldat %>% 
  group_by(package) %>% 
  summarise(total = sum(count)) %>% 
  ungroup() %>% 
  slice_max(order_by = total, n = ntop)

top5 <- top5sum_df %>% 
  pull(package)

top5_df <- dldat %>% 
  filter(package %in% top5) %>% 
  mutate(package = fct_reorder(package, count, function(x) -sum(x))) 
```

```{r download-barplot}
top5sum_df %>% 
  mutate(package = fct_reorder(package, total)) %>% 
  ggplot(aes(total, package)) +
  geom_col(aes(fill = package)) +
  labs(x = glue("Total downloads from {start} to {end}"),
       y = "Package") + 
  scale_x_continuous(labels = comma, expand = c(0, 0)) +
  myggtheme + 
  scale_fill_discrete_qualitative(rev = TRUE) + 
  guides(fill = FALSE)
```

```{r pkg-updates, cache = TRUE, dependson="doe-cran"}
pkg_url <- "https://cran.r-project.org/web/packages/{pkg}/index.html"
pkg_archive <- "https://cran.r-project.org/src/contrib/Archive/{pkg}/"
pkg_updates <- map(top5, function(pkg) {
  last_update <- read_html(glue(pkg_url)) %>% 
    html_table() %>% 
    .[[1]] %>% 
    filter(X1=="Published:") %>% 
    pull(X2) %>% 
    ymd()

  archive_dates <- tryCatch({ 
    read_html(glue(pkg_archive)) %>% 
      html_table() %>%
      .[[1]] %>% 
      pull(`Last modified`) %>% 
      ymd_hm() %>% 
      na.omit() %>% 
      as.Date()
  }, error = function(e) {
    NULL
  })
  c(archive_dates, last_update)
})
names(pkg_updates) <- top5

updates <- unlist(pkg_updates) %>% 
  enframe("package", "update") %>% 
  # unlist converts date to integers
  mutate(update = as.Date(update, origin = "1970-01-01"),
         # need to get rid of the numbers appended to pkg names
         package = str_extract(package, paste0(top5, collapse="|")),
         package = factor(package, levels = top5)) %>% 
  filter(between(update, start, end))
```

```{r download-timeplot, dependson="top5"}
ggplot(top5_df, aes(date, count, color = package)) +
  # add shadow lines
  geom_line(data = rename(top5_df, package2 = package), 
            color = "gray", aes(group = package2)) +
  # add date when package was updated
  geom_vline(data = updates, aes(xintercept = update),
             linetype = "dashed", color = "#79003e") + 
  # the trend line
  geom_line() +
  scale_y_log10() +
  facet_grid(package ~ .) + 
  labs(title = glue("Top 5 downloaded DoE packages from {start} to {end}")) + 
  scale_color_discrete_qualitative() +
  guides(color = FALSE) +
  myggtheme
```

## Introduction

`test`
\code{test}

\CRANpkg{agricolae}

\ctv{ExperimentalDesign}

\ctv{OfficialStatistics}

## Helping info to get started

Introductory section which may include references in parentheses
[@R], or cite a reference such as @R in the text.

## Section title in sentence case

Let's check fi this works Figure \@ref(fig:Rlogo).

This section may contain a figure such as Figure \ref{fig:Rlogo}.

```{r, Rlogo, echo=FALSE, fig.cap='The logo of R.', out.width='2in', fig.align='center', fig.pos='htbp'}
knitr::include_graphics(here::here('Rlogo.pdf'))
```

## Summary

This file is only a basic article template. For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](https://journal.r-project.org/share/author-guide.pdf).

### About this format and the R Journal requirements

`rticles::rjournal_article` will help you build the correct files requirements:

* A R file will be generated automatically using `knitr::purl` - see
https://bookdown.org/yihui/rmarkdown-cookbook/purl.html for more information.
* A tex file will be generated from this Rmd file and correctly included in
`RJwapper.tex` as expected to build `RJwrapper.pdf`.
* All figure files will be kept in the default rmarkdown `*_files` folder. This
happens because `keep_tex = TRUE` by default in `rticles::rjournal_article`
* Only the bib filename is to modifed. An example bib file is included in the
template (`RJreferences.bib`) and you will have to name your bib file as the
tex, R, and pdf files.

### 2021-02-07 paper.tex
initial with analysis from blog

--- 2021-02-07 arxiv/arxiv.tex
@@
-\documentclass{article}
+% !TeX root = RJwrapper.tex
+\title{Current status and propsects of R-packages for the design of
+experiments}
+\author{by Emi Tanaka}
 
-\usepackage{arxiv}
+\maketitle
 
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{lipsum}
-\usepackage{graphicx}
-
-\title{Capitalized Title Here}
-
-\author{
-    Emi Tanaka
-    \thanks{emitanaka.org}
-   \\
-    Department of Econometrics and Business Statistics \\
-    Monash University \\
-  Clayton, VIC 3800 \\
-  \texttt{\href{mailto:emi.tanaka@monash.edu}{\nolinkurl{emi.tanaka@monash.edu}}} \\
+\abstract{%
+An abstract of less than 150 words.
 }
 
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\begin{Schunk}
+\begin{table}
 
-% Pandoc citation processing
+\caption{\label{tab:bigram-title}The bigram of the R-package titles as provided in the DESCRIPTION file in CRAN.}
+\centering
+\begin{tabular}[t]{l|r}
+\hline
+Bigram & Count\\
+\hline
+optimal design & 10\\
+\hline
+experimental design & 8\\
+\hline
+clinical trial & 5\\
+\hline
+dose finding & 5\\
+\hline
+sequential design & 5\\
+\hline
+block design & 4\\
+\hline
+microarray experiment & 4\\
+\hline
+response surface & 4\\
+\hline
+\end{tabular}
+\end{table}
 
-\begin{document}
-\maketitle
+\end{Schunk}
 
-\def\tightlist{}
+\begin{Schunk}
+\begin{table}
 
-\begin{abstract}
-An abstract of less than 150 words.
-\end{abstract}
+\caption{\label{tab:bigram-desc}The bigram of the R-package descriptions as provided in the DESCRIPTION file in CRAN.}
+\centering
+\begin{tabular}[t]{lr}
+\toprule
+Bigram & Count\\
+\midrule
+experimental design & 11\\
+optimal design & 10\\
+package provide & 7\\
+response surface & 7\\
+factorial design & 6\\
+\addlinespace
+graphical user & 6\\
+user interface & 6\\
+block design & 5\\
+contour plot & 5\\
+design based & 5\\
+\addlinespace
+effect model & 5\\
+fractional factorial & 5\\
+microarray experiment & 5\\
+mixed effect & 5\\
+provide function & 5\\
+\addlinespace
+sample size & 5\\
+sequential design & 5\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\end{Schunk}
+
+\begin{Schunk}
+\begin{Soutput}
+#> `stat_bin()` using `bins = 30`. Pick better value with `binwidth`.
+\end{Soutput}
+
+\begin{center}\includegraphics{paper_files/figure-latex/download-hist-1} \end{center}
+
+\end{Schunk}
+
+\begin{Schunk}
+
+\begin{center}\includegraphics{paper_files/figure-latex/download-barplot-1} \end{center}
+
+\end{Schunk}
+
+\begin{Schunk}
+
+\begin{center}\includegraphics{paper_files/figure-latex/download-timeplot-1} \end{center}
+
+\end{Schunk}
 
@@
 
+\texttt{test} \code{test}
+
+\CRANpkg{agricolae}
+
+\ctv{ExperimentalDesign}
+
+\ctv{OfficialStatistics}
+
+\hypertarget{helping-info-to-get-started}{%
+\subsection{Helping info to get
+started}\label{helping-info-to-get-started}}
+
 Introductory section which may include references in parentheses
-({\textbf{???}}), or cite a reference such as ({\textbf{???}}) in the
-text.
+\citep{R}, or cite a reference such as \citet{R} in the text.
 
@@
 
+Let's check fi this works Figure @ref(fig:Rlogo).
+
 This section may contain a figure such as Figure \ref{fig:Rlogo}.
 
+\begin{Schunk}
 \begin{figure}[htbp]
@@
 
-\caption{The logo of R.}\label{fig:Rlogo}
+\caption[The logo of R]{The logo of R.}\label{fig:Rlogo}
 \end{figure}
-
-\hypertarget{another-section}{%
-\subsection{Another section}\label{another-section}}
-
-There will likely be several sections, perhaps including code snippets,
-such as:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{x \textless{}{-}}\StringTok{ }\DecValTok{1}\OperatorTok{:}\DecValTok{10}
-\KeywordTok{plot}\NormalTok{(x)}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{arxiv_files/figure-latex/unnamed-chunk-3-1.pdf}
+\end{Schunk}
 
@@
 
-\bibliographystyle{unsrt}
-\bibliography{references.bib}
+\bibliography{paper.bib}
 
-\end{document}
+\address{%
+Emi Tanaka\\
+Monash University\\%
+Monash University\\ Clayton campus, VIC 3800, Australia\\
+%
+\url{http://emitanaka.org/}%
+\\\textit{ORCiD: \href{https://orcid.org/0000-0002-1455-259X}{0000-0002-1455-259X}}%
+\\\href{mailto:emi.tanaka@monash.edu}{\nolinkurl{emi.tanaka@monash.edu}}
+}

### 2021-02-07 paper.R

@@ -1,0 +1,76 @@
+## ----setup, cache = FALSE, include = FALSE------------------------------------
+library(tidyverse)
+library(rvest)
+library(lubridate)
+library(cranlogs)
+library(glue)
+library(scales)
+library(colorspace)
+library(tidytext)
+library(pluralize)
+library(kableExtra)
+library(igraph)
+library(ggraph)
+
+knitr::opts_chunk$set(echo = FALSE, 
+                      cache = TRUE,
+                      cache.path = "cache/")
+
+myggtheme <- 
+  theme(panel.background = element_rect(fill = NA),
+        panel.grid = element_line(color = "#f6e5ee"),
+        axis.text = element_text(color = "#79003e"),
+        axis.line = element_line(color = "#79003e", size = 0.7),
+        axis.ticks.length = unit(1.4, "mm"),
+        axis.ticks = element_line(color = "#79003e", size = 0.7),
+        axis.title = element_text(color = "#79003e", face = "bold"),
+        strip.background = element_rect(color = "#79003e",
+                                        fill = "#AD0059"),
+        strip.text = element_text(color = "white"),
+        plot.title.position = "plot",
+        plot.title = element_text(color = "#79003e", face = "bold")) 
+
+
+## ----cran-titles--------------------------------------------------------------
+# Thanks to Dirk Eddelbuettel's answer on SO:
+# https://stackoverflow.com/questions/11560865/list-and-description-of-all-packages-in-cran-from-within-r
+url <- "http://cran.rstudio.com/web/packages/packages.rds"
+db <- readRDS(url(url)) %>% 
+  as.data.frame()
+
+
+## -----------------------------------------------------------------------------
+nanalysis <- db %>% 
+  filter(str_detect(tolower(Title), "analysis")) %>% 
+  nrow()
+
+ndesign <- db %>% 
+  filter(str_detect(tolower(Title), "design")) %>% 
+  nrow()
+
+
+## ----doe-cran-----------------------------------------------------------------
+dat_DoE <- read_html("https://cran.r-project.org/web/views/ExperimentalDesign.html")
+date_download <- Sys.Date()
+cran_names <- available.packages(contriburl = "http://cran.rstudio.com/") %>% 
+  rownames() %>% 
+  unique() # it should be unique
+doe_pkgs <- dat_DoE %>% 
[…]

### 2021-02-09 paper.Rmd
high level structure of the paper + some cosmetics for plots changed

@@
 abstract: >
-  An abstract of less than 150 words.
+  The critical role of data collection is well captured in the expression ``garbage in, garbage out'' -- in other words, if the collected data is rubbish then no analysis, however complex it may be, can make something out of it.
+The gold standard for data collection is through well-designed experiments.
+Re-running an experiment is generally expensive and in some cases difficult due to limited resources, The design of experiment is important in getting a good value out of the collected data.
+In this article I describe the current state of the R-packages for the design of experiments through textual analysis and download trends.
+I discuss also the software design of widely utilised R packages in the field and conclude with discussion of some future prospects for the field.
 preamble: |
@@
 ---
+
+# Introduction
+
+# Explorative Analysis
+
+## What are the common types of experimental designs?
+
+## How do packages interplay with each other?
+
+## Which packages are widely utilised?
+
+# Software Design
+
+# Propsects and Discussion
+
+\newpage
 
@@
                       fig.align='center', 
-                      fig.pos='htbp')
+                      fig.pos='htbp', 
+                      fig.width = 6,
+                      message = FALSE,
+                      warning = FALSE)
 
-myggtheme <- 
+theme_set(
   theme(panel.background = element_rect(fill = NA),
-        panel.grid = element_line(color = "#f6e5ee"),
-        axis.text = element_text(color = "#79003e"),
-        axis.line = element_line(color = "#79003e", size = 0.7),
+        panel.grid = element_line(color = "lightgray"),
+        axis.text = element_text(color = "black"),
+        axis.line = element_line(color = "black", size = 0.7),
         axis.ticks.length = unit(1.4, "mm"),
-        axis.ticks = element_line(color = "#79003e", size = 0.7),
-        axis.title = element_text(color = "#79003e", face = "bold"),
-        strip.background = element_rect(color = "#79003e",
-                                        fill = "#AD0059"),
+        axis.ticks = element_line(color = "black", size = 0.7),
+        axis.title = element_text(color = "black", face = "bold"),
+        strip.background = element_rect(color = "black",
+                                        fill = "black"),
         strip.text = element_text(color = "white"),
         plot.title.position = "plot",
-        plot.title = element_text(color = "#79003e", face = "bold")) 
+        plot.title = element_text(color = "black", face = "bold")))
 ```
@@
   kbl(caption = "The bigram of the R-package titles as provided in the DESCRIPTION file in CRAN.", 
-      col.names = c("Bigram", "Count")) 
+      col.names = c("Bigram", "Count"), 
+      booktabs = TRUE, 
+      linesep = "") 
 ```
@@
       col.names = c("Bigram", "Count"), 
-      booktabs = TRUE) 
+      booktabs = TRUE, 
+      linesep = "") 
 ```
@@
 ```{r cran-data}
-end <- Sys.Date() - 2 # usually 1-2 days are not available yet
-start <- end - years(5) + days(2)
+end <- Sys.Date() - 2 # usually 1-2 days data are not available yet
+start <- end - years(2) + days(2)
 dldat <- cran_downloads(doe_pkgs, from = start, to = end)
@@
 
-```{r download-hist}
+```{r download-hist, dependson = "cran-data"}
 dldat %>% 
@@
   scale_x_log10(label = comma) + 
-  myggtheme + 
   labs(x = glue("Total download counts from {start} to {end}"),
@@
 
-```{r top5}
-ntop <- 5
+```{r, dependson = "cran-data"}
+dldat %>% 
+  group_by(package) %>% 
+  summarise(total = sum(count)) %>% 
+  filter(total > 100000) %>% 
+  pull(package)
+```
+
+```{r top5, dependson = "cran-data"}
+ntop <- 10
 
@@
 
-```{r download-barplot}
+```{r download-barplot, dependson="cran-data"}
 top5sum_df %>% 
@@
   scale_x_continuous(labels = comma, expand = c(0, 0)) +
-  myggtheme + 
   scale_fill_discrete_qualitative(rev = TRUE) + 
@@
 
-```{r pkg-updates, cache = TRUE, dependson="doe-cran"}
+```{r pkg-updates, cache = TRUE, dependson=c("doe-cran", "top5")}
 pkg_url <- "https://cran.r-project.org/web/packages/{pkg}/index.html"
@@
   scale_color_discrete_qualitative() +
-  guides(color = FALSE) +
-  myggtheme
+  guides(color = FALSE) 
 ```

### 2021-02-09 paper.tex
high level structure of the paper + some cosmetics for plots changed

@@
 \abstract{%
-An abstract of less than 150 words.
+The critical role of data collection is well captured in the expression
+``garbage in, garbage out'\,' -- in other words, if the collected data
+is rubbish then no analysis, however complex it may be, can make
+something out of it. The gold standard for data collection is through
+well-designed experiments. Re-running an experiment is generally
+expensive and in some cases difficult due to limited resources, The
+design of experiment is important in getting a good value out of the
+collected data. In this article I describe the current state of the
+R-packages for the design of experiments through textual analysis and
+download trends. I discuss also the software design of widely utilised R
+packages in the field and conclude with discussion of some future
+prospects for the field.
 }
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+\hypertarget{explorative-analysis}{%
+\section{Explorative Analysis}\label{explorative-analysis}}
+
+\hypertarget{what-are-the-common-types-of-experimental-designs}{%
+\subsection{What are the common types of experimental
+designs?}\label{what-are-the-common-types-of-experimental-designs}}
+
+\hypertarget{how-do-packages-interplay-with-each-other}{%
+\subsection{How do packages interplay with each
+other?}\label{how-do-packages-interplay-with-each-other}}
+
+\hypertarget{which-packages-are-widely-utilised}{%
+\subsection{Which packages are widely
+utilised?}\label{which-packages-are-widely-utilised}}
+
+\hypertarget{software-design}{%
+\section{Software Design}\label{software-design}}
+
+\hypertarget{propsects-and-discussion}{%
+\section{Propsects and Discussion}\label{propsects-and-discussion}}
+
+\newpage
 
@@
 \centering
-\begin{tabular}[t]{l|r}
-\hline
+\begin{tabular}[t]{lr}
+\toprule
 Bigram & Count\\
-\hline
+\midrule
 optimal design & 10\\
-\hline
 experimental design & 8\\
-\hline
 clinical trial & 5\\
-\hline
 dose finding & 5\\
-\hline
 sequential design & 5\\
-\hline
 block design & 4\\
-\hline
 microarray experiment & 4\\
-\hline
 response surface & 4\\
-\hline
+\bottomrule
 \end{tabular}
@@
 factorial design & 6\\
-\addlinespace
 graphical user & 6\\
@@
 design based & 5\\
-\addlinespace
 effect model & 5\\
@@
 provide function & 5\\
-\addlinespace
 sample size & 5\\
@@
 \begin{Schunk}
-\begin{Soutput}
-#> `stat_bin()` using `bins = 30`. Pick better value with `binwidth`.
-\end{Soutput}
 
@@
 
+\end{Schunk}
+
+\begin{Schunk}
+\begin{Soutput}
+#>  [1] "agricolae"   "AlgDesign"   "conf.design" "DiceDesign"  "DiceKriging"
+#>  [6] "DoE.base"    "ez"          "FrF2"        "lhs"         "rsm"
+\end{Soutput}
 \end{Schunk}
@@
 
-\hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
+\hypertarget{introduction-1}{%
+\subsection{Introduction}\label{introduction-1}}
 

### 2021-03-05 paper.Rmd
abstract + intro

--- 2021-02-07 paper.Rmd
@@
 abstract: >
-  An abstract of less than 150 words.
+    Re-running an experiment is generally costly and in some cases impossible due to limited resources, so the design of experiment plays a critical role in increasing the quality of experimental data.
+In this article I describe the current state of the R-packages for the design of experiments through textual analysis and download trends.
+I discuss also the software design of widely utilised R packages in the field of experimental design and conclude with discussion of some future prospects for the field.
 preamble: |
@@
 library(ggraph)
+library(ggrepel)
+library(patchwork)
+library(ggwordcloud)
+tocache <- TRUE
 
 knitr::opts_chunk$set(echo = FALSE, 
-                      cache = TRUE,
+                      cache = FALSE,
                       cache.path = "cache/",
                       fig.align='center', 
-                      fig.pos='htbp')
+                      fig.pos='htbp', 
+                      fig.width = 6,
+                      message = FALSE,
+                      warning = FALSE)
 
-myggtheme <- 
+theme_set(
   theme(panel.background = element_rect(fill = NA),
-        panel.grid = element_line(color = "#f6e5ee"),
-        axis.text = element_text(color = "#79003e"),
-        axis.line = element_line(color = "#79003e", size = 0.7),
+        panel.grid = element_line(color = "lightgray"),
+        axis.text = element_text(color = "black"),
+        axis.line = element_line(color = "black", size = 0.7),
         axis.ticks.length = unit(1.4, "mm"),
-        axis.ticks = element_line(color = "#79003e", size = 0.7),
-        axis.title = element_text(color = "#79003e", face = "bold"),
-        strip.background = element_rect(color = "#79003e",
-                                        fill = "#AD0059"),
+        axis.ticks = element_line(color = "black", size = 0.7),
+        axis.title = element_text(color = "black", face = "bold"),
+        strip.background = element_rect(color = "black",
+                                        fill = "black"),
         strip.text = element_text(color = "white"),
         plot.title.position = "plot",
-        plot.title = element_text(color = "#79003e", face = "bold")) 
+        plot.title = element_text(color = "black", hjust = 0)))
 ```
 
+```{r ctv}
+doe_pkgs <- ctv:::.get_pkgs_from_ctv_or_repos("ExperimentalDesign", 
+                                              repos = "http://cran.rstudio.com/")[[1]]
+doe_pkgs <- c(doe_pkgs, "DeclareDesign")
+npkgs <- length(doe_pkgs)
+```
+
+```{r cran-data, cache = tocache}
+end <- Sys.Date() - 2 # usually 1-2 days data are not available yet
+start <- end - years(2) + days(2)
+dldat <- cran_downloads(doe_pkgs, from = start, to = end)
+```
+
+```{r cran-data2, cache = tocache}
+start2 <- end - years(5) + days(2)
+dldat2 <- cran_downloads(doe_pkgs, from = start2, to = end)
+```
+
+# Introduction
+
+The critical role of data collection is well captured in the expression "garbage in, garbage out" -- in other words, if the collected data is rubbish then no analysis, however complex it may be, can make something out of it.
+Methods for data collection can be dichotomised by the type of data collected -- namely, experimental or observational -- or alternatively, categorised into experimental design or survey design.
+Survey designs often, although not always, aim to collect observational data whilst experimental designs exclusively center on experimental data.
+This dichotomisation, to a great extent, is seen in CRAN task views where R-packages in experimental design are in \ctv{ExperimentalDesign} and R-packages in survey designs are in \ctv{OfficialStatistics}.
+
+In the CRAN task view of \ctv{ExperimentalDesign}, there are `r npkgs` R packages for experimental design and analysis of data from experiments, henceforth referred to as "DoE packages" in this paper.
+The sheer quantity and variation in the output experimental design in the R-packages are arguably unmatched with any other programming languages, e.g.
+in Python [@python], only a handful of libraries that generate design of experiment exist (namely `pyDOE`, `pyDOE2`, `dexpy`, `experimenter` and `GPdoemd`) with limited outputs.
+
+# Explorative Analysis
+
+```{r download-dist, dependson = "cran-data"}
+dldat_sum <- dldat %>% 
+  group_by(package) %>% 
+  summarise(total = sum(count)) %>% 
+  arrange(desc(total)) %>% 
+  mutate(rank = 1:n())
+
+dl1plot <- ggplot(dldat_sum, aes(x = "", y = total/1000)) + 
+  geom_violin() +
+  geom_boxplot(width = 0.1) + 
+  scale_y_log10(label = comma) + 
+  labs(x = "",
+       y = "Download counts in the last 2 year ('000s)", title = "(A)") + 
+  geom_text_repel(data = slice_max(dldat_sum, total, n = 13), 
+                  nudge_x = rep(c(0.2, -0.2, 0.2, -0.2, 0.15), c(3, 2, 2, 3, 3)),
+                  size = 10,
+                  aes(label = package)) + 
+  theme(axis.ticks.length.x = unit(0, "mm"))
+
+dldat_sum5 <- dldat2 %>% 
+  group_by(package) %>% 
+  summarise(total = sum(count)) %>% 
+  arrange(desc(total)) %>% 
+  mutate(rank = 1:n())
+
+top15 <- dldat_sum %>% 
+  left_join(dldat_sum5, by = "package") %>% 
+  slice_max(total.x + total.y, n = 15) %>% 
+  select(package, rank.x, rank.y) %>% 
+  pivot_longer(-package) %>% 
+  mutate(name = ifelse(name=="rank.x", "Last 2 years", "Last 5 years"))
+
+dl2plot <- ggplot(top15, aes(name, value, group = package)) +
+  geom_line() + 
+  geom_point() +
+  scale_y_reverse(breaks = 1:15) +
+  scale_x_discrete(expand = expansion(c(0.1, 0.5))) +
+  labs(x = "", y = "Ranking", title = "(B)") +
+  geom_text(data = filter(top15, name=="Last 5 years"),
+            aes(label = package), 
+            hjust = 0, 
+            nudge_x = 0.1,
+            size = 10) + 
+  theme(panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank())
+```
+
+```{r, fig.cap=glue("The above graph shows the total download counts from {start} to {end} of DoE packages."), fig.width = 6}
+dl1plot + dl2plot
+```
+
+```{r text, cache=tocache}
+url <- paste0(getOption("repos")["CRAN"], "web/packages/packages.rds")
+db <- readRDS(url(url)) %>% 
+  as.data.frame()
+
+doe_db <- db %>% 
+  filter(Package %in% doe_pkgs) %>% 
+  mutate(Description = str_replace_all(Description, "\n", " "),
+         Description = str_squish(Description),
+         Title = str_replace_all(Title, "\n", " "))
+```
+
+```{r}
+singularize2 <- function(x) {
+    res <- singularize(x)
+    res <- ifelse(res=="bia", "bias", res)
+    ifelse(res=="baye", "bayes", res)
+}
+stop_words_rex <- c(paste0("^", stop_words$word, "$"), "^doi", "^[0-9.]+$")
+get_ngram <- function(ngram, desc = c("Title", "Description")) {
+  desc <- match.arg(desc)
+  words <- paste0("word", 1:ngram)
+  doe_db %>% 
+    unnest_tokens(word, desc, token = "ngrams", n = ngram) %>% 
+    separate(word, words, sep = " ") %>% 
+    mutate(across(num_range("word", 1:ngram), singularize2)) %>% 
+    distinct(!!!map(c("Package", words), as.name)) %>% 
+    filter(if_all(num_range("word", 1:ngram), 
+                  ~!str_detect(., paste0(stop_words_rex, collapse = "|")))) %>% 
+    count(!!!map(words, as.name), sort = TRUE) %>% 
+    mutate(word = do.call("paste", map(words, ~eval(parse(text = .x))))) %>% 
+    select(word, n) 
+}
+```
+
+## What are the common types of experimental designs?
+
+## How do packages interplay with each other?
+
+## Which packages are widely utilised?
+
+# Software Design
+
+# Propsects and Discussion
+
+\newpage
+
 ```{r cran-titles}
-# Thanks to Dirk Eddelbuettel's answer on SO:
-# https://stackoverflow.com/questions/11560865/list-and-description-of-all-packages-in-cran-from-within-r
 url <- "http://cran.rstudio.com/web/packages/packages.rds"
@@
   as.data.frame()
-```
-
-```{r}
-nanalysis <- db %>% 
-  filter(str_detect(tolower(Title), "analysis")) %>% 
-  nrow()
-
-ndesign <- db %>% 
-  filter(str_detect(tolower(Title), "design")) %>% 
-  nrow()
-```
-
-```{r doe-cran}
-dat_DoE <- read_html("https://cran.r-project.org/web/views/ExperimentalDesign.html")
-date_download <- Sys.Date()
-cran_names <- available.packages(contriburl = "http://cran.rstudio.com/src/contrib") %>% 
-  rownames() %>% 
-  unique() # it should be unique
-doe_pkgs <- dat_DoE %>% 
-  html_nodes("li") %>% 
-  html_nodes("a") %>% 
-  html_text() %>% 
-  .[. %in% cran_names] %>% 
-  unique()
-
-dat_survey <- read_html("https://cran.r-project.org/web/views/OfficialStatistics.html")
-survey_pkgs <- dat_survey %>% 
-  html_nodes("li") %>% 
-  html_nodes("a") %>% 
-  html_text() %>% 
-  .[. %in% cran_names] %>% 
-  unique()
 ```
@@
 
-```{r bigram-title}
-bigram_tab(doe_db, Title) %>% 
+```{r bigram-title, results="asis"}
+btitle <- bigram_tab(doe_db, Title) %>% 
   filter(n > 3) %>% 
@@
   select(word, n) %>% 
-  kbl(caption = "The bigram of the R-package titles as provided in the DESCRIPTION file in CRAN.", 
-      col.names = c("Bigram", "Count")) 
-```
+  kbl(booktabs = TRUE, linesep = "")
 
-```{r bigram-desc}
-bigram_tab(doe_db, Description) %>% 
+bdesc <- bigram_tab(doe_db, Description) %>% 
   filter(n > 4) %>% 
@@
   select(word, n) %>% 
-  kbl(caption = "The bigram of the R-package descriptions as provided in the DESCRIPTION file in CRAN.", 
-      col.names = c("Bigram", "Count"), 
-      booktabs = TRUE) 
+  kbl(booktabs = TRUE, linesep = "")
+
+cat(c("\\begin{table}[h] \\centering ", 
+      btitle,
+    "\\hspace{1cm} \\centering ",
+      bdesc,
+    "\\caption{My tables} \\end{table}")) 
 ```
@@
 
-```{r cran-data}
-end <- Sys.Date() - 2 # usually 1-2 days are not available yet
-start <- end - years(5) + days(2)
-dldat <- cran_downloads(doe_pkgs, from = start, to = end)
-```
-
-```{r download-hist}
-dldat %>% 
-  group_by(package) %>% 
-  summarise(total = sum(count)) %>%
-  ggplot(aes(total)) + 
-  geom_histogram(color = "white", fill = "#AD0059") + 
-  scale_x_log10(label = comma) + 
-  myggtheme + 
-  labs(x = glue("Total download counts from {start} to {end}"),
-       y = "Number of packages") +
-  scale_y_continuous(expand = c(0, 0))
-```
-
-```{r top5}
-ntop <- 5
-
-top5sum_df <- dldat %>% 
-  group_by(package) %>% 
-  summarise(total = sum(count)) %>% 
-  ungroup() %>% 
-  slice_max(order_by = total, n = ntop)
-
-top5 <- top5sum_df %>% 
-  pull(package)
-
-top5_df <- dldat %>% 
-  filter(package %in% top5) %>% 
-  mutate(package = fct_reorder(package, count, function(x) -sum(x))) 
-```
-
-```{r download-barplot}
-top5sum_df %>% 
-  mutate(package = fct_reorder(package, total)) %>% 
-  ggplot(aes(total, package)) +
-  geom_col(aes(fill = package)) +
-  labs(x = glue("Total downloads from {start} to {end}"),
-       y = "Package") + 
-  scale_x_continuous(labels = comma, expand = c(0, 0)) +
-  myggtheme + 
-  scale_fill_discrete_qualitative(rev = TRUE) + 
-  guides(fill = FALSE)
-```
-
-```{r pkg-updates, cache = TRUE, dependson="doe-cran"}
-pkg_url <- "https://cran.r-project.org/web/packages/{pkg}/index.html"
-pkg_archive <- "https://cran.r-project.org/src/contrib/Archive/{pkg}/"
-pkg_updates <- map(top5, function(pkg) {
-  last_update <- read_html(glue(pkg_url)) %>% 
-    html_table() %>% 
-    .[[1]] %>% 
-    filter(X1=="Published:") %>% 
-    pull(X2) %>% 
-    ymd()
-
-  archive_dates <- tryCatch({ 
-    read_html(glue(pkg_archive)) %>% 
-      html_table() %>%
-      .[[1]] %>% 
-      pull(`Last modified`) %>% 
-      ymd_hm() %>% 
-      na.omit() %>% 
-      as.Date()
-  }, error = function(e) {
-    NULL
-  })
-  c(archive_dates, last_update)
-})
-names(pkg_updates) <- top5
-
-updates <- unlist(pkg_updates) %>% 
-  enframe("package", "update") %>% 
-  # unlist converts date to integers
-  mutate(update = as.Date(update, origin = "1970-01-01"),
-         # need to get rid of the numbers appended to pkg names
-         package = str_extract(package, paste0(top5, collapse="|")),
-         package = factor(package, levels = top5)) %>% 
-  filter(between(update, start, end))
-```
-
-```{r download-timeplot, dependson="top5"}
-ggplot(top5_df, aes(date, count, color = package)) +
-  # add shadow lines
-  geom_line(data = rename(top5_df, package2 = package), 
-            color = "gray", aes(group = package2)) +
-  # add date when package was updated
-  geom_vline(data = updates, aes(xintercept = update),
-             linetype = "dashed", color = "#79003e") + 
-  # the trend line
-  geom_line() +
-  scale_y_log10() +
-  facet_grid(package ~ .) + 
-  labs(title = glue("Top 5 downloaded DoE packages from {start} to {end}")) + 
-  scale_color_discrete_qualitative() +
-  guides(color = FALSE) +
-  myggtheme
-```
-
-## Introduction
-
-`test`
-\code{test}
-
 \CRANpkg{agricolae}
-
-\ctv{ExperimentalDesign}
-
-\ctv{OfficialStatistics}
 
@@
 This section may contain a figure such as Figure \ref{fig:Rlogo}.
-
-```{r, Rlogo, echo=FALSE, fig.cap='The logo of R.', out.width='2in', fig.align='center', fig.pos='htbp'}
-knitr::include_graphics(here::here('Rlogo.pdf'))
-```
 

### 2021-03-05 paper.tex
abstract + intro

@@
 \abstract{%
-The critical role of data collection is well captured in the expression
-``garbage in, garbage out'\,' -- in other words, if the collected data
-is rubbish then no analysis, however complex it may be, can make
-something out of it. The gold standard for data collection is through
-well-designed experiments. Re-running an experiment is generally
-expensive and in some cases difficult due to limited resources, The
-design of experiment is important in getting a good value out of the
-collected data. In this article I describe the current state of the
-R-packages for the design of experiments through textual analysis and
-download trends. I discuss also the software design of widely utilised R
-packages in the field and conclude with discussion of some future
+Re-running an experiment is generally costly and in some cases
+impossible due to limited resources, so the design of experiment plays a
+critical role in increasing the quality of experimental data. In this
+article I describe the current state of the R-packages for the design of
+experiments through textual analysis and download trends. I discuss also
+the software design of widely utilised R packages in the field of
+experimental design and conclude with discussion of some future
 prospects for the field.
@@
 
+The critical role of data collection is well captured in the expression
+``garbage in, garbage out'' -- in other words, if the collected data is
+rubbish then no analysis, however complex it may be, can make something
+out of it. Methods for data collection can be dichotomised by the type
+of data collected -- namely, experimental or observational -- or
+alternatively, categorised into experimental design or survey design.
+Survey designs often, although not always, aim to collect observational
+data whilst experimental designs are exclusively center on experimental
+data. This dichotomisation, to a great extent, is seen in CRAN task
+views where R-packages in experimental design are in
+\ctv{ExperimentalDesign} and R-packages in survey designs are in
+\ctv{OfficialStatistics}.
+
+In the CRAN task view of \ctv{ExperimentalDesign}, there are 113 R
+packages for experimental design and analysis of data from experiments,
+henceforth referred to as ``DoE packages'' in this paper. The sheer
+quantity and variation in the output experimental design in the
+R-packages are arguably unmatched with any other programming languages,
+e.g.~in Python \citep{python}, only a handful of libraries that generate
+design of experiment exist (namely \texttt{pyDOE}, \texttt{pyDOE2},
+\texttt{dexpy}, \texttt{experimenter} and \texttt{GPdoemd}) with limited
+outputs.
+
 \hypertarget{explorative-analysis}{%
 \section{Explorative Analysis}\label{explorative-analysis}}
+
+\begin{Schunk}
+\begin{figure}[htbp]
+
+{\centering \includegraphics{paper_files/figure-latex/unnamed-chunk-1-1}
+
+}
+
+\caption[The above graph shows the total download counts from 2019-02-09 to 2021-02-07 of DoE packages]{The above graph shows the total download counts from 2019-02-09 to 2021-02-07 of DoE packages.}\label{fig:unnamed-chunk-1}
+\end{figure}
+\end{Schunk}
 
@@
 
-\begin{Schunk}
-\begin{table}
-
-\caption{\label{tab:bigram-title}The bigram of the R-package titles as provided in the DESCRIPTION file in CRAN.}
-\centering
+\begin{table}[h] \centering  
 \begin{tabular}[t]{lr}
 \toprule
-Bigram & Count\\
+word & n\\
 \midrule
@@
 \bottomrule
-\end{tabular}
-\end{table}
-
-\end{Schunk}
-
-\begin{Schunk}
-\begin{table}
-
-\caption{\label{tab:bigram-desc}The bigram of the R-package descriptions as provided in the DESCRIPTION file in CRAN.}
-\centering
+\end{tabular} \hspace{1cm} \centering  
 \begin{tabular}[t]{lr}
 \toprule
-Bigram & Count\\
+word & n\\
 \midrule
@@
 package provide & 7\\
-response surface & 7\\
-factorial design & 6\\
 graphical user & 6\\
+response surface & 6\\
 user interface & 6\\
@@
 effect model & 5\\
-fractional factorial & 5\\
+factorial design & 5\\
 microarray experiment & 5\\
@@
 \bottomrule
-\end{tabular}
-\end{table}
-
-\end{Schunk}
-
-\begin{Schunk}
-
-\begin{center}\includegraphics{paper_files/figure-latex/download-hist-1} \end{center}
-
-\end{Schunk}
-
-\begin{Schunk}
-\begin{Soutput}
-#>  [1] "agricolae"   "AlgDesign"   "conf.design" "DiceDesign"  "DiceKriging"
-#>  [6] "DoE.base"    "ez"          "FrF2"        "lhs"         "rsm"
-\end{Soutput}
-\end{Schunk}
-
-\begin{Schunk}
-
-\begin{center}\includegraphics{paper_files/figure-latex/download-barplot-1} \end{center}
-
-\end{Schunk}
-
-\begin{Schunk}
-
-\begin{center}\includegraphics{paper_files/figure-latex/download-timeplot-1} \end{center}
-
-\end{Schunk}
-
-\hypertarget{introduction-1}{%
-\subsection{Introduction}\label{introduction-1}}
-
-\texttt{test} \code{test}
+\end{tabular} \caption{My tables} \end{table}
 
 \CRANpkg{agricolae}
-
-\ctv{ExperimentalDesign}
-
-\ctv{OfficialStatistics}
 
@@
 This section may contain a figure such as Figure \ref{fig:Rlogo}.
-
-\begin{Schunk}
-\begin{figure}[htbp]
-
-{\centering \includegraphics[width=2in]{/Users/etan0038/Dropbox/projects/paper-review-DoE-pkgs/Rlogo}
-
-}
-
-\caption[The logo of R]{The logo of R.}\label{fig:Rlogo}
-\end{figure}
-\end{Schunk}
 

### 2021-03-05 paper.Rmd
wordcloud

@@
 tocache <- TRUE
-
 knitr::opts_chunk$set(echo = FALSE, 
@@
                       cache.path = "cache/",
-                      fig.align='center', 
-                      fig.pos='htbp', 
+                      fig.align = 'center', 
+                      fig.pos = 'htbp', 
                       fig.width = 6,
@@
                   nudge_x = rep(c(0.2, -0.2, 0.2, -0.2, 0.15), c(3, 2, 2, 3, 3)),
-                  size = 10,
+                  size = 4,
                   aes(label = package)) + 
@@
             nudge_x = 0.1,
-            size = 10) + 
+            size = 4) + 
   theme(panel.grid.major = element_blank(),
@@
 
-```{r, fig.cap=glue("The above graph shows the total download counts from {start} to {end} of DoE packages."), fig.width = 6}
+```{r dlplots, fig.cap=glue("The above graph shows the total download counts from {start} to {end} of DoE packages."), fig.width = 6}
 dl1plot + dl2plot
@@
 ```{r text, cache=tocache}
-url <- paste0(getOption("repos")["CRAN"], "web/packages/packages.rds")
+url <- "http://cran.rstudio.com/web/packages/packages.rds"
 db <- readRDS(url(url)) %>% 
@@
 
-```{r}
+```{r fns}
 singularize2 <- function(x) {
@@
     distinct(!!!map(c("Package", words), as.name)) %>% 
-    filter(if_all(num_range("word", 1:ngram), 
+    # allow stop words in between
+    filter(if_all(num_range("word", c(1, ngram)), 
                   ~!str_detect(., paste0(stop_words_rex, collapse = "|")))) %>% 
@@
 }
+
+combine_dl <- function(ngram_df, desc) {
+  dldat_sum %>% 
+    left_join(doe_db, by = c("package" = "Package")) %>% 
+    select(package, total, {{desc}}) %>% 
+    mutate(bigram_title = map({{desc}}, ~filter(ngram_df, 
+                                              n > 1,
+                                              str_detect(tolower(.x), word)))) %>% 
+    unnest_longer(bigram_title) %>% 
+    unpack(bigram_title) %>% 
+    group_by(word) %>% 
+    summarise(n = unique(n),
+              pkgs = list(package),
+              total = sum(total)) %>% 
+    filter(!is.na(word)) %>% 
+    arrange(desc(total))
+}
+```
+
+```{r ngrams, cache=tocache, dependson="fns"}
+bigram_title <- get_ngram(2, "Title")
+bigram_desc <- get_ngram(2, "Description")
+trigram_desc <- get_ngram(3, "Description")
+```
+
+* Not much difference between title and description.
+* Just go with description and combine bi & tri.
+
+```{r wordcloud, fig.height = 3.3}
+dldat_comb_bigram_title <- combine_dl(bigram_title, Title)
+dldat_comb_bigram_desc <- combine_dl(bigram_desc, Title)
+dldat_comb_trigram_desc <- combine_dl(trigram_desc, Title)
+
+set.seed(1)
+ggplot(rbind(dldat_comb_bigram_desc,
+             dldat_comb_trigram_desc),
+       aes(label = word, size = total, color = n)) +
+  geom_text_wordcloud(shape = "square",
+                      show.legend = TRUE) +
+  scale_size_area(max_size = 8) +
+  guides(size = FALSE) +
+  scale_color_continuous(breaks = 1:11) +
+  theme(legend.position = "bottom")
 ```

### 2021-03-05 paper.tex
wordcloud

@@
 Survey designs often, although not always, aim to collect observational
-data whilst experimental designs are exclusively center on experimental
+data whilst experimental designs exclusively center on experimental
 data. This dichotomisation, to a great extent, is seen in CRAN task
@@
 
-{\centering \includegraphics{paper_files/figure-latex/unnamed-chunk-1-1}
+{\centering \includegraphics{paper_files/figure-latex/dlplots-1}
 
@@
 
-\caption[The above graph shows the total download counts from 2019-02-09 to 2021-02-07 of DoE packages]{The above graph shows the total download counts from 2019-02-09 to 2021-02-07 of DoE packages.}\label{fig:unnamed-chunk-1}
+\caption[The above graph shows the total download counts from 2019-03-06 to 2021-03-04 of DoE packages]{The above graph shows the total download counts from 2019-03-06 to 2021-03-04 of DoE packages.}\label{fig:dlplots}
 \end{figure}
+\end{Schunk}
+
+\begin{itemize}
+\tightlist
+\item
+  Not much difference between title and description.
+\item
+  Just go with description and combine bi \& tri.
+\end{itemize}
+
+\begin{Schunk}
+
+\begin{center}\includegraphics{paper_files/figure-latex/wordcloud-1} \end{center}
+
 \end{Schunk}

### 2021-03-23 paper.Rmd
widely used pkg

@@
 The critical role of data collection is well captured in the expression "garbage in, garbage out" -- in other words, if the collected data is rubbish then no analysis, however complex it may be, can make something out of it.
-Methods for data collection can be dichotomised by the type of data collected -- namely, experimental or observational -- or alternatively, categorised into experimental design or survey design.
+Methods for data collection can be dichotomised by the type of data collected -- namely, experimental or observational -- or alternatively, categorised as experimental design (including quasi-experimental design) or survey design.
+This dichotomisation, to a great extent, is seen in CRAN task views where R-packages in experimental design are in \ctv{ExperimentalDesign} and R-packages in survey designs are in \ctv{OfficialStatistics}.
 Survey designs often, although not always, aim to collect observational data whilst experimental designs exclusively center on experimental data.
-This dichotomisation, to a great extent, is seen in CRAN task views where R-packages in experimental design are in \ctv{ExperimentalDesign} and R-packages in survey designs are in \ctv{OfficialStatistics}.
+This paper is concerned with the latter.
 
@@
 in Python [@python], only a handful of libraries that generate design of experiment exist (namely `pyDOE`, `pyDOE2`, `dexpy`, `experimenter` and `GPdoemd`) with limited outputs.
+Thus, the study of DoE packages is also revealing into the current status of the field of experimental design.
 
-# Explorative Analysis
+The paper is organisd as follows. Section @\ref{eda}
+
+# Explorative analysis {#eda}
 
@@
 
-# Software Design
+Figure \ref{fig:dlplots} suggests that the \CRANpkg{AlgDesign} followed by \CRANpkg{agricolae} are far more downloaded than any other DoE pacakges.
+It is worth noting that \CRANpkg{agricolae} imports \CRANpkg{AlgDesign} thus any download of \CRANpkg{agricolae} likely results also in a download of \CRANpkg{AlgDesign}.
 
-# Propsects and Discussion
+# Software design
+
+# Discussion
 
@@
 
-\CRANpkg{agricolae}
-
 ## Helping info to get started

### 2021-03-24 analysis-download-logs.Rmd
analysis files

---
title: "Download logs"
author: "Emi Tanaka"
date: "24/03/2021"
output: html_document
---

```{r library, include=FALSE}
library(tidyverse)
library(rvest)
library(glue)
library(lubridate)
library(plotly)
library(cranlogs)
library(ggbeeswarm)
knitr::opts_chunk$set(echo = FALSE, 
                      cache = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      cache.path = "cache/")
```

```{r cache, eval = FALSE}
fs::dir_ls("cache") %>% 
  str_subset(".RData$") %>% 
  walk(load)
```

```{r ctv}
doe_pkgs <- ctv:::.get_pkgs_from_ctv_or_repos("ExperimentalDesign", 
                                              repos = "http://cran.rstudio.com/")[[1]]
survey_pkgs <- ctv:::.get_pkgs_from_ctv_or_repos("OfficialStatistics", 
                                              repos = "http://cran.rstudio.com/")[[1]]
npkgs <- length(doe_pkgs)
```

```{r all_pkgs}
url <- "http://cran.rstudio.com/web/packages/packages.rds"
db <- readRDS(url(url)) %>% 
  as.data.frame()
```

```{r extra-pkgs}
exclude_pkgs <- c("argonR", "card", "crispRdesignR", "genotypeR", "genepi", 
                  "agridat", "desplot", "DiceView", "DOS", "DOS2", "EngrExpt")
cdoe_pkgs <- db %>% 
  filter(str_detect(Title, "[dD]esign") | Package %in% doe_pkgs) %>% 
  filter(!Title %in% exclude_pkgs) %>% 
  pull(Package) 
```

```{r cran-data}
end <- Sys.Date() - 2 # usually 1-2 days data are not available yet
start <- end - years(2) + days(2)
dldat <- cran_downloads(doe_pkgs, from = start, to = end)
```

```{r summary}
dldat_sum <- dldat %>% 
  group_by(package) %>% 
  summarise(total = sum(count)) %>% 
  arrange(desc(total)) %>% 
  mutate(rank = 1:n())
```

```{r graph}
g1 <- ggplot(dldat_sum, aes(x = "", y = total/1000, label = package, text = rank)) + 
  geom_beeswarm() +
  scale_y_log10(label = comma) + 
  labs(x = "",
       y = "Download counts in the last 2 year ('000s)") + 
  theme(axis.ticks.length.x = unit(0, "mm"))
ggplotly(g1)
```

```{r updates}
pkg_url <- "https://cran.r-project.org/web/packages/{pkg}/index.html"
pkg_archive <- "https://cran.r-project.org/src/contrib/Archive/{pkg}/"
pkg_updates <- map(doe_pkgs, function(pkg) {
    last_update <- read_html(glue(pkg_url)) %>% 
      html_table() %>% 
      .[[1]] %>% 
      filter(X1=="Published:") %>% 
      pull(X2) %>% 
      ymd()

    archive_dates <- tryCatch({ 
        read_html(glue(pkg_archive)) %>% 
          html_table() %>%
          .[[1]] %>% 
          pull(`Last modified`) %>% 
          ymd_hm() %>% 
          na.omit() %>% 
          as.Date()
      }, error = function(e) {
        NULL
      })
    c(archive_dates, last_update)
  })
names(pkg_updates) <- doe_pkgs

updates <- unlist(pkg_updates) %>% 
  enframe("package", "update") %>% 
  # unlist converts date to integers
  mutate(update = as.Date(update, origin = "1970-01-01"),
         # need to get rid of the numbers appended to pkg names
         package = str_extract(package, paste0(doe_pkgs, collapse="|"))) 
```

```{r first}
pkgsum <- updates %>% 
  group_by(package) %>% 
  summarise(first = min(update),
            nupdates = n()) %>% 
  left_join(dldat_sum, by = "package")

g2 <- ggplot(pkgsum, aes(first, total, label = package, text = rank, size = nupdates)) + 
  geom_point(alpha = 0.4) + 
  scale_y_log10() + 
  #geom_smooth() +
  NULL
ggplotly(g2)
```

* Packages released after 2017 September doesn't seem to be listed in CRAN Task View

```{r}
g3 <- ggplot(pkgsum, aes(nupdates, total, label = package, text = rank)) + 
  geom_point(alpha = 0.4) + 
  scale_y_log10() + 
  scale_x_log10() + 
  #geom_smooth() +
  NULL
ggplotly(g3)
```

### 2021-03-24 analysis-word-cloud.Rmd
analysis files

--- 2021-03-05 paper.Rmd
@@
 ---
-title: Current status and propsects of R-packages for the design of experiments
-author:
-  # see ?rjournal_article for more information
-  - name: Emi Tanaka
-    affiliation: Monash University
-    address:
-    - Monash University
-    - Clayton campus, VIC 3800, Australia
-    url: http://emitanaka.org/
-    orcid: 0000-0002-1455-259X
-    email:  emi.tanaka@monash.edu
-abstract: >
-    Re-running an experiment is generally costly and in some cases impossible due to limited resources, so the design of experiment plays a critical role in increasing the quality of experimental data.
-In this article I describe the current state of the R-packages for the design of experiments through textual analysis and download trends.
-I discuss also the software design of widely utilised R packages in the field of experimental design and conclude with discussion of some future prospects for the field.
-preamble: |
-  % Any extra LaTeX you need in the preamble
-
-# per R journal requirement, the bib filename should be the same as the output 
-# tex file. Don't forget to rename the bib file and change this example value.
-bibliography: paper.bib
-
-output: rticles::rjournal_article
+title: "Word Cloud"
+author: "Emi Tanaka"
+date: "24/03/2021"
+output: html_document
 ---
 
-```{r setup, cache = FALSE, include = FALSE}
+```{r setup, include=FALSE}
 library(tidyverse)
 library(rvest)
+library(glue)
+library(pluralize)
 library(lubridate)
-library(cranlogs)
-library(glue)
-library(scales)
-library(colorspace)
-library(tidytext)
-library(pluralize)
-library(kableExtra)
-library(igraph)
-library(ggraph)
-library(ggrepel)
-library(patchwork)
+library(plotly)
 library(ggwordcloud)
-tocache <- TRUE
 knitr::opts_chunk$set(echo = FALSE, 
-                      cache = FALSE,
-                      cache.path = "cache/",
-                      fig.align = 'center', 
-                      fig.pos = 'htbp', 
-                      fig.width = 6,
+                      cache = TRUE,
+                      warning = FALSE,
                       message = FALSE,
-                      warning = FALSE)
-
-theme_set(
-  theme(panel.background = element_rect(fill = NA),
-        panel.grid = element_line(color = "lightgray"),
-        axis.text = element_text(color = "black"),
-        axis.line = element_line(color = "black", size = 0.7),
-        axis.ticks.length = unit(1.4, "mm"),
-        axis.ticks = element_line(color = "black", size = 0.7),
-        axis.title = element_text(color = "black", face = "bold"),
-        strip.background = element_rect(color = "black",
-                                        fill = "black"),
-        strip.text = element_text(color = "white"),
-        plot.title.position = "plot",
-        plot.title = element_text(color = "black", hjust = 0)))
-```
-
-```{r ctv}
-doe_pkgs <- ctv:::.get_pkgs_from_ctv_or_repos("ExperimentalDesign", 
-                                              repos = "http://cran.rstudio.com/")[[1]]
-doe_pkgs <- c(doe_pkgs, "DeclareDesign")
-npkgs <- length(doe_pkgs)
-```
-
-```{r cran-data, cache = tocache}
-end <- Sys.Date() - 2 # usually 1-2 days data are not available yet
-start <- end - years(2) + days(2)
-dldat <- cran_downloads(doe_pkgs, from = start, to = end)
-```
-
-```{r cran-data2, cache = tocache}
-start2 <- end - years(5) + days(2)
-dldat2 <- cran_downloads(doe_pkgs, from = start2, to = end)
-```
-
-# Introduction
-
-The critical role of data collection is well captured in the expression "garbage in, garbage out" -- in other words, if the collected data is rubbish then no analysis, however complex it may be, can make something out of it.
-Methods for data collection can be dichotomised by the type of data collected -- namely, experimental or observational -- or alternatively, categorised into experimental design or survey design.
-Survey designs often, although not always, aim to collect observational data whilst experimental designs exclusively center on experimental data.
-This dichotomisation, to a great extent, is seen in CRAN task views where R-packages in experimental design are in \ctv{ExperimentalDesign} and R-packages in survey designs are in \ctv{OfficialStatistics}.
-
-In the CRAN task view of \ctv{ExperimentalDesign}, there are `r npkgs` R packages for experimental design and analysis of data from experiments, henceforth referred to as "DoE packages" in this paper.
-The sheer quantity and variation in the output experimental design in the R-packages are arguably unmatched with any other programming languages, e.g.
-in Python [@python], only a handful of libraries that generate design of experiment exist (namely `pyDOE`, `pyDOE2`, `dexpy`, `experimenter` and `GPdoemd`) with limited outputs.
-
-# Explorative Analysis
-
-```{r download-dist, dependson = "cran-data"}
-dldat_sum <- dldat %>% 
-  group_by(package) %>% 
-  summarise(total = sum(count)) %>% 
-  arrange(desc(total)) %>% 
-  mutate(rank = 1:n())
-
-dl1plot <- ggplot(dldat_sum, aes(x = "", y = total/1000)) + 
-  geom_violin() +
-  geom_boxplot(width = 0.1) + 
-  scale_y_log10(label = comma) + 
-  labs(x = "",
-       y = "Download counts in the last 2 year ('000s)", title = "(A)") + 
-  geom_text_repel(data = slice_max(dldat_sum, total, n = 13), 
-                  nudge_x = rep(c(0.2, -0.2, 0.2, -0.2, 0.15), c(3, 2, 2, 3, 3)),
-                  size = 4,
-                  aes(label = package)) + 
-  theme(axis.ticks.length.x = unit(0, "mm"))
-
-dldat_sum5 <- dldat2 %>% 
-  group_by(package) %>% 
-  summarise(total = sum(count)) %>% 
-  arrange(desc(total)) %>% 
-  mutate(rank = 1:n())
-
-top15 <- dldat_sum %>% 
-  left_join(dldat_sum5, by = "package") %>% 
-  slice_max(total.x + total.y, n = 15) %>% 
-  select(package, rank.x, rank.y) %>% 
-  pivot_longer(-package) %>% 
-  mutate(name = ifelse(name=="rank.x", "Last 2 years", "Last 5 years"))
-
-dl2plot <- ggplot(top15, aes(name, value, group = package)) +
-  geom_line() + 
-  geom_point() +
-  scale_y_reverse(breaks = 1:15) +
-  scale_x_discrete(expand = expansion(c(0.1, 0.5))) +
-  labs(x = "", y = "Ranking", title = "(B)") +
-  geom_text(data = filter(top15, name=="Last 5 years"),
-            aes(label = package), 
-            hjust = 0, 
-            nudge_x = 0.1,
-            size = 4) + 
-  theme(panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank())
-```
-
-```{r dlplots, fig.cap=glue("The above graph shows the total download counts from {start} to {end} of DoE packages."), fig.width = 6}
-dl1plot + dl2plot
-```
-
-```{r text, cache=tocache}
-url <- "http://cran.rstudio.com/web/packages/packages.rds"
-db <- readRDS(url(url)) %>% 
-  as.data.frame()
-
-doe_db <- db %>% 
-  filter(Package %in% doe_pkgs) %>% 
-  mutate(Description = str_replace_all(Description, "\n", " "),
-         Description = str_squish(Description),
-         Title = str_replace_all(Title, "\n", " "))
+                      cache.path = "cache/")
 ```
@@
 
-```{r ngrams, cache=tocache, dependson="fns"}
+```{r ngrams, }
 bigram_title <- get_ngram(2, "Title")
@@
 ```
-
-* Not much difference between title and description.
-* Just go with description and combine bi & tri.
 
@@
 ```
-
-## What are the common types of experimental designs?
-
-## How do packages interplay with each other?
-
-## Which packages are widely utilised?
-
-# Software Design
-
-# Propsects and Discussion
-
-\newpage
-
-```{r cran-titles}
-url <- "http://cran.rstudio.com/web/packages/packages.rds"
-db <- readRDS(url(url)) %>% 
-  as.data.frame()
-```
-
-```{r bigram}
-stop_words_ext <- c(stop_words$word, "doi")
-
-doe_db <- db %>% 
-  filter(Package %in% doe_pkgs) %>% 
-  mutate(Description = str_replace_all(Description, "\n", " "),
-         Description = str_squish(Description),
-         Title = str_replace_all(Title, "\n", " "))
-
-bigram_tab <- function(data, col) {
-  data %>% 
-    unnest_tokens(word, {{col}}, token = "ngrams", n = 2) %>% 
-    separate(word, c("word1", "word2"), sep = " ") %>% 
-    mutate(word1 = singularize(word1),
-           word2 = singularize(word2)) %>% 
-    # don't count the same bigram within the same package
-    distinct(Package, word1, word2) %>% 
-    filter(!word1 %in% stop_words_ext,
-           !word2 %in% stop_words_ext,
-           !str_detect(word1, "^[0-9.]+$"),
-           !str_detect(word2, "^[0-9.]+$")) %>% 
-    count(word1, word2, sort = TRUE)  
-}
-```
-
-```{r bigram-title, results="asis"}
-btitle <- bigram_tab(doe_db, Title) %>% 
-  filter(n > 3) %>% 
-  mutate(word = paste(word1, word2)) %>% 
-  select(word, n) %>% 
-  kbl(booktabs = TRUE, linesep = "")
-
-bdesc <- bigram_tab(doe_db, Description) %>% 
-  filter(n > 4) %>% 
-  mutate(word = paste(word1, word2)) %>% 
-  select(word, n) %>% 
-  kbl(booktabs = TRUE, linesep = "")
-
-cat(c("\\begin{table}[h] \\centering ", 
-      btitle,
-    "\\hspace{1cm} \\centering ",
-      bdesc,
-    "\\caption{My tables} \\end{table}")) 
-```
-
-```{r}
-doe_imports <- doe_db %>% 
-  mutate(Depends = str_replace_all(Depends, "\n", " "),
-         Depends = str_replace_all(Depends, fixed("("), " ("),
-         Imports = str_replace_all(Imports, "\n", " "),
-         Imports = str_replace_all(Imports, fixed("("), " ("),
-         imports = str_c(Depends, Imports, sep = ","),
-         imports = str_split(imports, ","),
-         imports = map(imports, ~{
-           str_squish(.x) %>% 
-             word() %>% 
-             .[.!=""]}
-         ),
-         imports_doe = map(imports, ~.x[.x %in% doe_pkgs])) %>% 
-  select(Package, imports_doe) %>% 
-  unnest_longer(imports_doe) %>% 
-  filter(!is.na(imports_doe)) %>% 
-  rename(from = imports_doe, to = Package) %>% 
-  select(from, to)
-```
-
-```{r doe-network, eval = FALSE}
-graph_from_data_frame(doe_imports) %>% 
-  ggraph(layout = 'fr') +
-  geom_edge_link(aes(start_cap = label_rect(node1.name),
-                     end_cap = label_rect(node2.name)), 
-                 arrow = arrow(length = unit(2, 'mm')),
-                 color = "#79003e") + 
-  geom_node_text(aes(label = name),
-                 color = "#79003e") +
-  theme(panel.background = element_rect(fill = "#f6e5ee",
-                                        color = "#79003e"),
-        plot.margin = margin(20, 20, 20, 20))
-```
-
-\CRANpkg{agricolae}
-
-## Helping info to get started
-
-Introductory section which may include references in parentheses
-[@R], or cite a reference such as @R in the text.
-
-## Section title in sentence case
-
-Let's check fi this works Figure \@ref(fig:Rlogo).
-
-This section may contain a figure such as Figure \ref{fig:Rlogo}.
-
-## Summary
-
-This file is only a basic article template.
-For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](https://journal.r-project.org/share/author-guide.pdf).
-
-### About this format and the R Journal requirements
-
-`rticles::rjournal_article` will help you build the correct files requirements:
-
-* A R file will be generated automatically using `knitr::purl` - see
-https://bookdown.org/yihui/rmarkdown-cookbook/purl.html for more information.
-* A tex file will be generated from this Rmd file and correctly included in
-`RJwapper.tex` as expected to build `RJwrapper.pdf`.
-* All figure files will be kept in the default rmarkdown `*_files` folder. This
-happens because `keep_tex = TRUE` by default in `rticles::rjournal_article`
-* Only the bib filename is to modifed. An example bib file is included in the
-template (`RJreferences.bib`) and you will have to name your bib file as the
-tex, R, and pdf files.

### 2021-03-30 analysis-pkg-selection.Rmd
selection of rpkgs

---
title: "Selection of R-packages for review"
author: "Emi Tanaka"
date: "24/03/2021"
output: html_document
---

```{r library, include=FALSE}
library(tidyverse)
library(magrittr)
library(glue)
library(kableExtra)
library(lubridate)
knitr::opts_chunk$set(echo = FALSE, 
                      cache = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      cache.path = "cache/")
```

```{r cache, eval = FALSE}
fs::dir_ls("cache") %>% 
  str_subset(".RData$") %>% 
  walk(load)
```

```{r ctv}
doe_pkgs <- ctv:::.get_pkgs_from_ctv_or_repos("ExperimentalDesign", 
                                              repos = "http://cran.rstudio.com/")[[1]]
survey_pkgs <- ctv:::.get_pkgs_from_ctv_or_repos("OfficialStatistics", 
                                              repos = "http://cran.rstudio.com/")[[1]]
npkgs <- length(doe_pkgs)
```

```{r all_pkgs}
url <- "http://cran.rstudio.com/web/packages/packages.rds"
db <- readRDS(url(url)) %>% 
  as.data.frame()%>% 
  mutate(Description = str_replace_all(Description, "\n", " "),
         Description = str_squish(Description),
         Title = str_replace_all(Title, "\n", " "))
```

```{r extra-pkgs}
#exclude_pkgs <- c("argonR", "card", "crispRdesignR", "genotypeR", "genepi", "recipes", "agridat", "desplot", "DiceView", "DOS", "DOS2", "EngrExpt")
cdoe_pkgs <- db %>% 
  filter(str_detect(Title, "[dD]esign") | Package %in% doe_pkgs) %>% 
  #filter(!Title %in% exclude_pkgs) %>% 
  pull(Package) 
```

```{r}
# installation issues
exclude_pkgs <- c("osmgeosample", "secrdesign", "tanaka", "bioOED", "doebioresearch", "ssizeRNA", "SurvGSD")
design_pkgs <- db %>% 
  filter(str_detect(Title, "[dD]esign") | 
           Package %in% doe_pkgs) %>% 
  mutate(taskview_doe = ifelse(Package %in% doe_pkgs, "yes", "no"),
         taskview_survey = ifelse(Package %in% survey_pkgs, "yes", "no"),
         criteria = case_when(str_detect(tolower(paste(Title, Description)),
                                         "(experiment|block design|factorial design|clinical trial|optimal design|response surface design|stepped wedge design|split-plot design|phase i|crossover design|cluster randomi[sz]ed|non-inferiority trial)") ~ "DoE",
                              TRUE ~ "No")) %>% 
  arrange(desc(criteria)) %>% 
  dplyr::select(Package, Title, Description, taskview_doe, taskview_survey, criteria) %>% 
  filter(!Package %in% exclude_pkgs)

write_csv(design_pkgs, file = "output/design_classification.csv")
```
```{r, include = FALSE, cache = FALSE}
pacman::p_load(char = design_pkgs$Package, install = FALSE, update = FALSE)
```

```{css, echo = FALSE}
.tooltip {
  position: relative;
  display: inline-block;
  border-bottom: 1px dotted black;
  opacity: 1;
}

.tooltip .tooltiptext {
  visibility: hidden;
  width: 120px;
  background-color: black;
  color: #fff;
  text-align: center;
  border-radius: 6px;
  padding: 5px 0;

  /* Position the tooltip */
  position: absolute;
  z-index: 1;
}

.tooltip:hover .tooltiptext {
  visibility: visible;
}
```

```{r}
# https://stackoverflow.com/questions/8918753/r-help-page-as-object
Rd2list <- function(Rd){
    names(Rd) <- substring(sapply(Rd, attr, "Rd_tag"),2)
    temp_args <- Rd$arguments

    Rd$arguments <- NULL
    myrd <- lapply(Rd, unlist)
    myrd <- lapply(myrd, paste, collapse="")

    temp_args <- temp_args[sapply(temp_args , attr, "Rd_tag") == "\\item"]
    temp_args <- lapply(temp_args, lapply, paste, collapse="")
    temp_args <- lapply(temp_args, "names<-", c("arg", "description"))
    myrd$arguments <- temp_args
    return(myrd)
}

getHelpList <- function(fn, pkg){
    thefile <- do.call("help", list(topic = fn, package = pkg))
    myrd <- utils:::.getHelpFile(thefile)
    Rd2list(myrd)$description %>% 
      str_replace_all("\\n", "") %>% 
      str_squish() %>% 
      str_trim() 
}
```

```{r}
design_pkgs %>% 
    filter(criteria == "DoE") %>% 
    mutate(fns = map(Package, ~as.character(lsf.str(glue("package:{.x}"))))) %>%
    mutate(func = map_chr(fns, ~paste(glue("<div class='tooltip'>{.x}<span class='tooltiptext'>getHelpList(x, {Package})</span></div>"), collapse = ", ")))   %>% 
  dplyr::select(-fns, -criteria) %>% 
  DT::datatable(escape = FALSE)

design_pkgs %>% 
  filter(criteria == "No") %>% 
  mutate(fns = map(Package, ~lsf.str(glue("package:{.x}")))) %>% 
  mutate(func = glue_collapse(map_chr(fns, function(x) glue("<div class='tooltip'>{x}<span class='tooltiptext'>{getHelpList(x, Package)}</span></div>")), ", ")) %>% 
  dplyr::select(-fns, -criteria) %>% 
  DT::datatable(escape = FALSE)

```

```{r}
fdoe_pkgs <- design_pkgs %>% 
  filter(criteria=="DoE") %>% 
  pull(Package)
```

### 2021-08-01 analysis/analysis-pkg-selection.Rmd
paper folder

---
title: "Selection of R-packages for review"
author: "Emi Tanaka"
date: "24/03/2021"
output: html_document
---

```{r library, include=FALSE}
library(tidyverse)
library(magrittr)
library(glue)
library(kableExtra)
library(lubridate)
knitr::opts_chunk$set(echo = FALSE, 
                      cache = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      cache.path = "cache/")
```

```{r cache, eval = FALSE}
fs::dir_ls("cache") %>% 
  str_subset(".RData$") %>% 
  walk(load)
```

```{r ctv}
doe_pkgs <- ctv:::.get_pkgs_from_ctv_or_repos("ExperimentalDesign", 
                                              repos = "http://cran.rstudio.com/")[[1]]
survey_pkgs <- ctv:::.get_pkgs_from_ctv_or_repos("OfficialStatistics", 
                                              repos = "http://cran.rstudio.com/")[[1]]
npkgs <- length(doe_pkgs)
```

```{r all_pkgs}
url <- "http://cran.rstudio.com/web/packages/packages.rds"
db <- readRDS(url(url)) %>% 
  as.data.frame()%>% 
  mutate(Description = str_replace_all(Description, "\n", " "),
         Description = str_squish(Description),
         Title = str_replace_all(Title, "\n", " "))
```

```{r extra-pkgs}
#exclude_pkgs <- c("argonR", "card", "crispRdesignR", "genotypeR", "genepi", "recipes", "agridat", "desplot", "DiceView", "DOS", "DOS2", "EngrExpt")
cdoe_pkgs <- db %>% 
  filter(str_detect(Title, "[dD]esign") | Package %in% doe_pkgs) %>% 
  #filter(!Title %in% exclude_pkgs) %>% 
  pull(Package) 
```

```{r}
# installation issues
exclude_pkgs <- c("osmgeosample", "secrdesign", "tanaka", "bioOED", "doebioresearch", "ssizeRNA", "SurvGSD", "RcmdrPlugin.SCDA")
design_pkgs <- db %>% 
  filter(str_detect(Title, "[dD]esign") | 
           Package %in% doe_pkgs) %>% 
  mutate(taskview_doe = ifelse(Package %in% doe_pkgs, "yes", "no"),
         taskview_survey = ifelse(Package %in% survey_pkgs, "yes", "no"),
         criteria = case_when(str_detect(tolower(paste(Title, Description)),
                                         "(experiment|block design|factorial design|clinical trial|optimal design|response surface design|stepped wedge design|split-plot design|phase i|crossover design|cluster randomi[sz]ed|non-inferiority trial|supersaturated design|resolvable design|sequential design|sequential enrichment design|space-filling design|latin hypercube|dose[- ]finding|dose-escalation)") ~ "DoE",
                              TRUE ~ "No")) %>% 
  arrange(desc(criteria)) %>% 
  dplyr::select(Package, Title, Description, taskview_doe, taskview_survey, criteria) %>% 
  filter(!Package %in% exclude_pkgs)

write_csv(design_pkgs, file = "output/design_classification.csv")
```

```{r, include = FALSE, cache = FALSE}
pacman::p_load(char = design_pkgs$Package, install = FALSE, update = FALSE)
```

```{css, echo = FALSE}
.tooltip {
  position: relative;
  display: inline-block;
  border-bottom: 1px dotted black;
  opacity: 1;
}

.tooltip .tooltiptext {
  visibility: hidden;
  width: 120px;
  background-color: black;
  color: #fff;
  text-align: center;
  border-radius: 6px;
  padding: 5px 0;

  /* Position the tooltip */
  position: absolute;
  z-index: 1;
}

.tooltip:hover .tooltiptext {
  visibility: visible;
}
```

```{r}
# https://stackoverflow.com/questions/8918753/r-help-page-as-object
Rd2list <- function(Rd){
    names(Rd) <- substring(sapply(Rd, attr, "Rd_tag"),2)
    temp_args <- Rd$arguments

    Rd$arguments <- NULL
    myrd <- lapply(Rd, unlist)
    myrd <- lapply(myrd, paste, collapse="")

    temp_args <- temp_args[sapply(temp_args , attr, "Rd_tag") == "\\item"]
    temp_args <- lapply(temp_args, lapply, paste, collapse="")
    temp_args <- lapply(temp_args, "names<-", c("arg", "description"))
    myrd$arguments <- temp_args
    return(myrd)
}

getHelpList <- function(fns, pkg){
  map_chr(fns, function(fn) {
    thefile <- do.call("help", list(topic = fn, package = pkg))
    tryCatch(myrd <- utils:::.getHelpFile(thefile),
             error = function(...) "")
    if(exists("myrd")) {
      Rd2list(myrd)$description %>% 
        str_replace_all("\\n", "") %>% 
        str_squish() %>% 
        str_trim()     
    } else {
      ""
    }
  })
}
```

## Results {.tabset}

### Adaptive

```{r}
exclude_pkgs2 <- c(
  "augmentedRCBD", # analysis only
  "BayesCTDesign" # power and sample size calcuation only
)
include_pkgs <- c(
  "binseqtest", # Exact Binary Sequential Designs and Analysis
  "esDesign",
  "factorial2x2",
  "FSSF",
  "PharmPow"
)

design_pkgs %>% 
    filter(criteria == "DoE") %>% 
    filter(str_detect(tolower(Description), "sequential|adaptive|sample size")) %>% 
    mutate(fns = map(Package, ~as.character(lsf.str(glue("package:{.x}"))))) %>%
    mutate(func = map2_chr(fns, Package, ~paste0("<div class='tooltip'>", .x, "<span class='tooltiptext'>", getHelpList(.x, .y), "</span></div>", collapse = ", ")))  %>% 
  mutate(Package = glue("<a href='https://cran.r-project.org/web/packages/{Package}/index.html'>{Package}</a>")) %>% 
  dplyr::select(-fns, -criteria) %>% 
  DT::datatable(escape = FALSE)
```

### DoE

```{r}
design_pkgs %>% 
    filter(criteria == "DoE") %>% 
    filter(!str_detect(tolower(Description), "sequential |adaptive|sample size")) %>% 
    mutate(fns = map(Package, ~as.character(lsf.str(glue("package:{.x}"))))) %>%
    mutate(func = map2_chr(fns, Package, ~paste0("<div class='tooltip'>", .x, "<span class='tooltiptext'>", getHelpList(.x, .y), "</span></div>", collapse = ", ")))  %>% 
  mutate(Package = glue("<a href='https://cran.r-project.org/web/packages/{Package}/index.html'>{Package}</a>")) %>% 
  dplyr::select(-fns, -criteria) %>% 
  DT::datatable(escape = FALSE)
```
### Nope

```{r}
design_pkgs %>% 
  filter(criteria == "No") %>% 
    mutate(fns = map(Package, ~as.character(lsf.str(glue("package:{.x}"))))) %>%
    mutate(func = map2_chr(fns, Package, ~paste0("<div class='tooltip'>", .x, "<span class='tooltiptext'>", getHelpList(.x, .y), "</span></div>", collapse = ", ")))  %>% 
    mutate(Package = glue("<a href='https://cran.r-project.org/web/packages/{Package}/index.html'>{Package}</a>")) %>% 
  dplyr::select(-fns, -criteria) %>% 
  DT::datatable(escape = FALSE)

```

## {-}

```{r}
fdoe_pkgs <- design_pkgs %>% 
  filter(criteria=="DoE") %>% 
  pull(Package)
```

### 2021-08-01 paper/index.Rmd
paper folder

--- 2021-03-23 paper.Rmd
@@
 author:
-  # see ?rjournal_article for more information
   - name: Emi Tanaka
     affiliation: Monash University
-    address:
-    - Monash University
-    - Clayton campus, VIC 3800, Australia
-    url: http://emitanaka.org/
-    orcid: 0000-0002-1455-259X
-    email:  emi.tanaka@monash.edu
 abstract: >
@@
 
-output: rticles::rjournal_article
+output: 
+  bookdown::html_document2:
+    theme: paper
 ---
@@
 
-Figure \ref{fig:dlplots} suggests that the \CRANpkg{AlgDesign} followed by \CRANpkg{agricolae} are far more downloaded than any other DoE pacakges.
+Figure \ref{fig:dlplots} suggests that the \CRANpkg{AlgDesign} followed by \CRANpkg{agricolae} are far more downloaded than any other DoE packages.
 It is worth noting that \CRANpkg{agricolae} imports \CRANpkg{AlgDesign} thus any download of \CRANpkg{agricolae} likely results also in a download of \CRANpkg{AlgDesign}.
@@
 # Discussion
+
+An experiment involves running a number of steps:
+
+1. Defining the hypothesis or question
+2. Formulating methods to test hypothesis or answer question 
+3. Planning the experimental protocol:
+  * Determining experimental resources, 
+  * specifying the data collection procedure,
+  * constructing the experimental design layout, and
+  * proposing a model for analysis.
+4. Collect data 
+5. Analyse data, interpret result and conclusion
 
@@
 ```
-
-## Helping info to get started
-
-Introductory section which may include references in parentheses
-[@R], or cite a reference such as @R in the text.
-
-## Section title in sentence case
-
-Let's check fi this works Figure \@ref(fig:Rlogo).
-
-This section may contain a figure such as Figure \ref{fig:Rlogo}.
-
-## Summary
-
-This file is only a basic article template.
-For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](https://journal.r-project.org/share/author-guide.pdf).
-
-### About this format and the R Journal requirements
-
-`rticles::rjournal_article` will help you build the correct files requirements:
-
-* A R file will be generated automatically using `knitr::purl` - see
-https://bookdown.org/yihui/rmarkdown-cookbook/purl.html for more information.
-* A tex file will be generated from this Rmd file and correctly included in
-`RJwapper.tex` as expected to build `RJwrapper.pdf`.
-* All figure files will be kept in the default rmarkdown `*_files` folder. This
-happens because `keep_tex = TRUE` by default in `rticles::rjournal_article`
-* Only the bib filename is to modifed. An example bib file is included in the
-template (`RJreferences.bib`) and you will have to name your bib file as the
-tex, R, and pdf files.

### 2021-08-01 paper/rjournal/paper-rjournal.Rmd
paper folder

---
author:
  # see ?rjournal_article for more information
  # Dai Dai to add her own name here
  - name: Emi Tanaka
    affiliation: Monash University
    address:
    - Monash University
    - Clayton campus, VIC 3800, Australia
    url: http://emitanaka.org/
    orcid: 0000-0002-1455-259X
    email:  emi.tanaka@monash.edu
preamble: |
  % Any extra LaTeX you need in the preamble

# per R journal requirement, the bib filename should be the same as the output 
# tex file. Don't forget to rename the bib file and change this example value.
bibliography: ../paper.bib

output: rticles::rjournal_article
---

```{r, include = FALSE}
paper <- here::here("paper/index.Rmd")
meta <- rmarkdown::yaml_front_matter(paper)
```

---
title: "`r meta$title`"
abstract: >
  `r meta$abstract`
---

```{r child=paper}
```

### 2021-08-01 analysis/d01-rpkg-info.R
d01 rpkg info

@@ -1,0 +1,52 @@
+library(cranlogs)
+library(tidyverse)
+library(yaml)
+library(here)
+url <- "http://cran.rstudio.com/web/packages/packages.rds"
+db <- readRDS(url(url)) %>% 
+  as.data.frame()%>% 
+  mutate(Description = str_replace_all(Description, "\n", " "),
+         Description = str_squish(Description),
+         Title = str_replace_all(Title, "\n", " "))
+
+doe_pkgs <- ctv:::.get_pkgs_from_ctv_or_repos("ExperimentalDesign", 
+                                              repos = "http://cran.rstudio.com/")[[1]]
+dl <- cran_downloads(packages = doe_pkgs,
+                     from = "2020-01-01",
+                     to = "2020-12-31")
+dlsum <- dl %>% 
+  group_by(package) %>% 
+  summarise(total2020 = sum(count)) %>% 
+  arrange(desc(total2020)) %>% 
+  left_join(select(db, 
+                   package = Package, 
+                   depends = Depends,
+                   imports = Imports,
+                   suggest = Suggests,
+                   author = Author,
+                   date = Date,
+                   version = Version,
+                   title = Title,
+                   description = Description),
+            by = "package") %>% 
+  mutate(author = str_replace_all(author, "(\\(.+\\)|\\n)", ""),
+         author = str_squish(author))
+
+out <- pmap(dlsum, function(package, total2020, depends, imports, suggests, author, date, version, title, description) {
+  x <- list(list(download2020 = as.integer(total2020),
+                 depends = ifelse(is.na(depends), "", depends), 
+                 imports = ifelse(is.na(imports), "", imports), 
+                 suggests = ifelse(is.na(suggests), "", suggests), 
+                 author = ifelse(is.na(author), "", author), 
+                 date = ifelse(is.na(date), "", date), 
+                 version = ifelse(is.na(version), "", version), 
+                 title = ifelse(is.na(title), "", title), 
+                 description = ifelse(is.na(description), "", description),
+                 type = '',
+                 main = ''))
+  names(x) <- package
+  x
+})
+saveRDS(out, file = here("data/derived/doe-pkgs.rds"))
+out <- readRDS(here::here("data/derived/doe-pkgs.rds"))
+yaml::write_yaml(out, here::here("data/input/doe-pkgs.yml"))

### 2021-08-01 analysis/d01-rpkg-info.R
original - as not to accidentally overwrite

@@ -49,4 +49,4 @@ out <- pmap(dlsum, function(package, total2020, depends, imports, suggests, auth
 })
 saveRDS(out, file = here("data/derived/doe-pkgs.rds"))
 out <- readRDS(here::here("data/derived/doe-pkgs.rds"))
-yaml::write_yaml(out, here::here("data/input/doe-pkgs.yml"))
+yaml::write_yaml(out, here::here("data/input/doe-pkgs-original.yml"))

### 2021-08-01 analysis/d01-rpkg-info.R
data column

@@ -43,7 +43,8 @@ out <- pmap(dlsum, function(package, total2020, depends, imports, suggests, auth
                  title = ifelse(is.na(title), "", title), 
                  description = ifelse(is.na(description), "", description),
                  type = '',
-                 main = ''))
+                 main = '',
+                 data = ''))
   names(x) <- package
   x
 })

### 2021-08-16 analysis/doe-pkgs-analysis.Rmd
Transform the yaml into dataframe.

---
title: "doe-pkgs-analysis"
author: "Dewi Lestari Amaliah - 31251587"
date: "15/08/2021"
output: html_document
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(yaml)
library(reshape)
library(tidyverse)
library(lubridate)
```

```{r}
#load the yml file and transform it as data frame
pkg_df <- yaml.load_file(here::here("data/input/doe-pkgs.yml")) %>%
  melt()

#tidy the data frame
pkg_tidy <- pkg_df %>%
  pivot_wider(names_from = L3,
              values_from = value)

# pivot longer the data frame by its type 
pkg_tidy_long_type <- pkg_tidy %>%
  unnest(type) %>%
  unnest(date) %>%
  unnest(domain) %>%
  mutate(year = year(ymd(date)))
```

```{r}
# summarise the packages based on its type
pkg_tidy_long_type %>%
  filter(relevance != "no") %>%
  group_by(type) %>%
  count() %>%
  arrange(desc(n))
```
```{r}
#summarise the package based on its date of published 
pkg_tidy_long_type %>%
  filter(relevance != "no") %>%
  group_by(year) %>%
  count() %>%
  ggplot(aes(x = year, y = n)) +
  geom_line() +
  geom_point()
```

```{r}
pkg_tidy_long_type %>%
  filter(relevance != "no") %>%
  group_by(domain) %>%
  count() %>%
  arrange(desc(n))
```

```{r}
pkg_tidy_long_main <- pkg_tidy %>%
  unnest(main) %>%
  separate(main, into = c("funct_name", "funct_type", "funct_usage"), sep = ' - ')
```

```{r}
pkg_tidy_long_main %>%
  filter(relevance != "no") %>%
  group_by(funct_type) %>%
  count() %>%
  arrange(desc(n))
```

### 2021-08-29 analysis/doe-pkgs-analysis.Rmd
Add network graph of the packages.

@@
 library(lubridate)
+library(splitstackshape)
+library(igraph)
+library(geomnet)
+library(ggnetwork)
+library(tidygraph)
+library(ggraph)
 ```
@@
 ```
+## Network analysis
+
+### Create the depends/imports/suggest data frame
+
+```{r}
+# make list of doe package 
+list_pkg <- as.list(pkg_tidy)$L2
+
+# create function to extract the depends, imports, and suggests 
+get_pkg <- function(var){
+  # select only the package name and depends/imports/suggests
+  tmp_df <- pkg_tidy %>%
+    select(L2, var) %>%
+    # separate every depends/imports/suggests
+    cSplit(var, sep = " ,")
+
+  df <- tmp_df %>%
+    pivot_longer(c(2:ncol(tmp_df)), names_to = "type", values_to = "pkg") %>%
+    filter(!is.na(pkg)) %>%
+    # check if the depends/imports/suggests is also a DoE package (is in list_pkg)
+    mutate(doe_pkg_related = ifelse(pkg %in% list_pkg, "yes", "no")) %>%
+    # filter only package whose depends/imports/suggests is a DoE
+    filter(doe_pkg_related == "yes") %>%
+    # tidy the data to only have the type: depends/imports/suggests
+    cSplit("type", sep = "_") %>%
+    select(-type_2) %>%
+    rename(type = type_1)
+
+  return(df)
+}
+```
+
+### Create network data
+
+```{r}
+
+# create data frame from the above function
+depend_df <- get_pkg("depends")
+import_df <- get_pkg("imports")
+suggest_df <- get_pkg("suggests")
+
+# bind data frames
+pkg_edges <- rbind(depend_df, import_df, suggest_df) %>%
+  select(-doe_pkg_related)
+
+# extract unique package that has depends/imports/suggests
+from <- pkg_edges %>%
+  distinct(L2)
+
+# extract unique package that being depends/imports/suggests
+to <- pkg_edges %>%
+  distinct(pkg)
+
+# create nodes from the two previous dataframe
+pkg_nodes <- full_join(from, to, by = c("L2" = "pkg"))
+
+pkg_nodes_full <- pkg_tidy %>%
+  select(L2)
+
+# create the network package 
+pkg_net <- pkg_edges %>%
+  full_join(pkg_nodes, by = "L2")
+```
+
+## Create network using geomnet
+
+```{r}
+set.seed(12345)
+ggplot(pkg_net, aes(from_id = L2, to_id = pkg)) +
+  geom_net(layout.alg = "kamadakawai",
+    size = 2, 
+    labelon = TRUE, 
+    vjust = -0.6, 
+    directed = FALSE, 
+    fontsize = 2)
+```
+
+This method results error, and only work with the edges only. However, even with the example data, the error stil occurs (may be the bug in the package?)
+
+```{r}
+## Example of issue in geomnet
+
+edges <- madmen$edges
+nodes <- madmen$vertices
+
+madmen_net <- fortify(as.edgedf(edges), nodes)
+
+set.seed(12345)
+ggplot(madmen_net, aes(from_id = from_id, to_id = to_id)) +
+  geom_net(layout.alg = "kamadakawai",
+    size = 2, 
+    labelon = TRUE, 
+    vjust = -0.6, 
+    directed = FALSE, 
+    fontsize = 2)
+```
+
+## Create network using igraph
+
+```{r}
+routes_igraph <- graph_from_data_frame(d = pkg_edges, vertices = pkg_nodes, directed = TRUE)
+
+set.seed(1234567)
+plot(routes_igraph, edge.arrow.size = 0.2, layout = layout_with_graphopt)
+```
+
+## Create network using tidygraph and ggraph
+
+```{r}
+routes_tidy <- routes_igraph %>% as_tbl_graph() %>%
+  activate(edges)
+
+routes_tidy
+```
+
+```{r}
+ggraph(routes_tidy, layout = "stress") + 
+  geom_edge_link(aes(color = type),
+                 arrow = arrow(length = unit(1, 'mm')), 
+                 end_cap = circle(1, 'mm')) + 
+  geom_node_point() + 
+  geom_node_text(aes(label = name), repel = TRUE, size = 2) +
+  theme_graph()
+```

### 2021-08-29 analysis/doe-pkgs-analysis.Rmd
Add heatmap.

@@
 
-This method results error, and only work with the edges only. However, even with the example data, the error stil occurs (may be the bug in the package?)
+Note: This method results error, and only works with the edges only. However, even with the example data, the error stil occurs (may be the bug in the package?)
 
@@
 ```
+Notes: The label overlapped, it is hard to see the network
 
@@
 ```{r}
+# create the input based on the igraph object
 routes_tidy <- routes_igraph %>% as_tbl_graph() %>%
@@
 ```
+
+# Network graph with all the packages in CRAN task view
+
+```{r}
+routes_igraph_full <- graph_from_data_frame(d = pkg_edges, vertices = pkg_nodes_full, directed = TRUE)
+
+routes_tidy_full <- routes_igraph_full %>% as_tbl_graph() %>%
+  activate(edges)
+
+routes_tidy_full
+```
+
+```{r}
+ggraph(routes_tidy_full, layout = "stress") + 
+  geom_edge_link(aes(color = type),
+                 arrow = arrow(length = unit(1, 'mm')), 
+                 end_cap = circle(1, 'mm')) + 
+  geom_node_point() + 
+  geom_node_text(aes(label = name), repel = TRUE, size = 2) +
+  theme_graph()
+```
+
+Note: Hard to read. Hence, with all of the package in CRAN task view, I will make a heatmap instead.
+
+```{r}
+pkg.mat = get.adjacency(routes_igraph_full, sparse = FALSE)
+heatmap(pkg.mat[,109:1],Rowv = NA, Colv = NA,scale="none")
+```

### 2021-09-06 analysis/doe-pkgs-analysis.Rmd
Added explanation of packages interconnection.

@@
 library(ggraph)
+library(networkD3)
+library(plotly)
+library(qdapRegex)
 ```
@@
   pivot_wider(names_from = L3,
-              values_from = value)
+              values_from = value) %>%
+  mutate(author_ulrike = as.factor(ifelse(str_detect(author, "Ulrike"), 1, 0)))
 
@@
   tmp_df <- pkg_tidy %>%
-    select(L2, var) %>%
+    select(L2, author_ulrike, var) %>%
     # separate every depends/imports/suggests
@@
   df <- tmp_df %>%
-    pivot_longer(c(2:ncol(tmp_df)), names_to = "type", values_to = "pkg") %>%
+    pivot_longer(c(3:ncol(tmp_df)), names_to = "type", values_to = "pkg") %>%
     filter(!is.na(pkg)) %>%
@@
 pkg_edges <- rbind(depend_df, import_df, suggest_df) %>%
-  select(-doe_pkg_related)
+  select(-doe_pkg_related) %>%
+  select(pkg, L2, type, author_ulrike)
 
@@
 # create nodes from the two previous dataframe
-pkg_nodes <- full_join(from, to, by = c("L2" = "pkg"))
+pkg_nodes <- full_join(from, to, by = c("L2" = "pkg")) %>%
+  left_join(pkg_tidy, by = "L2") %>%
+  select(L2, author_ulrike)
 
-pkg_nodes_full <- pkg_tidy %>%
+pkg_nodes_full <- pkg_tidy%>%
   select(L2)
@@
                  end_cap = circle(1, 'mm')) + 
-  geom_node_point() + 
+  geom_node_point(aes(color = author_ulrike)) + 
+  scale_color_manual(values = c("#707B7C", "#F5B041"), labels = c("no", "yes")) +
   geom_node_text(aes(label = name), repel = TRUE, size = 2) +
   theme_graph()
+
 ```
@@
 ```
+
+Next:
+1. Color the nodes based on author (Binary variable whether the pkg is authored by Ulrike or not)
+2. Using interactive graph with D3
+3. Write on the paper
+
+# Interactive Plot
+
+```{r}
+nodes_d3 <- pkg_nodes %>%
+  rowid_to_column("id") %>%
+  mutate(id = id - 1)
+
+edges_d3 <-routes_tidy %>%
+  activate(edges) %>%
+  data.frame() %>%
+  mutate(from = from - 1,
+         to = to - 1)
+
+forceNetwork(Links = edges_d3, Nodes = nodes_d3, Source = "from", Target = "to", 
+             NodeID = "L2", Group = "author_ulrike", 
+             opacity = 1, fontSize = 16, zoom = TRUE)
+```

### 2021-09-06 paper/package_interconnection.Rmd
Added explanation of packages interconnection.

---
title: "Package Interconnection"
author: "Dewi Lestari Amaliah"
date: "03/09/2021"
output: html_document
---

```{r setup-section4, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(yaml)
library(reshape)
library(tidyverse)
library(lubridate)
library(splitstackshape)
library(igraph)
library(tidygraph)
library(ggraph)
library(networkD3)
```

```{r network-data}
pkg_df <- yaml.load_file(here::here("data/input/doe-pkgs.yml")) %>%
  melt()

#tidy the data frame
pkg_tidy <- pkg_df %>%
  pivot_wider(names_from = L3,
              values_from = value) %>%
  mutate(author_ulrike = as.factor(ifelse(str_detect(author, "Ulrike"), 1, 0)))

# make list of doe package 
list_pkg <- as.list(pkg_tidy)$L2

# create function to extract the depends, imports, and suggests 
get_pkg <- function(var){
  # select only the package name and depends/imports/suggests
  tmp_df <- pkg_tidy %>%
    select(L2, author_ulrike, var) %>%
    # separate every depends/imports/suggests
    cSplit(var, sep = " ,")

  df <- tmp_df %>%
    pivot_longer(c(3:ncol(tmp_df)), names_to = "type", values_to = "pkg") %>%
    filter(!is.na(pkg)) %>%
    # check if the depends/imports/suggests is also a DoE package (is in list_pkg)
    mutate(doe_pkg_related = ifelse(pkg %in% list_pkg, "yes", "no")) %>%
    # filter only package whose depends/imports/suggests is a DoE
    filter(doe_pkg_related == "yes") %>%
    # tidy the data to only have the type: depends/imports/suggests
    cSplit("type", sep = "_") %>%
    select(-type_2) %>%
    rename(type = type_1)

  return(df)
}

# create data frame from the above function
depend_df <- get_pkg("depends")
import_df <- get_pkg("imports")
suggest_df <- get_pkg("suggests")

# bind data frames
pkg_edges <- rbind(depend_df, import_df, suggest_df) %>%
  select(-doe_pkg_related) %>%
  select(pkg, L2, type, author_ulrike)

# extract unique package that has depends/imports/suggests
from <- pkg_edges %>%
  distinct(L2)

# extract unique package that being depends/imports/suggests
to <- pkg_edges %>%
  distinct(pkg)

# create nodes from the two previous dataframe
pkg_nodes <- full_join(from, to, by = c("L2" = "pkg")) %>%
  left_join(pkg_tidy, by = "L2") %>%
  select(L2, author_ulrike)

# create igraph object
routes_igraph <- graph_from_data_frame(d = pkg_edges, vertices = pkg_nodes, directed = TRUE)

# create the input based on the igraph object
routes_tidy <- routes_igraph %>% as_tbl_graph() %>%
  activate(edges)
```

Interaction between packages could indicate of how developers utilize the functionality of one or more packages to create new features in their packages. According to @StanstrupJan2019TmTi, dependency and connectivity in terms of reusing codes and inheriting objects could reflect a well-connected and interacting community. In this paper, we observe the interaction limited to only between packages that are listed as DoE packages. Further, the type of interaction include the depended, imported, and suggested packages. Out of `r nrow(pkg_tidy)`, there are only `r nrow(pkg_nodes)` packages that interplay each other. To observe these interaction, we use network visualisation using `igraph` [@igraph], `tidygraph` [@tidygraph], and `ggraph` [@ggraph].

```{r network-graph, fig.cap = "Interaction network between DoE R packages. The color of edges reflect the type of interaction, i.e., depends, imports, and suggests. Nodes' color indicates wether the package is authored by Ulrike Groumping or not."}
# create the graph
ggraph(routes_tidy, layout = "stress") + 
  geom_edge_link(aes(color = type),
                 arrow = arrow(length = unit(1, 'mm')), 
                 end_cap = circle(1, 'mm')) + 
  geom_node_point(aes(color = author_ulrike)) + 
  scale_color_manual(values = c("#707B7C", "#F5B041"), labels = c("no", "yes")) +
  geom_node_text(aes(label = name), repel = TRUE, size = 2) +
  theme_graph()
```
Figure \@fig(network-graph) shows the interplay of DoE packages. In the previous section, we noticed that  Ulrike Groemping authored many DoE packages. We also noticed that one of her package, DoE base, is among the most downloaded DoE package on CRAN. Hence, we categorise the package whether it is authored by her or not. The network conveys that there are 5 clusters of interaction with one big cluster.

The most downloaded DoE packages, such as `AlgDesign`, `rsm`, and `lhs` are observed to support many other packages. Dependent packages seems to use their functions to support their feature. We also observe that  `DoE.base` and `FrF2`, which are authored by Groemping, are also imported, depended, and suggested by many other packages because of their basic functionality in experimental design. Moreover, some of interacting packages, such as `DoE.wrapper` and `DoE.MIParray` are also developed by Groemping.

Further, packages regarding computer experiment (`DiceKriging`, `DiceEval`, `DiceView`, and `DiceDesign`) happen to interplay with each other so that they appear to form a sub-cluster. The four smaller clusters interact because of the similarity in their type and domain. For example `dfcm` and `dfpk` are used for dose-finding experiments. Regarding the similarity of its objective, although `FMC` aims to generate factorial design, this package does not interact with the packages in the big cluster, even though some packages in this cluster, such as `DoE.base`, also involve factorial design. This package instead imports `minimalRSD`, which is developed by the same author.

Based on this network, we can say that the interconnection of DoE package is due to functionality dependence, similarity of type or domain, and because they are developed by the same author/s.

```{r}
#create nodes for interactive network
nodes_d3 <- pkg_nodes %>%
  rowid_to_column("id") %>%
  mutate(id = id - 1)

#create edges for interactive network
edges_d3 <-routes_tidy %>%
  activate(edges) %>%
  data.frame() %>%
  mutate(from = from - 1,
         to = to - 1)

# interactive network
forceNetwork(Links = edges_d3, Nodes = nodes_d3, Source = "from", Target = "to", 
             NodeID = "L2", Group = "author_ulrike", 
             opacity = 1, fontSize = 16, zoom = TRUE)
```

# Create the categorisation based on main author 
# create the network by facetting by author 
# if there are many authors then create the facet for example if the author write >= 3 packages. 
# not only focus in the Ulrike alone

### 2021-09-10 analysis/doe-pkgs-analysis.Rmd
tidy the author data.

@@
   tmp_df <- pkg_tidy %>%
-    select(L2, author_ulrike, var) %>%
+    select(L2, author, var) %>%
     # separate every depends/imports/suggests
@@
   select(-doe_pkg_related) %>%
-  select(pkg, L2, type, author_ulrike)
+  select(pkg, L2, type, author)
 
@@
 # create nodes from the two previous dataframe
-pkg_nodes <- full_join(from, to, by = c("L2" = "pkg")) %>%
+pkg_nodes_tmp <- full_join(from, to, by = c("L2" = "pkg")) %>%
   left_join(pkg_tidy, by = "L2") %>%
-  select(L2, author_ulrike)
+  select(L2, author)
 
+# tidy author 
+aut_tidy <- pkg_nodes_tmp %>%
+  separate_rows(author, sep=", (?=[A-Z])") %>%
+  filter(!str_detect(author, "ctb") & !str_detect(author, "National") & !str_detect(author, "Columbia"))
+
+aut_tidy$author <- rm_angle(aut_tidy$author, replacement = "")
+aut_tidy$author <- rm_square(aut_tidy$author, replacement = "")
+
+aut_tidy <- aut_tidy %>%
+  mutate(author = ifelse(str_detect(author, "Contributor"), "Yves Deville",
+                         ifelse(str_detect(author, "based"), "Ernesto Barrios",
+                                ifelse(str_detect(author, "contributions"), "John Lawson", author))))
+
+aut_group <- aut_tidy %>%
+  group_by(author) %>%
+  count() %>%
+  arrange(desc(n)) %>%
+  mutate(aut_gr = as.factor(ifelse(n > 1, author, "other")))
+
+aut_tidy_join <- left_join(aut_tidy, aut_group, by = "author") %>%
+  group_by(L2) %>%
+  arrange(desc(n), .by_group = TRUE)
+
+pkg_nodes <- aut_tidy_join[!duplicated(aut_tidy_join$L2), ] %>%
+  select(L2, aut_gr)
+
+# nodes for all of the author in DoE packages
 pkg_nodes_full <- pkg_tidy%>%
@@
 
-## Create network using geomnet
-
-```{r}
-set.seed(12345)
-ggplot(pkg_net, aes(from_id = L2, to_id = pkg)) +
-  geom_net(layout.alg = "kamadakawai",
-    size = 2, 
-    labelon = TRUE, 
-    vjust = -0.6, 
-    directed = FALSE, 
-    fontsize = 2)
-```
-
-Note: This method results error, and only works with the edges only. However, even with the example data, the error stil occurs (may be the bug in the package?)
-
-```{r}
-## Example of issue in geomnet
-
-edges <- madmen$edges
-nodes <- madmen$vertices
-
-madmen_net <- fortify(as.edgedf(edges), nodes)
-
-set.seed(12345)
-ggplot(madmen_net, aes(from_id = from_id, to_id = to_id)) +
-  geom_net(layout.alg = "kamadakawai",
-    size = 2, 
-    labelon = TRUE, 
-    vjust = -0.6, 
-    directed = FALSE, 
-    fontsize = 2)
-```
-
 ## Create network using igraph
@@
 
-set.seed(1234567)
-plot(routes_igraph, edge.arrow.size = 0.2, layout = layout_with_graphopt)
+# set.seed(1234567)
+# plot(routes_igraph, edge.arrow.size = 0.2, layout = layout_with_graphopt)
+# Notes: The label overlapped, it is hard to see the network
 ```
-Notes: The label overlapped, it is hard to see the network
 
@@
                  end_cap = circle(1, 'mm')) + 
-  geom_node_point(aes(color = author_ulrike)) + 
-  scale_color_manual(values = c("#707B7C", "#F5B041"), labels = c("no", "yes")) +
+  geom_node_point(aes(color = aut_gr)) + 
   geom_node_text(aes(label = name), repel = TRUE, size = 2) +
@@
 forceNetwork(Links = edges_d3, Nodes = nodes_d3, Source = "from", Target = "to", 
-             NodeID = "L2", Group = "author_ulrike", 
+             NodeID = "L2", Group = "aut_gr", 
              opacity = 1, fontSize = 16, zoom = TRUE)

### 2021-09-10 analysis/doe-pkgs-analysis.Rmd
Test some visualisation by author.

@@
 output: html_document
+runtime: shiny
 ---
@@
 library(qdapRegex)
+library(shiny)
+library(visNetwork)
 ```
@@
   theme_graph()
+```
 
+```{r}
+aut_tidy_1st_aut <- aut_tidy_join[!duplicated(aut_tidy_join$L2), ] %>%
+  select(L2, author)
+
+aut_gr <- aut_tidy_1st_aut %>%
+  group_by(author) %>%
+  count() %>%
+  ungroup() %>%
+  arrange(desc(n)) %>%
+  mutate(aut_gr = as.factor(ifelse(n > 1, author, "other")))
+
+pkg_nodes_1aut <- left_join(aut_tidy_1st_aut, aut_gr, by = "author") %>%
+  select(L2, aut_gr) %>%
+  mutate(first_aut = factor(aut_gr, levels = c("Ulrike Groemping",
+                                           "John Lawson",
+                                           "Robert B. Gramacy",
+                                           "Delphine Dupuy",
+                                           "Ernesto Barrios",
+                                           "Kevin Wright",
+                                           "Peng Liu",
+                                           "Shwetank Lall",
+                                           "other"))) %>%
+  select(L2, first_aut)
+
+routes_igraph_1aut <- graph_from_data_frame(d = pkg_edges, vertices = pkg_nodes_1aut, directed = TRUE)
+
+# create the input based on the igraph object
+routes_tidy_1aut <- routes_igraph_1aut %>% as_tbl_graph() %>%
+  activate(edges)
+
+ggraph(routes_tidy_1aut, layout = "stress") + 
+  geom_edge_link(aes(color = type),
+                 arrow = arrow(length = unit(1, 'mm')), 
+                 end_cap = circle(1, 'mm')) + 
+  geom_node_point(aes(color = first_aut)) + 
+  geom_node_text(aes(label = name), repel = TRUE, size = 2) +
+  theme_graph()
 ```
@@
 
-# Interactive Plot
+# Interactive Plot with visnetwork
 
 ```{r}
-nodes_d3 <- pkg_nodes %>%
+nodes_d3 <- pkg_nodes_1aut %>%
   rowid_to_column("id") %>%
@@
 
-edges_d3 <-routes_tidy %>%
+edges_d3 <-routes_tidy_1aut %>%
   activate(edges) %>%
@@
 forceNetwork(Links = edges_d3, Nodes = nodes_d3, Source = "from", Target = "to", 
-             NodeID = "L2", Group = "aut_gr", 
+             NodeID = "L2", Group = "first_aut", 
              opacity = 1, fontSize = 16, zoom = TRUE)
 ```
+
+```{r}
+# shinyApp(
+#   ui = fluidPage(
+#     selectInput("first_aut", "Choose the author:",
+#                     choices = levels(pkg_nodes_1aut$first_aut)),
+#       mainPanel(plotOutput(outputId = "net"))
+#     ),
+#   
+#   server = function(input, output){
+#     
+#     dat <- reactive({
+#       if (input$first_aut == "Ulrike Groemping"){
+#         aut <- pkg_nodes_1aut %>%
+#           mutate(author = as.factor(ifelse(first_aut == "Ulrike Groemping", "Ulrike Groemping", "Other")))
+#       }
+#       
+#       if (input$first_aut == "John Lawson"){
+#         aut <- pkg_nodes_1aut %>%
+#           mutate(author = as.factor(ifelse(first_aut == "John Lawson", "John Lawson", "Other")))
+#       }
+#       
+#       if (input$first_aut == "Robert B. Gramacy"){
+#         aut <- pkg_nodes_1aut %>%
+#           mutate(author = as.factor(ifelse(first_aut == "Robert B. Gramacy", "Robert B. Gramacy", "Other")))
+#       }
+#       
+#       if (input$first_aut == "Delphine Dupuy"){
+#         aut <- pkg_nodes_1aut %>%
+#           mutate(author = as.factor(ifelse(first_aut == "Delphine Dupuy", "Delphine Dupuy", "Other")))
+#       }
+#       
+#       if (input$first_aut == "Ernesto Barrios"){
+#         aut <- pkg_nodes_1aut %>%
+#           mutate(author = as.factor(ifelse(first_aut == "Ernesto Barrios", "Ernesto Barrios", "Other")))
+#       }
+#       
+#       if (input$first_aut == "Kevin Wright"){
+#         aut <- pkg_nodes_1aut %>%
+#           mutate(author = as.factor(ifelse(first_aut == "Kevin Wright", "Kevin Wright", "Other")))
+#       }
+#       
+#        if (input$first_aut == "Peng Liu"){
+#         aut <- pkg_nodes_1aut %>%
+#           mutate(author = as.factor(ifelse(first_aut == "Peng Liu", "Peng Liu", "Other")))
+#        }
+#       
+#       if (input$first_aut == "Shwetank Lall"){
+#         aut <- pkg_nodes_1aut %>%
+#           mutate(author = as.factor(ifelse(first_aut == "Shwetank Lall", "Shwetank Lall", "Other")))
+#       }
+#       
+#       if (input$first_aut == "other"){
+#         aut <- pkg_nodes_1aut %>%
+#           mutate(author = as.factor(ifelse(first_aut == "other", "Other authors", "Other with > 1 packages")))
+#       }
+#       
+#       graph_from_data_frame(d = pkg_edges, vertices = aut, directed = TRUE) %>%
+#         as_tbl_graph() %>% activate(edges)
+#     })
+#     
+#     
+# 
+#     output$net <- renderPlot(
+#       ggraph(dat, layout = "stress") +
+#         geom_edge_link(aes(color = type),
+#                        arrow = arrow(length = unit(2, 'mm')),
+#                        end_cap = circle(1, 'mm')) +
+#         geom_node_point(aes(color = author), size = 3) +
+#         geom_node_text(aes(label = name), repel = TRUE, size = 3.5) +
+#         scale_color_manual(values = c("black", "#6C3483")) +
+#         theme_graph()
+#     )
+#   },
+#   
+#   options = list(height = 1500,
+#                  width = 1500)
+#   
+# )
+```

### 2021-09-11 analysis/doe-pkgs-analysis.Rmd
Add interactive network based on author.

@@
 ```{r}
-
 # create data frame from the above function
@@
 
-# tidy author 
+#### tidy author
+
+## separate author 
 aut_tidy <- pkg_nodes_tmp %>%
+  # separate each aut/ctb/cph for each package
   separate_rows(author, sep=", (?=[A-Z])") %>%
+  # filter only role aut
   filter(!str_detect(author, "ctb") & !str_detect(author, "National") & !str_detect(author, "Columbia"))
 
+# replace character inside angle and square bracket
 aut_tidy$author <- rm_angle(aut_tidy$author, replacement = "")
@@
 
+# remove contributors written as sentence, also remove cph 
 aut_tidy <- aut_tidy %>%
@@
 
+# extract author who wrote >1 packages
 aut_group <- aut_tidy %>%
@@
   arrange(desc(n)) %>%
-  mutate(aut_gr = as.factor(ifelse(n > 1, author, "other")))
+  filter(n > 1)
 
+## tidy author in pkg_nodes_tmp 
+pkg_nodes_tmp$author <- rm_angle(pkg_nodes_tmp$author, replacement = "")
+pkg_nodes_tmp$author <- rm_square(pkg_nodes_tmp$author, replacement = "")
+
+```
+
+```{r}
 aut_tidy_join <- left_join(aut_tidy, aut_group, by = "author") %>%
@@
 ```
+
+> Note
+
+Some packages are authored by the same persons. For Example, Dice
 

### 2021-09-11 paper/package_interconnection.Rmd
Add interactive network based on author.

@@
 date: "03/09/2021"
-output: html_document
+output: bookdown::html_document2
+bibliography: paper.bib
 ---
@@
 library(networkD3)
+library(visNetwork)
+library(qdapRegex)
 ```
@@
   pivot_wider(names_from = L3,
-              values_from = value) %>%
-  mutate(author_ulrike = as.factor(ifelse(str_detect(author, "Ulrike"), 1, 0)))
+              values_from = value)
 
@@
 # create nodes from the two previous dataframe
-pkg_nodes <- full_join(from, to, by = c("L2" = "pkg")) %>%
+pkg_nodes_tmp <- full_join(from, to, by = c("L2" = "pkg")) %>%
   left_join(pkg_tidy, by = "L2") %>%
-  select(L2, author_ulrike)
+  select(L2, author)
+
+### get author and count how many packages they write as the first author
+### there are many authors who wrote >1 packages but in the same packages, here, we just extract the first author
+
+# get first author 
+pkg_nodes_tmp$author <- rm_angle(pkg_nodes_tmp$author, replacement = "")
+pkg_nodes_tmp$author <- rm_square(pkg_nodes_tmp$author, replacement = "")
+pkg_nodes_tmp$author <- rm_white_punctuation(pkg_nodes_tmp$author, replacement = "")
+pkg_nodes_tmp$author <- ifelse(str_detect(pkg_nodes_tmp$author, "based"), "Ernesto Barrios", pkg_nodes_tmp$author)
+
+pkg_nodes_tmp <- pkg_nodes_tmp %>%
+  cSplit("author", sep = " ,") %>%
+  select(c(1:2)) %>%
+  rename(first_aut = author_1)
+
+aut_gr <- pkg_nodes_tmp %>%
+  group_by(first_aut) %>%
+  count() %>%
+  ungroup() %>%
+  arrange(desc(n)) %>%
+  right_join(pkg_nodes_tmp, by = "first_aut")
+
+pkg_nodes_active <- aut_gr %>%
+  mutate(first_author = case_when(first_aut == "Ulrike Groemping" ~ "Ulrike Groemping",
+                                 first_aut == "John Lawson" ~ "John Lawson",
+                                 first_aut == "Robert B. Gramacy" ~ "Robert B. Gramacy",
+                                 first_aut == "Ernesto Barrios" ~ "Ernesto Barrios",
+                                 first_aut == "Kevin Wright" ~ "Kevin Wright",
+                                 first_aut == "Shwetank Lall" ~ "Shwetank Lall",
+                                 n == 1 ~ "Other")) %>%
+  mutate(first_author = factor(first_author, levels = c("Ulrike Groemping",
+                                                        "John Lawson",
+                                                        "Robert B. Gramacy",
+                                                        "Ernesto Barrios",
+                                                        "Kevin Wright",
+                                                        "Shwetank Lall",
+                                                        "Other"))) %>%
+  select(L2, first_author)
 
 # create igraph object
-routes_igraph <- graph_from_data_frame(d = pkg_edges, vertices = pkg_nodes, directed = TRUE)
+routes_igraph <- graph_from_data_frame(d = pkg_edges, vertices = pkg_nodes_active, directed = TRUE)
 
@@
   activate(edges)
+
+# create the nodes for interactive network
+nodes <- pkg_nodes_active %>%
+  rowid_to_column("id") %>%
+  rename(label = L2) %>%
+  mutate(font.size = 25)
+
+#create edges for interactive network
+edges <-routes_tidy %>%
+  activate(edges) %>%
+  data.frame()
+
+edges$color <- case_when(edges$type == "imports" ~ "green",
+                         edges$type == "depends" ~ "red",
+                         edges$type == "suggests" ~ "blue")
+
+ledges <- data.frame(color = c("green", "red", "blue"), 
+                     label = c("imports", "depends", "suggests"))
+
 ```
@@
 Further, the type of interaction include the depended, imported, and suggested packages.
-Out of `r nrow(pkg_tidy)`, there are only `r nrow(pkg_nodes)` packages that interplay each other.
+Out of `r nrow(pkg_tidy)`, there are only `r nrow(pkg_nodes_active)` packages that interplay each other.
 To observe these interaction, we use network visualisation using `igraph` [@igraph], `tidygraph` [@tidygraph], and `ggraph` [@ggraph].
 
-```{r network-graph, fig.cap = "Interaction network between DoE R packages.
+```{r int-plot}
+visNetwork(nodes, edges, height = "500px", width = "100%") %>%
+  visOptions(selectedBy = "first_author") %>%
+  visLayout(randomSeed = 123) %>%
+  visEdges(shadow = FALSE,
+           arrows =list(to = list(enabled = TRUE, scaleFactor = 2))) %>%
+  visLegend(useGroups = F, addEdges = ledges)
+```
+
+```{r network-graph, eval = FALSE, fig.cap = "Interaction network between DoE R packages.
 The color of edges reflect the type of interaction, i.e., depends, imports, and suggests.
@@
                  end_cap = circle(1, 'mm')) + 
-  geom_node_point(aes(color = author_ulrike)) + 
-  scale_color_manual(values = c("#707B7C", "#F5B041"), labels = c("no", "yes")) +
+  geom_node_point(aes(color = first_author)) + 
   geom_node_text(aes(label = name), repel = TRUE, size = 2) +
@@
 
-```{r}
-#create nodes for interactive network
-nodes_d3 <- pkg_nodes %>%
-  rowid_to_column("id") %>%
-  mutate(id = id - 1)
-
-#create edges for interactive network
-edges_d3 <-routes_tidy %>%
-  activate(edges) %>%
-  data.frame() %>%
-  mutate(from = from - 1,
-         to = to - 1)
-
-# interactive network
-forceNetwork(Links = edges_d3, Nodes = nodes_d3, Source = "from", Target = "to", 
-             NodeID = "L2", Group = "author_ulrike", 
-             opacity = 1, fontSize = 16, zoom = TRUE)
-```
-
 # Create the categorisation based on main author 

### 2021-09-11 paper/package_interconnection.Rmd
Modify the static network plot.

@@
 ```
+
+# Interaction
 
@@
 
+```{r network-graph, fig.cap = "Interaction network between DoE R packages.
+The color of edges reflect the type of interaction, i.e., depends, imports, and suggests.
+Nodes' color indicates the authors who wrote more than 2 DoE packages as the first author.", fig.width=12, fig.height = 8}
+
+pkg_nodes_static <- aut_gr %>%
+  mutate(first_author = case_when(first_aut == "Ulrike Groemping" ~ "Ulrike Groemping",
+                                 first_aut == "John Lawson" ~ "John Lawson",
+                                 first_aut == "Robert B. Gramacy" ~ "Robert B. Gramacy",
+                                 n <=2 ~ "Other")) %>%
+  mutate(first_author = factor(first_author, levels = c("Ulrike Groemping",
+                                                        "John Lawson",
+                                                        "Robert B. Gramacy",
+                                                        "Other"))) %>%
+  select(L2, first_author)
+
+# create igraph object
+routes_igraph_static <- graph_from_data_frame(d = pkg_edges, vertices = pkg_nodes_static, directed = TRUE)
+
+# create the input based on the igraph object
+routes_tidy_static <- routes_igraph_static %>% as_tbl_graph() %>%
+  activate(edges)
+
+# create the graph
+ggraph(routes_tidy_static, layout = "stress") + 
+  geom_edge_link(aes(color = type),
+                 arrow = arrow(length = unit(3, 'mm')), 
+                 end_cap = circle(3, 'mm')) + 
+  geom_node_point(aes(color = first_author), size = 4) + 
+  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
+  scale_color_manual(values = c("blue", "grey50", "tomato", "magenta")) +
+  theme_graph()
+```
+
 ```{r int-plot}
@@
 
-```{r network-graph, eval = FALSE, fig.cap = "Interaction network between DoE R packages.
-The color of edges reflect the type of interaction, i.e., depends, imports, and suggests.
-Nodes' color indicates wether the package is authored by Ulrike Groumping or not."}
-# create the graph
-ggraph(routes_tidy, layout = "stress") + 
-  geom_edge_link(aes(color = type),
-                 arrow = arrow(length = unit(1, 'mm')), 
-                 end_cap = circle(1, 'mm')) + 
-  geom_node_point(aes(color = first_author)) + 
-  geom_node_text(aes(label = name), repel = TRUE, size = 2) +
-  theme_graph()
-```
 Figure \@fig(network-graph) shows the interplay of DoE packages.
@@
 Based on this network, we can say that the interconnection of DoE package is due to functionality dependence, similarity of type or domain, and because they are developed by the same author/s.
-
-# Create the categorisation based on main author 
-# create the network by facetting by author 
-# if there are many authors then create the facet for example if the author write >= 3 packages. 
-# not only focus in the Ulrike alone

### 2021-09-12 paper/package_interconnection.Rmd
Add faceted network graph.

@@
 
-pkg_nodes_active <- aut_gr %>%
+pkg_nodes <- aut_gr %>%
   mutate(first_author = case_when(first_aut == "Ulrike Groemping" ~ "Ulrike Groemping",
@@
 # create igraph object
-routes_igraph <- graph_from_data_frame(d = pkg_edges, vertices = pkg_nodes_active, directed = TRUE)
+routes_igraph <- graph_from_data_frame(d = pkg_edges, vertices = pkg_nodes, directed = TRUE)
 
@@
 # create the nodes for interactive network
-nodes <- pkg_nodes_active %>%
+nodes <- pkg_nodes %>%
   rowid_to_column("id") %>%
@@
 Further, the type of interaction include the depended, imported, and suggested packages.
-Out of `r nrow(pkg_tidy)`, there are only `r nrow(pkg_nodes_active)` packages that interplay each other.
+Out of `r nrow(pkg_tidy)`, there are only `r nrow(pkg_nodes)` packages that interplay each other.
 To observe these interaction, we use network visualisation using `igraph` [@igraph], `tidygraph` [@tidygraph], and `ggraph` [@ggraph].
@@
 
-pkg_nodes_static <- aut_gr %>%
-  mutate(first_author = case_when(first_aut == "Ulrike Groemping" ~ "Ulrike Groemping",
-                                 first_aut == "John Lawson" ~ "John Lawson",
-                                 first_aut == "Robert B. Gramacy" ~ "Robert B. Gramacy",
-                                 n <=2 ~ "Other")) %>%
-  mutate(first_author = factor(first_author, levels = c("Ulrike Groemping",
-                                                        "John Lawson",
-                                                        "Robert B. Gramacy",
-                                                        "Other"))) %>%
-  select(L2, first_author)
+# create the graph
+ggraph(routes_tidy, layout = "kk") + 
+  geom_edge_link(aes(color = type),
+                 arrow = arrow(length = unit(2, 'mm')), 
+                 end_cap = circle(2, 'mm')) + 
+  geom_node_point(aes(color = first_author), size = 4) + 
+  geom_node_text(aes(label = name), repel = TRUE, size = 3.1) +
+  scale_color_manual(values = c("blue", "tomato", "magenta", "grey50", "#48C9B0", "#C0392B", "#F1C40F")) +
+  theme_graph() 
+```
 
-# create igraph object
-routes_igraph_static <- graph_from_data_frame(d = pkg_edges, vertices = pkg_nodes_static, directed = TRUE)
+```{r, fig.width=12, fig.height = 6}
 
-# create the input based on the igraph object
-routes_tidy_static <- routes_igraph_static %>% as_tbl_graph() %>%
+pkg_nodes_facet <- pkg_nodes %>%
+  filter(first_author %in% c("Ulrike Groemping", "John Lawson", "Robert B. Gramacy"))
+
+pkg_edges_facet <- pkg_edges %>%
+  filter(L2 %in% pkg_nodes_facet$L2 &
+           pkg %in% pkg_nodes_facet$L2)
+
+routes_igraph_facet <- graph_from_data_frame(d = pkg_edges_facet, 
+                                             vertices = pkg_nodes_facet, 
+                                             directed = TRUE)
+
+# # create the input based on the igraph object
+routes_tidy_facet <- routes_igraph_facet %>% as_tbl_graph() %>%
   activate(edges)
 
-# create the graph
-ggraph(routes_tidy_static, layout = "stress") + 
+ggraph(routes_tidy_facet, layout = "kk") + 
   geom_edge_link(aes(color = type),
@@
   geom_node_point(aes(color = first_author), size = 4) +
-  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
-  scale_color_manual(values = c("blue", "grey50", "tomato", "magenta")) +
-  theme_graph()
+  geom_node_text(aes(label = name), repel = TRUE, size = 3.5) +
+  theme_graph() +
+  facet_nodes(~first_author) +
+  scale_color_manual(values = c("tomato", "#48C9B0", "#F1C40F"), guide = "none") +
+  th_foreground(foreground = "grey80",  border = TRUE) 
 ```

### 2021-09-12 paper/package_interconnection.Rmd
Modify faceted network.

@@
 
-```{r, fig.width=12, fig.height = 6}
+```{r, fig.width=12, fig.height = 7}
 
 pkg_nodes_facet <- pkg_nodes %>%
-  filter(first_author %in% c("Ulrike Groemping", "John Lawson", "Robert B. Gramacy"))
+  filter(first_author %in% c("Ulrike Groemping", "John Lawson", "Robert B. Gramacy", 
+                             "Kevin Wright", "Shwetank Lall", "Ernesto Barrios"))
 
@@
   facet_nodes(~first_author) +
-  scale_color_manual(values = c("tomato", "#48C9B0", "#F1C40F"), guide = "none") +
+  scale_color_manual(values = c("blue", "tomato", "magenta", "#48C9B0", "#C0392B", "#F1C40F"), 
+                     guide = "none") +
   th_foreground(foreground = "grey80",  border = TRUE) 

### 2021-09-12 analysis/doe-pkgs-analysis.Rmd
Compare interconnection with other CRAN Task View.

@@
 
+# Comparison to other CRAN TASK VIEW
+
+## Survey
+
 ```{r}
-# shinyApp(
-#   ui = fluidPage(
-#     selectInput("first_aut", "Choose the author:",
-#                     choices = levels(pkg_nodes_1aut$first_aut)),
-#       mainPanel(plotOutput(outputId = "net"))
-#     ),
-#   
-#   server = function(input, output){
-#     
-#     dat <- reactive({
-#       if (input$first_aut == "Ulrike Groemping"){
-#         aut <- pkg_nodes_1aut %>%
-#           mutate(author = as.factor(ifelse(first_aut == "Ulrike Groemping", "Ulrike Groemping", "Other")))
-#       }
-#       
-#       if (input$first_aut == "John Lawson"){
-#         aut <- pkg_nodes_1aut %>%
-#           mutate(author = as.factor(ifelse(first_aut == "John Lawson", "John Lawson", "Other")))
-#       }
-#       
-#       if (input$first_aut == "Robert B. Gramacy"){
-#         aut <- pkg_nodes_1aut %>%
-#           mutate(author = as.factor(ifelse(first_aut == "Robert B. Gramacy", "Robert B. Gramacy", "Other")))
-#       }
-#       
-#       if (input$first_aut == "Delphine Dupuy"){
-#         aut <- pkg_nodes_1aut %>%
-#           mutate(author = as.factor(ifelse(first_aut == "Delphine Dupuy", "Delphine Dupuy", "Other")))
-#       }
-#       
-#       if (input$first_aut == "Ernesto Barrios"){
-#         aut <- pkg_nodes_1aut %>%
-#           mutate(author = as.factor(ifelse(first_aut == "Ernesto Barrios", "Ernesto Barrios", "Other")))
-#       }
-#       
-#       if (input$first_aut == "Kevin Wright"){
-#         aut <- pkg_nodes_1aut %>%
-#           mutate(author = as.factor(ifelse(first_aut == "Kevin Wright", "Kevin Wright", "Other")))
-#       }
-#       
-#        if (input$first_aut == "Peng Liu"){
-#         aut <- pkg_nodes_1aut %>%
-#           mutate(author = as.factor(ifelse(first_aut == "Peng Liu", "Peng Liu", "Other")))
-#        }
-#       
-#       if (input$first_aut == "Shwetank Lall"){
-#         aut <- pkg_nodes_1aut %>%
-#           mutate(author = as.factor(ifelse(first_aut == "Shwetank Lall", "Shwetank Lall", "Other")))
-#       }
-#       
-#       if (input$first_aut == "other"){
-#         aut <- pkg_nodes_1aut %>%
-#           mutate(author = as.factor(ifelse(first_aut == "other", "Other authors", "Other with > 1 packages")))
-#       }
-#       
-#       graph_from_data_frame(d = pkg_edges, vertices = aut, directed = TRUE) %>%
-#         as_tbl_graph() %>% activate(edges)
-#     })
-#     
-#     
-# 
-#     output$net <- renderPlot(
-#       ggraph(dat, layout = "stress") +
-#         geom_edge_link(aes(color = type),
-#                        arrow = arrow(length = unit(2, 'mm')),
-#                        end_cap = circle(1, 'mm')) +
-#         geom_node_point(aes(color = author), size = 3) +
-#         geom_node_text(aes(label = name), repel = TRUE, size = 3.5) +
-#         scale_color_manual(values = c("black", "#6C3483")) +
-#         theme_graph()
-#     )
-#   },
-#   
-#   options = list(height = 1500,
-#                  width = 1500)
-#   
-# )
+list_pkg_survey <- ctv:::.get_pkgs_from_ctv_or_repos("OfficialStatistics", 
+                                                repos = "http://cran.rstudio.com/")[[1]]
+url <- "http://cran.rstudio.com/web/packages/packages.rds"
+db <- readRDS(url(url)) %>% 
+  as.data.frame()
+survey_pkg <- db %>%
+  filter(Package %in% list_pkg_survey) %>%
+  select(Package, Depends, Imports, Suggests)
+
+get_pkg_survey <- function(var){
+  # select only the package name and depends/imports/suggests
+  tmp_df <- survey_pkg %>%
+    select(Package, var) %>%
+    # separate every depends/imports/suggests
+    cSplit(var, sep = " ,")
+
+  df <- tmp_df %>%
+    pivot_longer(c(2:ncol(tmp_df)), names_to = "type", values_to = "pkg") %>%
+    filter(!is.na(pkg)) %>%
+    # check if the depends/imports/suggests is also a DoE package (is in list_pkg)
+    mutate(survey_pkg_related = ifelse(pkg %in% list_pkg_survey, "yes", "no")) %>%
+    # filter only package whose depends/imports/suggests is a DoE
+    filter(survey_pkg_related == "yes") %>%
+    # tidy the data to only have the type: depends/imports/suggests
+    cSplit("type", sep = "_") %>%
+    select(-type_2) %>%
+    rename(type = type_1)
+
+  return(df)
+}
+
+survey_depends <- get_pkg_survey("Depends")
+survey_imports <- get_pkg_survey("Imports")
+survey_suggests <- get_pkg_survey("Suggests")
+
+pkg_edges_survey <- rbind(survey_depends, survey_imports, survey_suggests)
+
+survey_from <- pkg_edges_survey %>%
+  distinct(Package)
+
+# extract unique package that being depends/imports/suggests
+survey_to <- pkg_edges_survey %>%
+  distinct(pkg)
+
+# create nodes from the two previous dataframe
+pkg_nodes_survey <- full_join(survey_from, survey_to, by = c("Package" = "pkg"))
 ```
+
+## Timeseries
+
+```{r}
+list_pkg_ts <- ctv:::.get_pkgs_from_ctv_or_repos("TimeSeries", 
+                                                repos = "http://cran.rstudio.com/")[[1]]
+
+ts_pkg <- db %>%
+  filter(Package %in% list_pkg_ts) %>%
+  select(Package, Depends, Imports, Suggests)
+
+get_pkg_ts <- function(var){
+  # select only the package name and depends/imports/suggests
+  tmp_df <- ts_pkg %>%
+    select(Package, var) %>%
+    # separate every depends/imports/suggests
+    cSplit(var, sep = " ,")
+
+  df <- tmp_df %>%
+    pivot_longer(c(2:ncol(tmp_df)), names_to = "type", values_to = "pkg") %>%
+    filter(!is.na(pkg)) %>%
+    # check if the depends/imports/suggests is also a DoE package (is in list_pkg)
+    mutate(survey_pkg_related = ifelse(pkg %in% list_pkg_ts, "yes", "no")) %>%
+    # filter only package whose depends/imports/suggests is a DoE
+    filter(survey_pkg_related == "yes") %>%
+    # tidy the data to only have the type: depends/imports/suggests
+    cSplit("type", sep = "_") %>%
+    select(-type_2) %>%
+    rename(type = type_1)
+
+  return(df)
+}
+
+ts_depends <- get_pkg_ts("Depends")
+ts_imports <- get_pkg_ts("Imports")
+ts_suggests <- get_pkg_ts("Suggests")
+
+pkg_edges_ts <- rbind(ts_depends, ts_imports, ts_suggests)
+
+ts_from <- pkg_edges_ts %>%
+  distinct(Package)
+
+# extract unique package that being depends/imports/suggests
+ts_to <- pkg_edges_ts %>%
+  distinct(pkg)
+
+# create nodes from the two previous dataframe
+pkg_nodes_ts <- full_join(ts_from, ts_to, by = c("Package" = "pkg"))
+```

### 2021-09-12 analysis/doe-pkgs-analysis.Rmd
Add comparison with other CRAN task views.

@@
 ```
+
+## Create Sankey Diagram for packages' type
+
+```{r}
+sankey_data <- pkg_tidy_long_type %>%
+  select(year, domain, type) %>%
+  group_by(domain, type) %>%
+  count() %>%
+  ungroup()
+```

### 2021-09-12 paper/package_interconnection.Rmd
Compare interconnection with other CRAN Task View.
Add comparison with other CRAN task views.

@@
 Further, the type of interaction include the depended, imported, and suggested packages.
-Out of `r nrow(pkg_tidy)`, there are only `r nrow(pkg_nodes)` packages that interplay each other.
-To observe these interaction, we use network visualisation using `igraph` [@igraph], `tidygraph` [@tidygraph], and `ggraph` [@ggraph].
+Out of `r nrow(pkg_tidy)`, there are only `r nrow(pkg_nodes)` (`r round(nrow(pkg_nodes)/nrow(pkg_tidy)*100,2)`%) packages that interplay each other.
+This number is relatively lower compared to the percentage of interaction in the packages in Survey and Time Series Analysis CRAN Task View.
+We found that `r round(73/132*100,2)`% Survey packages that depends, imports, and suggests each other, while these interconnection is observed in `r round(212/338*100,2)`% Time Series Analysis packages.
+
+Further, we observe the interaction in DoE package using network visualisation using `igraph` [@igraph], `tidygraph` [@tidygraph], and `ggraph` [@ggraph].
 
@@
 
-```{r, fig.width=12, fig.height = 7}
+```{r, fig.width=12, fig.height = 7, fig.cap= "Package network based on first authors who created more than 2 DoE packages. Packages authored by the same person tend to interplay each other."}
 
@@
 
+Figure \@ref(fig:network-graph) shows the interplay of DoE packages.
+The most downloaded DoE packages, such as `AlgDesign`, `rsm`, and `lhs` are observed to support many other packages.
+Dependent packages seems to use their functions to support their feature.
+We also observe that  `DoE.base` and `FrF2`, which are created by the same author, Ulrike Groemping, are also imported, depended, and suggested by many other packages because of their basic functionality in experimental design.
+Moreover, some of interacting packages, such as `DoE.wrapper` and `DoE.MIParray` are also developed by Groemping.
+
+Further, packages regarding computer experiment (`DiceKriging`, `DiceEval`, `DiceView`, and `DiceDesign`) happen to interplay with each other so that they appear to form a sub-cluster.
+The four smaller clusters interact because of the similarity in their type and domain.
+For example `dfcm` and `dfpk` are used for dose-finding experiments.
+Regarding the similarity of its objective, although `FMC` aims to generate factorial design, this package does not interact with the packages in the big cluster, even though some packages in this cluster, such as `DoE.base`, also involve factorial design.
+This package instead imports `minimalRSD`, which is developed by the same author.
+This is also observed in the smaller network involves `dynaTree`, `plgp`, and `tgp`.
+These three interplaying package are authored by the same author as well.
+
+Based on this network, we can say that the interconnection of DoE package is due to functionality dependence, similarity of type or domain, and because they are developed by the same author/s.
+
 ```{r int-plot}
@@
 ```
-
-Figure \@fig(network-graph) shows the interplay of DoE packages.
-In the previous section, we noticed that Ulrike Groemping authored many DoE packages.
-We also noticed that one of her package, DoE base, is among the most downloaded DoE package on CRAN.
-Hence, we categorise the package whether it is authored by her or not.
-The network conveys that there are 5 clusters of interaction with one big cluster.
-
-The most downloaded DoE packages, such as `AlgDesign`, `rsm`, and `lhs` are observed to support many other packages.
-Dependent packages seems to use their functions to support their feature.
-We also observe that  `DoE.base` and `FrF2`, which are authored by Groemping, are also imported, depended, and suggested by many other packages because of their basic functionality in experimental design.
-Moreover, some of interacting packages, such as `DoE.wrapper` and `DoE.MIParray` are also developed by Groemping.
-
-Further, packages regarding computer experiment (`DiceKriging`, `DiceEval`, `DiceView`, and `DiceDesign`) happen to interplay with each other so that they appear to form a sub-cluster.
-The four smaller clusters interact because of the similarity in their type and domain.
-For example `dfcm` and `dfpk` are used for dose-finding experiments.
-Regarding the similarity of its objective, although `FMC` aims to generate factorial design, this package does not interact with the packages in the big cluster, even though some packages in this cluster, such as `DoE.base`, also involve factorial design.
-This package instead imports `minimalRSD`, which is developed by the same author.
-
-Based on this network, we can say that the interconnection of DoE package is due to functionality dependence, similarity of type or domain, and because they are developed by the same author/s.

### 2021-09-12 analysis/doe-pkgs-analysis.Rmd
Create Sankey diagram for package type and domain.

@@
   ungroup()
+
+domain <- sankey_data %>%
+  distinct(domain) %>%
+  rename(node = domain)
+
+type <- sankey_data %>%
+  distinct(node = type)
+
+db_sankey <- rbind(domain, type)
+
+routes_sankey <- graph_from_data_frame(d = sankey_data, vertices = db_sankey, directed = TRUE)
+
+routes_sankey_tidy <- routes_sankey %>% as_tbl_graph() %>%
+  activate(edges)
+
+nodes_sankey <- db_sankey %>%
+  rowid_to_column("id") %>%
+  mutate(id = id - 1)
+
+edges_sankey <-routes_sankey_tidy %>%
+  activate(edges) %>%
+  data.frame() %>%
+  mutate(from = from - 1,
+         to = to - 1)
+
+sankeyNetwork(Links = edges_sankey, Nodes = nodes_sankey, Source = 'from', 
+                         Target = 'to', Value = 'n', NodeID = 'node',
+                         units = 'packages')
 ```

### 2021-09-17 paper/package_interconnection.Rmd
Incorporates author with >1 packages eventho not 1st aut.

@@
 ### there are many authors who wrote >1 packages but in the same packages, here, we just extract the first author
+```
 
-# get first author 
-pkg_nodes_tmp$author <- rm_angle(pkg_nodes_tmp$author, replacement = "")
-pkg_nodes_tmp$author <- rm_square(pkg_nodes_tmp$author, replacement = "")
-pkg_nodes_tmp$author <- rm_white_punctuation(pkg_nodes_tmp$author, replacement = "")
-pkg_nodes_tmp$author <- ifelse(str_detect(pkg_nodes_tmp$author, "based"), "Ernesto Barrios", pkg_nodes_tmp$author)
+```{r network-data}
 
-pkg_nodes_tmp <- pkg_nodes_tmp %>%
-  cSplit("author", sep = " ,") %>%
-  select(c(1:2)) %>%
-  rename(first_aut = author_1)
+## separate author 
+aut_tidy <- pkg_nodes_tmp %>%
+  # separate each aut/ctb/cph for each package
+  separate_rows(author, sep=", (?=[A-Z])") %>%
+  # filter only role aut and filter out cph
+  filter(!str_detect(author, "ctb") & !str_detect(author, "National") & !str_detect(author, "Columbia"))
 
-aut_gr <- pkg_nodes_tmp %>%
-  group_by(first_aut) %>%
+# replace character inside angle and square bracket
+aut_tidy$author <- rm_angle(aut_tidy$author, replacement = "")
+aut_tidy$author <- rm_square(aut_tidy$author, replacement = "")
+
+# remove contributors written as sentence
+aut_tidy <- aut_tidy %>%
+  mutate(author = ifelse(str_detect(author, "Contributor"), "Yves Deville",
+                         ifelse(str_detect(author, "based"), "Ernesto Barrios",
+                                ifelse(str_detect(author, "contributions"), "John Lawson", author))))
+
+# extract author who wrote >1 packages
+aut_group <- aut_tidy %>%
+  group_by(author) %>%
   count() %>%
-  ungroup() %>%
-  arrange(desc(n)) %>%
-  right_join(pkg_nodes_tmp, by = "first_aut")
+  arrange(desc(n))
 
-pkg_nodes <- aut_gr %>%
-  mutate(first_author = case_when(first_aut == "Ulrike Groemping" ~ "Ulrike Groemping",
-                                 first_aut == "John Lawson" ~ "John Lawson",
-                                 first_aut == "Robert B. Gramacy" ~ "Robert B. Gramacy",
-                                 first_aut == "Ernesto Barrios" ~ "Ernesto Barrios",
-                                 first_aut == "Kevin Wright" ~ "Kevin Wright",
-                                 first_aut == "Shwetank Lall" ~ "Shwetank Lall",
-                                 n == 1 ~ "Other")) %>%
-  mutate(first_author = factor(first_author, levels = c("Ulrike Groemping",
-                                                        "John Lawson",
-                                                        "Robert B. Gramacy",
-                                                        "Ernesto Barrios",
-                                                        "Kevin Wright",
-                                                        "Shwetank Lall",
-                                                        "Other"))) %>%
-  select(L2, first_author)
+aut_num <- filter(aut_group, n > 1)
+
+aut_tidy_join <- left_join(aut_tidy, aut_group, by = "author") %>%
+  group_by(L2) %>%
+  arrange(desc(n), .by_group = TRUE)
+
+pkg_nodes <- aggregate(author ~ L2, aut_tidy_join, FUN = toString) %>%
+  mutate(note = ifelse(str_detect(author, "Ulrike"), "Groemping",
+                       ifelse(str_detect(L2, "Dice"), "Dupuy, Roustant, Richet, Deville",
+                              ifelse(str_detect(author, "Gramacy"), "Gramacy, Taddy",
+                                     ifelse(str_detect(author, "Shwetank"), "Lall, Vhargese",
+                                            ifelse(str_detect(author, "Kevin Wright"), "Wright",
+                                                   ifelse(str_detect(author, "Peng Liu"), "Liu",
+                                                          ifelse(str_detect(author, "Ernesto"), "Barrios", "Other"))))))))
+
+# # get first author 
+# pkg_nodes_tmp$author <- rm_angle(pkg_nodes_tmp$author, replacement = "")
+# pkg_nodes_tmp$author <- rm_square(pkg_nodes_tmp$author, replacement = "")
+# pkg_nodes_tmp$author <- rm_white_punctuation(pkg_nodes_tmp$author, replacement = "")
+# pkg_nodes_tmp$author <- ifelse(str_detect(pkg_nodes_tmp$author, "based"), "Ernesto Barrios", pkg_nodes_tmp$author)
+# 
+# pkg_nodes_tmp <- pkg_nodes_tmp %>%
+#   cSplit("author", sep = " ,") %>%
+#   select(c(1:2)) %>%
+#   rename(first_aut = author_1) 
+# 
+# aut_gr <- pkg_nodes_tmp %>%
+#   group_by(first_aut) %>%
+#   count() %>%
+#   ungroup() %>%
+#   arrange(desc(n)) %>%
+#   right_join(pkg_nodes_tmp, by = "first_aut")
+# 
+# 
+# pkg_nodes <- aut_gr %>%
+#   mutate(first_author = case_when(
+#                                  n == 1 ~ "Other")) %>%
+#   mutate(first_author = factor(first_author, levels = c("Ulrike Groemping",
+#                                                         "John Lawson",
+#                                                         "Robert B. Gramacy",
+#                                                         "Ernesto Barrios",
+#                                                         "Kevin Wright",
+#                                                         "Shwetank Lall",
+#                                                         "Other"))) %>%
+#   select(L2, first_author)
 
@@
                  end_cap = circle(2, 'mm')) + 
-  geom_node_point(aes(color = first_author), size = 4) + 
+  geom_node_point(aes(color = note), size = 4) + 
   geom_node_text(aes(label = name), repel = TRUE, size = 3.1) +
-  scale_color_manual(values = c("blue", "tomato", "magenta", "grey50", "#48C9B0", "#C0392B", "#F1C40F")) +
+  #scale_color_manual(values = c("blue", "tomato", "magenta", "grey50", "#48C9B0", "#C0392B", "#F1C40F")) +
   theme_graph() 
@@
 pkg_nodes_facet <- pkg_nodes %>%
-  filter(first_author %in% c("Ulrike Groemping", "John Lawson", "Robert B. Gramacy", 
-                             "Kevin Wright", "Shwetank Lall", "Ernesto Barrios"))
+  filter(note != "Other")
 
@@
                  end_cap = circle(3, 'mm')) + 
-  geom_node_point(aes(color = first_author), size = 4) +
+  geom_node_point(aes(color = note), size = 4) +
   geom_node_text(aes(label = name), repel = TRUE, size = 3.5) +
   theme_graph() +
-  facet_nodes(~first_author) +
-  scale_color_manual(values = c("blue", "tomato", "magenta", "#48C9B0", "#C0392B", "#F1C40F"), 
-                     guide = "none") +
+  facet_nodes(~note) +
+  #scale_color_manual(values = c("blue", "tomato", "magenta", "#48C9B0", "#C0392B", "#F1C40F"), 
+                    # guide = "none") +
   th_foreground(foreground = "grey80",  border = TRUE) 
@@
 visNetwork(nodes, edges, height = "500px", width = "100%") %>%
-  visOptions(selectedBy = "first_author") %>%
+  visOptions(selectedBy = "note") %>%
   visLayout(randomSeed = 123) %>%

### 2021-09-19 paper/package_interconnection.Rmd
modify node label.

@@
 
-```{r network-data}
+```{r network-data-aut}
 
@@
                                                    ifelse(str_detect(author, "Peng Liu"), "Liu",
-                                                          ifelse(str_detect(author, "Ernesto"), "Barrios", "Other"))))))))
+                                                          ifelse(str_detect(author, "Ernesto"), "Barrios", "Other")))))))) %>%
+  mutate(name_add = ifelse(L2 == "plgp", "plgp (G)",
+                     ifelse(L2 == "tgp", "tgp (G,T)",
+                            ifelse(L2 == "dynaTree", "dynaTree (G,T)",
+                                   ifelse(L2 == "DiceDesign", "DiceDesign (Du,Ro)",
+                                          ifelse(L2 == "DiceEval", "DiceEval (Du)",
+                                                 ifelse(L2 == "DiceKriging", "DiceKriging (Ro;De;Ri)",
+                                                        ifelse(L2 == "DiceView", "DiceView (De;Ri)", L2))))))))
 
@@
   rowid_to_column("id") %>%
-  rename(label = L2) %>%
+  rename(label = name_add) %>%
   mutate(font.size = 25)
@@
   geom_node_point(aes(color = note), size = 4) + 
-  geom_node_text(aes(label = name), repel = TRUE, size = 3.1) +
+  geom_node_text(aes(label = name_add), repel = TRUE, size = 3.1) +
   #scale_color_manual(values = c("blue", "tomato", "magenta", "grey50", "#48C9B0", "#C0392B", "#F1C40F")) +
@@
   geom_node_point(aes(color = note), size = 4) +
-  geom_node_text(aes(label = name), repel = TRUE, size = 3.5) +
+  geom_node_text(aes(label = name_add), repel = TRUE, size = 3.5) +
   theme_graph() +
@@
                     # guide = "none") +
-  th_foreground(foreground = "grey80",  border = TRUE) 
+  th_foreground(foreground = "grey80",  border = TRUE) +
+  theme(legend.position = "none")
 ```

### 2021-09-19 analysis/doe-pkgs-analysis.Rmd
Modify the writings on network.

@@
 library(visNetwork)
+library(forcats)
 ```
@@
 ```
+
 ```{r}

### 2021-09-19 paper/package_interconnection.Rmd
Modify the writings on network.

@@
 The color of edges reflect the type of interaction, i.e., depends, imports, and suggests.
-Nodes' color indicates the authors who wrote more than 2 DoE packages as the first author.", fig.width=12, fig.height = 8}
+Nodes' color indicates the authors who wrote more than 2 DoE packages.", fig.width=12, fig.height = 8}
 
@@
 
-```{r, fig.width=12, fig.height = 7, fig.cap= "Package network based on first authors who created more than 2 DoE packages. Packages authored by the same person tend to interplay each other."}
+```{r, fig.width=12, fig.height = 7, fig.cap= "Package network based on authors who created more than 2 DoE packages. Packages authored by the same person tend to interplay each other."}
 
@@
 
-ggraph(routes_tidy_facet, layout = "kk") + 
+ggraph(routes_tidy_facet, layout = "stress") + 
   geom_edge_link(aes(color = type),
@@
 Moreover, some of interacting packages, such as `DoE.wrapper` and `DoE.MIParray` are also developed by Groemping.
+Further, packages regarding computer experiment (`DiceKriging`, `DiceEval`, `DiceView`, and `DiceDesign`) happen to interplay with each other so that they appear to form a sub-cluster.
+These packages also written by authors that interplay each other.
 
-Further, packages regarding computer experiment (`DiceKriging`, `DiceEval`, `DiceView`, and `DiceDesign`) happen to interplay with each other so that they appear to form a sub-cluster.
 The four smaller clusters interact because of the similarity in their type and domain.

### 2021-09-26 analysis/doe-pkgs-analysis.Rmd
Modify sankey and network diagram.

@@
   unnest(domain) %>%
+  unnest(type2) %>%
   mutate(year = year(ymd(date)))
+
+pkg_type <- pkg_tidy %>%
+  unnest(type2) %>%
+  unnest(domain)
 ```
@@
 # summarise the packages based on its type
-pkg_tidy_long_type %>%
-  filter(relevance != "no") %>%
-  group_by(type) %>%
+pkg_type %>%
+  group_by(type2) %>%
   count() %>%
@@
 ```{r}
-sankey_data <- pkg_tidy_long_type %>%
-  select(year, domain, type) %>%
-  group_by(domain, type) %>%
+sankey_data <- pkg_type %>%
+  select(domain, type2) %>%
+  group_by(domain, type2) %>%
   count() %>%
@@
 type <- sankey_data %>%
-  distinct(node = type)
+  distinct(node = type2)
 

### 2021-09-26 paper/package_interconnection.Rmd
Modify sankey and network diagram.

@@
   geom_node_text(aes(label = name_add), repel = TRUE, size = 3.1) +
-  #scale_color_manual(values = c("blue", "tomato", "magenta", "grey50", "#48C9B0", "#C0392B", "#F1C40F")) +
+  scale_color_manual(values = c("blue", "tomato", "magenta", "#48C9B0", "#C0392B", "#F1C40F", "grey50", "green")) +
   theme_graph() 
@@
   facet_nodes(~note) +
-  #scale_color_manual(values = c("blue", "tomato", "magenta", "#48C9B0", "#C0392B", "#F1C40F"), 
-                    # guide = "none") +
-  th_foreground(foreground = "grey80",  border = TRUE) +
-  theme(legend.position = "none")
+  scale_color_manual(values = c("blue", "tomato", "magenta", "#48C9B0", 
+                                  "#C0392B", "#F1C40F", "green"), 
+                     guide = "none") +
+  th_foreground(foreground = "grey80",  border = TRUE) 
 ```

### 2021-10-03 analysis/doe-pkgs-analysis.Rmd
Add year in sankey plot.

--- 2021-09-19 analysis/doe-pkgs-analysis.Rmd
@@
   mutate(year = year(ymd(date)))
+
 ```
@@
 # summarise the packages based on its type
-pkg_tidy_long_type %>%
-  filter(relevance != "no") %>%
-  group_by(type) %>%
-  count() %>%
-  arrange(desc(n))
+pkg_type <- pkg_tidy_long_type %>%
+  group_by(L2) %>%
+  slice_head(n = 1) %>%
+  mutate(year = factor(year, level = c("2009", "2011", "2012", "2013",
+                                       "2014", "2015", "2016", "2017",
+                                       "2018", "2019", "2020", "2021")))
 ```
@@
 ```{r}
-sankey_data <- pkg_tidy_long_type %>%
-  select(year, domain, type) %>%
+sankey_data <- pkg_type %>%
+  select(domain, type) %>%
   group_by(domain, type) %>%
@@
 ```
+
+```{r}
+sankey_data <- pkg_type %>%
+  select(domain, year) %>%
+  group_by(domain, year) %>%
+  count() %>%
+  ungroup()
+
+domain <- sankey_data %>%
+  distinct(domain) %>%
+  rename(node = domain)
+
+year <- sankey_data %>%
+  distinct(node = year)
+
+db_sankey <- rbind(domain, year)
+
+routes_sankey <- graph_from_data_frame(d = sankey_data, vertices = db_sankey, directed = TRUE)
+
+routes_sankey_tidy <- routes_sankey %>% as_tbl_graph() %>%
+  activate(edges)
+
+nodes_sankey <- db_sankey %>%
+  rowid_to_column("id") %>%
+  mutate(id = id - 1)
+
+edges_sankey <-routes_sankey_tidy %>%
+  activate(edges) %>%
+  data.frame() %>%
+  mutate(from = from - 1,
+         to = to - 1)
+
+sankeyNetwork(Links = edges_sankey, Nodes = nodes_sankey, Source = 'from', 
+                         Target = 'to', Value = 'n', NodeID = 'node',
+                         units = 'packages')
+```

### 2021-10-04 paper/rjournal/paper-rjournal.tex
fix commands so it works in html + latex

--- 2021-03-05 paper.tex
@@
 of data collected -- namely, experimental or observational -- or
-alternatively, categorised into experimental design or survey design.
-Survey designs often, although not always, aim to collect observational
-data whilst experimental designs exclusively center on experimental
-data. This dichotomisation, to a great extent, is seen in CRAN task
-views where R-packages in experimental design are in
-\ctv{ExperimentalDesign} and R-packages in survey designs are in
-\ctv{OfficialStatistics}.
+alternatively, categorised as experimental design (including
+quasi-experimental design) or survey design. This dichotomisation, to a
+great extent, is seen in CRAN task views where R-packages in
+experimental design are in \ctv{ExperimentalDesign} and R-packages in
+survey designs are in \ctv{OfficialStatistics}. Survey designs often,
+although not always, aim to collect observational data whilst
+experimental designs exclusively center on experimental data. This paper
+is concerned with the latter.
 
-In the CRAN task view of \ctv{ExperimentalDesign}, there are 113 R
+In the CRAN task view of \ctv{ExperimentalDesign}, there are 110 R
 packages for experimental design and analysis of data from experiments,
@@
 \texttt{dexpy}, \texttt{experimenter} and \texttt{GPdoemd}) with limited
-outputs.
+outputs. Thus, the study of DoE packages is also revealing into the
+current status of the field of experimental design.
 
-\hypertarget{explorative-analysis}{%
-\section{Explorative Analysis}\label{explorative-analysis}}
+The paper is organisd as follows. Section @ref(eda)
+
+\hypertarget{eda}{%
+\section{Explorative analysis}\label{eda}}
 
@@
 
-{\centering \includegraphics{paper_files/figure-latex/dlplots-1}
+{\centering \includegraphics{figures/dlplots-1}
 
@@
 
-\caption[The above graph shows the total download counts from 2019-03-06 to 2021-03-04 of DoE packages]{The above graph shows the total download counts from 2019-03-06 to 2021-03-04 of DoE packages.}\label{fig:dlplots}
+\caption[The above graph shows the total download counts from 2019-10-04 to 2021-10-02 of DoE packages]{The above graph shows the total download counts from 2019-10-04 to 2021-10-02 of DoE packages.}\label{fig:dlplots}
 \end{figure}
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/wordcloud-1} \end{center}
+\begin{center}\includegraphics{figures/wordcloud-1} \end{center}
 
@@
 
+Figure \ref{fig:dlplots} suggests that the \CRANpkg{AlgDesign} followed
+by \CRANpkg{agricolae} are far more downloaded than any other DoE
+packages. It is worth noting that \CRANpkg{agricolae} imports
+\CRANpkg{AlgDesign} thus any download of \CRANpkg{agricolae} likely
+results also in a download of \CRANpkg{AlgDesign}.
+
 \hypertarget{software-design}{%
-\section{Software Design}\label{software-design}}
+\section{Software design}\label{software-design}}
 
-\hypertarget{propsects-and-discussion}{%
-\section{Propsects and Discussion}\label{propsects-and-discussion}}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+An experiment involves running a number of steps:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Defining the hypothesis or question
+\item
+  Formulating methods to test hypothesis or answer question
+\item
+  Planning the experimental protocol:
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  Determining experimental resources,
+\item
+  specifying the data collection procedure,
+\item
+  constructing the experimental design layout, and
+\item
+  proposing a model for analysis.
+\end{itemize}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{3}
+\tightlist
+\item
+  Collect data
+\item
+  Analyse data, interpret result and conclusion
+\end{enumerate}
 
@@
 \midrule
-optimal design & 10\\
+optimal design & 9\\
 experimental design & 8\\
@@
 dose finding & 5\\
-sequential design & 5\\
 block design & 4\\
@@
 response surface & 4\\
+sequential design & 4\\
 \bottomrule
@@
 \midrule
-experimental design & 11\\
+experimental design & 10\\
 optimal design & 10\\
@@
 sample size & 5\\
-sequential design & 5\\
 \bottomrule
@@
 
-\CRANpkg{agricolae}
-
-\hypertarget{helping-info-to-get-started}{%
-\subsection{Helping info to get
-started}\label{helping-info-to-get-started}}
-
-Introductory section which may include references in parentheses
-\citep{R}, or cite a reference such as \citet{R} in the text.
-
-\hypertarget{section-title-in-sentence-case}{%
-\subsection{Section title in sentence
-case}\label{section-title-in-sentence-case}}
-
-Let's check fi this works Figure @ref(fig:Rlogo).
-
-This section may contain a figure such as Figure \ref{fig:Rlogo}.
-
-\hypertarget{summary}{%
-\subsection{Summary}\label{summary}}
-
-This file is only a basic article template. For full details of
-\emph{The R Journal} style and information on how to prepare your
-article for submission, see the
-\href{https://journal.r-project.org/share/author-guide.pdf}{Instructions
-for Authors}.
-
-\hypertarget{about-this-format-and-the-r-journal-requirements}{%
-\subsubsection{About this format and the R Journal
-requirements}\label{about-this-format-and-the-r-journal-requirements}}
-
-\texttt{rticles::rjournal\_article} will help you build the correct
-files requirements:
-
-\begin{itemize}
-\tightlist
-\item
-  A R file will be generated automatically using \texttt{knitr::purl} -
-  see \url{https://bookdown.org/yihui/rmarkdown-cookbook/purl.html} for
-  more information.
-\item
-  A tex file will be generated from this Rmd file and correctly included
-  in \texttt{RJwapper.tex} as expected to build \texttt{RJwrapper.pdf}.
-\item
-  All figure files will be kept in the default rmarkdown
-  \texttt{*\_files} folder. This happens because
-  \texttt{keep\_tex\ =\ TRUE} by default in
-  \texttt{rticles::rjournal\_article}
-\item
-  Only the bib filename is to modifed. An example bib file is included
-  in the template (\texttt{RJreferences.bib}) and you will have to name
-  your bib file as the tex, R, and pdf files.
-\end{itemize}
-
-\bibliography{paper.bib}
+\bibliography{../paper.bib}
 

### 2021-10-04 paper/rjournal/paper-rjournal.R

@@ -1,23 +1,11 @@
----
-title: Current status and propsects of R-packages for the design of experiments
-author:
-  - name: Emi Tanaka
-    affiliation: Monash University
-abstract: >
-    Re-running an experiment is generally costly and in some cases impossible due to limited resources, so the design of experiment plays a critical role in increasing the quality of experimental data.  In this article I describe the current state of the R-packages for the design of experiments through textual analysis and download trends. I discuss also the software design of widely utilised R packages in the field of experimental design and conclude with discussion of some future prospects for the field.
-preamble: |
-  % Any extra LaTeX you need in the preamble
-  
-# per R journal requirement, the bib filename should be the same as the output 
-# tex file. Don't forget to rename the bib file and change this example value.
-bibliography: paper.bib
+## ---- include = FALSE---------------------------------------------------------
+paper <- here::here("paper/index.Rmd")
+meta <- rmarkdown::yaml_front_matter(paper)
 
-output: 
-  bookdown::html_document2:
-    theme: paper
----
 
-```{r setup, cache = FALSE, include = FALSE}
+## ----child=paper--------------------------------------------------------------
+
+## ----setup, cache = FALSE, include = FALSE------------------------------------
 library(tidyverse)
 library(rvest)
 library(lubridate)
@@ -64,42 +52,27 @@ ctv <- function(view) {
 CRANpkg <- function(pkg) {
   ifelse(knitr::is_html_output(), glue("[{pkg}](https://cran.r-project.org/web/packages/{pkg}/index.html)"), glue("\\CRANpkg{[pkg]}", .open = "[", .close = "]"))
 }
-ref <- function(x) {
-  ifelse(knitr::is_html_output(), glue("\\@ref({x})"), glue("\\ref{[x]}", .open = "[", .close = "]"))
-}
-```
 
-```{r ctv}
+
+## ----ctv----------------------------------------------------------------------
 doe_pkgs <- ctv:::.get_pkgs_from_ctv_or_repos("ExperimentalDesign", 
                                               repos = "http://cran.rstudio.com/")[[1]]
 doe_pkgs <- c(doe_pkgs, "DeclareDesign")
 npkgs <- length(doe_pkgs)
-```
 
-```{r cran-data, cache = tocache}
+
+## ----cran-data, cache = tocache-----------------------------------------------
[…]

### 2021-10-10 paper/package_interconnection.Rmd
Add inline code to the text.

@@
 
+```{r}
+# interaction in survey CRAN task view
+
+list_pkg_survey <- ctv:::.get_pkgs_from_ctv_or_repos("OfficialStatistics", 
+                                                repos = "http://cran.rstudio.com/")[[1]]
+url <- "http://cran.rstudio.com/web/packages/packages.rds"
+db <- readRDS(url(url)) %>% 
+  as.data.frame()
+survey_pkg <- db %>%
+  filter(Package %in% list_pkg_survey) %>%
+  select(Package, Depends, Imports, Suggests)
+
+get_pkg_survey <- function(var){
+  # select only the package name and depends/imports/suggests
+  tmp_df <- survey_pkg %>%
+    select(Package, var) %>%
+    # separate every depends/imports/suggests
+    cSplit(var, sep = " ,")
+
+  df <- tmp_df %>%
+    pivot_longer(c(2:ncol(tmp_df)), names_to = "type", values_to = "pkg") %>%
+    filter(!is.na(pkg)) %>%
+    # check if the depends/imports/suggests is also a DoE package (is in list_pkg)
+    mutate(survey_pkg_related = ifelse(pkg %in% list_pkg_survey, "yes", "no")) %>%
+    # filter only package whose depends/imports/suggests is a DoE
+    filter(survey_pkg_related == "yes") %>%
+    # tidy the data to only have the type: depends/imports/suggests
+    cSplit("type", sep = "_") %>%
+    select(-type_2) %>%
+    rename(type = type_1)
+
+  return(df)
+}
+
+survey_depends <- get_pkg_survey("Depends")
+survey_imports <- get_pkg_survey("Imports")
+survey_suggests <- get_pkg_survey("Suggests")
+
+pkg_edges_survey <- rbind(survey_depends, survey_imports, survey_suggests)
+
+survey_from <- pkg_edges_survey %>%
+  distinct(Package)
+
+# extract unique package that being depends/imports/suggests
+survey_to <- pkg_edges_survey %>%
+  distinct(pkg)
+
+# create nodes from the two previous dataframe
+pkg_nodes_survey <- full_join(survey_from, survey_to, by = c("Package" = "pkg"))
+
+# Interaction in Time Series Analysis CRAN task view
+
+list_pkg_ts <- ctv:::.get_pkgs_from_ctv_or_repos("TimeSeries", 
+                                                repos = "http://cran.rstudio.com/")[[1]]
+
+ts_pkg <- db %>%
+  filter(Package %in% list_pkg_ts) %>%
+  select(Package, Depends, Imports, Suggests)
+
+get_pkg_ts <- function(var){
+  # select only the package name and depends/imports/suggests
+  tmp_df <- ts_pkg %>%
+    select(Package, var) %>%
+    # separate every depends/imports/suggests
+    cSplit(var, sep = " ,")
+
+  df <- tmp_df %>%
+    pivot_longer(c(2:ncol(tmp_df)), names_to = "type", values_to = "pkg") %>%
+    filter(!is.na(pkg)) %>%
+    # check if the depends/imports/suggests is also a DoE package (is in list_pkg)
+    mutate(survey_pkg_related = ifelse(pkg %in% list_pkg_ts, "yes", "no")) %>%
+    # filter only package whose depends/imports/suggests is a DoE
+    filter(survey_pkg_related == "yes") %>%
+    # tidy the data to only have the type: depends/imports/suggests
+    cSplit("type", sep = "_") %>%
+    select(-type_2) %>%
+    rename(type = type_1)
+
+  return(df)
+}
+
+ts_depends <- get_pkg_ts("Depends")
+ts_imports <- get_pkg_ts("Imports")
+ts_suggests <- get_pkg_ts("Suggests")
+
+pkg_edges_ts <- rbind(ts_depends, ts_imports, ts_suggests)
+
+ts_from <- pkg_edges_ts %>%
+  distinct(Package)
+
+# extract unique package that being depends/imports/suggests
+ts_to <- pkg_edges_ts %>%
+  distinct(pkg)
+
+# create nodes from the two previous dataframe
+pkg_nodes_ts <- full_join(ts_from, ts_to, by = c("Package" = "pkg"))
+```
+
 Interaction between packages could indicate of how developers utilize the functionality of one or more packages to create new features in their packages.
@@
 This number is relatively lower compared to the percentage of interaction in the packages in Survey and Time Series Analysis CRAN Task View.
-We found that `r round(73/132*100,2)`% Survey packages that depends, imports, and suggests each other, while these interconnection is observed in `r round(212/338*100,2)`% Time Series Analysis packages.
+We found that `r round(nrow(pkg_nodes_survey)/nrow(survey_pkg)*100,2)`% Survey packages that depends, imports, and suggests each other, while these interconnection is observed in `r round(nrow(nrow(pkg_nodes_ts)/nrow(ts_pkg)*100,2)`% Time Series Analysis packages.
 

### 2021-10-10 paper/package_interconnection.Rmd
Change the ifelse code to be case_when.

@@
 pkg_nodes <- aggregate(author ~ L2, aut_tidy_join, FUN = toString) %>%
-  mutate(note = ifelse(str_detect(author, "Ulrike"), "Groemping",
-                       ifelse(str_detect(L2, "Dice"), "Dupuy, Roustant, Richet, Deville",
-                              ifelse(str_detect(author, "Gramacy"), "Gramacy, Taddy",
-                                     ifelse(str_detect(author, "Shwetank"), "Lall, Vhargese",
-                                            ifelse(str_detect(author, "Kevin Wright"), "Wright",
-                                                   ifelse(str_detect(author, "Peng Liu"), "Liu",
-                                                          ifelse(str_detect(author, "Ernesto"), "Barrios", "Other")))))))) %>%
-  mutate(name_add = ifelse(L2 == "plgp", "plgp (G)",
-                     ifelse(L2 == "tgp", "tgp (G,T)",
-                            ifelse(L2 == "dynaTree", "dynaTree (G,T)",
-                                   ifelse(L2 == "DiceDesign", "DiceDesign (Du,Ro)",
-                                          ifelse(L2 == "DiceEval", "DiceEval (Du)",
-                                                 ifelse(L2 == "DiceKriging", "DiceKriging (Ro;De;Ri)",
-                                                        ifelse(L2 == "DiceView", "DiceView (De;Ri)", L2))))))))
+  mutate(note = case_when(str_detect(author, "Ulrike") ~ "Groemping",
+                          str_detect(L2, "Dice") ~ "Dupuy, Roustant, Richet, Deville",
+                          str_detect(author, "Gramacy") ~ "Gramacy, Taddy",
+                          str_detect(author, "Shwetank") ~ "Lall, Vhargese",
+                          str_detect(author, "Kevin Wright") ~ "Wright",
+                          str_detect(author, "Peng Liu") ~ "Liu",
+                          str_detect(author, "Ernesto") ~ "Barrios")) %>%
+  mutate(note = ifelse(is.na(note), "Other", note)) %>%
+  mutate(name_add = case_when(L2 == "plgp" ~ "plgp (G)",
+                              L2 == "tgp" ~ "tgp (G,T)",
+                              L2 == "dynaTree" ~ "dynaTree (G,T)",
+                              L2 == "DiceDesign" ~ "DiceDesign (Du,Ro)",
+                              L2 == "DiceEval" ~ "DiceEval (Du)",
+                              L2 == "DiceKriging" ~ "DiceKriging (Ro;De;Ri)",
+                              L2 == "DiceView" ~ "DiceView (De;Ri)")) %>%
+  mutate(name_add = ifelse(is.na(name_add), L2, name_add))
 

### 2021-10-24 tidy_author.Rmd
Add author tidying file.

---
title: "tidy_author"
author: "Dewi Amaliah"
date: "24/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(qdapRegex)
library(entity)
```

```{r}
url <- "http://cran.rstudio.com/web/packages/packages.rds"
db <- readRDS(url(url)) %>% 
  as.data.frame()

ctvs <- map_dfr(ctv::available.views(repos = "http://cran.rstudio.com/"), 
                  ~tibble(name = .x[["name"]],
                          topic = .x[["topic"]],
                          pkgs = map(name, ~ctv:::.get_pkgs_from_ctv_or_repos(.x, 
                                                repos = "http://cran.rstudio.com/")[[1]])))

ctvdf <- ctvs %>% 
  unnest_longer(pkgs) %>% 
  left_join(select(db, Package, Author, Depends, Imports, Suggests), 
            by = c("pkgs" = "Package")) %>% 
  mutate(across(c(Depends, Imports, Suggests),
                ~str_split(.x, ", "))) %>% 
  mutate(across(c(Depends, Imports, Suggests),
                ~map(.x, function(apkg) str_extract(apkg, "^[A-Za-z0-9]+")))) %>%
  # remove angle bracket
  mutate(author_tidy = rm_angle(Author, replacement = ""),
         # remove square bracket
         author_tidy = rm_square(author_tidy, replacement = ""),
         # remove bracket
         author_tidy = rm_bracket(author_tidy, replacement = ""),
         author_tidy = str_replace(author_tidy, "&", ","),
         author_tidy = str_replace(author_tidy, '\\\\', ""),
         # remove white space before comma
         author_tidy = rm_white_comma(author_tidy, replacement = ""),
         # replace "and" with comma
         author_tidy = str_replace(author_tidy, " and", ","),
         # remove extra comma 
         author_tidy = str_replace(author_tidy, ",,", ","),
         author_tidy2 = ifelse(str_detect(author_tidy, "riginal") | 
                                 str_detect(author_tidy, " by") | 
                                 str_detect(author_tidy, "with") | 
                                 str_detect(author_tidy, " under"), 
                               person_entity(author_tidy), 
                               author_tidy))

ctvdf_sentences <- ctvdf %>%
  filter(str_detect(author_tidy, "riginal") | str_detect(author_tidy, " by") | str_detect(author_tidy, "with") | str_detect(author_tidy, " under"))

ctvdf_sentences_test <- ctvdf_sentences %>%
  mutate(author_tidy2 = person_entity(author_tidy)) %>%
  mutate(author_tidy3 = str_replace(author_tidy, "with contributions from", ","),
         author_tidy3 = str_replace(author_tidy, "Original by", ""))

ctvdf_with <- ctvdf %>%
  filter(str_detect(author_tidy, "with"))

# try to filter English words

print(ctvdf_with$author_tidy)
```

### 2021-10-31 tidy_author.Rmd
Identify author written in sentences to get key words to remove.

@@
                 ~map(.x, function(apkg) str_extract(apkg, "^[A-Za-z0-9]+")))) %>%
-  # remove angle bracket
-  mutate(author_tidy = rm_angle(Author, replacement = ""),
-         # remove square bracket
-         author_tidy = rm_square(author_tidy, replacement = ""),
-         # remove bracket
-         author_tidy = rm_bracket(author_tidy, replacement = ""),
-         author_tidy = str_replace(author_tidy, "&", ","),
-         author_tidy = str_replace(author_tidy, '\\\\', ""),
-         # remove white space before comma
-         author_tidy = rm_white_comma(author_tidy, replacement = ""),
-         # replace "and" with comma
-         author_tidy = str_replace(author_tidy, " and", ","),
-         # remove extra comma 
-         author_tidy = str_replace(author_tidy, ",,", ","),
-         author_tidy2 = ifelse(str_detect(author_tidy, "riginal") | 
-                                 str_detect(author_tidy, " by") | 
-                                 str_detect(author_tidy, "with") | 
-                                 str_detect(author_tidy, " under"), 
-                               person_entity(author_tidy), 
-                               author_tidy))
+  mutate(author_tidy = str_replace_all(Author, "\\s+", " "),
+         author_tidy = str_replace_all(author_tidy, "<.*?>", ""),
+         author_tidy = str_replace_all(author_tidy, "\\[.*?\\]", ""),
+         author_tidy = str_replace_all(author_tidy, "\\(.*?\\)", ""),
+         author_tidy = str_replace_all(author_tidy, "(&| and)", ","),
+         author_tidy = str_replace_all(author_tidy, "\\(.*\\)", ""),
+         author_tidy = str_replace_all(author_tidy, "\\.*$", ""),
+         author_tidy = str_replace_all(author_tidy, ",,", ","),
+         author_tidy = str_replace_all(author_tidy, " \\)", ""),
+         author_tidy = str_replace_all(author_tidy, "\\s*,", ","),
+         author_tidy = str_squish(author_tidy))
+
+# words_to_delete <- c("Original by ",
+#                      "Compiled by ",
+#                      "is the sole author, creator of the BOOM project.
+Some code in the BOOM libraries has been modified from other open source projects.
+These include Cephes, NEWUOA, a modified version of the R math libraries .
+Original copyright notices have been maintained in all source files.
+In these cases, copyright claimed by Steven L.
+Scott is limited to modifications made to the original code.
+Google claims copyright for code written while Steven L.
+Scott was employed at Google from 2008 - 2018, but BOOM is not an officially supported Google project",
+#                      " original C code for ARMS by ",
+#                      "with contributions from ",
+#                      "originally written by ",
+#                      " changes, packaged by ",
+#                      " With considerable contributions by ",
+#                      " Ported to S-PLUS by ",
+#                      "except dchud.f, dchdd.f, which were written by ",
+#                      " with contributions from ",
+#                      "based on Fortran code by")
 
 ctvdf_sentences <- ctvdf %>%
-  filter(str_detect(author_tidy, "riginal") | str_detect(author_tidy, " by") | str_detect(author_tidy, "with") | str_detect(author_tidy, " under"))
+  filter(str_detect(author_tidy, "riginal") | str_detect(author_tidy, " by ") | str_detect(author_tidy, "with") |
+           str_detect(author_tidy, "under")| str_detect(author_tidy, "sole") | 
+           str_detect(author_tidy,"written") | str_detect(author_tidy, "port") | str_detect(author_tidy, " work")|
+           str_detect(author_tidy,"suggestion")|str_detect(author_tidy,"code"))
 
@@
 
-print(ctvdf_with$author_tidy)
+print(ctvdf_sentences$author_tidy)
 ```

### 2021-10-31 tidy_author.Rmd
tidy the author written in sentences form.

@@
          author_tidy = str_replace_all(author_tidy, " \\)", ""),
+         author_tidy = str_replace_all(author_tidy, "\\\\", ""),
          author_tidy = str_replace_all(author_tidy, "\\s*,", ","),
@@
 
-# words_to_delete <- c("Original by ",
-#                      "Compiled by ",
+# words_to_delete <- c("Compiled by ",
 #                      "is the sole author, creator of the BOOM project.
@@
 
+library(yaml)
+
 ctvdf_sentences <- ctvdf %>%
-  filter(str_detect(author_tidy, "riginal") | str_detect(author_tidy, " by ") | str_detect(author_tidy, "with") |
-           str_detect(author_tidy, "under")| str_detect(author_tidy, "sole") | 
-           str_detect(author_tidy,"written") | str_detect(author_tidy, "port") | str_detect(author_tidy, " work")|
-           str_detect(author_tidy,"suggestion")|str_detect(author_tidy,"code"))
-
-ctvdf_sentences_test <- ctvdf_sentences %>%
-  mutate(author_tidy2 = person_entity(author_tidy)) %>%
-  mutate(author_tidy3 = str_replace(author_tidy, "with contributions from", ","),
-         author_tidy3 = str_replace(author_tidy, "Original by", ""))
-
-ctvdf_with <- ctvdf %>%
-  filter(str_detect(author_tidy, "with"))
-
-# try to filter English words
+  filter(str_detect(author_tidy, "riginal") | 
+           str_detect(author_tidy, " by ") | 
+           str_detect(author_tidy, "with ") |
+           str_detect(author_tidy, " under ") | 
+           str_detect(author_tidy, "sole") | 
+           str_detect(author_tidy,"written") | 
+           str_detect(author_tidy, "port") | 
+           str_detect(author_tidy, " work")|
+           str_detect(author_tidy,"suggestion") | 
+           str_detect(author_tidy,"code") | 
+           str_detect(author_tidy, " from ")) %>%
+  select(pkgs, author_tidy) %>%
+  distinct()
 
 print(ctvdf_sentences$author_tidy)
+
+# manually tidy the author in sentences due to the variability
+
+tidy_aut <- tribble(~tidy_aut,
+                    "Kjetil Halvorsen",
+                    "Steven L. Scott",
+                    "Giovanni Petris, Luca Tardella, Wally Gilks",
+                    "Joseph L. Schafer",
+                    "David I. Hastie, Silvia Liverani, Sylvia Richardson, Aurore J. Lavigne, Lucy Leigh, Lamiae Azizi, Xi Liu, Ruizhu Huang, Austin Gratton, Wei Jing",
+                    "Simon Jackman, Alex Tahk, Achim Zeileis, Christina Maimone, Jim Fearon, Zoe Meers",
+                    "Andrew Gelman, Sibylle Sturtz, Uwe Ligges, Gregor Gorjanc, Jouni Kerman",
+                    "Madeleine Thompson, G. W. Stewart",
+                    "Severine Demeyer, Alexandre Allard, Bertrand Iooss",
+                    "Thomas Lumley, Alan Miller",
+                    "Alessandra R. Brazzale, Ruggero Bellio, Douglas Bates",
+                    "Barry W. Brown, James Lovato, Kathy Russel, Kjetil B Halvorsen",
+                    "Daniel G. Muenz, Ray Griner, Huichao Chen, Lijuan Deng, Sachiko Miyahara, Scott R. Evans, Lingling Li, Hajime Uno, Laura M. Smeaton",
+                    "Michal Juraska, Peter B. Gilbert, Xiaomin Lu, Min Zhang, Marie Davidian, Anastasios A. Tsiatis",
+                    "Peter Langfelder, Bin Zhang, Steve Horvath",
+                    "Campo Elias Pardo, Pedro Cesar del Campo, Camilo Jose Torres, Ivan Diaz, Mauricio Sadinle, Jhonathan Medina",
+                    "Fionn Murtagh, R development team, Peter Langfelder",
+                    "Steve Su, Diethelm Wuertz, Martin Maechler, Rmetrics core team members, Juha Karvanen,
+                    Robert King, Benjamin Dean, R core team",
+                    "Peter Macdonald, Juan Du",
+                    "Ulric Lund, Claudio Agostinelli",
+                    "Marius Hofert, Ivan Kojadinovic, Martin Maechler, Jun Yan, Johanna G. Nešlehová, Rebecca Morger",
+                    "Alec Stephenson, Chris Ferro",
+                    "Alessandra R. Brazzale, Douglas Bates",
+                    "Yohan Chalabi, Christophe Dutang, Petr Savicky, Diethelm Wuertz, M. Matsumoto, M. Saito, D. Knuth.", 
+                    "Robert King, Emily Anderson",
+                    "Stefan Wilhelm, Manjunath B G",
+                    "Rob Tibshirani, Friedrich Leisch",
+                    "Roger Koenker, Stephen Portnoy, Pin Tian Ng, Blaise Melly, Achim Zeileis, Philip Grosjean, Cleve Moler, Yousef Saad, Victor Chernozhukov, Ivan Fernandez-Val, Brian D Ripley", 
+                    "Ariane Straub, Torsten Hothorn", 
+                    "Stéphane Dray, Anne-Béatrice Dufour, Jean Thioulouse, with contributions from Thibaut Jombart, Sandrine Pavoine, Jean R.
+Lobry, Sébastien Ollier, Daniel Borcard, Pierre Legendre, Stéphanie Bougeard, Aurélie Siberchicot, Daniel Chessel",
+                    "C.J.F. ter Braak, A.P. Schaffers, Gavin L. Simpson, Ron Wehrens, Bjorn-Helge Mevik",
+                    "Stephen Milborrow, Trevor Hastie, Rob Tibshirani, Thomas Lumley", 
+                    "Janet E. Heffernan, Alec G. Stephenson",
+                    "Jeff Laake, Devin Johnson, Paul Conn, Jay Rotella",
+                    "Trevor Hastie, Robert Tibshirani, Friedrich Leisch, Kurt Hornik, Brian D. Ripley,  Balasubramanian Narasimhan",
+                    "Leo Breiman, Adele Cutler, Andy Liaw, Matthew Wiener",
+                    "Jeff Laake, Eldar Rakhimberdiev, Ben Augustine, Daniel Turek, Brett McClintock, Bret Collier, Jay Rotella, David Pavlacky, Andrew Paul, Luke Eberhart-Phillips, Jake Ivan, Connor Wood",
+                    "Bertrand Iooss, Sebastien Da Veiga, Alexandre Janon, Gilles Pujol, Baptiste Broto, Khalid Boumhaout, Thibault Delage, Reda El Amri, Jana Fruth, Laurent Gilquin, Joseph Guillaume, Margot Herin, Marouane Il Idrissi, Loic Le Gratiet, Paul Lemaitre, Amandine Marrel, Anouar Meynaoui, Barry L.
+Nelson, Filippo Monari, Roelof Oomen, Oldrich Rakovec, Bernardo Ramos, Olivier Roustant, Eunhye Song, Jeremy Staum, Roman Sueur, Taieb Touati, Vanessa Verges, Frank Weber",
+                    "Ernesto Barrios, Daniel Meyer",
+                    "Douglas Bates, Karen A.F. Copeland",
+                    "John Lawson, G. Vining",
+                    "Jeff Enos, David Kane, Kyle Campbell, Daniel Gerlanc, Aaron Schwartz, Daniel Suo, Alexei Colin, Luyi Zhao",
+                    "Matt Dowle, Arun Srinivasan, Jan Gorecki, Michael Chirico, Pasha Stetsenko, Tom Short, Steve Lianoglou, Eduard Antonyan, Markus Bonsch, Hugh Parsonage, Scott Ritchie, Kun Ren, Xianying Tan, Rick Saporta, Otto Seiskari, Xianghui Dong, Michel Lang, Watal Iwasaki, Seth Wenchel, Karl Broman, Tobias Schmidt, David Arenburg, Ethan Smith, Francois Cocquemas, Matthieu Gomez, Philippe Chataignon, Nello Blaser, Dmitry Selivanov, Andrey Riabushenko, Cheng Lee, Declan Groves, Daniel Possenriede, Felipe Parages, Denes Toth, Mus Yaramaz-David, Ayappan Perumal, James Sams, Martin Morgan, Michael Quinn, Javrucebo, Marc Outins, Roy Storey, Manish Saraswat, Morgan Jacob, Michael Schubmehl, Davis Vaughan, Toby Hocking, Leonardo Silvestri, Tyson Barrett, Jim Hester, Anthony Damico, Sebastian Freundt, David Simons, Elliott Sales de Andrade, Cole Miller, Jens Peder Meldgaard, Vaclav Tlapak, Kevin Ushey, Dirk Eddelbuettel, Ben Schwen",
+                    "Jan Beran, Brandon Whitcher, Martin Maechler",
+                    "Sarah Brockhaus, David Ruegamer, Almond Stoecker, Torsten Hothorn", 
+                    "Alessio Boattini, Federico C. F. Calboli, Vincente Canto Cassola, Martin Maechler",  
+                    "Jing Hua Zhao, Kurt Hornik, Brian Ripley",
+                    "Gregory Warnes, Gregor Gorjanc, Friedrich Leisch, Michael Man",
+                    "Jon Clayden, K. Kosako, K. Takata",
+                    "K. Burkett, B. McNeney, J. Graham, Zhijian Chen",
+                    "John Maindonald, Randall Johnson",
+                    "Thomas Hoffmann, Christoph Lange",
+                    "Karl W Broman, Hao Wu, Gary Churchill, Saunak Sen, Danny Arends, Robert Corty, Timothee Flutre, Ritsert Jansen, Pjotr Prins, Lars Ronnegard, Rohan Shah, Laura Shannon, Quoc Tran, Aaron Wolen, Brian Yandell, R Core Team",
+                    "Augustin Luna, Kristin K. Nicodemus,",
+                    "David W. Scott, Albrecht Gebhardt, Stephen Kaluzny",
+                    "Dan Carr, Nicholas Lewin-Koh, Martin Maechler, Deepayan Sarkar",
+                    "Martin Maechler, Kurt Hornik, Brian Ripley",
+                    "Nele Schuwirth, Peter Reichert, Simone Langhans, Christian Michel", 
+                    "Michael U. Kemp, E. Emiel van Loon, Judy Shamoun-Baranes, Willem Bouten",
+                    "Justin Lokhorst, Bill Venables, Berwin Turlach, Martin Maechler",  
+                    "Zhu Wang, Achim Zeileis, Simon Jackman, Brian Ripley, Patrick Breheny",
+                    "Marijke Welvaert, Joke Durnez, Beatrijs Moerkerke, Yves Rosseel, Geert Verdoolaege",
+                    "Thomas Quinn, Kristin Nicodemus, Glaucia Martins, Carlos Maia", 
+                    "Mark Stevenson, Evan Sergeant, Telmo Nunes, Cord Heuer, Jonathon Marshall, Javier Sanchez, Ron Thornton, Jeno Reiczigel, Jim Robison-Cox, Paola Sebastiani, Peter Solymos, Kazuki Yoshida, Geoff Jones, Sarah Pirikahu, Simon Firestone, Ryan Kyle, Johann Popp, Mathew Jay, Charles Reynard, Allison Cheung, Nagendra Singanallur, Aniko Szabo",
+                    "Ted Harding, Fernando Tusell, Joseph L. Schafer",
+                    "Il-Youp Kwak, Wuming Gong",
+                    "Joseph L. Schafer", 
+                    "Alvaro A. Novo, Joseph L. Schafer",
+                    "Joseph L. Schafer",
+                    "Christian Buchta, Michael Hahsler, Daniel Diaz",
+                    "Naisyin Wang, Adrian Raftery, Chris Fraley",
+                    "Gregory R. Warnes, Ben Bolker, Thomas Lumley, Randall C. Johnson, Intramural Research Program", 
+                    "Marek Gagolewski, Bartek Tartanus",
+                    "Bill Venables, Kurt Hornik, Georgi Boshnakov",
+                    "John P. Nolan, Alan Genz",
+                    "R. J. Renka, Albrecht Gebhardt, Stephen Eglen, Sergei Zuyev, Denis White",
+                    "Stephen Tueller",
+                    "Josh Pasek, Alex Tahk, R core team, Gene Culter, Marcus Schwemmle",
+                    "Ben B. Hansen, Mark Fredrickson, Josh Buckner, Josh Errickson, Adam Rauh, Peter Solenberger, Dimitri P. Bertsekas, Paul Tseng",
+                    "Berwin A. Turlach, Andreas Weingessel, Cleve Moler)",
+                    "Charles J. Geyer, Glen D. Meeden, Komei Fukuda",
+                    "Jon Clayden, Ivan Zelinka",
+                    "Guillaume Guenard, Pierre Legendre",
+                    "Julien Clavel, Aaron King, Emmanuel Paradis",
+                    "Claudia van Borkulo, Sacha Epskamp, Alexander Robitzsch",
+                    "Seung W. Choi, Laura E. Gibbons, Paul K. Crane",
+                    "Kenneth Knoblauch, Laurence T. Maloney, Laurence T. Maloney, J. N. Yang",
+                    "Anton Grau Larsen, Jacob Lunding, Christoph Ellersgaard, Stefan Andrade",
+                    "Benjamin Hofner",
+                    "Eliot J B McIntire, Alex M Chubaty, Tati Micheletti, Ceres Barros, Ian Eddy, Minister of Natural Resources Canada",
+                    "Eliot J B McIntire, the Minister of Natural Resources Canada",
+                    "Johan Zetterqvist, Arvid Sjölander, Alexander Ploner",
+                    "Rob J Hyndman, Heather Booth, Leonie Tickle, John Maindonald",
+                    "Vincent J Carey, Thomas Lumley, Brian Ripley, Cleve Moler",
+                    "James E. Helmreich, Robert M. Pruzek, KuangNan Xiong",
+                    "Adrian Bowman, Adelchi Azzalini, B. D. Ripley, Adrian Bowman, Adelchi Azzalini",
+                    "Clement Calenge, Scott Fortmann-Roe",
+                    "Clement Calenge, Mathieu Basille",
+                    "Clement Calenge, Stephane Dray, Manuela Royer",
+                    "Clement Calenge, Mathieu Basille",
+                    "Virgilio Gómez-Rubio, Juan Ferrándiz-Ferragud, Antonio López-Quílez, Roger Bivand",
+                    "Marcelino de la Cruz Rot, Philip M. Dixon, Jose M. Blanco-Moreno",
+                    "Richard Detomasi, Ministerio de Desarrollo Social, Infraestrutura de Datos Espaciales, Servicio Geográfico Militar, Ministerio de Transporte y Obras Públicas, Instituto Nacional de Estadística,  Instituto Geográfico Militar, Ministerio de Vivienda, Ordenamiento Territorial y Medio Ambiente",
+                    "Edzer Pebesma, Jon Olav Skoien, Olivier Baume, A.Chorti, D.T. Hristopulos, H. Kazianka, S.J. Melles, G. Spiliopoulos",
+                    "Christian Graul, Francois Guillem",
+                    "Richard A. Becker, Allan R. Wilks, Ray Brownrigg",
+                    "Doug McIlroy, Ray Brownrigg, Thomas P Minka, Roger Bivand",
+                    "Richard A. Becker, Allan R. Wilks, Ray Brownrigg, Thomas P Minka, Alex Deckmyn",
+                    "Ian Fellows, Jan Peter Stotz",
+                    "Seong-Yun Hong, David O'Sullivan, Benjamin Jarvis, Changlock Choi",
+                    "James J. Majure, Iowa State University, Albrecht Gebhardt",
+                    "Moreno I. Coco, Nicholas D. Duran, Rick Dale, Denis O'Hora, Michael J. Spivey.",
+                    "Mai Zhou, Yifan Yang",
+                    "William H. Barton, Dr. Mai Zhou",
+                    "Isabelle Clerc-Urmès, Michel Grzebyk, Guy Hédelin, CENSUR working survival group",
+                    "Klein, Moeschberger, Jun Yan",
+                    "Gareth Ambler, Axel Benner",
+                    "Kenneth Hess, R. Gentleman",
+                    "Patrick J. Heagerty, Paramita Saha-Chaudhuri",
+                    "Patrick J. Heagerty, Paramita Saha-Chaudhuri",
+                    "Thomas Scheike, Torben Martinussen, Jeremy Silver, Klaus Holst",
+                    "F.L. Ramsey, D.W. Schafer, Jeannie Sifneos, Berwin A. Turlach, Nicholas Horton, Kate Aloisio, Ruobing Zhang, Randall Pruim",
+                    "F.L. Ramsey, D.W. Schafer, Jeannie Sifneos, Berwin A. Turlach, Nicholas Horton, Linda Loi, Kate Aloisio, Ruobing Zhang, Randall Pruim",
+                    "Vasilis Dakos, S.R. Carpenter, T. Cline, L. Lahti",
+                    "Thomas Ruf, Press et al.",
+                    "Mustafa Gokce Baydogan",
+                    "Roger D. Peng, Aidan McDermott",
+                    "Dirk Eddelbuettel, Bill Evans, Mike Birdgeneau, Henrik Bengtsson, Seth Wenchel, Colin Gillespie, Chan-Yub Park",
+                    "Bob Rudis, Jay Jacobs, David Severski, Quinn Weber, Konrad Karczewski, Shinya Uryu, Gregory Jefferis, Ed Niles, Rick Saporta, Jonathan Sidi, Matt Kaye, Xinye Li, Andrie de Vries",
+                    "Andrew Thomas, Dave Lunn, David Spiegelhalter, Nicky Best, Uwe Ligges, Sibylle Sturtz, Andrew Gelman, Gregor Gorjanc, Chris Jackson. Linux port, Chris Jackson",
+                    "Andrew Gelman, Sibylle Sturtz, Uwe Ligges, Gregor Gorjanc, Jouni Kerman, Neal Thomas")
+
+tidy_aut2 <- cbind(ctvdf_sentences, tidy_aut) %>%
+  select(pkgs, tidy_aut)
+
+ctvdf_author <- left_join(ctvdf, tidy_aut2, by = "pkgs") %>%
+  mutate(tidy_aut = if_else(is.na(tidy_aut), author_tidy, tidy_aut),
+         tidy_aut = str_replace_all(tidy_aut, ",,", ","))
 ```

### 2021-11-21 tidy_author.Rmd
Manually inspecting and tidying the author column.

@@
          author_tidy = str_replace_all(author_tidy, "\\(.*?\\)", ""),
+         author_tidy = str_replace_all(author_tidy, "\\(.*\\)", ""),
          author_tidy = str_replace_all(author_tidy, "(&| and)", ","),
-         author_tidy = str_replace_all(author_tidy, "\\(.*\\)", ""),
          author_tidy = str_replace_all(author_tidy, "\\.*$", ""),
@@
 
-library(yaml)
-
 ctvdf_sentences <- ctvdf %>%
@@
            str_detect(author_tidy,"code") | 
-           str_detect(author_tidy, " from ")) %>%
+           str_detect(author_tidy, " from ") |
+           str_detect(author_tidy, " for ") |
+           str_detect(author_tidy, "FORTRAN") |
+           str_detect(pkgs, "fBasics") |
+           str_detect(pkgs, "RobPer") |
+           str_detect(pkgs, "dfpk") |
+           str_detect(pkgs, "nor1") |
+           str_detect(pkgs, "mkin") |
+           str_detect(author_tidy, "Imperial College of Science, Technology, Medicine") |
+           str_detect(pkgs, "RH2") |
+           str_detect(pkgs, "DPQ") |
+           str_detect(pkgs, "RDieHarder") |
+           str_detect(pkgs, "sfsmisc") |
+           str_detect(pkgs, "oce") |
+           str_detect(pkgs, "DiceKriging") |
+           str_detect(pkgs, "fracdiff") |
+           str_detect(pkgs, "RcppQuantuccia") |
+           str_detect(pkgs, "dbstats") |
+           str_detect(pkgs, "drake") |
+           str_detect(pkgs, "proffer") |
+           str_detect(pkgs, "targets") |
+           str_detect(pkgs, "GSODR") |
+           str_detect(pkgs, "MODIStsp") |
+           str_detect(pkgs, "baggr") |
+           str_detect(pkgs, "EvidenceSynthesis") |
+           str_detect(pkgs, "SAMURAI") |
+           str_detect(pkgs, "crfsuite") |
+           str_detect(pkgs, "udpipe") |
+           str_detect(pkgs, "Rmpfr") |
+           str_detect(pkgs, "EdSurvey") |
+           str_detect(pkgs, "rrcov3way") |
+           str_detect(pkgs, "RcppDE") |
+           str_detect(pkgs, "irr") |
+           str_detect(pkgs, "irtDemo") |
+           str_detect(pkgs, "kcirt") |
+           str_detect(pkgs, "MLCIRTwithin") |
+           str_detect(pkgs, "MultiLCIRT") |
+           str_detect(pkgs, "HTMLUtils") |
+           str_detect(pkgs, "workflowr") |
+           str_detect(pkgs, "landscapemetrics") |
+           str_detect(pkgs, "bentcableAR") |
+           str_detect(pkgs, "BMTAR") |
+           str_detect(pkgs, "MARSS") |
+           str_detect(pkgs, "MultipleBubbles")) %>%
   select(pkgs, author_tidy) %>%
@@
 
+```
+
+```{r}
 # manually tidy the author in sentences due to the variability
@@
                     "Alessandra R. Brazzale, Ruggero Bellio, Douglas Bates",
+                    "Anita M. Thieler, Jonathan Rathjens, Roland Fried, Brenton R. Clarke, Uwe Ligges, Matias Salibian-Barrera, Gert Willems, Victor Yohai",
                     "Barry W. Brown, James Lovato, Kathy Russel, Kjetil B Halvorsen",
+                    "Artemis Toumazi, Sarah Zohar, Moreno Ursino, Trustees of Columbia University, French National Institute of Health and Medical Research",
+                    "Charlie Casper, Oscar A. Perez",
                     "Daniel G. Muenz, Ray Griner, Huichao Chen, Lijuan Deng, Sachiko Miyahara, Scott R. Evans, Lingling Li, Hajime Uno, Laura M. Smeaton",
@@
                     "Peter Macdonald, Juan Du",
+                    "Martin Maechler, Friedrich Leisch, Erik Jørgensen",
+                    "Adelchi Azzalini, Giovanna Menardi, Adelchi Azzalini, Tiziana Rosolin", 
+                    "G. Grothendieck, Thomas Mueller",
+                    "Rich FitzJohn, Wes Hinsley, Imperial College of Science, Technology and Medicine",
+                    "Johannes Ranke, Katrin Lindenberger, René Lehmann, Eurofins Regulatory AG",
                     "Ulric Lund, Claudio Agostinelli",
                     "Marius Hofert, Ivan Kojadinovic, Martin Maechler, Jun Yan, Johanna G. Nešlehová, Rebecca Morger",
+                    "Martin Maechler, Morten Welinder, Wolfgang Viechtbauer, Ross Ihaka, Marius Hofert, R-core, R Foundation",
                     "Alec Stephenson, Chris Ferro",
+                    "Diethelm Wuertz, Tobias Setz, Yohan Chalabi, Martin Maechler",
                     "Alessandra R. Brazzale, Douglas Bates",
                     "Yohan Chalabi, Christophe Dutang, Petr Savicky, Diethelm Wuertz, M. Matsumoto, M. Saito, D. Knuth.", 
+                    "Dirk Eddelbuettel, Robert G Brown, David Bauer",
+                    "Martin Maechler, Werner Stahel, Andreas Ruckstuhl, Christian Keller, Kjetil Halvorsen, Alain Hauser, iChristoph Buser, Lorenz Gygax, Bill Venables, Tony Plate, Isabelle Flückiger, Marcel Wolbers, Markus Keller, Sandrine Dudoit, Jane Fridlyand, Greg Snow, Henrik Aa.
+Nielsen, Vincent Carey, Ben Bolker, Philippe Grosjean, Frédéric Ibanez, Caterina Savi, Charles Geyer, Jens Oehlschlägel",
                     "Robert King, Emily Anderson",
@@
                     "Rob Tibshirani, Friedrich Leisch",
+                    "Douglas Bates, Martin Maechler, Timothy A. Davis, Jens Oehlschlägel, Jason Riedy, Regents of the University of California, R Core Team",
                     "Roger Koenker, Stephen Portnoy, Pin Tian Ng, Blaise Melly, Achim Zeileis, Philip Grosjean, Cleve Moler, Yousef Saad, Victor Chernozhukov, Ivan Fernandez-Val, Brian D Ripley", 
                     "Ariane Straub, Torsten Hothorn", 
-                    "Stéphane Dray, Anne-Béatrice Dufour, Jean Thioulouse, with contributions from Thibaut Jombart, Sandrine Pavoine, Jean R.
+                    "Stéphane Dray, Anne-Béatrice Dufour, Jean Thioulouse, Thibaut Jombart, Sandrine Pavoine, Jean R.
 Lobry, Sébastien Ollier, Daniel Borcard, Pierre Legendre, Stéphanie Bougeard, Aurélie Siberchicot, Daniel Chessel",
@@
                     "Trevor Hastie, Robert Tibshirani, Friedrich Leisch, Kurt Hornik, Brian D. Ripley,  Balasubramanian Narasimhan",
+                    "Dan Kelley, Clark Richards, Chantelle Layton, British Geological Survey",
                     "Leo Breiman, Adele Cutler, Andy Liaw, Matthew Wiener",
@@
                     "Ernesto Barrios, Daniel Meyer",
+                    "Olivier Roustant, David Ginsbourger, Yves Deville, Clement Chevalier, Yann Richet",
                     "Douglas Bates, Karen A.F. Copeland",
@@
                     "Matt Dowle, Arun Srinivasan, Jan Gorecki, Michael Chirico, Pasha Stetsenko, Tom Short, Steve Lianoglou, Eduard Antonyan, Markus Bonsch, Hugh Parsonage, Scott Ritchie, Kun Ren, Xianying Tan, Rick Saporta, Otto Seiskari, Xianghui Dong, Michel Lang, Watal Iwasaki, Seth Wenchel, Karl Broman, Tobias Schmidt, David Arenburg, Ethan Smith, Francois Cocquemas, Matthieu Gomez, Philippe Chataignon, Nello Blaser, Dmitry Selivanov, Andrey Riabushenko, Cheng Lee, Declan Groves, Daniel Possenriede, Felipe Parages, Denes Toth, Mus Yaramaz-David, Ayappan Perumal, James Sams, Martin Morgan, Michael Quinn, Javrucebo, Marc Outins, Roy Storey, Manish Saraswat, Morgan Jacob, Michael Schubmehl, Davis Vaughan, Toby Hocking, Leonardo Silvestri, Tyson Barrett, Jim Hester, Anthony Damico, Sebastian Freundt, David Simons, Elliott Sales de Andrade, Cole Miller, Jens Peder Meldgaard, Vaclav Tlapak, Kevin Ushey, Dirk Eddelbuettel, Ben Schwen",
+                    "Martin Maechler, Chris Fraley, Friedrich Leisch, Valderio Reisen, Artur Lemonte, Rob Hyndman",
                     "Jan Beran, Brandon Whitcher, Martin Maechler",
+                    "Dirk Eddelbuettel",
+                    "Eva Boj, Adria Caballe, Pedro Delicado, Josep Fortiana",
                     "Sarah Brockhaus, David Ruegamer, Almond Stoecker, Torsten Hothorn", 
@@
                     "Gregory Warnes, Gregor Gorjanc, Friedrich Leisch, Michael Man",
-                    "Jon Clayden, K. Kosako, K. Takata",
                     "K. Burkett, B. McNeney, J. Graham, Zhijian Chen",
@@
                     "Martin Maechler, Kurt Hornik, Brian Ripley",
+                    "William Michael Landau, Alex Axthelm, Jasper Clarkberg, Kirill Müller, Ben Bond-Lamberty, Tristan Mahr, Miles McBain, Noam Ross, Ellis Hughes, Matthew Mark Strasiotto, Ben Marwick, Peter Slaughter, Eli Lilly and Company",
+                    "William Michael Landau, Eli Lilly and Company",
+                    "William Michael Landau, Matthew T. Warkentin, Samantha Oliver, Tristan Mahr, Eli Lilly and Company",
+                    "David Bronaugh",
                     "Nele Schuwirth, Peter Reichert, Simone Langhans, Christian Michel", 
+                    "Adam H. Sparks, Tomislav Hengl, Andrew Nelson, Hugh Parsonage, Taras Kaduk, Gwenael Giboire, Łukasz Pawlik, Ross Darnell, Tyler Widdison",
+                    "Lorenzo Busetto, Luigi Ranghetti, Leah Wasser, Jeff Hanson, Babak Naimi",
                     "Michael U. Kemp, E. Emiel van Loon, Judy Shamoun-Baranes, Willem Bouten",
+                    "Tobias Pilz, Potsdam Institute for Climate Impact Research",
+                    "Ernest Guevarra, Valid International, Water and Sanitation for the Urban Poor",
                     "Justin Lokhorst, Bill Venables, Berwin Turlach, Martin Maechler",  
@@
                     "Marijke Welvaert, Joke Durnez, Beatrijs Moerkerke, Yves Rosseel, Geert Verdoolaege",
+                    "Witold Wiecek, Rachael Meager, Brice Green, Trustees of Columbia University",
                     "Thomas Quinn, Kristin Nicodemus, Glaucia Martins, Carlos Maia", 
                     "Mark Stevenson, Evan Sergeant, Telmo Nunes, Cord Heuer, Jonathon Marshall, Javier Sanchez, Ron Thornton, Jeno Reiczigel, Jim Robison-Cox, Paola Sebastiani, Peter Solymos, Kazuki Yoshida, Geoff Jones, Sarah Pirikahu, Simon Firestone, Ryan Kyle, Johann Popp, Mathew Jay, Charles Reynard, Allison Cheung, Nagendra Singanallur, Aniko Szabo",
+                    "Martijn Schuemie, Marc A. Suchard, Observational Health Data Science and Informatics",
+                    "Noory Y. Kim, Shrikant I. Bangdiwala, Gerald Gartlehner",
                     "Ted Harding, Fernando Tusell, Joseph L. Schafer",
@@
                     "Alvaro A. Novo, Joseph L. Schafer",
+                    "Steven M.
+Boker, Michael C.
+Neale, Hermine H.
+Maes, Michael J.
+Wilde, Michael Spiegel, Timothy R.
+Brick, Ryne Estabrook, Timothy C.
+Bates, Paras Mehta, Timo von Oertzen, Ross J.
+Gore, Michael D.
+Hunter, Daniel C.
+Hackett, Julian Karch, Andreas M.
+Brandmaier, Joshua N.
+Pritikin, Mahsa Zahery, Robert M.
+Kirkpatrick, Yang Wang, Ben Goodrich, Charles Driver, Massachusetts Institute of Technology, S.
+G.
+Johnson, Association for Computing Machinery, Dieter Kraft, Stefan Wilhelm, Sarah Medland, Carl F.
+Falk, Matt Keller, Manjunath B G, The Regents of the University of California, Lester Ingber, Wong Shao Voon, Juan Palacios, Jiang Yang, Gael Guennebaud, Jitse Niesen",
                     "Joseph L. Schafer",
@@
                     "Gregory R. Warnes, Ben Bolker, Thomas Lumley, Randall C. Johnson, Intramural Research Program",
+                    "Jan Wijffels, BNOSAC, Naoaki Okazaki, Bob Jenkins, Jorge Nocedal, Jesse Long",
+                    "Jon Clayden, K. Kosako, K. Takata",
                     "Marek Gagolewski, Bartek Tartanus",
+                    "Jan Wijffels, BNOSAC, , Milan Straka, Jana Straková, Institute of Formal and Applied Linguistics, Faculty of Mathematics and Physics, Charles University in Prague",
+                    "Hiroshi Akima, Albrecht Gebhardt, Thomas Petzold, Martin Maechler, YYYY Association for Computing Machinery, Inc.",
                     "Bill Venables, Kurt Hornik, Georgi Boshnakov",
+                    "Yixuan Qiu, Jiali Mei, authors of the ARPACK library",
+                    "Douglas Bates, Dirk Eddelbuettel, Romain Francois, Yixuan Qiu",
+                    "Martin Maechler, Richard M. Heiberger, John C. Nash, Hans W. Borchers",
                     "John P. Nolan, Alan Genz",
                     "R. J. Renka, Albrecht Gebhardt, Stephen Eglen, Sergei Zuyev, Denis White",
+                    "Paul Bailey, Ahmad Emad, Huade Huo, Michael Lee, Yuqi Liao, Alex Lishinski, Trang Nguyen, Qingshu Xie, Jiao Yu, Ting Zhang, Eric Buehler, Sun-joo Lee, Emmanuel Sikali, Jeppe Bundsgaard, Ren C'deBaca, Anders Astrup Christensen",
                     "Stephen Tueller",
+                    "Valentin Todorov, Maria Anna Di Palma, Michele Gallo",
                     "Josh Pasek, Alex Tahk, R core team, Gene Culter, Marcus Schwemmle",
                     "Ben B. Hansen, Mark Fredrickson, Josh Buckner, Josh Errickson, Adam Rauh, Peter Solenberger, Dimitri P. Bertsekas, Paul Tseng",
-                    "Berwin A. Turlach, Andreas Weingessel, Cleve Moler)",
+                    "Berwin A. Turlach, Andreas Weingessel, Cleve Moler",
                     "Charles J. Geyer, Glen D. Meeden, Komei Fukuda",
+                    "Dirk Eddelbuettel",
                     "Jon Clayden, Ivan Zelinka",
@@
                     "Julien Clavel, Aaron King, Emmanuel Paradis",
+                    "Matthias Gamer, Jim Lemon, Ian Fellows, Puspendra Singh",
+                    "Metin Bulus, Wes Bonifay",
                     "Claudia van Borkulo, Sacha Epskamp, Alexander Robitzsch",
+                    "Dave Zes, Jimmy Lewis, Dana Landis",
                     "Seung W. Choi, Laura E. Gibbons, Paul K. Crane",
-                    "Kenneth Knoblauch, Laurence T. Maloney, Laurence T. Maloney, J. N. Yang",
+                    "Francesco Bartolucci, Silvia Bacci",
+                    "Kenneth Knoblauch, Laurence T. Maloney, J. N. Yang",
+                    "Francesco Bartolucci, Silvia Bacci, Michela Gnaldi",
                     "Anton Grau Larsen, Jacob Lunding, Christoph Ellersgaard, Stefan Andrade",
+                    "Markus Loecher",
+                    "Rich FitzJohn, Robert Ashton, Alex Hill, Martin Eden, Wes Hinsley, Emma Russell, James Thompson, Imperial College of Science, Technology and Medicine",
                     "Benjamin Hofner",
@@
                     "Eliot J B McIntire, the Minister of Natural Resources Canada",
+                    "John Blischak, Peter Carbonetto, Matthew Stephens, Luke Zappia, Pierre Formont, Tim Trice, Jiaxiang Li, Michael J. Kane, Anh Tran, Sydney Purdue",
                     "Johan Zetterqvist, Arvid Sjölander, Alexander Ploner",
+                    "Martin Maechler, Werner A. Stahel, Rolf Turner, Ueli Oetliker, Tobias Schoch",
                     "Rob J Hyndman, Heather Booth, Leonie Tickle, John Maindonald",
@@
                     "Marcelino de la Cruz Rot, Philip M. Dixon, Jose M. Blanco-Moreno",
+                    "Douglas Nychka, Reinhard Furrer, John Paige, Stephan Sain, Florian Gerber, Matthew Iverson, University Corporation for Atmospheric Research",
+                    "Rafael H.
+M.
+Pereira, Caio Nogueira Goncalves, Paulo Henrique Fernandes de Araujo, Guilherme Duarte Carvalho, Rodrigo Almeida de Arruda, Igor Nascimento, Barbara Santiago Pedreira da Costa, Welligtton Silva Cavedo, Pedro R.
+Andrade, Alan da Silva, Carlos Kauê Vieira Braga, Carl Schmertmann, Alessandro Samuel-Rosa, Daniel Ferreira, Ipea-Institue for Applied Economic Research",
                     "Richard Detomasi, Ministerio de Desarrollo Social, Infraestrutura de Datos Espaciales, Servicio Geográfico Militar, Ministerio de Transporte y Obras Públicas, Instituto Nacional de Estadística,  Instituto Geográfico Militar, Ministerio de Vivienda, Ordenamiento Territorial y Medio Ambiente",
                     "Edzer Pebesma, Jon Olav Skoien, Olivier Baume, A.Chorti, D.T. Hristopulos, H. Kazianka, S.J. Melles, G. Spiliopoulos",
+                    "Maximillian H.K.
+Hesselbarth, Marco Sciaini, Jakub Nowosad, Sebastian Hanss, Laura J.
+Graham, Jeffrey Hollister, Kimberly A.
+With, Florian Privé, Project Nayuki, Matt Strimas-Mackey",
                     "Christian Graul, Francois Guillem",
@@
                     "Ian Fellows, Jan Peter Stotz",
+                    "Andrea Gilardi, Robin Lovelace, Barry Rowlingson, Salva Fernández, Nicholas Potter",
+                    "Robert J.
+Hijmans, Jacob van Etten, Michael Sumner, Joe Cheng, Dan Baston, Andrew Bevan, Roger Bivand, Lorenzo Busetto, Mort Canty, Ben Fasoli, David Forrest, Aniruddha Ghosh, Duncan Golicher, Josh Gray, Jonathan A.
+Greenberg, Paul Hiemstra, Kassel Hingee, Institute for Mathematics Applied Geosciences, Charles Karney, Matteo Mattiuzzi, Steven Mosher, Babak Naimi, Jakub Nowosad, Edzer Pebesma, Oscar Perpinan Lamigueiro, Etienne B.
+Racine, Barry Rowlingson, Ashton Shortridge, Bill Venables, Rafael Wueest",
                     "Seong-Yun Hong, David O'Sullivan, Benjamin Jarvis, Changlock Choi",
@@
                     "F.L. Ramsey, D.W. Schafer, Jeannie Sifneos, Berwin A. Turlach, Nicholas Horton, Linda Loi, Kate Aloisio, Ruobing Zhang, Randall Pruim",
+                    "Grace Chiu, CSIRO",
+                    "Valeria Bejarano Salcedo, Sergio Alejandro Calderon Villanueva, Andrey Duvan Rincon Torres",
                     "Vasilis Dakos, S.R. Carpenter, T. Cline, L. Lahti",
@@
                     "Mustafa Gokce Baydogan",
+                    "Eli Holmes, Eric Ward, Mark Scheuerell, Kellie Wills, NOAA",
+                    "Pedro Araujo, Gustavo Lacerda, Peter C.B. Phillips, Shu-Ping Shi",
                     "Roger D. Peng, Aidan McDermott",
@@
          tidy_aut = str_replace_all(tidy_aut, ",,", ","))
+
+#write_csv(ctvdf_author, "author_pkg.csv")
 ```

### 2022-03-10 paper/index.Rmd
Added explanation of packages interconnection.
add download trend plot

@@
 abstract: >
-    Re-running an experiment is generally costly and in some cases impossible due to limited resources, so the design of experiment plays a critical role in increasing the quality of experimental data.
-In this article I describe the current state of the R-packages for the design of experiments through textual analysis and download trends.
-I discuss also the software design of widely utilised R packages in the field of experimental design and conclude with discussion of some future prospects for the field.
-preamble: |
-  % Any extra LaTeX you need in the preamble
-
-# per R journal requirement, the bib filename should be the same as the output 
-# tex file. Don't forget to rename the bib file and change this example value.
+    Re-running an experiment is generally costly and in some cases impossible due to limited resources, so the design of an experiment plays a critical role in increasing the quality of experimental data.
+In this article we describe the current state of the R-packages for the design of experiments through textual analysis and assessment of download trends.
+We discuss also the software design of widely utilised R packages in the field of experimental design and conclude with discussion of some future prospects for the field.
 bibliography: paper.bib
@@
                       fig.pos = 'htbp', 
+                      fig.path = "figures/",
                       fig.width = 6,
@@
         strip.background = element_rect(color = "black",
-                                        fill = "black"),
-        strip.text = element_text(color = "white"),
+                                        fill = "white"),
+        strip.text = element_text(color = "black", 
+                                  margin = margin(5, 5, 5, 5)),
         plot.title.position = "plot",
         plot.title = element_text(color = "black", hjust = 0)))
+
+ctv <- function(view) {
+  ifelse(knitr::is_html_output(), glue("[{view}](https://cran.r-project.org/web/views/{view}.html)"), glue("\\ctv{[view]}", .open = "[", .close = "]"))
+}
+CRANpkg <- function(pkg) {
+  ifelse(knitr::is_html_output(), glue("[{pkg}](https://cran.r-project.org/web/packages/{pkg}/index.html)"), glue("\\CRANpkg{[pkg]}", .open = "[", .close = "]"))
+}
+ref <- function(x) {
+  ifelse(knitr::is_html_output(), glue("\\@ref({x})"), glue("\\ref{[x]}", .open = "[", .close = "]"))
+}
 ```
@@
 ```{r cran-data, cache = tocache}
-end <- Sys.Date() - 2 # usually 1-2 days data are not available yet
-start <- end - years(2) + days(2)
+end <- Sys.Date() - 3 # usually 1-2 days data are not available yet
+start <- end - years(2)
 dldat <- cran_downloads(doe_pkgs, from = start, to = end)
@@
 ```{r cran-data2, cache = tocache}
-start2 <- end - years(5) + days(2)
+start2 <- end - years(5) 
 dldat2 <- cran_downloads(doe_pkgs, from = start2, to = end)
+```
+
+```{r data-summary}
+dldat_sum <- dldat %>% 
+  group_by(package) %>% 
+  summarise(total = sum(count)) %>% 
+  arrange(desc(total)) %>% 
+  mutate(rank = 1:n())
+
+dldat_sum5 <- dldat2 %>% 
+  group_by(package) %>% 
+  summarise(total = sum(count)) %>% 
+  arrange(desc(total)) %>% 
+  mutate(rank = 1:n())
+
+top_2years <- pull(dldat_sum, package)
+top_5years <- pull(dldat_sum5, package)
+```
+
+```{r pkg-updates, cache = tocache}
+ntop <- 8
+pkg_url <- "https://cran.r-project.org/web/packages/{pkg}/index.html"
+pkg_archive <- "https://cran.r-project.org/src/contrib/Archive/{pkg}/"
+pkg_updates <- map(top_2years[1:ntop], function(pkg) {
+    last_update <- read_html(glue(pkg_url)) %>% 
+      html_table() %>% 
+      .[[1]] %>% 
+      filter(X1=="Published:") %>% 
+      pull(X2) %>% 
+      ymd()
+
+    archive_dates <- tryCatch({ 
+        read_html(glue(pkg_archive)) %>% 
+          html_table() %>%
+          .[[1]] %>% 
+          pull(`Last modified`) %>% 
+          ymd_hm() %>% 
+          na.omit() %>% 
+          as.Date()
+      }, error = function(e) {
+        NULL
+      })
+    c(archive_dates, last_update)
+  })
+names(pkg_updates) <- top_2years[1:ntop]
+
+updates <- unlist(pkg_updates) %>% 
+  enframe("package", "update") %>% 
+  # unlist converts date to integers
+  mutate(update = as.Date(update, origin = "1970-01-01"),
+         # need to get rid of the numbers appended to pkg names
+         package = str_extract(package, paste0(top_2years[1:ntop], collapse="|")),
+         package = factor(package, levels = top_2years[1:ntop])) %>% 
+  filter(between(update, start2, end))
+
 ```
@@
 Methods for data collection can be dichotomised by the type of data collected -- namely, experimental or observational -- or alternatively, categorised as experimental design (including quasi-experimental design) or survey design.
-This dichotomisation, to a great extent, is seen in CRAN task views where R-packages in experimental design are in \ctv{ExperimentalDesign} and R-packages in survey designs are in \ctv{OfficialStatistics}.
-Survey designs often, although not always, aim to collect observational data whilst experimental designs exclusively center on experimental data.
+This dichotomisation, to a great extent, is seen in [CRAN task views](https://cran.r-project.org/web/views/) where R-packages in experimental design are in `r ctv("ExperimentalDesign")` and R-packages in survey designs are in `r ctv("OfficialStatistics")`.
+Survey designs often, although not always, aim to collect observational data whilst experimental designs exclusively center on experimental data where the subject and experimental factors tend to be in the control of the experimenter.
 This paper is concerned with the latter.
 
-In the CRAN task view of \ctv{ExperimentalDesign}, there are `r npkgs` R packages for experimental design and analysis of data from experiments, henceforth referred to as "DoE packages" in this paper.
+In the CRAN task view of `r ctv("ExperimentalDesign")`, there are `r npkgs` R packages for experimental design and analysis of data from experiments, henceforth referred to as "**DoE packages**" in this paper.
 The sheer quantity and variation in the output experimental design in the R-packages are arguably unmatched with any other programming languages, e.g.
@@
 
-The paper is organisd as follows. Section @\ref{eda}
+The paper is organised as follows.
+Section \@ref(eda) explores some analysis of the most downloaded DoE packages, followed by discussion of the discussion of the software design of these packages in Section \@ref(design), and concluding with a discussion in Section \@ref(discussion) of future prospects.
 
@@
 
+The Comprehensive R Archive Network (CRAN) is a network of servers located across the world that store R and R-packages.
+
+```{r download-trend, fig.width = 10,  dependson = "cran-data", fig.height = 8}
+dldat2 %>% 
+  filter(package %in% top_2years[1:ntop]) %>% 
+  mutate(package = factor(package, levels = top_2years[1:ntop])) %>%
+ggplot(aes(date, count + 1)) +
+  geom_line(data = rename(dldat2, pkg = package),
+            color = "grey80",
+            aes(group = pkg)) +
+  geom_line(data = ~rename(., pkg = package),
+            color = "grey60",
+            aes(group = pkg)) +
+  geom_vline(data = updates, 
+             aes(xintercept = update),
+             color = "red", linetype = "dashed") +
+  geom_line() +
+  scale_y_log10() +
+  scale_x_date(expand = c(0, 0)) + 
+  facet_grid(package ~ .) +
+  theme(legend.position = "bottom")
+```
+
 ```{r download-dist, dependson = "cran-data"}
-dldat_sum <- dldat %>% 
-  group_by(package) %>% 
-  summarise(total = sum(count)) %>% 
-  arrange(desc(total)) %>% 
-  mutate(rank = 1:n())
 
@@
        y = "Download counts in the last 2 year ('000s)", title = "(A)") + 
-  geom_text_repel(data = slice_max(dldat_sum, total, n = 13), 
+  geom_text_repel(data = ~slice_max(., total, n = 13), 
                   nudge_x = rep(c(0.2, -0.2, 0.2, -0.2, 0.15), c(3, 2, 2, 3, 3)),
@@
   theme(axis.ticks.length.x = unit(0, "mm"))
-
-dldat_sum5 <- dldat2 %>% 
-  group_by(package) %>% 
-  summarise(total = sum(count)) %>% 
-  arrange(desc(total)) %>% 
-  mutate(rank = 1:n())
 
@@
   labs(x = "", y = "Ranking", title = "(B)") +
-  geom_text(data = filter(top15, name=="Last 5 years"),
+  geom_text(data = ~filter(., name=="Last 5 years"),
             aes(label = package), 
@@
 
-Figure \ref{fig:dlplots} suggests that the \CRANpkg{AlgDesign} followed by \CRANpkg{agricolae} are far more downloaded than any other DoE packages.
-It is worth noting that \CRANpkg{agricolae} imports \CRANpkg{AlgDesign} thus any download of \CRANpkg{agricolae} likely results also in a download of \CRANpkg{AlgDesign}.
+Figure `r ref("fig:dlplots")` suggests that the `r CRANpkg("AlgDesign")` followed by `r CRANpkg("agricolae")` are far more downloaded than any other DoE packages.
+It is worth noting that `r CRANpkg("agricolae")` imports `r CRANpkg("AlgDesign")` thus any download of `r CRANpkg("agricolae")` likely results also in a download of `r CRANpkg("AlgDesign")`.
 
-# Software design
+# Software design {#design}
 
-# Discussion
+# Discussion {#discussion}
+
+The analysis showed that ...
+.
+There are several limitations to the exploratory analysis.
+These include that the download data are limited to the RStudio CRAN download logs, one of XXX mirrors (albeit most popular), and the study is based on CRAN packages listed in the DoE CRAN task view -- there may well be some R-packages that are not on CRAN or part of the task view.
+Further exploration could expand the study for packages that may be outside this list.
 

### 2022-03-10 paper/rjournal/paper-rjournal.tex
add download trend plot

@@
 
-In the CRAN task view of \ctv{ExperimentalDesign}, there are 110 R
+In the CRAN task view of \ctv{ExperimentalDesign}, there are 107 R
 packages for experimental design and analysis of data from experiments,
@@
 
-\caption[The above graph shows the total download counts from 2019-10-04 to 2021-10-02 of DoE packages]{The above graph shows the total download counts from 2019-10-04 to 2021-10-02 of DoE packages.}\label{fig:dlplots}
+\caption[The above graph shows the total download counts from 2020-03-10 to 2022-03-08 of DoE packages]{The above graph shows the total download counts from 2020-03-10 to 2022-03-08 of DoE packages.}\label{fig:dlplots}
 \end{figure}
@@
 \midrule
-optimal design & 9\\
 experimental design & 8\\
+optimal design & 7\\
 clinical trial & 5\\
@@
 experimental design & 10\\
-optimal design & 10\\
-package provide & 7\\
-graphical user & 6\\
+optimal design & 8\\
 response surface & 6\\
-user interface & 6\\
 block design & 5\\
 contour plot & 5\\
-design based & 5\\
 effect model & 5\\
 factorial design & 5\\
+graphical user & 5\\
 microarray experiment & 5\\
 mixed effect & 5\\
-provide function & 5\\
+package provide & 5\\
 sample size & 5\\
+user interface & 5\\
 \bottomrule
@@
 %
-\url{http://emitanaka.org/}%
-\\\textit{ORCiD: \href{https://orcid.org/0000-0002-1455-259X}{0000-0002-1455-259X}}%
-\\\href{mailto:emi.tanaka@monash.edu}{\nolinkurl{emi.tanaka@monash.edu}}
+\url{http://emitanaka.org/}\\%
+\textit{ORCiD: \href{https://orcid.org/0000-0002-1455-259X}{0000-0002-1455-259X}}\\%
+\href{mailto:emi.tanaka@monash.edu}{\nolinkurl{emi.tanaka@monash.edu}}%
 }

### 2022-03-14 paper/paper.Rmd
implement targets in the workflow

--- 2021-08-01 paper/index.Rmd
@@
 ```{r setup, cache = FALSE, include = FALSE}
+library(targets)
+library(glue)
 library(tidyverse)
-library(rvest)
-library(lubridate)
-library(cranlogs)
-library(glue)
-library(scales)
-library(colorspace)
-library(tidytext)
-library(pluralize)
-library(kableExtra)
-library(igraph)
-library(ggraph)
-library(ggrepel)
-library(patchwork)
-library(ggwordcloud)
-tocache <- TRUE
 knitr::opts_chunk$set(echo = FALSE, 
@@
 
-theme_set(
-  theme(panel.background = element_rect(fill = NA),
-        panel.grid = element_line(color = "lightgray"),
-        axis.text = element_text(color = "black"),
-        axis.line = element_line(color = "black", size = 0.7),
-        axis.ticks.length = unit(1.4, "mm"),
-        axis.ticks = element_line(color = "black", size = 0.7),
-        axis.title = element_text(color = "black", face = "bold"),
-        strip.background = element_rect(color = "black",
-                                        fill = "black"),
-        strip.text = element_text(color = "white"),
-        plot.title.position = "plot",
-        plot.title = element_text(color = "black", hjust = 0)))
-```
-
-```{r ctv}
-doe_pkgs <- ctv:::.get_pkgs_from_ctv_or_repos("ExperimentalDesign", 
-                                              repos = "http://cran.rstudio.com/")[[1]]
-doe_pkgs <- c(doe_pkgs, "DeclareDesign")
-npkgs <- length(doe_pkgs)
-```
-
-```{r cran-data, cache = tocache}
-end <- Sys.Date() - 2 # usually 1-2 days data are not available yet
-start <- end - years(2) + days(2)
-dldat <- cran_downloads(doe_pkgs, from = start, to = end)
-```
-
-```{r cran-data2, cache = tocache}
-start2 <- end - years(5) + days(2)
-dldat2 <- cran_downloads(doe_pkgs, from = start2, to = end)
+ctv <- function(view) {
+  ifelse(knitr::is_html_output(), glue("[{view}](https://cran.r-project.org/web/views/{view}.html)"), glue("\\ctv{[view]}", .open = "[", .close = "]"))
+}
+CRANpkg <- function(pkg) {
+  ifelse(knitr::is_html_output(), glue("[{pkg}](https://cran.r-project.org/web/packages/{pkg}/index.html)"), glue("\\CRANpkg{[pkg]}", .open = "[", .close = "]"))
+}
+ref <- function(x) {
+  ifelse(knitr::is_html_output(), glue("\\@ref({x})"), glue("\\ref{[x]}", .open = "[", .close = "]"))
+}
 ```
@@
 Methods for data collection can be dichotomised by the type of data collected -- namely, experimental or observational -- or alternatively, categorised as experimental design (including quasi-experimental design) or survey design.
-This dichotomisation, to a great extent, is seen in CRAN task views where R-packages in experimental design are in \ctv{ExperimentalDesign} and R-packages in survey designs are in \ctv{OfficialStatistics}.
-Survey designs often, although not always, aim to collect observational data whilst experimental designs exclusively center on experimental data.
+This dichotomisation, to a great extent, is seen in [CRAN task views](https://cran.r-project.org/web/views/) (a volunteered maintained list of R-packages by topic) where R-packages in experimental design are in `r ctv("ExperimentalDesign")` and R-packages in survey designs are in `r ctv("OfficialStatistics")`.
+Survey designs often, although not always, aim to collect observational data whilst experimental designs exclusively center on experimental data where the subject and experimental factors tend to be in the control of the experimenter.
 This paper is concerned with the latter.
-
-In the CRAN task view of \ctv{ExperimentalDesign}, there are `r npkgs` R packages for experimental design and analysis of data from experiments, henceforth referred to as "DoE packages" in this paper.
-The sheer quantity and variation in the output experimental design in the R-packages are arguably unmatched with any other programming languages, e.g.
-in Python [@python], only a handful of libraries that generate design of experiment exist (namely `pyDOE`, `pyDOE2`, `dexpy`, `experimenter` and `GPdoemd`) with limited outputs.
-Thus, the study of DoE packages is also revealing into the current status of the field of experimental design.
-
-The paper is organisd as follows. Section @\ref{eda}
-
-# Explorative analysis {#eda}
-
-```{r download-dist, dependson = "cran-data"}
-dldat_sum <- dldat %>% 
-  group_by(package) %>% 
-  summarise(total = sum(count)) %>% 
-  arrange(desc(total)) %>% 
-  mutate(rank = 1:n())
-
-dl1plot <- ggplot(dldat_sum, aes(x = "", y = total/1000)) + 
-  geom_violin() +
-  geom_boxplot(width = 0.1) + 
-  scale_y_log10(label = comma) + 
-  labs(x = "",
-       y = "Download counts in the last 2 year ('000s)", title = "(A)") + 
-  geom_text_repel(data = slice_max(dldat_sum, total, n = 13), 
-                  nudge_x = rep(c(0.2, -0.2, 0.2, -0.2, 0.15), c(3, 2, 2, 3, 3)),
-                  size = 4,
-                  aes(label = package)) + 
-  theme(axis.ticks.length.x = unit(0, "mm"))
-
-dldat_sum5 <- dldat2 %>% 
-  group_by(package) %>% 
-  summarise(total = sum(count)) %>% 
-  arrange(desc(total)) %>% 
-  mutate(rank = 1:n())
-
-top15 <- dldat_sum %>% 
-  left_join(dldat_sum5, by = "package") %>% 
-  slice_max(total.x + total.y, n = 15) %>% 
-  select(package, rank.x, rank.y) %>% 
-  pivot_longer(-package) %>% 
-  mutate(name = ifelse(name=="rank.x", "Last 2 years", "Last 5 years"))
-
-dl2plot <- ggplot(top15, aes(name, value, group = package)) +
-  geom_line() + 
-  geom_point() +
-  scale_y_reverse(breaks = 1:15) +
-  scale_x_discrete(expand = expansion(c(0.1, 0.5))) +
-  labs(x = "", y = "Ranking", title = "(B)") +
-  geom_text(data = filter(top15, name=="Last 5 years"),
-            aes(label = package), 
-            hjust = 0, 
-            nudge_x = 0.1,
-            size = 4) + 
-  theme(panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank())
-```
-
-```{r dlplots, fig.cap=glue("The above graph shows the total download counts from {start} to {end} of DoE packages."), fig.width = 6}
-dl1plot + dl2plot
-```
-
-```{r text, cache=tocache}
-url <- "http://cran.rstudio.com/web/packages/packages.rds"
-db <- readRDS(url(url)) %>% 
-  as.data.frame()
-
-doe_db <- db %>% 
-  filter(Package %in% doe_pkgs) %>% 
-  mutate(Description = str_replace_all(Description, "\n", " "),
-         Description = str_squish(Description),
-         Title = str_replace_all(Title, "\n", " "))
-```
-
-```{r fns}
-singularize2 <- function(x) {
-    res <- singularize(x)
-    res <- ifelse(res=="bia", "bias", res)
-    ifelse(res=="baye", "bayes", res)
-}
-stop_words_rex <- c(paste0("^", stop_words$word, "$"), "^doi", "^[0-9.]+$")
-get_ngram <- function(ngram, desc = c("Title", "Description")) {
-  desc <- match.arg(desc)
-  words <- paste0("word", 1:ngram)
-  doe_db %>% 
-    unnest_tokens(word, desc, token = "ngrams", n = ngram) %>% 
-    separate(word, words, sep = " ") %>% 
-    mutate(across(num_range("word", 1:ngram), singularize2)) %>% 
-    distinct(!!!map(c("Package", words), as.name)) %>% 
-    # allow stop words in between
-    filter(if_all(num_range("word", c(1, ngram)), 
-                  ~!str_detect(., paste0(stop_words_rex, collapse = "|")))) %>% 
-    count(!!!map(words, as.name), sort = TRUE) %>% 
-    mutate(word = do.call("paste", map(words, ~eval(parse(text = .x))))) %>% 
-    select(word, n) 
-}
-
-combine_dl <- function(ngram_df, desc) {
-  dldat_sum %>% 
-    left_join(doe_db, by = c("package" = "Package")) %>% 
-    select(package, total, {{desc}}) %>% 
-    mutate(bigram_title = map({{desc}}, ~filter(ngram_df, 
-                                              n > 1,
-                                              str_detect(tolower(.x), word)))) %>% 
-    unnest_longer(bigram_title) %>% 
-    unpack(bigram_title) %>% 
-    group_by(word) %>% 
-    summarise(n = unique(n),
-              pkgs = list(package),
-              total = sum(total)) %>% 
-    filter(!is.na(word)) %>% 
-    arrange(desc(total))
-}
-```
-
-```{r ngrams, cache=tocache, dependson="fns"}
-bigram_title <- get_ngram(2, "Title")
-bigram_desc <- get_ngram(2, "Description")
-trigram_desc <- get_ngram(3, "Description")
-```
-
-* Not much difference between title and description.
-* Just go with description and combine bi & tri.
-
-```{r wordcloud, fig.height = 3.3}
-dldat_comb_bigram_title <- combine_dl(bigram_title, Title)
-dldat_comb_bigram_desc <- combine_dl(bigram_desc, Title)
-dldat_comb_trigram_desc <- combine_dl(trigram_desc, Title)
-
-set.seed(1)
-ggplot(rbind(dldat_comb_bigram_desc,
-             dldat_comb_trigram_desc),
-       aes(label = word, size = total, color = n)) +
-  geom_text_wordcloud(shape = "square",
-                      show.legend = TRUE) +
-  scale_size_area(max_size = 8) +
-  guides(size = FALSE) +
-  scale_color_continuous(breaks = 1:11) +
-  theme(legend.position = "bottom")
-```
-
-## What are the common types of experimental designs?
-
-## How do packages interplay with each other?
-
-## Which packages are widely utilised?
-
-Figure \ref{fig:dlplots} suggests that the \CRANpkg{AlgDesign} followed by \CRANpkg{agricolae} are far more downloaded than any other DoE packages.
-It is worth noting that \CRANpkg{agricolae} imports \CRANpkg{AlgDesign} thus any download of \CRANpkg{agricolae} likely results also in a download of \CRANpkg{AlgDesign}.
-
-# Software design
-
-# Discussion
-
-An experiment involves running a number of steps:
-
-1. Defining the hypothesis or question
-2. Formulating methods to test hypothesis or answer question 
-3. Planning the experimental protocol:
-  * Determining experimental resources, 
-  * specifying the data collection procedure,
-  * constructing the experimental design layout, and
-  * proposing a model for analysis.
-4. Collect data 
-5. Analyse data, interpret result and conclusion
-
-\newpage
-
-```{r cran-titles}
-url <- "http://cran.rstudio.com/web/packages/packages.rds"
-db <- readRDS(url(url)) %>% 
-  as.data.frame()
-```
-
-```{r bigram}
-stop_words_ext <- c(stop_words$word, "doi")
-
-doe_db <- db %>% 
-  filter(Package %in% doe_pkgs) %>% 
-  mutate(Description = str_replace_all(Description, "\n", " "),
-         Description = str_squish(Description),
-         Title = str_replace_all(Title, "\n", " "))
-
-bigram_tab <- function(data, col) {
-  data %>% 
-    unnest_tokens(word, {{col}}, token = "ngrams", n = 2) %>% 
-    separate(word, c("word1", "word2"), sep = " ") %>% 
-    mutate(word1 = singularize(word1),
-           word2 = singularize(word2)) %>% 
-    # don't count the same bigram within the same package
-    distinct(Package, word1, word2) %>% 
-    filter(!word1 %in% stop_words_ext,
-           !word2 %in% stop_words_ext,
-           !str_detect(word1, "^[0-9.]+$"),
-           !str_detect(word2, "^[0-9.]+$")) %>% 
-    count(word1, word2, sort = TRUE)  
-}
-```
-
-```{r bigram-title, results="asis"}
-btitle <- bigram_tab(doe_db, Title) %>% 
-  filter(n > 3) %>% 
-  mutate(word = paste(word1, word2)) %>% 
-  select(word, n) %>% 
-  kbl(booktabs = TRUE, linesep = "")
-
-bdesc <- bigram_tab(doe_db, Description) %>% 
-  filter(n > 4) %>% 
-  mutate(word = paste(word1, word2)) %>% 
-  select(word, n) %>% 
-  kbl(booktabs = TRUE, linesep = "")
-
-cat(c("\\begin{table}[h] \\centering ", 
-      btitle,
-    "\\hspace{1cm} \\centering ",
-      bdesc,
-    "\\caption{My tables} \\end{table}")) 
-```
-
-```{r}
-doe_imports <- doe_db %>% 
-  mutate(Depends = str_replace_all(Depends, "\n", " "),
-         Depends = str_replace_all(Depends, fixed("("), " ("),
-         Imports = str_replace_all(Imports, "\n", " "),
-         Imports = str_replace_all(Imports, fixed("("), " ("),
-         imports = str_c(Depends, Imports, sep = ","),
-         imports = str_split(imports, ","),
-         imports = map(imports, ~{
-           str_squish(.x) %>% 
-             word() %>% 
-             .[.!=""]}
-         ),
-         imports_doe = map(imports, ~.x[.x %in% doe_pkgs])) %>% 
-  select(Package, imports_doe) %>% 
-  unnest_longer(imports_doe) %>% 
-  filter(!is.na(imports_doe)) %>% 
-  rename(from = imports_doe, to = Package) %>% 
-  select(from, to)
-```
-
-```{r doe-network, eval = FALSE}
-graph_from_data_frame(doe_imports) %>% 
-  ggraph(layout = 'fr') +
-  geom_edge_link(aes(start_cap = label_rect(node1.name),
-                     end_cap = label_rect(node2.name)), 
-                 arrow = arrow(length = unit(2, 'mm')),
-                 color = "#79003e") + 
-  geom_node_text(aes(label = name),
-                 color = "#79003e") +
-  theme(panel.background = element_rect(fill = "#f6e5ee",
-                                        color = "#79003e"),
-        plot.margin = margin(20, 20, 20, 20))
-```

### 2022-03-14 paper/R/functions.R

@@ -1,0 +1,114 @@
+
+plot_setup <- function() {
+  theme_set(
+    theme(panel.background = element_rect(fill = NA),
+          panel.grid = element_line(color = "lightgray"),
+          axis.text = element_text(color = "black"),
+          axis.line = element_line(color = "black", size = 0.7),
+          axis.ticks.length = unit(1.4, "mm"),
+          axis.ticks = element_line(color = "black", size = 0.7),
+          axis.title = element_text(color = "black", face = "bold"),
+          strip.background = element_rect(color = "black",
+                                          fill = "white"),
+          strip.text = element_text(color = "black", 
+                                    margin = margin(5, 5, 5, 5)),
+          plot.title.position = "plot",
+          plot.title = element_text(color = "black", hjust = 0)))
+}
+
+
+
+
+ctv_list <- function(ctv) {
+  ctv:::.get_pkgs_from_ctv_or_repos(ctv, 
+                                    repos = "http://cran.rstudio.com/")[[1]]  
+}
+
+cran_downloads_duration <- function(pkgs, duration) {
+  end <- Sys.Date() - 3 
+  start <- end - years(duration)
+  cran_downloads(pkgs, from = start, to = end)  
+}
+
+cran_downloads_rank <- function(data) {
+  data %>% 
+    group_by(package) %>% 
+    summarise(total = sum(count)) %>% 
+    arrange(desc(total)) %>% 
+    mutate(rank = 1:n())
+}
+
+
+
+
+pkg_updates <- function(pkgs, duration) {
+  pkg_url <- "https://cran.r-project.org/web/packages/{pkg}/index.html"
+  pkg_archive <- "https://cran.r-project.org/src/contrib/Archive/{pkg}/"
+  res <- map(pkgs, function(pkg) {
+    last_update <- read_html(glue(pkg_url)) %>% 
+      html_table() %>% 
+      .[[1]] %>% 
+      filter(X1=="Published:") %>% 
+      pull(X2) %>% 
+      ymd()
+    
+    archive_dates <- tryCatch({ 
+      read_html(glue(pkg_archive)) %>% 
+        html_table() %>%
+        .[[1]] %>% 
+        pull(`Last modified`) %>% 
+        ymd_hm() %>% 
+        na.omit() %>% 
+        as.Date()
+    }, error = function(e) {
+      NULL
+    })
+  c(archive_dates, last_update)
+  })
+  names(res) <- pkgs
+  
+  unlist(res) %>% 
[…]

### 2022-03-14 paper/_targets.R

@@ -1,0 +1,27 @@
+library(targets)
+library(ggplot2)
+library(tarchetypes)
+source("R/functions.R")
+tar_option_set(packages = c("tidyverse", "rvest", "lubridate",
+                            "cranlogs", "glue", "scales", "colorspace",
+                            "tidytext", "pluralize", "kableExtra", 
+                            "igraph", "ggraph", "ggrepel", "patchwork",
+                            "ggwordcloud"))
+
+plot_setup()
+
+list(
+  tar_target(doe_pkgs,  ctv_list("ExperimentalDesign")),
+  tar_target(npkgs, length(doe_pkgs)),
+  tar_target(doe_db, pkg_db(doe_pkgs)),
+  tar_target(cran_download_data2, cran_downloads_duration(doe_pkgs, 2)),
+  tar_target(cran_download_data5, cran_downloads_duration(doe_pkgs, 5)),
+  tar_target(cran_rank2, cran_downloads_rank(cran_download_data2)),
+  tar_target(cran_rank5, cran_downloads_rank(cran_download_data5)),
+  tar_target(ntop, 8),
+  tar_target(top2yrs, cran_rank2$package[1:ntop]),
+  tar_target(top5yrs, cran_rank5$package[1:ntop]),
+  tar_target(updates2, pkg_updates(top2yrs, duration = 5)),
+  tar_target(plot_download_trend, download_trend(cran_rank5, updates2)),
+  tar_render(report, "paper.Rmd")
+)

### 2022-04-12 paper/paper.Rmd
update targets with stl model

@@
   - name: Emi Tanaka
+    affiliation: Monash University
+  - name: Dewi Amaliah 
     affiliation: Monash University
@@
 }
+
+tar_load(npkgs)
 ```
@@
 This paper is concerned with the latter.
+
+In the CRAN task view of `r ctv("ExperimentalDesign")`, there are `r npkgs` R packages for experimental design and analysis of data from experiments, henceforth referred to as "**DoE packages**" in this paper.
+The sheer quantity and variation in the output experimental design in the R-packages are arguably unmatched with any other programming languages, e.g.
+in Python [@python], only a handful of libraries that generate design of experiment exist (namely `pyDOE`, `pyDOE2`, `dexpy`, `experimenter` and `GPdoemd`) with limited outputs.
+Thus, the study of DoE packages is also revealing into the current status of the field of experimental design.
+
+The paper is organised as follows.
+Section \@ref(eda) explores some analysis of the most downloaded DoE packages, followed by discussion of the discussion of the software design of these packages in Section \@ref(design), and concluding with a discussion in Section \@ref(discussion) of future prospects.
+
+# Explorative analysis {#eda}
+
+The Comprehensive R Archive Network (CRAN) is a network of servers located across the world that store R and R-packages.
+
+```{r download-trend, fig.width = 10,  dependson = "cran-data", fig.height = 8}
+tar_read(plot_download_trend)
+```

### 2022-04-12 paper/R/functions.R

@@ -25,9 +25,18 @@ ctv_list <- function(ctv) {
 }
 
 cran_downloads_duration <- function(pkgs, duration) {
-  end <- Sys.Date() - 3 
+  end <- as.Date("2022-03-01")  
   start <- end - years(duration)
   cran_downloads(pkgs, from = start, to = end)  
+}
+
+cran_ed_scrub_duration <- function(duration) {
+  end <- as.Date("2022-03-01")  
+  start <- end - years(duration)
+  cranscrub::ctvExperimentalDesign %>% 
+    rename(count = n_unique) %>% 
+    select(-n_total) %>% 
+    dplyr::filter(between(date, start, end))
 }
 
 cran_downloads_rank <- function(data) {
@@ -36,6 +45,19 @@ cran_downloads_rank <- function(data) {
     summarise(total = sum(count)) %>% 
     arrange(desc(total)) %>% 
     mutate(rank = 1:n())
+}
+
+stl_model <- function(data, pkg, updates, trend, season) {
+  data %>% 
+    as_tsibble(index = date, key = package) %>% 
+    filter(package == pkg) %>% 
+    model(STL(count ~ trend(window = trend) + season(window = season),
+              robust = TRUE)) %>% 
+    components() %>% 
+    autoplot() + 
+    geom_vline(data = filter(updates, package == pkg), 
+               aes(xintercept = update), color = "red")
+  
 }
 
 
@@ -74,17 +96,17 @@ pkg_updates <- function(pkgs, duration) {
            # need to get rid of the numbers appended to pkg names
            package = str_extract(package, paste0(pkgs, collapse="|")),
            package = factor(package, levels = pkgs)) %>% 
-    filter(update >= Sys.Date() - years(duration))
+    filter(update >= as.Date("2022-03-01") - years(duration))
 }
 
 
-download_trend <- function(datsum, updates) {
+download_trend <- function(data, updates) {
   top_pkgs <- unique(updates$package)
-  datsum %>% 
+  data %>% 
     filter(package %in% top_pkgs) %>% 
     mutate(package = factor(package, levels = top_pkgs)) %>%
     ggplot(aes(date, count + 1)) +
-    geom_line(data = rename(datsum, pkg = package),
+    geom_line(data = rename(data, pkg = package),
               color = "grey80",
               aes(group = pkg)) +
     geom_line(data = ~rename(., pkg = package),

### 2022-04-12 paper/_targets.R

@@ -6,7 +6,7 @@ tar_option_set(packages = c("tidyverse", "rvest", "lubridate",
                             "cranlogs", "glue", "scales", "colorspace",
                             "tidytext", "pluralize", "kableExtra", 
                             "igraph", "ggraph", "ggrepel", "patchwork",
-                            "ggwordcloud"))
+                            "ggwordcloud", "fable"))
 
 plot_setup()
 
@@ -16,12 +16,17 @@ list(
   tar_target(doe_db, pkg_db(doe_pkgs)),
   tar_target(cran_download_data2, cran_downloads_duration(doe_pkgs, 2)),
   tar_target(cran_download_data5, cran_downloads_duration(doe_pkgs, 5)),
+  tar_target(cran_scrub_data2, cran_ed_scrub_duration(2)),
+  tar_target(cran_scrub_data5, cran_ed_scrub_duration(5)),
   tar_target(cran_rank2, cran_downloads_rank(cran_download_data2)),
   tar_target(cran_rank5, cran_downloads_rank(cran_download_data5)),
+  tar_target(cran_srank2, cran_downloads_rank(cran_scrub_data2)),
+  tar_target(cran_srank5, cran_downloads_rank(cran_scrub_data5)),
   tar_target(ntop, 8),
   tar_target(top2yrs, cran_rank2$package[1:ntop]),
   tar_target(top5yrs, cran_rank5$package[1:ntop]),
   tar_target(updates2, pkg_updates(top2yrs, duration = 5)),
-  tar_target(plot_download_trend, download_trend(cran_rank5, updates2)),
+  tar_target(plot_download_trend, download_trend(cran_download_data5, updates2)),
+  tar_target(plot_scrub_trend, download_trend(cran_scrub_data5, updates2)),
   tar_render(report, "paper.Rmd")
 )

### 2022-04-13 paper/R/functions.R
downloads by year

@@ -47,11 +47,22 @@ cran_downloads_rank <- function(data) {
     mutate(rank = 1:n())
 }
 
-stl_model <- function(data, pkg, updates, trend, season) {
+cran_downloads_rank_by_year <- function(data) {
   data %>% 
-    as_tsibble(index = date, key = package) %>% 
+    mutate(year = year(date)) %>% 
+    group_by(package, year) %>% 
+    summarise(total = sum(count)) %>% 
+    arrange(desc(total), year) %>% 
+    group_by(year) %>% 
+    mutate(rank = 1:n()) %>% 
+    ungroup()
+}
+
+plot_stl_model <- function(data, pkg, updates, trend, speriod1, speriod2) {
+  data %>% 
     filter(package == pkg) %>% 
-    model(STL(count ~ trend(window = trend) + season(window = season),
+    as_tsibble(index = date) %>% 
+    model(STL(count ~ trend(window = !!trend) + season(period = !!speriod1) + season(period = !!speriod2),
               robust = TRUE)) %>% 
     components() %>% 
     autoplot() + 
@@ -60,7 +71,15 @@ stl_model <- function(data, pkg, updates, trend, season) {
   
 }
 
-
+plot_download_distribution <- function(data) {
+  data %>% 
+    mutate(year = as.factor(year)) %>% 
+    ggplot(aes(x = year, y = total)) + 
+    geom_violin() +
+    geom_boxplot(width = 0.1) + 
+    labs(x = "Year", y = "Downloads") + 
+    scale_y_log10(label = comma) 
+}
 
 
 pkg_updates <- function(pkgs, duration) {

### 2022-04-13 paper/_targets.R

@@ -6,7 +6,7 @@ tar_option_set(packages = c("tidyverse", "rvest", "lubridate",
                             "cranlogs", "glue", "scales", "colorspace",
                             "tidytext", "pluralize", "kableExtra", 
                             "igraph", "ggraph", "ggrepel", "patchwork",
-                            "ggwordcloud", "fable"))
+                            "ggwordcloud", "fable", "feasts", "lubridate"))
 
 plot_setup()
 
@@ -20,13 +20,19 @@ list(
   tar_target(cran_scrub_data5, cran_ed_scrub_duration(5)),
   tar_target(cran_rank2, cran_downloads_rank(cran_download_data2)),
   tar_target(cran_rank5, cran_downloads_rank(cran_download_data5)),
+  tar_target(cran_rank_yearly, cran_downloads_rank_by_year(cran_download_data5)),
   tar_target(cran_srank2, cran_downloads_rank(cran_scrub_data2)),
   tar_target(cran_srank5, cran_downloads_rank(cran_scrub_data5)),
+  tar_target(cran_srank_yearly, cran_downloads_rank_by_year(cran_scrub_data5)),
   tar_target(ntop, 8),
   tar_target(top2yrs, cran_rank2$package[1:ntop]),
   tar_target(top5yrs, cran_rank5$package[1:ntop]),
   tar_target(updates2, pkg_updates(top2yrs, duration = 5)),
+  tar_target(updates_all, pkg_updates(doe_pkgs, duration = Inf)),
+  tar_target(plot_download_dist, plot_download_distribution(cran_srank_yearly)),
   tar_target(plot_download_trend, download_trend(cran_download_data5, updates2)),
   tar_target(plot_scrub_trend, download_trend(cran_scrub_data5, updates2)),
+  tar_target(plot_stl_7wy, plot_stl_model(cran_scrub_data5, "agricolae", updates2, 7, "week", "year")),
+
   tar_render(report, "paper.Rmd")
 )

### 2022-04-14 paper/R/functions.R
adding gini index to plots

@@ -1,7 +1,6 @@
 
 plot_setup <- function() {
-  theme_set(
-    theme(panel.background = element_rect(fill = NA),
+  theme(panel.background = element_rect(fill = NA),
           panel.grid = element_line(color = "lightgray"),
           axis.text = element_text(color = "black"),
           axis.line = element_line(color = "black", size = 0.7),
@@ -13,7 +12,7 @@ plot_setup <- function() {
           strip.text = element_text(color = "black", 
                                     margin = margin(5, 5, 5, 5)),
           plot.title.position = "plot",
-          plot.title = element_text(color = "black", hjust = 0)))
+          plot.title = element_text(color = "black", hjust = 0))
 }
 
 
@@ -47,15 +46,18 @@ cran_downloads_rank <- function(data) {
     mutate(rank = 1:n())
 }
 
-cran_downloads_rank_by_year <- function(data) {
+cran_downloads_rank_by_year <- function(data, first_release) {
   data %>% 
     mutate(year = year(date)) %>% 
     group_by(package, year) %>% 
     summarise(total = sum(count)) %>% 
     arrange(desc(total), year) %>% 
+    left_join(first_release, by = "package") %>% 
     group_by(year) %>% 
     mutate(rank = 1:n()) %>% 
-    ungroup()
+    ungroup() %>% 
+    filter(year > year(first)) %>% 
+    filter(year != year(Sys.Date()))
 }
 
 plot_stl_model <- function(data, pkg, updates, trend, speriod1, speriod2) {
@@ -71,18 +73,62 @@ plot_stl_model <- function(data, pkg, updates, trend, speriod1, speriod2) {
   
 }
 
-plot_download_distribution <- function(data) {
+plot_download_distribution <- function(data, ginidata) {
   data %>% 
     mutate(year = as.factor(year)) %>% 
     ggplot(aes(x = year, y = total)) + 
     geom_violin() +
     geom_boxplot(width = 0.1) + 
     labs(x = "Year", y = "Downloads") + 
-    scale_y_log10(label = comma) 
+    scale_y_log10(label = comma) +
+    geom_label(data = mutate(ginidata, year = as.factor(year)), 
+              aes(y = 900, label = scales::percent(gini, 0.1)))
+}
+
+plot_lorenz_curve <- function(data, year) {
+  
+  dat <- data %>% 
+    filter(year == .env$year) 
+    
[…]

### 2022-04-14 paper/_targets.R

@@ -6,9 +6,9 @@ tar_option_set(packages = c("tidyverse", "rvest", "lubridate",
                             "cranlogs", "glue", "scales", "colorspace",
                             "tidytext", "pluralize", "kableExtra", 
                             "igraph", "ggraph", "ggrepel", "patchwork",
-                            "ggwordcloud", "fable", "feasts", "lubridate"))
+                            "ggwordcloud", "fable", "feasts", "lubridate",
+                            "ineq"))
 
-plot_setup()
 
 list(
   tar_target(doe_pkgs,  ctv_list("ExperimentalDesign")),
@@ -20,19 +20,29 @@ list(
   tar_target(cran_scrub_data5, cran_ed_scrub_duration(5)),
   tar_target(cran_rank2, cran_downloads_rank(cran_download_data2)),
   tar_target(cran_rank5, cran_downloads_rank(cran_download_data5)),
-  tar_target(cran_rank_yearly, cran_downloads_rank_by_year(cran_download_data5)),
   tar_target(cran_srank2, cran_downloads_rank(cran_scrub_data2)),
   tar_target(cran_srank5, cran_downloads_rank(cran_scrub_data5)),
-  tar_target(cran_srank_yearly, cran_downloads_rank_by_year(cran_scrub_data5)),
+  tar_target(updates_all, pkg_updates(doe_pkgs)),
+  tar_target(first_release, updates_all %>% 
+               group_by(package) %>% 
+               summarise(first = min(update)) %>% 
+               ungroup()),
+  tar_target(cran_rank_yearly, cran_downloads_rank_by_year(cran_download_data5, first_release)),
+  tar_target(cran_scrub_rank_yearly, cran_downloads_rank_by_year(cran_scrub_data5, first_release)),
   tar_target(ntop, 8),
   tar_target(top2yrs, cran_rank2$package[1:ntop]),
   tar_target(top5yrs, cran_rank5$package[1:ntop]),
-  tar_target(updates2, pkg_updates(top2yrs, duration = 5)),
-  tar_target(updates_all, pkg_updates(doe_pkgs, duration = Inf)),
-  tar_target(plot_download_dist, plot_download_distribution(cran_srank_yearly)),
-  tar_target(plot_download_trend, download_trend(cran_download_data5, updates2)),
-  tar_target(plot_scrub_trend, download_trend(cran_scrub_data5, updates2)),
[…]

### 2022-04-15 paper/R/functions.R
add bigrams and trigrams to target

@@ -203,3 +203,44 @@ pkg_db <- function(pkgs) {
            Title = str_replace_all(Title, "\n", " "))  
 }
 
+
+
+
+singularize2 <- function(x) {
+  res <- singularize(x)
+  res <- ifelse(res=="bia", "bias", res)
+  ifelse(res=="baye", "bayes", res)
+}
+stop_words_rex <- c(paste0("^", stop_words$word, "$"), "^doi", "^[0-9.]+$")
+get_ngram <- function(ngram, desc = c("Title", "Description")) {
+  desc <- match.arg(desc)
+  words <- paste0("word", 1:ngram)
+  doe_db %>% 
+    unnest_tokens(word, desc, token = "ngrams", n = ngram) %>% 
+    separate(word, words, sep = " ") %>% 
+    mutate(across(num_range("word", 1:ngram), singularize2)) %>% 
+    distinct(!!!map(c("Package", words), as.name)) %>% 
+    # allow stop words in between
+    filter(if_all(num_range("word", c(1, ngram)), 
+                  ~!str_detect(., paste0(stop_words_rex, collapse = "|")))) %>% 
+    count(!!!map(words, as.name), sort = TRUE) %>% 
+    mutate(word = do.call("paste", map(words, ~eval(parse(text = .x))))) %>% 
+    select(word, n) 
+}
+
+combine_dl <- function(ngram_df, desc) {
+  dldat_sum %>% 
+    left_join(doe_db, by = c("package" = "Package")) %>% 
+    select(package, total, {{desc}}) %>% 
+    mutate(bigram_title = map({{desc}}, ~filter(ngram_df, 
+                                                n > 1,
+                                                str_detect(tolower(.x), word)))) %>% 
+    unnest_longer(bigram_title) %>% 
+    unpack(bigram_title) %>% 
+    group_by(word) %>% 
+    summarise(n = unique(n),
+              pkgs = list(package),
+              total = sum(total)) %>% 
+    filter(!is.na(word)) %>% 
+    arrange(desc(total))
+}

### 2022-04-15 paper/_targets.R

@@ -43,6 +43,15 @@ list(
   tar_target(plot_download_trend, download_trend(cran_download_data5, updates2) + plot_setup()),
   tar_target(plot_scrub_trend, download_trend(cran_scrub_data5, updates2) + plot_setup()),
   tar_target(plot_stl_7wy, plot_stl_model(cran_scrub_data5, "agricolae", updates2, 7, "week", "year") + plot_setup()),
+  
+  tar_target(bigram_title, get_ngram(2, "Title")),
+  tar_target(bigram_desc, get_ngram(2, "Description")),
+  tar_target(trigram_desc, get_ngram(3, "Description")),
+  
+  tar_target(dldat_comb_bigram_title,  combine_dl(bigram_title, Title)),
+  tar_target(dldat_comb_bigram_desc,  combine_dl(bigram_desc, Title)),
+  tar_target(dldat_comb_trigram_desc,  combine_dl(trigram_desc, Title)),
+  
 
   tar_render(report, "paper.Rmd")
 )

### 2022-05-02 arxiv/arxiv.Rmd
move paper folder to root

@@
     email: emi.tanaka@monash.edu
-bibliography: references.bib
+bibliography: ../paper.bib
 biblio-style: unsrt
@@
 ```{r, include = FALSE}
-paper <- here::here("paper.Rmd")
+paper <- here::here("paper/index.Rmd")
 meta <- rmarkdown::yaml_front_matter(paper)

### 2022-05-02 index.Rmd
move paper folder to root

--- 2022-03-10 paper/index.Rmd
@@
 
-theme_set(
-  theme(panel.background = element_rect(fill = NA),
-        panel.grid = element_line(color = "lightgray"),
-        axis.text = element_text(color = "black"),
-        axis.line = element_line(color = "black", size = 0.7),
-        axis.ticks.length = unit(1.4, "mm"),
-        axis.ticks = element_line(color = "black", size = 0.7),
-        axis.title = element_text(color = "black", face = "bold"),
-        strip.background = element_rect(color = "black",
-                                        fill = "white"),
-        strip.text = element_text(color = "black", 
-                                  margin = margin(5, 5, 5, 5)),
-        plot.title.position = "plot",
-        plot.title = element_text(color = "black", hjust = 0)))
+npkgs <- tar_read(npkgs)
 
@@
 
-```{r ctv}
-doe_pkgs <- ctv:::.get_pkgs_from_ctv_or_repos("ExperimentalDesign", 
-                                              repos = "http://cran.rstudio.com/")[[1]]
-doe_pkgs <- c(doe_pkgs, "DeclareDesign")
-npkgs <- length(doe_pkgs)
-```
-
-```{r cran-data, cache = tocache}
-end <- Sys.Date() - 3 # usually 1-2 days data are not available yet
-start <- end - years(2)
-dldat <- cran_downloads(doe_pkgs, from = start, to = end)
-```
-
-```{r cran-data2, cache = tocache}
-start2 <- end - years(5) 
-dldat2 <- cran_downloads(doe_pkgs, from = start2, to = end)
-```
-
-```{r data-summary}
-dldat_sum <- dldat %>% 
-  group_by(package) %>% 
-  summarise(total = sum(count)) %>% 
-  arrange(desc(total)) %>% 
-  mutate(rank = 1:n())
-
-dldat_sum5 <- dldat2 %>% 
-  group_by(package) %>% 
-  summarise(total = sum(count)) %>% 
-  arrange(desc(total)) %>% 
-  mutate(rank = 1:n())
-
-top_2years <- pull(dldat_sum, package)
-top_5years <- pull(dldat_sum5, package)
-```
-
-```{r pkg-updates, cache = tocache}
-ntop <- 8
-pkg_url <- "https://cran.r-project.org/web/packages/{pkg}/index.html"
-pkg_archive <- "https://cran.r-project.org/src/contrib/Archive/{pkg}/"
-pkg_updates <- map(top_2years[1:ntop], function(pkg) {
-    last_update <- read_html(glue(pkg_url)) %>% 
-      html_table() %>% 
-      .[[1]] %>% 
-      filter(X1=="Published:") %>% 
-      pull(X2) %>% 
-      ymd()
-
-    archive_dates <- tryCatch({ 
-        read_html(glue(pkg_archive)) %>% 
-          html_table() %>%
-          .[[1]] %>% 
-          pull(`Last modified`) %>% 
-          ymd_hm() %>% 
-          na.omit() %>% 
-          as.Date()
-      }, error = function(e) {
-        NULL
-      })
-    c(archive_dates, last_update)
-  })
-names(pkg_updates) <- top_2years[1:ntop]
-
-updates <- unlist(pkg_updates) %>% 
-  enframe("package", "update") %>% 
-  # unlist converts date to integers
-  mutate(update = as.Date(update, origin = "1970-01-01"),
-         # need to get rid of the numbers appended to pkg names
-         package = str_extract(package, paste0(top_2years[1:ntop], collapse="|")),
-         package = factor(package, levels = top_2years[1:ntop])) %>% 
-  filter(between(update, start2, end))
-
-```
-
 # Introduction
@@
 Methods for data collection can be dichotomised by the type of data collected -- namely, experimental or observational -- or alternatively, categorised as experimental design (including quasi-experimental design) or survey design.
-This dichotomisation, to a great extent, is seen in [CRAN task views](https://cran.r-project.org/web/views/) where R-packages in experimental design are in `r ctv("ExperimentalDesign")` and R-packages in survey designs are in `r ctv("OfficialStatistics")`.
+This dichotomisation, to a great extent, is seen in [CRAN task views](https://cran.r-project.org/web/views/) (a volunteered maintained list of R-packages by topic) where R-packages in experimental design are in `r ctv("ExperimentalDesign")` and R-packages in survey designs are in `r ctv("OfficialStatistics")`.
 Survey designs often, although not always, aim to collect observational data whilst experimental designs exclusively center on experimental data where the subject and experimental factors tend to be in the control of the experimenter.
@@
 ```{r download-trend, fig.width = 10,  dependson = "cran-data", fig.height = 8}
-dldat2 %>% 
-  filter(package %in% top_2years[1:ntop]) %>% 
-  mutate(package = factor(package, levels = top_2years[1:ntop])) %>%
-ggplot(aes(date, count + 1)) +
-  geom_line(data = rename(dldat2, pkg = package),
-            color = "grey80",
-            aes(group = pkg)) +
-  geom_line(data = ~rename(., pkg = package),
-            color = "grey60",
-            aes(group = pkg)) +
-  geom_vline(data = updates, 
-             aes(xintercept = update),
-             color = "red", linetype = "dashed") +
-  geom_line() +
-  scale_y_log10() +
-  scale_x_date(expand = c(0, 0)) + 
-  facet_grid(package ~ .) +
-  theme(legend.position = "bottom")
+tar_read(plot_download_trend)
 ```
 
-```{r download-dist, dependson = "cran-data"}
+```{r download-dist, dependson = "cran-data", eval = FALSE}
 
@@
 
-```{r dlplots, fig.cap=glue("The above graph shows the total download counts from {start} to {end} of DoE packages."), fig.width = 6}
+```{r dlplots, fig.cap=glue("The above graph shows the total download counts from {start} to {end} of DoE packages."), eval = FALSE, fig.width = 6}
 dl1plot + dl2plot
@@
 
-```{r text, cache=tocache}
-url <- "http://cran.rstudio.com/web/packages/packages.rds"
-db <- readRDS(url(url)) %>% 
-  as.data.frame()
-
-doe_db <- db %>% 
-  filter(Package %in% doe_pkgs) %>% 
-  mutate(Description = str_replace_all(Description, "\n", " "),
-         Description = str_squish(Description),
-         Title = str_replace_all(Title, "\n", " "))
-```
-
-```{r fns}
+```{r fns, eval = FALSE}
 singularize2 <- function(x) {
@@
 
-```{r ngrams, cache=tocache, dependson="fns"}
+```{r ngrams, cache=tocache, dependson="fns", eval = FALSE}
 bigram_title <- get_ngram(2, "Title")
@@
 
-\newpage
-
-```{r cran-titles}
-url <- "http://cran.rstudio.com/web/packages/packages.rds"
-db <- readRDS(url(url)) %>% 
-  as.data.frame()
-```
-
-```{r bigram}
+```{r bigram, eval = FALSE}
 stop_words_ext <- c(stop_words$word, "doi")
@@
 
-```{r bigram-title, results="asis"}
+```{r bigram-title, results="asis", eval = FALSE}
 btitle <- bigram_tab(doe_db, Title) %>% 
@@
 
-```{r}
+```{r, eval = FALSE}
 doe_imports <- doe_db %>% 

### 2022-05-02 paper.Rmd
move paper folder to root

--- 2021-03-05 paper.Rmd
@@
 author:
-  # see ?rjournal_article for more information
   - name: Emi Tanaka
     affiliation: Monash University
-    address:
-    - Monash University
-    - Clayton campus, VIC 3800, Australia
-    url: http://emitanaka.org/
-    orcid: 0000-0002-1455-259X
-    email:  emi.tanaka@monash.edu
+  - name: Dewi Amaliah 
+    affiliation: Monash University
 abstract: >
-    Re-running an experiment is generally costly and in some cases impossible due to limited resources, so the design of experiment plays a critical role in increasing the quality of experimental data.
-In this article I describe the current state of the R-packages for the design of experiments through textual analysis and download trends.
-I discuss also the software design of widely utilised R packages in the field of experimental design and conclude with discussion of some future prospects for the field.
-preamble: |
-  % Any extra LaTeX you need in the preamble
-
-# per R journal requirement, the bib filename should be the same as the output 
-# tex file. Don't forget to rename the bib file and change this example value.
+    Re-running an experiment is generally costly and in some cases impossible due to limited resources, so the design of an experiment plays a critical role in increasing the quality of experimental data.
+In this article we describe the current state of the R-packages for the design of experiments through textual analysis and assessment of download trends.
+We discuss also the software design of widely utilised R packages in the field of experimental design and conclude with discussion of some future prospects for the field.
 bibliography: paper.bib
 
-output: rticles::rjournal_article
+output: 
+  bookdown::html_document2:
+    theme: paper
 ---
@@
 ```{r setup, cache = FALSE, include = FALSE}
+library(targets)
+library(glue)
 library(tidyverse)
-library(rvest)
-library(lubridate)
-library(cranlogs)
-library(glue)
-library(scales)
-library(colorspace)
-library(tidytext)
-library(pluralize)
-library(kableExtra)
-library(igraph)
-library(ggraph)
-library(ggrepel)
-library(patchwork)
-library(ggwordcloud)
-tocache <- TRUE
-
 knitr::opts_chunk$set(echo = FALSE, 
@@
                       cache.path = "cache/",
-                      fig.align='center', 
-                      fig.pos='htbp', 
+                      fig.align = 'center', 
+                      fig.pos = 'htbp', 
+                      fig.path = "figures/",
                       fig.width = 6,
@@
 
-theme_set(
-  theme(panel.background = element_rect(fill = NA),
-        panel.grid = element_line(color = "lightgray"),
-        axis.text = element_text(color = "black"),
-        axis.line = element_line(color = "black", size = 0.7),
-        axis.ticks.length = unit(1.4, "mm"),
-        axis.ticks = element_line(color = "black", size = 0.7),
-        axis.title = element_text(color = "black", face = "bold"),
-        strip.background = element_rect(color = "black",
-                                        fill = "black"),
-        strip.text = element_text(color = "white"),
-        plot.title.position = "plot",
-        plot.title = element_text(color = "black", hjust = 0)))
-```
+ctv <- function(view) {
+  ifelse(knitr::is_html_output(), glue("[{view}](https://cran.r-project.org/web/views/{view}.html)"), glue("\\ctv{[view]}", .open = "[", .close = "]"))
+}
+CRANpkg <- function(pkg) {
+  ifelse(knitr::is_html_output(), glue("[{pkg}](https://cran.r-project.org/web/packages/{pkg}/index.html)"), glue("\\CRANpkg{[pkg]}", .open = "[", .close = "]"))
+}
+ref <- function(x) {
+  ifelse(knitr::is_html_output(), glue("\\@ref({x})"), glue("\\ref{[x]}", .open = "[", .close = "]"))
+}
 
-```{r ctv}
-doe_pkgs <- ctv:::.get_pkgs_from_ctv_or_repos("ExperimentalDesign", 
-                                              repos = "http://cran.rstudio.com/")[[1]]
-doe_pkgs <- c(doe_pkgs, "DeclareDesign")
-npkgs <- length(doe_pkgs)
-```
-
-```{r cran-data, cache = tocache}
-end <- Sys.Date() - 2 # usually 1-2 days data are not available yet
-start <- end - years(2) + days(2)
-dldat <- cran_downloads(doe_pkgs, from = start, to = end)
-```
-
-```{r cran-data2, cache = tocache}
-start2 <- end - years(5) + days(2)
-dldat2 <- cran_downloads(doe_pkgs, from = start2, to = end)
+tar_load(npkgs)
 ```
@@
 The critical role of data collection is well captured in the expression "garbage in, garbage out" -- in other words, if the collected data is rubbish then no analysis, however complex it may be, can make something out of it.
-Methods for data collection can be dichotomised by the type of data collected -- namely, experimental or observational -- or alternatively, categorised into experimental design or survey design.
-Survey designs often, although not always, aim to collect observational data whilst experimental designs exclusively center on experimental data.
-This dichotomisation, to a great extent, is seen in CRAN task views where R-packages in experimental design are in \ctv{ExperimentalDesign} and R-packages in survey designs are in \ctv{OfficialStatistics}.
+Methods for data collection can be dichotomised by the type of data collected -- namely, experimental or observational -- or alternatively, categorised as experimental design (including quasi-experimental design) or survey design.
+This dichotomisation, to a great extent, is seen in [CRAN task views](https://cran.r-project.org/web/views/) (a volunteered maintained list of R-packages by topic) where R-packages in experimental design are in `r ctv("ExperimentalDesign")` and R-packages in survey designs are in `r ctv("OfficialStatistics")`.
+Survey designs often, although not always, aim to collect observational data whilst experimental designs exclusively center on experimental data where the subject and experimental factors tend to be in the control of the experimenter.
+This paper is concerned with the latter.
 
-In the CRAN task view of \ctv{ExperimentalDesign}, there are `r npkgs` R packages for experimental design and analysis of data from experiments, henceforth referred to as "DoE packages" in this paper.
+In the CRAN task view of `r ctv("ExperimentalDesign")`, there are `r npkgs` R packages for experimental design and analysis of data from experiments, henceforth referred to as "**DoE packages**" in this paper.
 The sheer quantity and variation in the output experimental design in the R-packages are arguably unmatched with any other programming languages, e.g.
 in Python [@python], only a handful of libraries that generate design of experiment exist (namely `pyDOE`, `pyDOE2`, `dexpy`, `experimenter` and `GPdoemd`) with limited outputs.
+Thus, the study of DoE packages is also revealing into the current status of the field of experimental design.
 
-# Explorative Analysis
+The paper is organised as follows.
+Section \@ref(eda) explores some analysis of the most downloaded DoE packages, followed by discussion of the discussion of the software design of these packages in Section \@ref(design), and concluding with a discussion in Section \@ref(discussion) of future prospects.
 
-```{r download-dist, dependson = "cran-data"}
-dldat_sum <- dldat %>% 
-  group_by(package) %>% 
-  summarise(total = sum(count)) %>% 
-  arrange(desc(total)) %>% 
-  mutate(rank = 1:n())
+# Explorative analysis {#eda}
 
-dl1plot <- ggplot(dldat_sum, aes(x = "", y = total/1000)) + 
-  geom_violin() +
-  geom_boxplot(width = 0.1) + 
-  scale_y_log10(label = comma) + 
-  labs(x = "",
-       y = "Download counts in the last 2 year ('000s)", title = "(A)") + 
-  geom_text_repel(data = slice_max(dldat_sum, total, n = 13), 
-                  nudge_x = rep(c(0.2, -0.2, 0.2, -0.2, 0.15), c(3, 2, 2, 3, 3)),
-                  size = 10,
-                  aes(label = package)) + 
-  theme(axis.ticks.length.x = unit(0, "mm"))
+The Comprehensive R Archive Network (CRAN) is a network of servers located across the world that store R and R-packages.
 
-dldat_sum5 <- dldat2 %>% 
-  group_by(package) %>% 
-  summarise(total = sum(count)) %>% 
-  arrange(desc(total)) %>% 
-  mutate(rank = 1:n())
-
-top15 <- dldat_sum %>% 
-  left_join(dldat_sum5, by = "package") %>% 
-  slice_max(total.x + total.y, n = 15) %>% 
-  select(package, rank.x, rank.y) %>% 
-  pivot_longer(-package) %>% 
-  mutate(name = ifelse(name=="rank.x", "Last 2 years", "Last 5 years"))
-
-dl2plot <- ggplot(top15, aes(name, value, group = package)) +
-  geom_line() + 
-  geom_point() +
-  scale_y_reverse(breaks = 1:15) +
-  scale_x_discrete(expand = expansion(c(0.1, 0.5))) +
-  labs(x = "", y = "Ranking", title = "(B)") +
-  geom_text(data = filter(top15, name=="Last 5 years"),
-            aes(label = package), 
-            hjust = 0, 
-            nudge_x = 0.1,
-            size = 10) + 
-  theme(panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank())
+```{r download-share, fig.width = 10, fig.height = 8, fig.cap = "The distribution of the number of downloads for each DoE package by year; any package that was released in the same year was removed in that year so that each download count is for the full year.
+The label on the bottom of the plot shows the Gini index for downloads.
+The Gini index is consistent a little above 60% each year indicating that there a relatively small number of packages are downloaded most.
+"}
+tar_read(plot_download_dist)
 ```
 
-```{r, fig.cap=glue("The above graph shows the total download counts from {start} to {end} of DoE packages."), fig.width = 6}
-dl1plot + dl2plot
+```{r rank-over-time, fig.width = 10, fig.height = 8, fig.cap = "The plot shows the rank of top 10 downloaded packages by year.
+Most of the top 10 ranking packages have been in top 10 for the last 5 years with `lhs` steadily climbing up the ranks."}
+tar_read(plot_rank_dist)
 ```
-
-```{r text, cache=tocache}
-url <- paste0(getOption("repos")["CRAN"], "web/packages/packages.rds")
-db <- readRDS(url(url)) %>% 
-  as.data.frame()
-
-doe_db <- db %>% 
-  filter(Package %in% doe_pkgs) %>% 
-  mutate(Description = str_replace_all(Description, "\n", " "),
-         Description = str_squish(Description),
-         Title = str_replace_all(Title, "\n", " "))
-```
-
-```{r}
-singularize2 <- function(x) {
-    res <- singularize(x)
-    res <- ifelse(res=="bia", "bias", res)
-    ifelse(res=="baye", "bayes", res)
-}
-stop_words_rex <- c(paste0("^", stop_words$word, "$"), "^doi", "^[0-9.]+$")
-get_ngram <- function(ngram, desc = c("Title", "Description")) {
-  desc <- match.arg(desc)
-  words <- paste0("word", 1:ngram)
-  doe_db %>% 
-    unnest_tokens(word, desc, token = "ngrams", n = ngram) %>% 
-    separate(word, words, sep = " ") %>% 
-    mutate(across(num_range("word", 1:ngram), singularize2)) %>% 
-    distinct(!!!map(c("Package", words), as.name)) %>% 
-    filter(if_all(num_range("word", 1:ngram), 
-                  ~!str_detect(., paste0(stop_words_rex, collapse = "|")))) %>% 
-    count(!!!map(words, as.name), sort = TRUE) %>% 
-    mutate(word = do.call("paste", map(words, ~eval(parse(text = .x))))) %>% 
-    select(word, n) 
-}
-```
-
-## What are the common types of experimental designs?
-
-## How do packages interplay with each other?
-
-## Which packages are widely utilised?
-
-# Software Design
-
-# Propsects and Discussion
-
-\newpage
-
-```{r cran-titles}
-url <- "http://cran.rstudio.com/web/packages/packages.rds"
-db <- readRDS(url(url)) %>% 
-  as.data.frame()
-```
-
-```{r bigram}
-stop_words_ext <- c(stop_words$word, "doi")
-
-doe_db <- db %>% 
-  filter(Package %in% doe_pkgs) %>% 
-  mutate(Description = str_replace_all(Description, "\n", " "),
-         Description = str_squish(Description),
-         Title = str_replace_all(Title, "\n", " "))
-
-bigram_tab <- function(data, col) {
-  data %>% 
-    unnest_tokens(word, {{col}}, token = "ngrams", n = 2) %>% 
-    separate(word, c("word1", "word2"), sep = " ") %>% 
-    mutate(word1 = singularize(word1),
-           word2 = singularize(word2)) %>% 
-    # don't count the same bigram within the same package
-    distinct(Package, word1, word2) %>% 
-    filter(!word1 %in% stop_words_ext,
-           !word2 %in% stop_words_ext,
-           !str_detect(word1, "^[0-9.]+$"),
-           !str_detect(word2, "^[0-9.]+$")) %>% 
-    count(word1, word2, sort = TRUE)  
-}
-```
-
-```{r bigram-title, results="asis"}
-btitle <- bigram_tab(doe_db, Title) %>% 
-  filter(n > 3) %>% 
-  mutate(word = paste(word1, word2)) %>% 
-  select(word, n) %>% 
-  kbl(booktabs = TRUE, linesep = "")
-
-bdesc <- bigram_tab(doe_db, Description) %>% 
-  filter(n > 4) %>% 
-  mutate(word = paste(word1, word2)) %>% 
-  select(word, n) %>% 
-  kbl(booktabs = TRUE, linesep = "")
-
-cat(c("\\begin{table}[h] \\centering ", 
-      btitle,
-    "\\hspace{1cm} \\centering ",
-      bdesc,
-    "\\caption{My tables} \\end{table}")) 
-```
-
-```{r}
-doe_imports <- doe_db %>% 
-  mutate(Depends = str_replace_all(Depends, "\n", " "),
-         Depends = str_replace_all(Depends, fixed("("), " ("),
-         Imports = str_replace_all(Imports, "\n", " "),
-         Imports = str_replace_all(Imports, fixed("("), " ("),
-         imports = str_c(Depends, Imports, sep = ","),
-         imports = str_split(imports, ","),
-         imports = map(imports, ~{
-           str_squish(.x) %>% 
-             word() %>% 
-             .[.!=""]}
-         ),
-         imports_doe = map(imports, ~.x[.x %in% doe_pkgs])) %>% 
-  select(Package, imports_doe) %>% 
-  unnest_longer(imports_doe) %>% 
-  filter(!is.na(imports_doe)) %>% 
-  rename(from = imports_doe, to = Package) %>% 
-  select(from, to)
-```
-
-```{r doe-network, eval = FALSE}
-graph_from_data_frame(doe_imports) %>% 
-  ggraph(layout = 'fr') +
-  geom_edge_link(aes(start_cap = label_rect(node1.name),
-                     end_cap = label_rect(node2.name)), 
-                 arrow = arrow(length = unit(2, 'mm')),
-                 color = "#79003e") + 
-  geom_node_text(aes(label = name),
-                 color = "#79003e") +
-  theme(panel.background = element_rect(fill = "#f6e5ee",
-                                        color = "#79003e"),
-        plot.margin = margin(20, 20, 20, 20))
-```
-
-\CRANpkg{agricolae}
-
-## Helping info to get started
-
-Introductory section which may include references in parentheses
-[@R], or cite a reference such as @R in the text.
-
-## Section title in sentence case
-
-Let's check fi this works Figure \@ref(fig:Rlogo).
-
-This section may contain a figure such as Figure \ref{fig:Rlogo}.
-
-## Summary
-
-This file is only a basic article template.
-For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](https://journal.r-project.org/share/author-guide.pdf).
-
-### About this format and the R Journal requirements
-
-`rticles::rjournal_article` will help you build the correct files requirements:
-
-* A R file will be generated automatically using `knitr::purl` - see
-https://bookdown.org/yihui/rmarkdown-cookbook/purl.html for more information.
-* A tex file will be generated from this Rmd file and correctly included in
-`RJwapper.tex` as expected to build `RJwrapper.pdf`.
-* All figure files will be kept in the default rmarkdown `*_files` folder. This
-happens because `keep_tex = TRUE` by default in `rticles::rjournal_article`
-* Only the bib filename is to modifed. An example bib file is included in the
-template (`RJreferences.bib`) and you will have to name your bib file as the
-tex, R, and pdf files.

### 2022-05-02 R/functions.R

@@ -15,13 +15,21 @@ plot_setup <- function() {
           plot.title = element_text(color = "black", hjust = 0))
 }
 
-
-
-
 ctv_list <- function(ctv) {
   ctv:::.get_pkgs_from_ctv_or_repos(ctv, 
                                     repos = "http://cran.rstudio.com/")[[1]]  
 }
+
+get_pkg_updates <- function(pkgs, message = FALSE) {
+  res <- lapply(pkgs, function(pkg) {
+    message(paste("Package", pkg))
+    pkgsearch::cran_package_history(pkg) %>% 
+      select(package = Package, update = date) %>% 
+      mutate(update = as.Date(update))
+  })
+  do.call("rbind", res)
+}
+
 
 cran_downloads_duration <- function(pkgs, duration) {
   end <- as.Date("2022-03-01")  
@@ -127,48 +135,6 @@ plot_rank_distribution <- function(data, nrank = 10, stat) {
   }
 }
 
-
-pkg_updates <- function(pkgs) {
-  pkg_url <- "https://cran.r-project.org/web/packages/{pkg}/index.html"
-  pkg_archive <- "https://cran.r-project.org/src/contrib/Archive/{pkg}/"
-  res <- map(pkgs, function(pkg) {
-    last_update <- read_html(glue(pkg_url)) %>% 
-      html_table() %>% 
-      .[[1]] %>% 
-      filter(X1=="Published:") %>% 
-      pull(X2) %>% 
-      ymd()
-    
-    archive_dates <- tryCatch({ 
-      read_html(glue(pkg_archive)) %>% 
-        html_table() %>%
-        .[[1]] %>% 
-        pull(`Last modified`) %>% 
-        ymd_hm() %>% 
-        na.omit() %>% 
-        as.Date()
-    }, error = function(e) {
-      NULL
-    })
-    c(archive_dates, last_update)
-  })
-  names(res) <- pkgs
-  
-  unlist(res) %>% 
-    enframe("package", "update") %>% 
-    # unlist converts date to integers
-    mutate(update = as.Date(update, origin = "1970-01-01"),
-           # need to get rid of the numbers appended to pkg names
-           package = str_extract(package, paste0(pkgs, collapse="|")),
-           package = factor(package, levels = pkgs)) 
-}
-
-pkg_updates_duration <- function(pkgs, duration) {
-  pkg_updates(pkgs) %>% 
[…]

### 2022-05-02 _targets.R

@@ -7,13 +7,14 @@ tar_option_set(packages = c("tidyverse", "rvest", "lubridate",
                             "tidytext", "pluralize", "kableExtra", 
                             "igraph", "ggraph", "ggrepel", "patchwork",
                             "ggwordcloud", "fable", "feasts", "lubridate",
-                            "ineq"))
+                            "ineq", "cranscrub"))
 
 
 list(
   tar_target(doe_pkgs,  ctv_list("ExperimentalDesign")),
   tar_target(npkgs, length(doe_pkgs)),
-  tar_target(doe_db, pkg_db(doe_pkgs)),
+  tar_target(doe_db, cranscrub::pkg_db(doe_pkgs)),
+  tar_target(updates_all, get_pkg_updates(doe_pkgs)),
   tar_target(cran_download_data2, cran_downloads_duration(doe_pkgs, 2)),
   tar_target(cran_download_data5, cran_downloads_duration(doe_pkgs, 5)),
   tar_target(cran_scrub_data2, cran_ed_scrub_duration(2)),
@@ -22,7 +23,6 @@ list(
   tar_target(cran_rank5, cran_downloads_rank(cran_download_data5)),
   tar_target(cran_srank2, cran_downloads_rank(cran_scrub_data2)),
   tar_target(cran_srank5, cran_downloads_rank(cran_scrub_data5)),
-  tar_target(updates_all, pkg_updates(doe_pkgs)),
   tar_target(first_release, updates_all %>% 
                group_by(package) %>% 
                summarise(first = min(update)) %>% 
@@ -44,13 +44,13 @@ list(
   tar_target(plot_scrub_trend, download_trend(cran_scrub_data5, updates2) + plot_setup()),
   tar_target(plot_stl_7wy, plot_stl_model(cran_scrub_data5, "agricolae", updates2, 7, "week", "year") + plot_setup()),
   
-  tar_target(bigram_title, get_ngram(2, "Title")),
-  tar_target(bigram_desc, get_ngram(2, "Description")),
-  tar_target(trigram_desc, get_ngram(3, "Description")),
-  
-  tar_target(dldat_comb_bigram_title,  combine_dl(bigram_title, Title)),
-  tar_target(dldat_comb_bigram_desc,  combine_dl(bigram_desc, Title)),
-  tar_target(dldat_comb_trigram_desc,  combine_dl(trigram_desc, Title)),
[…]

### 2022-05-31 other/tidy_author-ET.Rmd
significant update to the paper

---
title: "tidy_author"
author: "Dewi Amaliah"
date: "24/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#library(qdapRegex)
#library(entity)
```

```{r}
url <- "http://cran.rstudio.com/web/packages/packages.rds"
db <- readRDS(url(url)) %>% 
  as.data.frame()

ctvs <- map_dfr(ctv::available.views(repos = "http://cran.rstudio.com/"), 
                  ~tibble(name = .x[["name"]],
                          topic = .x[["topic"]],
                          pkgs = map(name, ~ctv:::.get_pkgs_from_ctv_or_repos(.x, 
                                                repos = "http://cran.rstudio.com/")[[1]])))

words_to_delete <- c(" with", " by", " under", " contributions?", " is", " the", 
                     " author", " [Oo]riginal(ly)?", " C")

ctvdf <- ctvs %>% 
  unnest_longer(pkgs) %>% 
  left_join(select(db, Package, Author, Depends, Imports, Suggests), 
            by = c("pkgs" = "Package")) %>% 
  mutate(across(c(Depends, Imports, Suggests),
                ~str_split(.x, ", "))) %>% 
  mutate(across(c(Depends, Imports, Suggests),
                ~map(.x, function(apkg) str_extract(apkg, "^[A-Za-z0-9]+")))) %>%

  mutate(
    # remove angle bracket
    author_tidy = str_replace_all(Author, "\\n", ""),
    # remove square bracket
    author_tidy = str_replace_all(author_tidy, "<.+>", ""),
    author_tidy = str_replace_all(author_tidy, "\\[.+\\]", ""),
    # remove bracket
    author_tidy = str_replace_all(author_tidy, "\\(.*\\)", ""),
    author_tidy = str_replace_all(author_tidy, "[.]", ""),
    author_tidy = str_replace_all(author_tidy, "(&| and)", ","),
    author_tidy = str_replace_all(author_tidy, ",,", ","),
    author_tidy = str_replace(author_tidy, " ,$", ""),
    author_tidy = str_replace_all(author_tidy, 
                                  paste0("(", paste(words_to_delete, collapse = "|"), ")"),
                                  ""),
    author_tidy = str_squish(author_tidy)) %>% 
  select(Author, author_tidy)

tidytext::stop_words

ctvdf_sentences <- ctvdf %>%
  filter(str_detect(author_tidy, "riginal") | str_detect(author_tidy, " by") | str_detect(author_tidy, "with") | str_detect(author_tidy, " under"))

ctvdf_sentences_test <- ctvdf_sentences %>%
  mutate(author_tidy2 = person_entity(author_tidy)) %>%
  mutate(author_tidy3 = str_replace(author_tidy, "with contributions from", ","),
         author_tidy3 = str_replace(author_tidy, "Original by", ""))

ctvdf_with <- ctvdf %>%
  filter(str_detect(author_tidy, "with"))

# try to filter English words

print(ctvdf_with$author_tidy)
```

### 2022-05-31 paper.Rmd
significant update to the paper

@@
 ---
-title: Current status and propsects of R-packages for the design of experiments
+title: Current status and prospects of R-packages for the design of experiments
 author:
@@
 bibliography: paper.bib
-
+knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'index.html')) })
+keywords: ["experimental design", "CRAN task view", "user interface"]
 output: 
@@
     theme: paper
+  # below is to have a JSS paper style template
+  pagedown::jss_paged:
+    template: assets/html-template.html
+    csl: assets/html-template.csl
+    self_contained: false
 ---
+
+```{css, echo = FALSE}
+h1, h2, h3, div > * {
+ font-family: "Helvetica Neue",Helvetica,Arial,sans-serif;
+}
+
+```
 
@@
 library(tidyverse)
+library(patchwork)
+library(tidygraph)
+library(ggraph)
+source("R/plot.R")
 knitr::opts_chunk$set(echo = FALSE, 
-                      cache = FALSE,
+                      cache = TRUE,
                       cache.path = "cache/",
@@
                       fig.width = 6,
+                      fig.height = 6.5,
                       message = FALSE,
@@
 ctv <- function(view) {
-  ifelse(knitr::is_html_output(), glue("[{view}](https://cran.r-project.org/web/views/{view}.html)"), glue("\\ctv{[view]}", .open = "[", .close = "]"))
+  ifelse(knitr::is_html_output(), glue("[{view}](https://cran.r-project.org/web/views/{view}.html)"), glue("ctv{[view]}", .open = "[", .close = "]"))
 }
@@
 tar_load(npkgs)
+tar_load(gini_yearly)
+tar_load(gini_yearly_ctv)
 ```
@@
 Methods for data collection can be dichotomised by the type of data collected -- namely, experimental or observational -- or alternatively, categorised as experimental design (including quasi-experimental design) or survey design.
-This dichotomisation, to a great extent, is seen in [CRAN task views](https://cran.r-project.org/web/views/) (a volunteered maintained list of R-packages by topic) where R-packages in experimental design are in `r ctv("ExperimentalDesign")` and R-packages in survey designs are in `r ctv("OfficialStatistics")`.
-Survey designs often, although not always, aim to collect observational data whilst experimental designs exclusively center on experimental data where the subject and experimental factors tend to be in the control of the experimenter.
-This paper is concerned with the latter.
+This dichotomisation, to a great extent, is seen in the [CRAN task views](https://cran.r-project.org/web/views/) (a volunteered maintained list of R-packages by topic) where R-packages in experimental design are in `r ctv("ExperimentalDesign")` and R-packages in survey designs are in `r ctv("OfficialStatistics")`.
+Experimental designs exclusively center on the collection of experimental data where the subject and experimental factors tend to be in the control of the experimenter.
+A subset of experimental designs are segregated into `r ctv("ClinicalTrials")`, where the focus is on clinical trials with primary interest in sample size calculations.
+This paper focuses on the packages in `r ctv("ExperimentalDesign")`, henceforth referred to as "**DoE packages**".
 
-In the CRAN task view of `r ctv("ExperimentalDesign")`, there are `r npkgs` R packages for experimental design and analysis of data from experiments, henceforth referred to as "**DoE packages**" in this paper.
+In the CRAN task view of `r ctv("ExperimentalDesign")`, there are `r npkgs` R packages for experimental design and analysis of data from experiments.
 The sheer quantity and variation in the output experimental design in the R-packages are arguably unmatched with any other programming languages, e.g.
@@
 The paper is organised as follows.
-Section \@ref(eda) explores some analysis of the most downloaded DoE packages, followed by discussion of the discussion of the software design of these packages in Section \@ref(design), and concluding with a discussion in Section \@ref(discussion) of future prospects.
+Section `r ref("data")` briefly describes the data source used for the analysis in Section \@ref(eda); Section \@ref(eda) presents some insights into the state of the current DoE packages; Section \@ref(design) discuss the software design of DoE packages, and we conclude with a discussion in Section \@ref(discussion) of future prospects.
 
-# Explorative analysis {#eda}
+# Data {#data}
 
-The Comprehensive R Archive Network (CRAN) is a network of servers located across the world that store R and R-packages.
+To study the DoE packages, we analyse data using three sources of data as described next.
 
-```{r download-share, fig.width = 10, fig.height = 8, fig.cap = "The distribution of the number of downloads for each DoE package by year; any package that was released in the same year was removed in that year so that each download count is for the full year.
-The label on the bottom of the plot shows the Gini index for downloads.
-The Gini index is consistent a little above 60% each year indicating that there a relatively small number of packages are downloaded most.
-"}
+## RStudio CRAN download logs
+
+The Comprehensive R Archive Network (CRAN) is a network of servers located across the world that store mirrored versions of R and R-packages.
+The most popular network is the RStudio mirror (the default server for those that use the RStudio IDE).
+The RStudio mirror is also the only server that provides a comprehensive daily download logs of R and R-packages since October 2012.
+The summary data can be easily accessed with the `cranlogs` package [@cranlogs].
+This paper uses the data from the beginning of 2013 to end of 2021 (a total of 9 years) for the DoE packages.
+
+## Package descriptions
+
+All CRAN packages have a title, description, package connections (suggests, dependency and imports) and more in the DESCRIPTION file.
+We use the text data from the title and description (as accessed in `r as.Date(file.info("_targets/objects/ngram_download_data")$mtime)`) in Section \@ref(topics).
+
+## CRAN task views
+
+The list of packages in each CRAN task views (as of `r as.Date(file.info("_targets/objects/ctv_pkgs_data")$mtime)`) are used to contrast the characteristics of DoE packages in Section \@ref(silo).
+
+# Explorative data analysis {#eda}
+
+All results presented are derived from exploratory data analysis of observational data; consequently, all interpretations are somewhat speculative and 
+may not be indicative of the true state of the field of experimental design.
+In particular, any analysis over time are confounded by the fact that the nature of users and package management have also changed over the years.
+It should be noted that some DoE packages may have been archived or removed from the task view over the years so any cross-sectional analysis presented may not reflect the set of all DoE packages at that particular time period (although we assume such incidences are low).
+
+A subset of DoE packages is not primarily about design of experiments but about the analysis of experimental data.
+A complete delineation of these packages is difficult as many have some functions that can aid decisions or constructions of experimental designs (and any categorisation is prone to our subjective bias) so we opted not to remove any DoE packages in the analysis.
+
+## Small, but diverse, set of packages are sufficient for most designs {#popular}
+
+There have been at least 50 DoE packages since 2013 but most of the downloads are concentrated in just a handful of packages.
+For example, Figure \@ref(fig:plot-lorenz) shows a Lorenz curve [@Lorenz1905-tc] for the total package downloads in 2021 of `r tar_read(gini_yearly) %>% filter(year==2021) %>% pull(n)` DoE packages (first released prior to 2021); we can see from Figure \@ref(fig:plot-lorenz) that bottom 90\% of DoE packages (in terms of total download count in 2021) only share about 30\% of total downloads across all DoE packages -- in another words, 70\% of the total downloads are due to 11 packages (10% of the DoE packages).
+
+```{r plot-lorenz, fig.height = 4, fig.width = 4, fig.cap = "Lorenz curve of the total download count for DoE packages in 2021. The red line corresponds to the line of perfect equality."}
+tar_read(plot_lorenz2021) +
+  plot_setup() +
+  labs(y = "Cumulative share of downloads (%)",
+       x = "Percentage of DoE packages") +
+  theme(plot.margin = margin(r = 20, t = 20)) +
+  scale_x_continuous(breaks = seq(0, 100, by = 10)) +
+  scale_y_continuous(breaks = seq(0, 100, by = 10))
+```
+
+ If we consider package downloads as a measure of "wealth", then we can consider using the Gini index [@Gini1921-mf] as a measure of download inequality across packages.
+The ratio of the red region over the total colored regions in Figure \@ref(fig:plot-lorenz) corresponds to the Gini index for 2021.
+A Gini index of 0\% indicates equality in downloads across packages while a value of 100\% indicates maximal inequality (all downloads are due to one package).
+In Figure \@ref(fig:download-share), we see that the distributions of the downloads each year have a heavy right tail with the Gini index ranging from `r scales::percent(min(gini_yearly$gini), 0.1)` to `r scales::percent(max(gini_yearly$gini), 0.1)` across the years 2013 to 2021, indicating that there is a high level of inequality of downloads across packages, particularly with more pronounced inequality in the last 5 years.
+
+```{r download-share, fig.width = 5.5, fig.height = 4, fig.cap = "The distribution of the number of downloads for each DoE package by year.
+Packages were removed in any year if it was released in that year or later so that each download count is for the full year.
+The label on the bottom of the plot shows the Gini index for downloads and the number of packages with full year of download count in the corresponding year.
+In the last 5 years, the Gini index is consistent above 60\\% each year indicating that most downloads are due to a relatively small number of packages."}
 tar_read(plot_download_dist)
@@
 
-```{r rank-over-time, fig.width = 10, fig.height = 8, fig.cap = "The plot shows the rank of top 10 downloaded packages by year.
-Most of the top 10 ranking packages have been in top 10 for the last 5 years with `lhs` steadily climbing up the ranks."}
+While in absolute terms the Gini index is high for `r ctv("ExperimentalDesign")`, the inequality is not as severe as other CRAN task views as shown in Figure \@ref(fig:fig-gini-all-ctvs).
+We can see in Figure \@ref(fig:fig-gini-all-ctvs) that the Gini index is generally increasing for DoE packages (as is generally the case for other CRAN task views although this is hard to see from this plot).
+
+```{r fig-gini-all-ctvs, fig.height = 8, fig.cap = "The points show the Gini index of the download counts by year facetted by CRAN task view with the color showing the number of packages.
+The grey line shows the distribution of the Gini index across years for all other CRAN task views.
+The facets are ordered by increasing value of the Gini index in 2021."}
+gini_yearly_ctv %>% 
+  ungroup() %>%  
+  arrange(ctv, year) %>% 
+  mutate(ctv = fct_reorder(ctv, gini, last),
+         year = factor(year),
+         gini = 100 * gini) %>%  
+  ggplot(aes(year, gini)) + 
+  geom_line(data = ~rename(.x, ctv2 = ctv), 
+            aes(group = ctv2), color = "grey") +
+  facet_wrap(~ctv, ncol = 4) + 
+  geom_line(aes(group = ctv)) + 
+  geom_point(aes(color = n))  +
+  plot_setup() +
+  labs(x = "Year", y = "Gini index (%)", color = "# of packages") +
+  guides(color = guide_colorbar(direction = "horizontal")) +
+  theme(legend.position = c(0.5, 0), 
+        legend.justification = "left",
+        strip.text = element_text(size = 7),
+        axis.text.x = element_text(angle = 90, vjust = 0.5)) +
+  scale_colour_distiller(palette = "YlOrRd", direction = 1) 
+```
+
+From Figures \@ref(fig:download-share) and \@ref(fig:fig-gini-all-ctvs), we could draw some indicative states of the field of experimental designs with possible counterfactual interpretations:
+
+* The increase in the number of packages that are not highly downloaded may mean that **_there are more packages to construct niche experimental designs_**.
+Some examples of these packages include `qtlDesign`, `PwrGSD` and `Crossover` made for QTL experiments, group sequential designs and crossover trials, respectively.
+These packages would naturally have a smaller number of potential users.
+Counterfactual to this, the increase could be due to other external factors, such as an increase in a number of skilled contributors, a change in CRAN policy or management to add packages (either to CRAN and/or task view), and/or that new packages are still yet to amass users.
+While there is an argument that low download counts are due to the low utility and/or quality of the packages, packages in CRAN task views are selected by expert maintainers; thus we can reasonably assume that any package listed in CRAN task view are of a decent utility and quality.
+* If the downloads are reflective of the experimental designs used in practice, **_small set of packages appear to be sufficient for most to construct the full set of designs of experiments needed in practice_**.
+Packages of course evolve and the top downloaded packages have had regular updates that may have broadened its scope.
+* While small subset of DoE packages are most frequently used, none of these packages are dominant in the field (judging from comparisons with other CRAN task views).
+This suggests that there are **_diverse approaches to designing experiments_**.
+This observation doesn't take into account other approaches to generating experimental designs, like the proprietary GUI software, CycDesignN, which may be widely used.
+
+## There is a lack of adaptation of new or innovative designs  {#ranking}
+
+We can see in Figure \@ref(fig:rank-over-time) that most of the top 10 ranking packages have been in the top 10 for the last 9 years with `lhs` steadily climbing up the ranks in the last few years.
+
+```{r rank-over-time, fig.width = 5.5, fig.height = 4.5, fig.cap = "The plot shows the rank of top 10 downloaded packages by year. "}
 tar_read(plot_rank_dist)
 ```
+
+```{r established-packages, fig.height = 5, fig.cap = "The above figure shows the total download (in log scale) of a package in the corresponding year against the first release date of the package.
+The blue line corresponds to the fit of a simple linear regression model.
+In any given year, "}
+tar_read(cran_rank_yearly) %>%
+  mutate(nupdates = map2_int(package, year, ~{
+    tar_read(updates_all) %>% 
+      filter(package == .x,
+             lubridate::year(update) <= .y) %>% 
+      nrow()
+  })) %>% 
+  ggplot(aes(first, total/1000)) +
+  geom_point(aes(color = nupdates)) + 
+  facet_wrap(~year, scale = "free_y") +
+  geom_smooth(se = FALSE, 
+              method = "lm", 
+              formula = y ~ x) +
+  scale_y_log10() + 
+  scale_color_continuous(trans = "log10") + 
+  plot_setup() +
+  labs(x = "First release date",
+       y = "Total download (1000s)",
+       color = "# of updates")
+```
+
+* The lack of change in the top ranking packages are indicative that **_there has been a lack of innovation in the field of experimental design_**.
+
+## Computer experiments are on the rise {#topics}
+
+```{r wordcloud-over-time, fig.width = 5.5, fig.height = 5, fig.cap = "The above figures shows the word cloud of bigrams from the title and descriptions of the CRAN packages.
+The color shows how often the bigram appears across the DoE packages and the size are relative to the total download count of the packages that contain the bigram."}
+tar_read(plot_wordcloud2) + 
+  scale_size_area(max_size = 8) +
+  scale_color_continuous(breaks = c(2, 4, 6, 8, 10, 12)) +
+  theme(legend.text = element_text(size = 10),
+        legend.title = element_text(size = 12))
+```
+
+## Development of experimental designs occur in silos {#silo}
+
+```{r ctv-summ-table}
+tar_read(ctv_summary_table) %>% 
+  knitr::kable(col.names = c("Name", "Topic", "# of packages", "Total # of contributors" , "Average # of contributors", "Intra-connectivity (%)"), 
+               caption = "A summary table for the CRAN task view that shows in order: the name of the task view, the full topic name, the total of packages, the total number of contributors, the average number of contributors, and the intra-connectivity.
+The intra-connectivity measures the percentage of packages that depends, suggest or imports at least one other package within the same task view.
+A low intra-connectivity suggests that development within the topic mostly occur in silos whilst high  intra-connectivity suggests that there are more interactions within the topic.",
+               digits = c(0, 0, 0, 1, 2, 0),
+               booktabs = TRUE ) %>%
+ kableExtra::kable_styling(latex_options = c("striped", "scale_down")) %>% 
+ kableExtra::column_spec(2, width = "12em") %>% 
+ kableExtra::column_spec(3:6, width = "5em")
+
+```
+
+```{r ctv-summ-plot, fig.width = 5.5, fig.height = 6, fig.cap = "The figure above is a visualisation of Table 1.
+The task views in the bottom-left corner are topics that are more indicative of contributors working in silos."}
+tar_read(ctv_summary_table) %>% 
+  ggplot(aes(avg_contributors, intra_perc)) +
+  geom_point(aes(size = ncontributors,
+                 color = n)) +
+  ggnewscale::new_scale_color() + 
+  ggrepel::geom_text_repel(aes(label = name,
+                               color = name=="ExperimentalDesign"),
+                           force = 150,
+                           max.iter = 20000) +
+  guides(color = "none") + 
+  scale_color_manual(values = c("black", "#CC3311")) + 
+  labs(x = "Average number of contributors",
+       y = "Intra-connectivity (%)",
+       size = "Total number of contributors",
+       color_new = "Total number of packages") +
+  plot_setup() +
+  theme(legend.position = "bottom",
+        legend.box = "vertical",
+        legend.box.just = "left")
+```
+
+```{r plot-doe-network, fig.width = 6, fig.height = 6, fig.cap  = "Package connections (depends, suggests and imports) within the DoE packages.
+DoE packages that do not depend, suggest or import another DoE package is not shown."}
+tar_read(plot_doe_network) +
+  theme(legend.position = "bottom")
+```
+
+# Interface design  {#design}
+
+In software design, there are two interface designs to consider: user interface (UI) and application programming interface (API).
+UI is concerned with the usage of the software by the user, while the API is concerned with how different programs interact and is predominately of the interest to the developer.
+In this section, we discuss these interface designs broadly, with examples illustrated from the top downloaded packages (as shown in Figure \@ref(fig:rank-over-time)).
+We exclude `ez` and `DiceKriging` from the discussion as the former  is predominately visualisation of experimental data and latter is about the analysis of computer experiments in addition to belonging to the same suite of packages as `DiceDesign`.
+
+## The case of specialised designs
+
+Computer experiments, as discussed in Section \@ref(topics), which generally involve space filling designs have been on the rise. The exemplar for this type of designs are `lhs` and `DiceDesign`.
+
+Factorial experiments offer a challenge in the construction and allocation of the treatment factors.
+This is reflected in a number of packages that specifically addresses this challenge.
+These include `DoE.base` [@DoE.base], `FrF2` [@FrF2] for fractional 2-level factorial designs, `conf.design` [@conf.design], `FrF2.catlg128` [@FrF2.catlg128] and `BHH2` [@BHH2].
+
+Response surface design, as illustrated by `rsm` [@rsm], is a type of experiment that involve variables with continuous measures.
+
+Finally, another specialised design is sequential design (also called adaptive sampling), which is best represented by `tgp` [@tgp] -- these require prior information, which are used to inform the next experimental design.
+Follow-up experiments, which can also be classified as a sequential design, is implemented by `BsMD` [@BsMD].
+
+## The case of menu functions
+
+`agricolae` [@agricolae]
+
+## The case of general designs
+
+`AlgDesign` [@AlgDesign]
+
+`DoE.wrapper` [@DoE.wrapper]
+
+# Discussion {#discussion}
+
+Volunteer maintained -- some experimental design packages may not be included in the CRAN task view.
+
+# Acknowledgement
+
+This paper uses the `targets` framework [@targets] for reproducibility, `knitr` [@knitr] and `rmarkdown` [@rmarkdown] for creating reproducible documents, `ggplot2` [@ggplot2] for visualisation and `kableExtra` [@kableExtra] for customising the table.
+All code to reproduce this paper is found at https://github.com/emitanaka/paper-DoE-review.
+
+# References

### 2022-05-31 R/functions.R

@@ -1,26 +1,19 @@
 
-plot_setup <- function() {
-  theme(panel.background = element_rect(fill = NA),
-          panel.grid = element_line(color = "lightgray"),
-          axis.text = element_text(color = "black"),
-          axis.line = element_line(color = "black", size = 0.7),
-          axis.ticks.length = unit(1.4, "mm"),
-          axis.ticks = element_line(color = "black", size = 0.7),
-          axis.title = element_text(color = "black", face = "bold"),
-          strip.background = element_rect(color = "black",
-                                          fill = "white"),
-          strip.text = element_text(color = "black", 
-                                    margin = margin(5, 5, 5, 5)),
-          plot.title.position = "plot",
-          plot.title = element_text(color = "black", hjust = 0))
-}
+
 
 ctv_list <- function(ctv) {
   ctv:::.get_pkgs_from_ctv_or_repos(ctv, 
                                     repos = "http://cran.rstudio.com/")[[1]]  
 }
 
-get_pkg_updates <- function(pkgs, message = FALSE) {
+pkg_first_release <- function(data) {
+  data %>% 
+    group_by(package) %>% 
+    summarise(first = min(update)) %>% 
+    ungroup()
+}
+
+pkg_update_dates <- function(pkgs, message = FALSE) {
   res <- lapply(pkgs, function(pkg) {
     message(paste("Package", pkg))
     pkgsearch::cran_package_history(pkg) %>% 
@@ -31,10 +24,14 @@ get_pkg_updates <- function(pkgs, message = FALSE) {
 }
 
 
-cran_downloads_duration <- function(pkgs, duration) {
-  end <- as.Date("2022-03-01")  
-  start <- end - years(duration)
-  cran_downloads(pkgs, from = start, to = end)  
+cran_downloads_duration <- function(pkgs, start_year) {
+  # it can have trouble downloading long periods so break it up by year
+  map_dfr(start_year:2021, ~{
+    start <- as.Date(paste0(.x, "-01-01"))
+    end <- as.Date(paste0(.x, "-12-31"))
+    cran_downloads(pkgs, from = start, to = end)  
+  })
+  
 }
 
 cran_ed_scrub_duration <- function(duration) {
@@ -68,94 +65,38 @@ cran_downloads_rank_by_year <- function(data, first_release) {
[…]

### 2022-05-31 R/plot.R

@@ -1,0 +1,181 @@
+
+plot_setup <- function() {
+  theme(panel.background = element_rect(fill = NA),
+        panel.grid = element_line(color = "lightgray"),
+        axis.text = element_text(color = "black"),
+        axis.line = element_line(color = "black", size = 0.7),
+        axis.ticks.length = unit(1.4, "mm"),
+        axis.ticks = element_line(color = "black", size = 0.7),
+        axis.title = element_text(color = "black", face = "bold"),
+        strip.background = element_rect(color = "black",
+                                        fill = "white"),
+        strip.text = element_text(color = "black", 
+                                  margin = margin(5, 5, 5, 5)),
+        plot.title.position = "plot",
+        plot.title = element_text(color = "black", hjust = 0))
+}
+
+plot_stl_model <- function(data, pkg, updates, trend, speriod1, speriod2) {
+  data %>% 
+    filter(package == pkg) %>% 
+    as_tsibble(index = date) %>% 
+    model(STL(count ~ trend(window = !!trend) + season(period = !!speriod1) + season(period = !!speriod2),
+              robust = TRUE)) %>% 
+    components() %>% 
+    autoplot() + 
+    geom_vline(data = filter(updates, package == pkg), 
+               aes(xintercept = update), color = "red")
+  
+}
+
+plot_download_distribution <- function(data, ginidata) {
+  data %>% 
+    mutate(year = as.factor(year)) %>% 
+    ggplot(aes(x = year, y = total)) + 
+    geom_violin() +
+    geom_boxplot(width = 0.1) + 
+    labs(x = "Year", y = "Downloads") + 
+    scale_y_log10(label = comma, expand = c(0.1, 0)) +
+    geom_label(data = mutate(ginidata, year = as.factor(year)), 
+               size = 3,
+               aes(y = 900, label = paste0(scales::percent(gini, 0.1),
+                                           "\nn=", n)))
+}
+
+plot_lorenz_curve <- function(data, year) {
+  
+  dat <- data %>% 
+    filter(year == .env$year)
+  
+  data <- data.frame(p = ineq::Lc(dat$total)$p, L = ineq::Lc(dat$total)$L) 
+  anndata <- data %>% 
+    arrange(desc(L)) %>% 
+    filter(L < 0.5) %>% 
+    slice(1)
+  ggplot(data, aes(p * 100, L * 100)) +
+    geom_abline(color = "red") +
+    geom_line() + 
+    geom_point() +
[…]

### 2022-05-31 _targets.R

@@ -2,56 +2,115 @@ library(targets)
 library(ggplot2)
 library(tarchetypes)
 source("R/functions.R")
+source("R/plot.R")
 tar_option_set(packages = c("tidyverse", "rvest", "lubridate",
                             "cranlogs", "glue", "scales", "colorspace",
                             "tidytext", "pluralize", "kableExtra", 
                             "igraph", "ggraph", "ggrepel", "patchwork",
                             "ggwordcloud", "fable", "feasts", "lubridate",
-                            "ineq", "cranscrub"))
+                            "ineq", "cranscrub", "tidygraph", "ggraph", 
+                            "rlang", "igraph"))
 
 
 list(
-  tar_target(doe_pkgs,  ctv_list("ExperimentalDesign")),
+  tar_target(doe_pkgs,  cranscrub::ctv_pkgs("ExperimentalDesign")),
+  tar_target(ctv_pkgs_df, cranscrub::ctv_table()),
   tar_target(npkgs, length(doe_pkgs)),
   tar_target(doe_db, cranscrub::pkg_db(doe_pkgs)),
-  tar_target(updates_all, get_pkg_updates(doe_pkgs)),
-  tar_target(cran_download_data2, cran_downloads_duration(doe_pkgs, 2)),
-  tar_target(cran_download_data5, cran_downloads_duration(doe_pkgs, 5)),
+  tar_target(updates_all, pkg_update_dates(doe_pkgs)),
+  tar_target(first_release, pkg_first_release(updates_all)),
+  
+  tar_target(cran_download_data2, cran_downloads_duration(doe_pkgs, 2020)),
+  tar_target(cran_download_data5, cran_downloads_duration(doe_pkgs, 2013)),
+  tar_target(cran_downloads_all_ctv, map_dfr(1:nrow(ctv_pkgs_df), ~{
+      cran_downloads_duration(ctv_pkgs_df$packages[[.x]], 2013) %>% 
+        mutate(ctv = ctv_pkgs_df$name[[.x]])
+    })),
+  tar_target(updates_all_ctv, map_dfr(1:nrow(ctv_pkgs_df), ~{
+    pkg_update_dates(ctv_pkgs_df$packages[[.x]])
+  })),
+  tar_target(first_release_all, pkg_first_release(updates_all_ctv)),
+  
+  
+  
+  # scrub data --- not much interesting difference 
   tar_target(cran_scrub_data2, cran_ed_scrub_duration(2)),
[…]

### 2022-05-31 rjournal/paper-rjournal.R

@@ -1,288 +1,0 @@
-## ---- include = FALSE---------------------------------------------------------
-paper <- here::here("paper/index.Rmd")
-meta <- rmarkdown::yaml_front_matter(paper)
-
-
-## ----child=paper--------------------------------------------------------------
-
-## ----setup, cache = FALSE, include = FALSE------------------------------------
-library(tidyverse)
-library(rvest)
-library(lubridate)
-library(cranlogs)
-library(glue)
-library(scales)
-library(colorspace)
-library(tidytext)
-library(pluralize)
-library(kableExtra)
-library(igraph)
-library(ggraph)
-library(ggrepel)
-library(patchwork)
-library(ggwordcloud)
-tocache <- TRUE
-knitr::opts_chunk$set(echo = FALSE, 
-                      cache = FALSE,
-                      cache.path = "cache/",
-                      fig.align = 'center', 
-                      fig.pos = 'htbp', 
-                      fig.path = "figures/",
-                      fig.width = 6,
-                      message = FALSE,
-                      warning = FALSE)
-
-theme_set(
-  theme(panel.background = element_rect(fill = NA),
-        panel.grid = element_line(color = "lightgray"),
-        axis.text = element_text(color = "black"),
-        axis.line = element_line(color = "black", size = 0.7),
-        axis.ticks.length = unit(1.4, "mm"),
-        axis.ticks = element_line(color = "black", size = 0.7),
-        axis.title = element_text(color = "black", face = "bold"),
-        strip.background = element_rect(color = "black",
-                                        fill = "black"),
-        strip.text = element_text(color = "white"),
-        plot.title.position = "plot",
-        plot.title = element_text(color = "black", hjust = 0)))
-
-ctv <- function(view) {
-  ifelse(knitr::is_html_output(), glue("[{view}](https://cran.r-project.org/web/views/{view}.html)"), glue("\\ctv{[view]}", .open = "[", .close = "]"))
-}
-CRANpkg <- function(pkg) {
-  ifelse(knitr::is_html_output(), glue("[{pkg}](https://cran.r-project.org/web/packages/{pkg}/index.html)"), glue("\\CRANpkg{[pkg]}", .open = "[", .close = "]"))
-}
-
-
[…]

### 2022-05-31 rjournal/paper.R

@@ -1,76 +1,0 @@
-## ----setup, cache = FALSE, include = FALSE------------------------------------
-library(tidyverse)
-library(rvest)
-library(lubridate)
-library(cranlogs)
-library(glue)
-library(scales)
-library(colorspace)
-library(tidytext)
-library(pluralize)
-library(kableExtra)
-library(igraph)
-library(ggraph)
-
-knitr::opts_chunk$set(echo = FALSE, 
-                      cache = TRUE,
-                      cache.path = "cache/")
-
-myggtheme <- 
-  theme(panel.background = element_rect(fill = NA),
-        panel.grid = element_line(color = "#f6e5ee"),
-        axis.text = element_text(color = "#79003e"),
-        axis.line = element_line(color = "#79003e", size = 0.7),
-        axis.ticks.length = unit(1.4, "mm"),
-        axis.ticks = element_line(color = "#79003e", size = 0.7),
-        axis.title = element_text(color = "#79003e", face = "bold"),
-        strip.background = element_rect(color = "#79003e",
-                                        fill = "#AD0059"),
-        strip.text = element_text(color = "white"),
-        plot.title.position = "plot",
-        plot.title = element_text(color = "#79003e", face = "bold")) 
-
-
-## ----cran-titles--------------------------------------------------------------
-# Thanks to Dirk Eddelbuettel's answer on SO:
-# https://stackoverflow.com/questions/11560865/list-and-description-of-all-packages-in-cran-from-within-r
-url <- "http://cran.rstudio.com/web/packages/packages.rds"
-db <- readRDS(url(url)) %>% 
-  as.data.frame()
-
-
-## -----------------------------------------------------------------------------
-nanalysis <- db %>% 
-  filter(str_detect(tolower(Title), "analysis")) %>% 
-  nrow()
-
-ndesign <- db %>% 
-  filter(str_detect(tolower(Title), "design")) %>% 
-  nrow()
-
-
-## ----doe-cran-----------------------------------------------------------------
-dat_DoE <- read_html("https://cran.r-project.org/web/views/ExperimentalDesign.html")
-date_download <- Sys.Date()
-cran_names <- available.packages(contriburl = "http://cran.rstudio.com/") %>% 
-  rownames() %>% 
-  unique() # it should be unique
-doe_pkgs <- dat_DoE %>% 
[…]

### 2022-06-03 paper.Rmd
added more in the EDA section -- separated some to supp materials

@@
     Re-running an experiment is generally costly and in some cases impossible due to limited resources, so the design of an experiment plays a critical role in increasing the quality of experimental data.
-In this article we describe the current state of the R-packages for the design of experiments through textual analysis and assessment of download trends.
+In this article we describe the current state of the R-packages for the design of experiments through exploratory data analysis of package downloads, package metadata, and comparison of characteristics with other topics.
 We discuss also the software design of widely utilised R packages in the field of experimental design and conclude with discussion of some future prospects for the field.
@@
 output: 
-  bookdown::html_document2: 
-    theme: paper
+  #bookdown::html_document2: 
+  #  theme: paper
   # below is to have a JSS paper style template
@@
 }
-
 ```
@@
 This dichotomisation, to a great extent, is seen in the [CRAN task views](https://cran.r-project.org/web/views/) (a volunteered maintained list of R-packages by topic) where R-packages in experimental design are in `r ctv("ExperimentalDesign")` and R-packages in survey designs are in `r ctv("OfficialStatistics")`.
-Experimental designs exclusively center on the collection of experimental data where the subject and experimental factors tend to be in the control of the experimenter.
+Experimental designs exclusively center on the collection of experimental data where many of the experimental factors tend to be in the control of the experimenter.
 A subset of experimental designs are segregated into `r ctv("ClinicalTrials")`, where the focus is on clinical trials with primary interest in sample size calculations.
-This paper focuses on the packages in `r ctv("ExperimentalDesign")`, henceforth referred to as "**DoE packages**".
+This paper focuses on the packages in `r ctv("ExperimentalDesign")`, henceforth referred to as "DoE packages".
 
 In the CRAN task view of `r ctv("ExperimentalDesign")`, there are `r npkgs` R packages for experimental design and analysis of data from experiments.
-The sheer quantity and variation in the output experimental design in the R-packages are arguably unmatched with any other programming languages, e.g.
-in Python [@python], only a handful of libraries that generate design of experiment exist (namely `pyDOE`, `pyDOE2`, `dexpy`, `experimenter` and `GPdoemd`) with limited outputs.
-Thus, the study of DoE packages is also revealing into the current status of the field of experimental design.
+The sheer quantity and variation of experimental designs in the R-packages are arguably unmatched with any other programming languages, e.g.
+in Python [@python], only a handful of libraries that generate design of experiment exist (namely `pyDOE`, `pyDOE2`, `dexpy`, `experimenter` and `GPdoemd`) with limited type of designs.
+Thus, the study of DoE packages can be revealing into the current status of the field of experimental design.
 
 The paper is organised as follows.
-Section `r ref("data")` briefly describes the data source used for the analysis in Section \@ref(eda); Section \@ref(eda) presents some insights into the state of the current DoE packages; Section \@ref(design) discuss the software design of DoE packages, and we conclude with a discussion in Section \@ref(discussion) of future prospects.
+Section \@ref(data) briefly describes the data source used for the analysis in Section \@ref(eda); Section \@ref(eda) presents some insights into the state of the current DoE packages by exploratory data analysis of package download data, text descriptions and comparisons with other CRAN task views; Section \@ref(design) discuss the interface designs of DoE packages, and we conclude with a discussion in Section \@ref(discussion) of future prospects.
 
@@
 
-All CRAN packages have a title, description, package connections (suggests, dependency and imports) and more in the DESCRIPTION file.
+All CRAN packages have a title, description, package connections (suggests, dependency and imports) and other meta-information in the DESCRIPTION file.
 We use the text data from the title and description (as accessed in `r as.Date(file.info("_targets/objects/ngram_download_data")$mtime)`) in Section \@ref(topics).
@@
 While in absolute terms the Gini index is high for `r ctv("ExperimentalDesign")`, the inequality is not as severe as other CRAN task views as shown in Figure \@ref(fig:fig-gini-all-ctvs).
-We can see in Figure \@ref(fig:fig-gini-all-ctvs) that the Gini index is generally increasing for DoE packages (as is generally the case for other CRAN task views although this is hard to see from this plot).
+We can see in Figure \@ref(fig:fig-gini-all-ctvs) that the Gini index is generally increasing for DoE packages (as is generally the case for other CRAN task views as shown in Figure S1 in the Supplementary Materials).
 
@@
 
+Figure \@ref(fig:established-packages) shows a moderate negative correlation between the first release date and the (log of) total download counts of DoE packages in any given year from 2013 to 2021.
+This suggest that in general, a package released earlier are more likely to be used today (possibly for legacy reasons or the general inertia for adoption of new packages).
+We can also see in Figure \@ref(fig:established-packages) the most downloaded packages were generally released in 2004 to 2010.
+
 ```{r established-packages, fig.height = 5, fig.cap = "The above figure shows the total download (in log scale) of a package in the corresponding year against the first release date of the package.
-The blue line corresponds to the fit of a simple linear regression model.
-In any given year, "}
+The blue line corresponds to the least squares fit of a simple linear regression model.
+The label in the right-hand upper corner shows the sample correlation coefficient between the first release date and the log, with base 10, of the total download count.
+The high-leverage point on the far left belongs to `conf.design`, authored by one of the earlier contributors to R."}
 tar_read(cran_rank_yearly) %>%
@@
   geom_point(aes(color = nupdates)) + 
+  geom_label(data = ~ .x %>% 
+              group_by(year) %>% 
+              summarise(R = cor(as.numeric(first), log10(total/1000)) %>% 
+                          scales::comma(0.01) %>% 
+                          paste0("r = ", .),
+                        y = quantile(total/1000, 0.99)),
+            aes(label = R, y = y, x = as.Date("2014-11-01")),
+            size = 3) +
   facet_wrap(~year, scale = "free_y") +
   geom_smooth(se = FALSE, 
-              method = "lm", 
+              method = lm, 
               formula = y ~ x) +
@@
        y = "Total download (1000s)",
-       color = "# of updates")
+       color = "# of updates") +
+  theme(legend.position = "bottom")
 ```
 
-* The lack of change in the top ranking packages are indicative that **_there has been a lack of innovation in the field of experimental design_**.
+The lack of change in the top ranking packages are indicative that there has been generally little need for new or innovative experimental design needs for the mass.
+We do however also see that the the top downloaded packages generally have more updates so it is possible that the packages have improved or broadened the scope of its usage.
 
-## Computer experiments are on the rise {#topics}
+## Computer experiments and optimal designs are of interest {#topics}
 
-```{r wordcloud-over-time, fig.width = 5.5, fig.height = 5, fig.cap = "The above figures shows the word cloud of bigrams from the title and descriptions of the CRAN packages.
-The color shows how often the bigram appears across the DoE packages and the size are relative to the total download count of the packages that contain the bigram."}
+Figure \@ref(fig:wordcloud-over-time) shows some common purposes of DoE packages, based on bigrams in the package title and description.
+A bigram is only shown as the frequency of single word was not insightful and there were not many trigrams common across packages.
+Not surprisingly, the bigram "experimental design" was the most common.
+More interestingly, "optimal design" appeared across different packages (indicated by the size of the word in Figure \@ref(fig:wordcloud-over-time)), and the bigrams "latin hypercube" and "computer experiment" used across a few packages that are downloaded frequently (indicated by the color of the word in Figure \@ref(fig:wordcloud-over-time)).
+Although there exists a separate `r ctv("ClinicalTrials")` task view, the DoE packages clearly include some packages that are of interest to clinical trials as shown by the size of the bigram "clinical trial" in Figure \@ref(fig:wordcloud-over-time).
+
+```{r wordcloud-over-time, fig.width = 5.5, fig.height = 5, fig.cap = "The above figure shows the word cloud of bigrams from the title and descriptions of the CRAN packages.
+The size shows how often the bigram appears across the DoE packages and the color are relative to the total download count in 2021 of the packages that contain the bigram."}
 tar_read(plot_wordcloud2) + 
-  scale_size_area(max_size = 8) +
-  scale_color_continuous(breaks = c(2, 4, 6, 8, 10, 12)) +
+  scale_size_area(max_size = 5) +
   theme(legend.text = element_text(size = 10),
-        legend.title = element_text(size = 12))
+        legend.title = element_text(size = 12)) +
+  labs(color = "Total downloads (%)")
 ```
@@
 
-```{r ctv-summ-table}
-tar_read(ctv_summary_table) %>% 
-  knitr::kable(col.names = c("Name", "Topic", "# of packages", "Total # of contributors" , "Average # of contributors", "Intra-connectivity (%)"), 
-               caption = "A summary table for the CRAN task view that shows in order: the name of the task view, the full topic name, the total of packages, the total number of contributors, the average number of contributors, and the intra-connectivity.
-The intra-connectivity measures the percentage of packages that depends, suggest or imports at least one other package within the same task view.
-A low intra-connectivity suggests that development within the topic mostly occur in silos whilst high  intra-connectivity suggests that there are more interactions within the topic.",
-               digits = c(0, 0, 0, 1, 2, 0),
-               booktabs = TRUE ) %>%
- kableExtra::kable_styling(latex_options = c("striped", "scale_down")) %>% 
- kableExtra::column_spec(2, width = "12em") %>% 
- kableExtra::column_spec(3:6, width = "5em")
+Figure \@ref(fig:ctv-summ-plot) shows that the `r ctv("ExperimentalDesign")` have the lowest average number of contributors among all CRAN task views.
+In addition, we can also see in Figure \@ref(fig:ctv-summ-plot)  that the `r ctv("ExperimentalDesign")` task view has one of the least intra-connectivity (the percentage of packages that make use of other packages within the same task view).
+The full connection of DoE packages with other DoE packages is shown in Figure \@ref(fig:plot-doe-network).
+These observations suggest that the field of experimental design is one of the least collaborative field and package development generally occur in silos.
 
-```
-
-```{r ctv-summ-plot, fig.width = 5.5, fig.height = 6, fig.cap = "The figure above is a visualisation of Table 1.
-The task views in the bottom-left corner are topics that are more indicative of contributors working in silos."}
+```{r ctv-summ-plot, fig.width = 5.5, fig.height = 6, fig.cap = "The above figure is a scatterplot of the intra-connectivity (the percentage of packages that depends, suggest or imports at least one other package within the same task view) and the average number of contributors for each CRAN task view.
+A low intra-connectivity suggests that development within the topic mostly occur in silos whilst high  intra-connectivity suggests that there are more interactions within the topic.
+The color shows the number of packages, the size of the point corresponds to the total number of contributors, and the text labels show the CRAN task view name.
+The label of ExperimentalDesign task view is colored red.
+The task views in the bottom-left corner are topics that are more indicative of contributors working in silos.
+The actual numerical values are show in Table S1 in the Supplementary Material."}
 tar_read(ctv_summary_table) %>% 
@@
 
+In this section, we present some examples the user interface of widely used specialised designs, namely, computer experiments, factorial experiments, response surface designs and sequential designs.
+
 Computer experiments, as discussed in Section \@ref(topics), which generally involve space filling designs have been on the rise. The exemplar for this type of designs are `lhs` and `DiceDesign`.
@@
 
-`agricolae` [@agricolae]
+The `agricolae` package [@agricolae] was the most downloaded DoE packages in 2013 to 2018 (Figure \@ref(fig:rank-over-time)) and possibly prior to 2013 (data unavailable).
+This package is the prime example of constructing designs based on menu functions, e.g.
+`design.crd()`, `design.rcbd()` and `design.split()` construct a completely randomised design, randomised complete block design and split-plot design, respectively.
+Users typically supply the treatment labels and the number of replications as argument to these functions.
 

### 2022-06-03 supp.Rmd
added more in the EDA section -- separated some to supp materials

---
title: Supplmentary material for "Current status and prospects of R-packages for the design of experiments"
author:
  - Emi Tanaka
  - Dewi Amaliah 
bibliography: paper.bib
keywords: ["experimental design", "CRAN task view", "user interface"]
output: 
  bookdown::pdf_document2
header-includes:
  - \renewcommand{\figurename}{Figure S}
  - \makeatletter
  - \def\fnum@figure{\figurename\thefigure}
  - \makeatother
---

```{css, echo = FALSE}
h1, h2, h3, div > * {
 font-family: "Helvetica Neue",Helvetica,Arial,sans-serif;
}

```

```{r setup, cache = FALSE, include = FALSE}
library(targets)
library(glue)
library(tidyverse)
library(patchwork)
library(tidygraph)
library(ggraph)
source("R/plot.R")
knitr::opts_chunk$set(echo = FALSE, 
                      cache = TRUE,
                      cache.path = "cache/",
                      fig.align = 'center', 
                      fig.pos = 'htbp', 
                      fig.path = "figures/",
                      fig.width = 6,
                      fig.height = 6.5,
                      message = FALSE,
                      warning = FALSE)

ref <- function(x) {
  ifelse(knitr::is_html_output(), glue("\\@ref({x})"), glue("\\ref{[x]}", .open = "[", .close = "]"))
}

tar_load(npkgs)
tar_load(gini_yearly)
tar_load(gini_yearly_ctv)
```

```{r fig-gini-all-ctvs, fig.height = 8, fig.cap = "The points show the Gini index of the download counts by year facetted by CRAN task view with the color showing the number of packages. The grey line shows the distribution of the Gini index across years for all other CRAN task views. The facets are ordered by increasing value of the Gini index in 2021."}
gini_yearly_ctv %>% 
  ungroup() %>%  
  arrange(ctv, year) %>% 
  mutate(ctv = fct_reorder(ctv, gini, last),
         year = factor(year),
         gini = 100 * gini) %>%  
  ggplot(aes(year, gini)) + 
  facet_wrap(~ctv, ncol = 4, scale = "free_y") + 
  geom_line(aes(group = ctv)) + 
  geom_point(aes(color = n))  +
  plot_setup() +
  labs(x = "Year", y = "Gini index (%)", color = "# of packages") +
  guides(color = guide_colorbar(direction = "horizontal")) +
  theme(legend.position = c(0.5, 0), 
        legend.justification = "left",
        strip.text = element_text(size = 7),
        axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  scale_colour_distiller(palette = "YlOrRd", direction = 1) 
```

```{r ctv-summ-table}
tar_read(ctv_summary_table) %>% 
  knitr::kable(col.names = c("Name", "Topic", "# of packages", "Total # of contributors" , "Average # of contributors", "Intra-connectivity (%)"), 
               caption = "A summary table for the CRAN task view that shows in order: the name of the task view, the full topic name, the total of packages, the total number of contributors, the average number of contributors, and the intra-connectivity. The intra-connectivity measures the percentage of packages that depends, suggest or imports at least one other package within the same task view. A low intra-connectivity suggests that development within the topic mostly occur in silos whilst high  intra-connectivity suggests that there are more interactions within the topic. The row is ordered by the average number of contributors.",
               digits = c(0, 0, 0, 1, 2, 0),
               booktabs = TRUE ) %>%
 kableExtra::kable_styling(latex_options = c("striped", "scale_down")) %>% 
 kableExtra::column_spec(2, width = "12em") %>% 
 kableExtra::column_spec(3:6, width = "5em")

```

### 2022-06-03 R/plot.R

@@ -145,18 +145,17 @@ plot_word_cloud_over_years <- function(data) {
 
 plot_word_cloud <- function(data) {
   data %>% 
-    filter(perc >= 1) %>% 
-    ggplot(aes(label = word, size = perc, color = n)) +
+    filter(perc >= 1, n > 1) %>% 
+    ggplot(aes(label = word, color = perc, size = n)) +
     geom_text_wordcloud(shape = "square",
-                        show.legend = TRUE,
-                        rm_outside = TRUE) +
+                        show.legend = TRUE) + #,
+                        #rm_outside = TRUE) +
     scale_size_area(max_size = 18) +
     guides(size = "none") +
-    colorspace::scale_color_continuous_sequential(breaks = round(seq(2, 13, length.out = 4)),
-                                                  palette = "Blues") +
+    #colorspace::scale_color_continuous_sequential(breaks = round(seq(2, 13, length.out = 4)),
+    #                                              palette = "Blues") +
     theme(legend.position = "bottom",
           text = element_text(size = 18)) +
-    labs(color = "# of packages") +
     guides(color = guide_colorbar(title.position = "top")) 
 }
 

### 2022-06-06 paper/arxiv/arxiv.Rmd
put the arxiv version under paper/

---
authors:
  - name: Emi Tanaka
    thanks: emitanaka.org
    department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: emi.tanaka@monash.edu
  - name: Dewi Amaliah
    department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
header-includes: |
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
bibliography: ../../paper.bib
biblio-style: unsrt
#preamble: >
output: 
  rticles::arxiv_article:
    keep_tex: true
---

```{r, include = FALSE}
paper <- here::here("paper.Rmd")
meta <- rmarkdown::yaml_front_matter(paper)
```

---
title: "`r meta$title`"
abstract: >
  `r meta$abstract`
---

```{r child=paper}
```

### 2022-06-06 paper/arxiv/arxiv.tex
put the arxiv version under paper/

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}

\title{Current status and prospects of R-packages for the design of
experiments}

\author{
    Emi Tanaka
    \thanks{emitanaka.org}
   \\
    Department of Econometrics and Business Statistics \\
    Monash University \\
  Clayton, VIC 3800 \\
  \texttt{\href{mailto:emi.tanaka@monash.edu}{\nolinkurl{emi.tanaka@monash.edu}}} \\
   \And
    Dewi Amaliah
   \\
    Department of Econometrics and Business Statistics \\
    Monash University \\
  Clayton, VIC 3800 \\
  \texttt{} \\
  }

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{colortbl}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}

\begin{document}
\maketitle

\begin{abstract}
Re-running an experiment is generally costly and in some cases
impossible due to limited resources, so the design of an experiment
plays a critical role in increasing the quality of experimental data. In
this article we describe the current state of the R-packages for the
design of experiments through exploratory data analysis of package
downloads, package metadata, and comparison of characteristics with
other topics. We discuss also the software design of widely utilised R
packages in the field of experimental design and conclude with
discussion of some future prospects for the field.
\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The critical role of data collection is well captured in the expression
``garbage in, garbage out'' -- in other words, if the collected data is
rubbish then no analysis, however complex it may be, can make something
out of it. Methods for data collection can be dichotomised by the type
of data collected -- namely, experimental or observational -- or
alternatively, categorised as experimental design (including
quasi-experimental design) or survey design. This dichotomisation, to a
great extent, is seen in the
\href{https://cran.r-project.org/web/views/}{CRAN task views} (a
volunteered maintained list of R-packages by topic) where R-packages in
experimental design are in ctv\{ExperimentalDesign\} and R-packages in
survey designs are in ctv\{OfficialStatistics\}. Experimental designs
exclusively center on the collection of experimental data where many of
the experimental factors tend to be in the control of the experimenter.
A subset of experimental designs are segregated into
ctv\{ClinicalTrials\}, where the focus is on clinical trials with
primary interest in sample size calculations. This paper focuses on the
packages in ctv\{ExperimentalDesign\}, henceforth referred to as ``DoE
packages''.

In the CRAN task view of ctv\{ExperimentalDesign\}, there are 114 R
packages for experimental design and analysis of data from experiments.
The sheer quantity and variation of experimental designs in the
R-packages are arguably unmatched with any other programming languages,
e.g.~in Python (Rossum 1995), only a handful of libraries that generate
design of experiment exist (namely \texttt{pyDOE}, \texttt{pyDOE2},
\texttt{dexpy}, \texttt{experimenter} and \texttt{GPdoemd}) with limited
type of designs. Thus, the study of DoE packages can be revealing into
the current status of the field of experimental design.

The paper is organised as follows. Section @ref(data) briefly describes
the data source used for the analysis in Section @ref(eda); Section
@ref(eda) presents some insights into the state of the current DoE
packages by exploratory data analysis of package download data, text
descriptions and comparisons with other CRAN task views; Section
@ref(design) discuss the interface designs of DoE packages, and we
conclude with a discussion in Section @ref(discussion) of future
prospects.

\hypertarget{data}{%
\section{Data}\label{data}}

To study the DoE packages, we analyse data using three sources of data
as described next.

\hypertarget{rstudio-cran-download-logs}{%
\subsection{RStudio CRAN download
logs}\label{rstudio-cran-download-logs}}

The Comprehensive R Archive Network (CRAN) is a network of servers
located across the world that store mirrored versions of R and
R-packages. The most popular network is the RStudio mirror (the default
server for those that use the RStudio IDE). The RStudio mirror is also
the only server that provides a comprehensive daily download logs of R
and R-packages since October 2012. The summary data can be easily
accessed with the \texttt{cranlogs} package (Csárdi 2019). This paper
uses the data from the beginning of 2013 to end of 2021 (a total of 9
years) for the DoE packages.

\hypertarget{package-descriptions}{%
\subsection{Package descriptions}\label{package-descriptions}}

All CRAN packages have a title, description, package connections
(suggests, dependency and imports) and other meta-information in the
DESCRIPTION file. We use the text data from the title and description
(as accessed in 2022-05-23) in Section @ref(topics).

\hypertarget{cran-task-views}{%
\subsection{CRAN task views}\label{cran-task-views}}

The list of packages in each CRAN task views (as of 2022-05-25) are used
to contrast the characteristics of DoE packages in Section @ref(silo).

\hypertarget{eda}{%
\section{Explorative data analysis}\label{eda}}

All results presented are derived from exploratory data analysis of
observational data; consequently, all interpretations are somewhat
speculative and may not be indicative of the true state of the field of
experimental design. In particular, any analysis over time are
confounded by the fact that the nature of users and package management
have also changed over the years. It should be noted that some DoE
packages may have been archived or removed from the task view over the
years so any cross-sectional analysis presented may not reflect the set
of all DoE packages at that particular time period (although we assume
such incidences are low).

A subset of DoE packages is not primarily about design of experiments
but about the analysis of experimental data. A complete delineation of
these packages is difficult as many have some functions that can aid
decisions or constructions of experimental designs (and any
categorisation is prone to our subjective bias) so we opted not to
remove any DoE packages in the analysis.

\hypertarget{popular}{%
\subsection{Small, but diverse, set of packages are sufficient for most
designs}\label{popular}}

There have been at least 50 DoE packages since 2013 but most of the
downloads are concentrated in just a handful of packages. For example,
Figure @ref(fig:plot-lorenz) shows a Lorenz curve (Lorenz 1905) for the
total package downloads in 2021 of 113 DoE packages (first released
prior to 2021); we can see from Figure @ref(fig:plot-lorenz) that bottom
90\% of DoE packages (in terms of total download count in 2021) only
share about 30\% of total downloads across all DoE packages -- in
another words, 70\% of the total downloads are due to 11 packages (10\%
of the DoE packages).

\begin{figure}[htbp]

{\centering \includegraphics{figures/plot-lorenz-1}

}

\caption{Lorenz curve of the total download count for DoE packages in 2021. The red line corresponds to the line of perfect equality.}\label{fig:plot-lorenz}
\end{figure}

If we consider package downloads as a measure of ``wealth'', then we can
consider using the Gini index (Gini 1921) as a measure of download
inequality across packages. The ratio of the red region over the total
colored regions in Figure @ref(fig:plot-lorenz) corresponds to the Gini
index for 2021. A Gini index of 0\% indicates equality in downloads
across packages while a value of 100\% indicates maximal inequality (all
downloads are due to one package). In Figure @ref(fig:download-share),
we see that the distributions of the downloads each year have a heavy
right tail with the Gini index ranging from 32.5\% to 68.3\% across the
years 2013 to 2021, indicating that there is a high level of inequality
of downloads across packages, particularly with more pronounced
inequality in the last 5 years.

\begin{figure}[htbp]

{\centering \includegraphics{figures/download-share-1}

}

\caption{The distribution of the number of downloads for each DoE package by year. Packages were removed in any year if it was released in that year or later so that each download count is for the full year. The label on the bottom of the plot shows the Gini index for downloads and the number of packages with full year of download count in the corresponding year. In the last 5 years, the Gini index is consistent above 60\% each year indicating that most downloads are due to a relatively small number of packages.}\label{fig:download-share}
\end{figure}

While in absolute terms the Gini index is high for
ctv\{ExperimentalDesign\}, the inequality is not as severe as other CRAN
task views as shown in Figure @ref(fig:fig-gini-all-ctvs). We can see in
Figure @ref(fig:fig-gini-all-ctvs) that the Gini index is generally
increasing for DoE packages (as is generally the case for other CRAN
task views as shown in Figure S1 in the Supplementary Materials).

\begin{figure}[htbp]

{\centering \includegraphics{figures/fig-gini-all-ctvs-1}

}

\caption{The points show the Gini index of the download counts by year facetted by CRAN task view with the color showing the number of packages. The grey line shows the distribution of the Gini index across years for all other CRAN task views. The facets are ordered by increasing value of the Gini index in 2021.}\label{fig:fig-gini-all-ctvs}
\end{figure}

From Figures @ref(fig:download-share) and @ref(fig:fig-gini-all-ctvs),
we could draw some indicative states of the field of experimental
designs with possible counterfactual interpretations:

\begin{itemize}
\tightlist
\item
  The increase in the number of packages that are not highly downloaded
  may mean that \textbf{\emph{there are more packages to construct niche
  experimental designs}}. Some examples of these packages include
  \texttt{qtlDesign}, \texttt{PwrGSD} and \texttt{Crossover} made for
  QTL experiments, group sequential designs and crossover trials,
  respectively. These packages would naturally have a smaller number of
  potential users. Counterfactual to this, the increase could be due to
  other external factors, such as an increase in a number of skilled
  contributors, a change in CRAN policy or management to add packages
  (either to CRAN and/or task view), and/or that new packages are still
  yet to amass users. While there is an argument that low download
  counts are due to the low utility and/or quality of the packages,
  packages in CRAN task views are selected by expert maintainers; thus
  we can reasonably assume that any package listed in CRAN task view are
  of a decent utility and quality.
\item
  If the downloads are reflective of the experimental designs used in
  practice, \textbf{\emph{small set of packages appear to be sufficient
  for most to construct the full set of designs of experiments needed in
  practice}}. Packages of course evolve and the top downloaded packages
  have had regular updates that may have broadened its scope.
\item
  While small subset of DoE packages are most frequently used, none of
  these packages are dominant in the field (judging from comparisons
  with other CRAN task views). This suggests that there are
  \textbf{\emph{diverse approaches to designing experiments}}. This
  observation doesn't take into account other approaches to generating
  experimental designs, like the proprietary GUI software, CycDesignN,
  which may be widely used.
\end{itemize}

\hypertarget{ranking}{%
\subsection{There is a lack of adaptation of new or innovative
designs}\label{ranking}}

We can see in Figure @ref(fig:rank-over-time) that most of the top 10
ranking packages have been in the top 10 for the last 9 years with
\texttt{lhs} steadily climbing up the ranks in the last few years.

\begin{figure}[htbp]

{\centering \includegraphics{figures/rank-over-time-1}

}

\caption{The plot shows the rank of top 10 downloaded packages by year. }\label{fig:rank-over-time}
\end{figure}

Figure @ref(fig:established-packages) shows a moderate negative
correlation between the first release date and the (log of) total
download counts of DoE packages in any given year from 2013 to 2021.
This suggest that in general, a package released earlier are more likely
to be used today (possibly for legacy reasons or the general inertia for
adoption of new packages). We can also see in Figure
@ref(fig:established-packages) the most downloaded packages were
generally released in 2004 to 2010.

\begin{figure}[htbp]

{\centering \includegraphics{figures/established-packages-1}

}

\caption{The above figure shows the total download (in log scale) of a package in the corresponding year against the first release date of the package. The blue line corresponds to the least squares fit of a simple linear regression model. The label in the right-hand upper corner shows the sample correlation coefficient between the first release date and the log, with base 10, of the total download count. The high-leverage point on the far left belongs to `conf.design`, authored by one of the earlier contributors to R.}\label{fig:established-packages}
\end{figure}

The lack of change in the top ranking packages are indicative that there
has been generally little need for new or innovative experimental design
needs for the mass. We do however also see that the the top downloaded
packages generally have more updates so it is possible that the packages
have improved or broadened the scope of its usage.

\hypertarget{topics}{%
\subsection{Computer experiments and optimal designs are of
interest}\label{topics}}

Figure @ref(fig:wordcloud-over-time) shows some common purposes of DoE
packages, based on bigrams in the package title and description. A
bigram is only shown as the frequency of single word was not insightful
and there were not many trigrams common across packages. Not
surprisingly, the bigram ``experimental design'' was the most common.
More interestingly, ``optimal design'' appeared across different
packages (indicated by the size of the word in Figure
@ref(fig:wordcloud-over-time)), and the bigrams ``latin hypercube'' and
``computer experiment'' used across a few packages that are downloaded
frequently (indicated by the color of the word in Figure
@ref(fig:wordcloud-over-time)). Although there exists a separate
ctv\{ClinicalTrials\} task view, the DoE packages clearly include some
packages that are of interest to clinical trials as shown by the size of
the bigram ``clinical trial'' in Figure @ref(fig:wordcloud-over-time).

\begin{figure}[htbp]

{\centering \includegraphics{figures/wordcloud-over-time-1}

}

\caption{The above figure shows the word cloud of bigrams from the title and descriptions of the CRAN packages. The size shows how often the bigram appears across the DoE packages and the color are relative to the total download count in 2021 of the packages that contain the bigram.}\label{fig:wordcloud-over-time}
\end{figure}

\hypertarget{silo}{%
\subsection{Development of experimental designs occur in
silos}\label{silo}}

Figure @ref(fig:ctv-summ-plot) shows that the ctv\{ExperimentalDesign\}
have the lowest average number of contributors among all CRAN task
views. In addition, we can also see in Figure @ref(fig:ctv-summ-plot)
that the ctv\{ExperimentalDesign\} task view has one of the least
intra-connectivity (the percentage of packages that make use of other
packages within the same task view). The full connection of DoE packages
with other DoE packages is shown in Figure @ref(fig:plot-doe-network).
These observations suggest that the field of experimental design is one
of the least collaborative field and package development generally occur
in silos.

\begin{figure}[htbp]

{\centering \includegraphics{figures/ctv-summ-plot-1}

}

\caption{The above figure is a scatterplot of the intra-connectivity (the percentage of packages that depends, suggest or imports at least one other package within the same task view) and the average number of contributors for each CRAN task view. A low intra-connectivity suggests that development within the topic mostly occur in silos whilst high  intra-connectivity suggests that there are more interactions within the topic. The color shows the number of packages, the size of the point corresponds to the total number of contributors, and the text labels show the CRAN task view name.  The label of ExperimentalDesign task view is colored red. The task views in the bottom-left corner are topics that are more indicative of contributors working in silos. The actual numerical values are show in Table S1 in the Supplementary Material.}\label{fig:ctv-summ-plot}
\end{figure}

\begin{figure}[htbp]

{\centering \includegraphics{figures/plot-doe-network-1}

}

\caption{Package connections (depends, suggests and imports) within the DoE packages. DoE packages that do not depend, suggest or import another DoE package is not shown.}\label{fig:plot-doe-network}
\end{figure}

\hypertarget{design}{%
\section{Interface design}\label{design}}

In software design, there are two interface designs to consider: user
interface (UI) and application programming interface (API). UI is
concerned with the usage of the software by the user, while the API is
concerned with how different programs interact and is predominately of
the interest to the developer. In this section, we discuss these
interface designs broadly, with examples illustrated from the top
downloaded packages (as shown in Figure @ref(fig:rank-over-time)). We
exclude \texttt{ez} and \texttt{DiceKriging} from the discussion as the
former is predominately visualisation of experimental data and latter is
about the analysis of computer experiments in addition to belonging to
the same suite of packages as \texttt{DiceDesign}.

\hypertarget{the-case-of-specialised-designs}{%
\subsection{The case of specialised
designs}\label{the-case-of-specialised-designs}}

In this section, we present some examples the user interface of widely
used specialised designs, namely, computer experiments, factorial
experiments, response surface designs and sequential designs.

Computer experiments, as discussed in Section @ref(topics), which
generally involve space filling designs have been on the rise. The
exemplar for this type of designs are \texttt{lhs} and
\texttt{DiceDesign}. For the \texttt{lhs} package, functions like
\texttt{randomLHS()}, \texttt{optimumLHS()}, and \texttt{maximinLHS()}

Factorial experiments offer a challenge in the construction and
allocation of the treatment factors. This is reflected in a number of
packages that specifically addresses this challenge. These include
\texttt{DoE.base} (Grömping 2018), \texttt{FrF2} (Grömping 2014) for
fractional 2-level factorial designs, \texttt{conf.design} (Venables
2013), \texttt{FrF2.catlg128} (Grömping 2022) and \texttt{BHH2} (Barrios
2016).

Response surface design, as illustrated by \texttt{rsm} (Lenth 2009), is
a type of experiment that involve variables with continuous measures.

Finally, another specialised design is sequential design (also called
adaptive sampling), which is best represented by \texttt{tgp} (Gramacy
and Taddy 2010) -- these require prior information, which are used to
inform the next experimental design. Follow-up experiments, which can
also be classified as a sequential design, is implemented by
\texttt{BsMD} (Barrios 2020).

\hypertarget{the-case-of-menu-functions}{%
\subsection{The case of menu
functions}\label{the-case-of-menu-functions}}

The \texttt{agricolae} package (de Mendiburu 2021) was the most
downloaded DoE packages in 2013 to 2018 (Figure
@ref(fig:rank-over-time)) and possibly prior to 2013 (data unavailable).
This package is the prime example of constructing designs based on menu
functions, e.g.~\texttt{design.crd()}, \texttt{design.rcbd()} and
\texttt{design.split()} construct a completely randomised design,
randomised complete block design and split-plot design, respectively.
Users typically supply the treatment labels and the number of
replications as argument to these functions.

\hypertarget{the-case-of-general-designs}{%
\subsection{The case of general
designs}\label{the-case-of-general-designs}}

\texttt{AlgDesign} (Wheeler 2022)

\texttt{DoE.wrapper} (Grömping 2020)

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Volunteer maintained -- some experimental design packages may not be
included in the CRAN task view.

\hypertarget{acknowledgement}{%
\section{Acknowledgement}\label{acknowledgement}}

This paper uses the \texttt{targets} framework (Landau 2021) for
reproducibility, \texttt{knitr} (Xie 2015) and \texttt{rmarkdown} (Xie,
Allaire, and Grolemund 2018) for creating reproducible documents,
\texttt{ggplot2} (Wickham 2016) for visualisation and
\texttt{kableExtra} (Zhu 2021) for customising the table. All code to
reproduce this paper is found at
\url{https://github.com/emitanaka/paper-DoE-review}.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-BHH2}{}}%
Barrios, Ernesto. 2016. \emph{Bhh2: Useful Functions for Box, Hunter and
Hunter II}. \url{https://CRAN.R-project.org/package=BHH2}.

\leavevmode\vadjust pre{\hypertarget{ref-BsMD}{}}%
---------. 2020. \emph{BsMD: Bayes Screening and Model Discrimination}.
\url{https://CRAN.R-project.org/package=BsMD}.

\leavevmode\vadjust pre{\hypertarget{ref-cranlogs}{}}%
Csárdi, Gábor. 2019. \emph{Cranlogs: Download Logs from the 'RStudio'
'CRAN' Mirror}. \url{https://CRAN.R-project.org/package=cranlogs}.

\leavevmode\vadjust pre{\hypertarget{ref-agricolae}{}}%
de Mendiburu, Felipe. 2021. \emph{Agricolae: Statistical Procedures for
Agricultural Research}.
\url{https://CRAN.R-project.org/package=agricolae}.

\leavevmode\vadjust pre{\hypertarget{ref-Gini1921-mf}{}}%
Gini, Corrado. 1921. {``Measurement of Inequality of Incomes.''}
\emph{The Economic Journal} 31 (121): 124--26.
\url{https://doi.org/10.2307/2223319}.

\leavevmode\vadjust pre{\hypertarget{ref-tgp}{}}%
Gramacy, Robert B., and Matthew Taddy. 2010. {``Categorical Inputs,
Sensitivity Analysis, Optimization and Importance Tempering with {tgp}
Version 2, an {R} Package for Treed Gaussian Process Models.''}
\emph{Journal of Statistical Software} 33 (6): 1--48.
\url{https://doi.org/10.18637/jss.v033.i06}.

\leavevmode\vadjust pre{\hypertarget{ref-FrF2}{}}%
Grömping, Ulrike. 2014. {``{R} Package {FrF2} for Creating and Analyzing
Fractional Factorial 2-Level Designs.''} \emph{Journal of Statistical
Software} 56 (1): 1--56. \url{https://www.jstatsoft.org/v56/i01/}.

\leavevmode\vadjust pre{\hypertarget{ref-DoE.base}{}}%
---------. 2018. {``{R} Package {DoE.base} for Factorial Experiments.''}
\emph{Journal of Statistical Software} 85 (5): 1--41.
\url{https://doi.org/10.18637/jss.v085.i05}.

\leavevmode\vadjust pre{\hypertarget{ref-DoE.wrapper}{}}%
---------. 2020. \emph{DoE.wrapper: Wrapper Package for Design of
Experiments Functionality}.
\url{https://CRAN.R-project.org/package=DoE.wrapper}.

\leavevmode\vadjust pre{\hypertarget{ref-FrF2.catlg128}{}}%
---------. 2022. \emph{FrF2.catlg128: Catalogues of Resolution IV 128
Run 2-Level Fractional Factorials up to 33 Factors That Do Have 5-Letter
Words}. \url{https://CRAN.R-project.org/package=FrF2.catlg128}.

\leavevmode\vadjust pre{\hypertarget{ref-targets}{}}%
Landau, William Michael. 2021. {``The Targets r Package: A Dynamic
Make-Like Function-Oriented Pipeline Toolkit for Reproducibility and
High-Performance Computing.''} \emph{Journal of Open Source Software} 6
(57): 2959. \url{https://doi.org/10.21105/joss.02959}.

\leavevmode\vadjust pre{\hypertarget{ref-rsm}{}}%
Lenth, Russell V. 2009. {``Response-Surface Methods in {R}, Using
{rsm}.''} \emph{Journal of Statistical Software} 32 (7): 1--17.
\url{https://doi.org/10.18637/jss.v032.i07}.

\leavevmode\vadjust pre{\hypertarget{ref-Lorenz1905-tc}{}}%
Lorenz, M O. 1905. {``Methods of Measuring the Concentration of
Wealth.''} \emph{Publications of the American Statistical Association} 9
(70): 209--19. \url{https://doi.org/10.2307/2276207}.

\leavevmode\vadjust pre{\hypertarget{ref-python}{}}%
Rossum, G. van. 1995. {``Python Tutorial.''} CS-R9526. Amsterdam:
Centrum voor Wiskunde en Informatica (CWI).

\leavevmode\vadjust pre{\hypertarget{ref-conf.design}{}}%
Venables, Bill. 2013. \emph{Conf.design: Construction of Factorial
Designs}. \url{https://CRAN.R-project.org/package=conf.design}.

\leavevmode\vadjust pre{\hypertarget{ref-AlgDesign}{}}%
Wheeler, Bob. 2022. \emph{AlgDesign: Algorithmic Experimental Design}.
\url{https://CRAN.R-project.org/package=AlgDesign}.

\leavevmode\vadjust pre{\hypertarget{ref-ggplot2}{}}%
Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data
Analysis}. Springer-Verlag New York.
\url{https://ggplot2.tidyverse.org}.

\leavevmode\vadjust pre{\hypertarget{ref-knitr}{}}%
Xie, Yihui. 2015. \emph{Dynamic Documents with {R} and Knitr}. 2nd ed.
Boca Raton, Florida: Chapman; Hall/CRC. \url{https://yihui.org/knitr/}.

\leavevmode\vadjust pre{\hypertarget{ref-rmarkdown}{}}%
Xie, Yihui, J. J. Allaire, and Garrett Grolemund. 2018. \emph{R
Markdown: The Definitive Guide}. Boca Raton, Florida: Chapman; Hall/CRC.
\url{https://bookdown.org/yihui/rmarkdown}.

\leavevmode\vadjust pre{\hypertarget{ref-kableExtra}{}}%
Zhu, Hao. 2021. \emph{kableExtra: Construct Complex Table with 'Kable'
and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.

\end{CSLReferences}

\bibliographystyle{unsrt}
\bibliography{../../paper.bib}

\end{document}

### 2022-06-07 paper.Rmd
section 1-3 draft

@@
 ---
-title: Current status and prospects of R-packages for the design of experiments
+title: Current state and prospects of R-packages for the design of experiments
 author:
@@
     Re-running an experiment is generally costly and in some cases impossible due to limited resources, so the design of an experiment plays a critical role in increasing the quality of experimental data.
-In this article we describe the current state of the R-packages for the design of experiments through exploratory data analysis of package downloads, package metadata, and comparison of characteristics with other topics.
+In this article we describe the current state of the R-packages for the design of experiments through an exploratory data analysis of package downloads, package metadata, and comparison of characteristics with other topics.
+We observe that experimental designs in practice appear to be sufficiently manufactured by a small number of packages and the development of experimental designs occur in silos.
 We discuss also the software design of widely utilised R packages in the field of experimental design and conclude with discussion of some future prospects for the field.
@@
 knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'index.html')) })
-keywords: ["experimental design", "CRAN task view", "user interface"]
+keywords: ["experimental design", "CRAN task view", "user interface", "application programming interface", "software design"]
 output: 
-  #bookdown::html_document2: 
-  #  theme: paper
+  bookdown::html_document2: 
+    theme: paper
   # below is to have a JSS paper style template
@@
 The critical role of data collection is well captured in the expression "garbage in, garbage out" -- in other words, if the collected data is rubbish then no analysis, however complex it may be, can make something out of it.
+A carefully crafted data collection scheme is therefore critical to optimise the information from data.
+Of all the data collection schemes, the field of experimental designs is specifically devoted to planning an experiment largely based on the founding principles by @Fisher1935-qc or an optimisation framework like those described in @Pukelsheim2006-sv.
+These experimental designs are often constructed with the aid of a statistical software, such as R [@R], so software usage can aid in understanding the practice of experimental design.
+Given the large gain that can be made from collecting higher quality experimental data, we benefit from understanding the current state of the experimental designs in practice.
+
+<!-- Unlike many other data collection schemes, experiments provides the greatest control and consequently, a greater opportunity to statistically optimise the information drawn from the experimental data.
+-->
+
 Methods for data collection can be dichotomised by the type of data collected -- namely, experimental or observational -- or alternatively, categorised as experimental design (including quasi-experimental design) or survey design.
-This dichotomisation, to a great extent, is seen in the [CRAN task views](https://cran.r-project.org/web/views/) (a volunteered maintained list of R-packages by topic) where R-packages in experimental design are in `r ctv("ExperimentalDesign")` and R-packages in survey designs are in `r ctv("OfficialStatistics")`.
-Experimental designs exclusively center on the collection of experimental data where many of the experimental factors tend to be in the control of the experimenter.
-A subset of experimental designs are segregated into `r ctv("ClinicalTrials")`, where the focus is on clinical trials with primary interest in sample size calculations.
-This paper focuses on the packages in `r ctv("ExperimentalDesign")`, henceforth referred to as "DoE packages".
+This dichotomisation, to a great extent, is seen in the [Comprehensive R Archive Network (CRAN) task views](https://cran.r-project.org/web/views/) (a volunteer maintained list of R-packages by topic; see Table S1 in the Supplementary Materials for the full list of available topics) where R-packages in experimental design are in `r ctv("ExperimentalDesign")` task view and R-packages in survey designs are in `r ctv("OfficialStatistics")` task view.
+A subset of experimental designs are segregated into `r ctv("ClinicalTrials")` task view, where the focus is on clinical trials with primary interest in sample size calculations.
+This paper focuses on the packages in `r ctv("ExperimentalDesign")` task view, henceforth referred to as "DoE packages".
 
-In the CRAN task view of `r ctv("ExperimentalDesign")`, there are `r npkgs` R packages for experimental design and analysis of data from experiments.
+In `r ctv("ExperimentalDesign")` task view, there are `r npkgs` R packages for experimental design and analysis of data from experiments.
 The sheer quantity and variation of experimental designs in the R-packages are arguably unmatched with any other programming languages, e.g.
 in Python [@python], only a handful of libraries that generate design of experiment exist (namely `pyDOE`, `pyDOE2`, `dexpy`, `experimenter` and `GPdoemd`) with limited type of designs.
-Thus, the study of DoE packages can be revealing into the current status of the field of experimental design.
+Thus, the study of DoE packages, based on quantitative and qualitative data, can give us an objective view to the state of the current experimental designs in practice.
+
+A utility of software can also be described in terms of its design, such as the user interface (UI) and application programming interface (API).
+UI is concerned with the interaction of the software by the user, while the API is concerned with how different programs interact and is predominately of the interest to the developer.
+These interface designs is an abstraction to specifying the desired experimental design and its choices enable how a user expresses the specification of the experimental design.
 
@@
 
-```{r fig-gini-all-ctvs, fig.height = 8, fig.cap = "The points show the Gini index of the download counts by year facetted by CRAN task view with the color showing the number of packages.
+```{r fig-gini-all-ctvs, fig.height = 4, fig.width = 4, fig.cap = "The points show the Gini index of the download counts by year for ExperimentalDesign CRAN task view with the color showing the number of packages.
 The grey line shows the distribution of the Gini index across years for all other CRAN task views.
-The facets are ordered by increasing value of the Gini index in 2021."}
+See Figure S2 in the Supplementary Material for the line graph for the Gini index across years by each CRAN task view."}
 gini_yearly_ctv %>% 
@@
          gini = 100 * gini) %>%  
-  ggplot(aes(year, gini)) + 
-  geom_line(data = ~rename(.x, ctv2 = ctv), 
-            aes(group = ctv2), color = "grey") +
-  facet_wrap(~ctv, ncol = 4) + 
-  geom_line(aes(group = ctv)) + 
-  geom_point(aes(color = n))  +
+  ggplot(aes(year, gini, group = ctv)) + 
+  geom_line(color = "grey") +
+  geom_line(data = ~filter(.x, ctv=="ExperimentalDesign")) + 
+  geom_point(data = ~filter(.x, ctv=="ExperimentalDesign"),
+             aes(color = n))  +
   plot_setup() +
-  labs(x = "Year", y = "Gini index (%)", color = "# of packages") +
+  labs(x = "Year", y = "Gini index (%)", color = "# of packages",
+       title = "ExperimentalDesign CRAN task view") +
   guides(color = guide_colorbar(direction = "horizontal")) +
-  theme(legend.position = c(0.5, 0), 
-        legend.justification = "left",
-        strip.text = element_text(size = 7),
-        axis.text.x = element_text(angle = 90, vjust = 0.5)) +
+  theme(legend.position = "bottom") +
   scale_colour_distiller(palette = "YlOrRd", direction = 1) 
@@
 
-Computer experiments, as discussed in Section \@ref(topics), which generally involve space filling designs have been on the rise. The exemplar for this type of designs are `lhs` and `DiceDesign`.
+Computer experiments, as discussed in Section \@ref(topics), which generally involve space filling designs, have been on the rise.
+The exemplar for this type of designs are `lhs` and `DiceDesign`.
+For the `lhs` package, functions like `randomLHS()`, `optimumLHS()`, and `maximinLHS()`, require users to specify the sample size ($n$) and the number of variables ($p$) then it generates a Latin hypercube sample, based on different optimisation schemes (in this case, random, S optimal and maxmin criteria, respectively; see package documentation for more details).
+Similarly for the `DiceDesign`, there is a comprehensive list of space-filling designs such as `dmaxDesign()`, `lhsDesign()`, `straussDesign()` and `wspDesign()` with input of $n$ and $p$ as before (among extra parameters for some) that implement algorithms that either maximise the entropy, produce a random Latin hypercube sample, based on Strauss process and WSP algorithm, respectively.
+These designs all output a $n \times p$ matrix with values between 0 and 1.
 

### 2022-06-07 R/functions.R

@@ -116,7 +116,9 @@ ngram_data <- function(data, ngram, desc = c("Title", "Description", "Both")) {
            Description = tolower(Description),
            # end of title and beginning of description could form a ngram so avoid
            # this with a dummy word
-           Both = paste(Title, DUMMYWORD, Description)) %>% 
+           Both = paste(Title, DUMMYWORD, Description)) %>%
+    # n_min = 2 -- decided not to do this and do it separately to make 
+    # word separation below easier
     unnest_tokens(word, desc, token = "ngrams", n = ngram) %>% 
     separate(word, words, sep = " ") %>% 
     mutate(across(num_range("word", 1:ngram), singularize2)) %>% 

### 2022-06-08 paper.Rmd
make emphasis on user interface (rather than software design) and expand acknowledgement

@@
     Re-running an experiment is generally costly and in some cases impossible due to limited resources, so the design of an experiment plays a critical role in increasing the quality of experimental data.
-In this article we describe the current state of the R-packages for the design of experiments through an exploratory data analysis of package downloads, package metadata, and comparison of characteristics with other topics.
+In this article we describe the current state of the R-packages for the design of experiments through an exploratory data analysis of package downloads, package metadata, and the comparison of characteristics with other topics.
 We observe that experimental designs in practice appear to be sufficiently manufactured by a small number of packages and the development of experimental designs occur in silos.
-We discuss also the software design of widely utilised R packages in the field of experimental design and conclude with discussion of some future prospects for the field.
+We discuss also the interface designs of widely utilised R packages in the field of experimental design and discuss its future prospects for the field.
 bibliography: paper.bib
@@
 }
+url <- function(x) {
+  ifelse(knitr::is_html_output(), x, glue("\\url{[x]}", .open = "[", .close = "]"))
+}
 
@@
 Methods for data collection can be dichotomised by the type of data collected -- namely, experimental or observational -- or alternatively, categorised as experimental design (including quasi-experimental design) or survey design.
-This dichotomisation, to a great extent, is seen in the [Comprehensive R Archive Network (CRAN) task views](https://cran.r-project.org/web/views/) (a volunteer maintained list of R-packages by topic; see Table S1 in the Supplementary Materials for the full list of available topics) where R-packages in experimental design are in `r ctv("ExperimentalDesign")` task view and R-packages in survey designs are in `r ctv("OfficialStatistics")` task view.
+This dichotomisation, to a great extent, is seen in the [Comprehensive R Archive Network (CRAN) task views](https://cran.r-project.org/web/views/) (a volunteer maintained list of R-packages by topic) where R-packages in experimental design are in `r ctv("ExperimentalDesign")` task view and R-packages in survey designs are in `r ctv("OfficialStatistics")` task view.
+The full list of available topics can be see in Table S1 in the Supplementary Materials and also accessed via the `ctv` package [@ctv].
 A subset of experimental designs are segregated into `r ctv("ClinicalTrials")` task view, where the focus is on clinical trials with primary interest in sample size calculations.
@@
 
-A utility of software can also be described in terms of its design, such as the user interface (UI) and application programming interface (API).
-UI is concerned with the interaction of the software by the user, while the API is concerned with how different programs interact and is predominately of the interest to the developer.
-These interface designs is an abstraction to specifying the desired experimental design and its choices enable how a user expresses the specification of the experimental design.
+A utility of software can also be described by its design to facilitate clear expression and interpretation of the desired experimental design.
+Certain programming language design can hinder or discourage development of a reliable program [@Wasserman1975-xr].
+The immense popularity of `tidyverse` [a collection of R-packages for various stages of data analysis that place enormous emphasis on the interface design by @Wickham2019-mj] is a testament to the impact an interface design can have in practice.
 
@@
 This suggests that there are **_diverse approaches to designing experiments_**.
-This observation doesn't take into account other approaches to generating experimental designs, like the proprietary GUI software, CycDesignN, which may be widely used.
+This observation doesn't take into account other approaches to generating experimental designs, like the proprietary software, CycDesignN [@cycdesignn] and SAS [@sas1985sas], which may be widely used.
 
@@
 In software design, there are two interface designs to consider: user interface (UI) and application programming interface (API).
-UI is concerned with the usage of the software by the user, while the API is concerned with how different programs interact and is predominately of the interest to the developer.
-In this section, we discuss these interface designs broadly, with examples illustrated from the top downloaded packages (as shown in Figure \@ref(fig:rank-over-time)).
+UI is concerned with the interaction of the software by the user, while the API is concerned with how different programs interact and is predominately of the interest to the developer.
+The UI design is an abstraction to specifying the desired experimental design and its choices enable how a user expresses the specification of the experimental design.
+
+In this section, we discuss these interface designs broadly, with examples illustrated from the top downloaded packages (shown in Figure \@ref(fig:rank-over-time)).
 We exclude `ez` and `DiceKriging` from the discussion as the former  is predominately visualisation of experimental data and latter is about the analysis of computer experiments in addition to belonging to the same suite of packages as `DiceDesign`.
@@
 
-This paper uses the `targets` framework [@targets] for reproducibility, `knitr` [@knitr] and `rmarkdown` [@rmarkdown] for creating reproducible documents, `ggplot2` [@ggplot2] for visualisation and `kableExtra` [@kableExtra] for customising the table.
-All code to reproduce this paper is found at https://github.com/emitanaka/paper-DoE-review.
+This paper uses the `targets` framework [@targets] for reproducibility, `knitr` [@knitr] and `rmarkdown` [@rmarkdown] for creating reproducible documents, `ggplot2` [@ggplot2], `ggraph` [@ggraph], `ggwordcloud` [@ggwordcloud] and `colorspace` [@colorspace] for visualisation, `kableExtra` [@kableExtra] for customising the table in the Supplementary Material and `tidyverse` [@Wickham2019-mj], `tidytext` [@tidytext], `pluralize` [@pluralize], and `ineq` [@ineq] for data processing and manipulation, and `cranlogs` [@cranlogs] and `ctv` [@ctv] for extracting data.
+All code to reproduce this paper is found at `r url("https://github.com/emitanaka/paper-DoE-review")`.
 

### 2022-06-10 paper.Rmd
latest

@@
 
-## The case of specialised designs
+`DoE.wrapper` [@DoE.wrapper]
 
-In this section, we present some examples the user interface of widely used specialised designs, namely, computer experiments, factorial experiments, response surface designs and sequential designs.
+## The case of space-filling and optimal designs
 
@@
 
+The `AlgDesign` [@AlgDesign] package offers three primary functions for generating optimal designs: `optBlock()`, `optFederov()` and `optMonteCarlo()`.
+These functions in general require a formula in terms of the supplied terms with the choice of criterion (D, A and I for the latter two).
+The difference lies in the search algorithm for the optimal design.
+
+## The case of factorial designs
+
 Factorial experiments offer a challenge in the construction and allocation of the treatment factors.
 This is reflected in a number of packages that specifically addresses this challenge.
-These include `DoE.base` [@DoE.base], `FrF2` [@FrF2] for fractional 2-level factorial designs, `conf.design` [@conf.design], `FrF2.catlg128` [@FrF2.catlg128] and `BHH2` [@BHH2].
+These include `DoE.base` [@DoE.base] for full factorial and (regular and irregular) orthogonal array designs via `fac.design()` and `oa.design()`, respectively, `FrF2` [@FrF2] for fractional 2-level factorial designs using `FrF2()`, `FrF2Large()` or `pb()` Plankett-Burnman designs [@Plackett1946-ly], `conf.design` [@conf.design] for symmetric confounded factorial designs via `conf.design()`, and `BHH2` [@BHH2] to generate full or fractional 2-level factorial design matrix via `ffDesMatrix()`.
+All these designs generally requires the user to input the number of factors and if the design is allowed to vary in the number of levels, the input of the number of levels for each factor.
+The construction of some of these designs are catalogue-based, e.g.
+`oa.design()` in `DoE.base` and the entire package `FrF2.catlg128` [@FrF2.catlg128] is a list of fractional 2-level factorial designs.
 
-Response surface design, as illustrated by `rsm` [@rsm], is a type of experiment that involve variables with continuous measures.
+Response surface design, as illustrated by `rsm` [@rsm], is a factorial experiment but the focus of it is to analyse using response surface modelling.
+The two well known response surface designs are Box-Behnken design [@Box1960-wr] and
+central-composite designs [@Box1951-ji] via `bbd()` and `ccd()` for central-composite designs providing formula.
 
-Finally, another specialised design is sequential design (also called adaptive sampling), which is best represented by `tgp` [@tgp] -- these require prior information, which are used to inform the next experimental design.
+## The case of sequential designs
+
+Another specialised design is sequential design (also called adaptive sampling), which is best represented by `tgp` [@tgp] -- these require prior information, which are used to inform the next experimental design using `tgp.design()` and `dopt.gp()` which requires users to supply candidate samples to sub-sample from and a model or design from a prior design.
 Follow-up experiments, which can also be classified as a sequential design, is implemented by `BsMD` [@BsMD].
+In `BsMD`, the follow-up design is determined by a model discriminant approach using `MD()`.
 
@@
 `design.crd()`, `design.rcbd()` and `design.split()` construct a completely randomised design, randomised complete block design and split-plot design, respectively.
-Users typically supply the treatment labels and the number of replications as argument to these functions.
-
-## The case of general designs
-
-`AlgDesign` [@AlgDesign]
-
-`DoE.wrapper` [@DoE.wrapper]
+Users typically supply the treatment labels (or number of treatments in the case of `design.split()`) and the number of replications as argument to these functions.
 

### 2022-06-11 paper.Rmd
discussion added

@@
 
-Volunteer maintained -- some experimental design packages may not be included in the CRAN task view.
+Through the exploratory data analysis of three data sources (package download logs, package metadata and CRAN task views) outlined in Section \@ref(data), we have observed that the the total download of DoE packages are concentrated only on a handful of R-packages although these represent a diverse set in comparison with other CRAN task views.
+Furthermore, the data suggests that the field of experimental design is the least collaborative field; out of all CRAN task views, it has one of the least average number of authors and the lowest intra-connectivity (any dependency on other packages within its own task view).
+There are a number of limitations and short coming in our exploratory data analysis.
+First, CRAN task views are volunteer maintained so some experimental design packages may not be included in the DoE packages.
+Second, we only use the RStudio CRAN mirror download, which may bias our observations.
+Third, our analysis are limited to R-packages alone, although the sheer quantity and variation of experimental design is unmatched with any other programming languages.
+Finally, all our statements should be treated as speculative rather than conclusive; the data are all observational so no conclusive, generalisable statement are possible.
+Regardless, the data driven nature of our analysis give an objective insight into the field of experimental design.
+
+The interface design (discussed in Section \@ref(design)) reveals that the most widely used DoE packages generally have functions that is 1) a menu format (i.e.
+the name of the function correspond to a particular recipe of the design or optimal search algorithm) and 2) context is often a second-thought (e.g.
+only the number of factors is needed and the package will assign pseudo factor names or an argument exists for optional input of character vector that corresponds to the factor names).
 

### 2022-06-11 paper/arxiv/arxiv.tex
discussion added

@@
 
-\title{Current status and prospects of R-packages for the design of
+\title{Current state and prospects of R-packages for the design of
 experiments}
@@
 this article we describe the current state of the R-packages for the
-design of experiments through exploratory data analysis of package
-downloads, package metadata, and comparison of characteristics with
-other topics. We discuss also the software design of widely utilised R
-packages in the field of experimental design and conclude with
-discussion of some future prospects for the field.
+design of experiments through an exploratory data analysis of package
+downloads, package metadata, and the comparison of characteristics with
+other topics. We observe that experimental designs in practice appear to
+be sufficiently manufactured by a small number of packages and the
+development of experimental designs occur in silos. We discuss also the
+interface designs of widely utilised R packages in the field of
+experimental design and discuss its future prospects for the field.
 \end{abstract}
@@
 rubbish then no analysis, however complex it may be, can make something
-out of it. Methods for data collection can be dichotomised by the type
-of data collected -- namely, experimental or observational -- or
-alternatively, categorised as experimental design (including
-quasi-experimental design) or survey design. This dichotomisation, to a
-great extent, is seen in the
-\href{https://cran.r-project.org/web/views/}{CRAN task views} (a
-volunteered maintained list of R-packages by topic) where R-packages in
-experimental design are in ctv\{ExperimentalDesign\} and R-packages in
-survey designs are in ctv\{OfficialStatistics\}. Experimental designs
-exclusively center on the collection of experimental data where many of
-the experimental factors tend to be in the control of the experimenter.
-A subset of experimental designs are segregated into
-ctv\{ClinicalTrials\}, where the focus is on clinical trials with
-primary interest in sample size calculations. This paper focuses on the
-packages in ctv\{ExperimentalDesign\}, henceforth referred to as ``DoE
+out of it. A carefully crafted data collection scheme is therefore
+critical to optimise the information from data. Of all the data
+collection schemes, the field of experimental designs is specifically
+devoted to planning an experiment largely based on the founding
+principles by Fisher (1935) or an optimisation framework like those
+described in Pukelsheim (2006). These experimental designs are often
+constructed with the aid of a statistical software, such as R (R Core
+Team 2021), so software usage can aid in understanding the practice of
+experimental design. Given the large gain that can be made from
+collecting higher quality experimental data, we benefit from
+understanding the current state of the experimental designs in practice.
+
+Methods for data collection can be dichotomised by the type of data
+collected -- namely, experimental or observational -- or alternatively,
+categorised as experimental design (including quasi-experimental design)
+or survey design. This dichotomisation, to a great extent, is seen in
+the \href{https://cran.r-project.org/web/views/}{Comprehensive R Archive
+Network (CRAN) task views} (a volunteer maintained list of R-packages by
+topic) where R-packages in experimental design are in
+ctv\{ExperimentalDesign\} task view and R-packages in survey designs are
+in ctv\{OfficialStatistics\} task view. The full list of available
+topics can be see in Table S1 in the Supplementary Materials and also
+accessed via the \texttt{ctv} package (Zeileis 2005). A subset of
+experimental designs are segregated into ctv\{ClinicalTrials\} task
+view, where the focus is on clinical trials with primary interest in
+sample size calculations. This paper focuses on the packages in
+ctv\{ExperimentalDesign\} task view, henceforth referred to as ``DoE
 packages''.
 
-In the CRAN task view of ctv\{ExperimentalDesign\}, there are 114 R
-packages for experimental design and analysis of data from experiments.
-The sheer quantity and variation of experimental designs in the
-R-packages are arguably unmatched with any other programming languages,
-e.g.~in Python (Rossum 1995), only a handful of libraries that generate
-design of experiment exist (namely \texttt{pyDOE}, \texttt{pyDOE2},
+In ctv\{ExperimentalDesign\} task view, there are 114 R packages for
+experimental design and analysis of data from experiments. The sheer
+quantity and variation of experimental designs in the R-packages are
+arguably unmatched with any other programming languages, e.g.~in Python
+(Rossum 1995), only a handful of libraries that generate design of
+experiment exist (namely \texttt{pyDOE}, \texttt{pyDOE2},
 \texttt{dexpy}, \texttt{experimenter} and \texttt{GPdoemd}) with limited
-type of designs. Thus, the study of DoE packages can be revealing into
-the current status of the field of experimental design.
+type of designs. Thus, the study of DoE packages, based on quantitative
+and qualitative data, can give us an objective view to the state of the
+current experimental designs in practice.
+
+A utility of software can also be described by its design to facilitate
+clear expression and interpretation of the desired experimental design.
+Certain programming language design can hinder or discourage development
+of a reliable program (Wasserman 1975). The immense popularity of
+\texttt{tidyverse} (a collection of R-packages for various stages of
+data analysis that place enormous emphasis on the interface design by
+Wickham et al. 2019) is a testament to the impact an interface design
+can have in practice.
 
@@
 
-\caption{The points show the Gini index of the download counts by year facetted by CRAN task view with the color showing the number of packages.
+\caption{The points show the Gini index of the download counts by year for ExperimentalDesign CRAN task view with the color showing the number of packages.
 The grey line shows the distribution of the Gini index across years for all other CRAN task views.
-The facets are ordered by increasing value of the Gini index in 2021.}\label{fig:fig-gini-all-ctvs}
+See Figure S2 in the Supplementary Material for the line graph for the Gini index across years by each CRAN task view.}\label{fig:fig-gini-all-ctvs}
 \end{figure}
@@
   observation doesn't take into account other approaches to generating
-  experimental designs, like the proprietary GUI software, CycDesignN,
+  experimental designs, like the proprietary software, CycDesignN
+  (Whittaker, Williams, and John 2022) and SAS (SAS Institute 1985),
   which may be widely used.
@@
 interface (UI) and application programming interface (API). UI is
-concerned with the usage of the software by the user, while the API is
-concerned with how different programs interact and is predominately of
-the interest to the developer. In this section, we discuss these
-interface designs broadly, with examples illustrated from the top
-downloaded packages (as shown in Figure @ref(fig:rank-over-time)). We
-exclude \texttt{ez} and \texttt{DiceKriging} from the discussion as the
-former is predominately visualisation of experimental data and latter is
-about the analysis of computer experiments in addition to belonging to
-the same suite of packages as \texttt{DiceDesign}.
+concerned with the interaction of the software by the user, while the
+API is concerned with how different programs interact and is
+predominately of the interest to the developer. The UI design is an
+abstraction to specifying the desired experimental design and its
+choices enable how a user expresses the specification of the
+experimental design.
 
-\hypertarget{the-case-of-specialised-designs}{%
-\subsection{The case of specialised
-designs}\label{the-case-of-specialised-designs}}
+In this section, we discuss these interface designs broadly, with
+examples illustrated from the top downloaded packages (shown in Figure
+@ref(fig:rank-over-time)). We exclude \texttt{ez} and
+\texttt{DiceKriging} from the discussion as the former is predominately
+visualisation of experimental data and latter is about the analysis of
+computer experiments in addition to belonging to the same suite of
+packages as \texttt{DiceDesign}.
 
-In this section, we present some examples the user interface of widely
-used specialised designs, namely, computer experiments, factorial
-experiments, response surface designs and sequential designs.
+\texttt{DoE.wrapper} (Grömping 2020)
+
+\hypertarget{the-case-of-space-filling-and-optimal-designs}{%
+\subsection{The case of space-filling and optimal
+designs}\label{the-case-of-space-filling-and-optimal-designs}}
 
 Computer experiments, as discussed in Section @ref(topics), which
-generally involve space filling designs have been on the rise. The
+generally involve space filling designs, have been on the rise. The
 exemplar for this type of designs are \texttt{lhs} and
 \texttt{DiceDesign}. For the \texttt{lhs} package, functions like
-\texttt{randomLHS()}, \texttt{optimumLHS()}, and \texttt{maximinLHS()}
+\texttt{randomLHS()}, \texttt{optimumLHS()}, and \texttt{maximinLHS()},
+require users to specify the sample size (\(n\)) and the number of
+variables (\(p\)) then it generates a Latin hypercube sample, based on
+different optimisation schemes (in this case, random, S optimal and
+maxmin criteria, respectively; see package documentation for more
+details). Similarly for the \texttt{DiceDesign}, there is a
+comprehensive list of space-filling designs such as
+\texttt{dmaxDesign()}, \texttt{lhsDesign()}, \texttt{straussDesign()}
+and \texttt{wspDesign()} with input of \(n\) and \(p\) as before (among
+extra parameters for some) that implement algorithms that either
+maximise the entropy, produce a random Latin hypercube sample, based on
+Strauss process and WSP algorithm, respectively. These designs all
+output a \(n \times p\) matrix with values between 0 and 1.
+
+The \texttt{AlgDesign} (Wheeler 2022) package offers three primary
+functions for generating optimal designs: \texttt{optBlock()},
+\texttt{optFederov()} and \texttt{optMonteCarlo()}. These functions in
+general require a formula in terms of the supplied terms with the choice
+of criterion (D, A and I for the latter two). The difference lies in the
+search algorithm for the optimal design.
+
+\hypertarget{the-case-of-factorial-designs}{%
+\subsection{The case of factorial
+designs}\label{the-case-of-factorial-designs}}
 
@@
 packages that specifically addresses this challenge. These include
-\texttt{DoE.base} (Grömping 2018), \texttt{FrF2} (Grömping 2014) for
-fractional 2-level factorial designs, \texttt{conf.design} (Venables
-2013), \texttt{FrF2.catlg128} (Grömping 2022) and \texttt{BHH2} (Barrios
-2016).
+\texttt{DoE.base} (Grömping 2018) for full factorial and (regular and
+irregular) orthogonal array designs via \texttt{fac.design()} and
+\texttt{oa.design()}, respectively, \texttt{FrF2} (Grömping 2014) for
+fractional 2-level factorial designs using \texttt{FrF2()},
+\texttt{FrF2Large()} or \texttt{pb()} Plankett-Burnman designs (Plackett
+and Burman 1946), \texttt{conf.design} (Venables 2013) for symmetric
+confounded factorial designs via \texttt{conf.design()}, and
+\texttt{BHH2} (Barrios 2016) to generate full or fractional 2-level
+factorial design matrix via \texttt{ffDesMatrix()}. All these designs
+generally requires the user to input the number of factors and if the
+design is allowed to vary in the number of levels, the input of the
+number of levels for each factor. The construction of some of these
+designs are catalogue-based, e.g.~\texttt{oa.design()} in
+\texttt{DoE.base} and the entire package \texttt{FrF2.catlg128}
+(Grömping 2022) is a list of fractional 2-level factorial designs.
 
 Response surface design, as illustrated by \texttt{rsm} (Lenth 2009), is
-a type of experiment that involve variables with continuous measures.
+a factorial experiment but the focus of it is to analyse using response
+surface modelling. The two well known response surface designs are
+Box-Behnken design (Box and Behnken 1960) and central-composite designs
+(Box and Wilson 1951) via \texttt{bbd()} and \texttt{ccd()} for
+central-composite designs providing formula.
 
-Finally, another specialised design is sequential design (also called
-adaptive sampling), which is best represented by \texttt{tgp} (Gramacy
-and Taddy 2010) -- these require prior information, which are used to
-inform the next experimental design. Follow-up experiments, which can
-also be classified as a sequential design, is implemented by
-\texttt{BsMD} (Barrios 2020).
+\hypertarget{the-case-of-sequential-designs}{%
+\subsection{The case of sequential
+designs}\label{the-case-of-sequential-designs}}
+
+Another specialised design is sequential design (also called adaptive
+sampling), which is best represented by \texttt{tgp} (Gramacy and Taddy
+2010) -- these require prior information, which are used to inform the
+next experimental design using \texttt{tgp.design()} and
+\texttt{dopt.gp()} which requires users to supply candidate samples to
+sub-sample from and a model or design from a prior design. Follow-up
+experiments, which can also be classified as a sequential design, is
+implemented by \texttt{BsMD} (Barrios 2020). In \texttt{BsMD}, the
+follow-up design is determined by a model discriminant approach using
+\texttt{MD()}.
 
@@
 randomised complete block design and split-plot design, respectively.
-Users typically supply the treatment labels and the number of
-replications as argument to these functions.
-
-\hypertarget{the-case-of-general-designs}{%
-\subsection{The case of general
-designs}\label{the-case-of-general-designs}}
-
-\texttt{AlgDesign} (Wheeler 2022)
-
-\texttt{DoE.wrapper} (Grömping 2020)
+Users typically supply the treatment labels (or number of treatments in
+the case of \texttt{design.split()}) and the number of replications as
+argument to these functions.
 
@@
 
-Volunteer maintained -- some experimental design packages may not be
-included in the CRAN task view.
+Through the exploratory data analysis of three data sources (package
+download logs, package metadata and CRAN task views) outlined in Section
+@ref(data), we have observed that the the total download of DoE packages
+are concentrated only on a handful of R-packages although these
+represent a diverse set in comparison with other CRAN task views.
+Furthermore, the data suggests that the field of experimental design is
+the least collaborative field; out of all CRAN task views, it has one of
+the least average number of authors and the lowest intra-connectivity
+(any dependency on other packages within its own task view). There are a
+number of limitations and short coming in our exploratory data analysis.
+First, CRAN task views are volunteer maintained so some experimental
+design packages may not be included in the DoE packages. Second, we only
+use the RStudio CRAN mirror download, which may bias our observations.
+Third, our analysis are limited to R-packages alone, although the sheer
+quantity and variation of experimental design is unmatched with any
+other programming languages. Finally, all our statements should be
+treated as speculative rather than conclusive; the data are all
+observational so no conclusive, generalisable statement are possible.
+Regardless, the data driven nature of our analysis give an objective
+insight into the field of experimental design.
+
+The interface design (discussed in Section @ref(design)) reveals that
+the most widely used DoE packages generally have functions that is 1) a
+menu format (i.e.~the name of the function correspond to a particular
+recipe of the design or optimal search algorithm) and 2) context is
+often a second-thought (e.g.~only the number of factors is needed and
+the package will assign pseudo factor names or an argument exists for
+optional input of character vector that corresponds to the factor
+names).
 
@@
 Allaire, and Grolemund 2018) for creating reproducible documents,
-\texttt{ggplot2} (Wickham 2016) for visualisation and
-\texttt{kableExtra} (Zhu 2021) for customising the table. All code to
-reproduce this paper is found at
-\url{https://github.com/emitanaka/paper-DoE-review}.
+\texttt{ggplot2} (Wickham 2016), \texttt{ggraph} (Pedersen 2021),
+\texttt{ggwordcloud} (Le Pennec and Slowikowski 2019) and
+\texttt{colorspace} (Zeileis et al. 2020) for visualisation,
+\texttt{kableExtra} (Zhu 2021) for customising the table in the
+Supplementary Material and \texttt{tidyverse} (Wickham et al. 2019),
+\texttt{tidytext} (Silge and Robinson 2016), \texttt{pluralize} (Rudis
+and Embrey 2020), and \texttt{ineq} (Zeileis 2014) for data processing
+and manipulation, and \texttt{cranlogs} (Csárdi 2019) and \texttt{ctv}
+(Zeileis 2005) for extracting data. All code to reproduce this paper is
+found at \url{https://github.com/emitanaka/paper-DoE-review}.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Box1960-wr}{}}%
+Box, G E P, and D W Behnken. 1960. {``Some New Three Level Designs for
+the Study of Quantitative Variables.''} \emph{Technometrics: A Journal
+of Statistics for the Physical, Chemical, and Engineering Sciences} 2
+(4): 455--75. \url{https://doi.org/10.2307/1266454}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Box1951-ji}{}}%
+Box, G E P, and K B Wilson. 1951. {``On the Experimental Attainment of
+Optimum Conditions.''} \emph{Journal of the Royal Statistical Society.
+Series B, Statistical Methodology} 13 (1): 1--45.
+
 \leavevmode\vadjust pre{\hypertarget{ref-cranlogs}{}}%
@@
 \url{https://CRAN.R-project.org/package=agricolae}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Fisher1935-qc}{}}%
+Fisher, Ronald. 1935. \emph{The Design of Experiments}. Oliver; Boyd.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-ggwordcloud}{}}%
+Le Pennec, Erwan, and Kamil Slowikowski. 2019. \emph{Ggwordcloud: A Word
+Cloud Geom for 'Ggplot2'}.
+\url{https://CRAN.R-project.org/package=ggwordcloud}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-rsm}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-ggraph}{}}%
+Pedersen, Thomas Lin. 2021. \emph{Ggraph: An Implementation of Grammar
+of Graphics for Graphs and Networks}.
+\url{https://CRAN.R-project.org/package=ggraph}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Plackett1946-ly}{}}%
+Plackett, R L, and J P Burman. 1946. {``The Design of Optimum
+Multifactorial Experiments.''} \emph{Biometrika} 33 (4): 302--25.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Pukelsheim2006-sv}{}}%
+Pukelsheim, Friedrich. 2006. \emph{Optimal Design of Experiments}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-R}{}}%
+R Core Team. 2021. \emph{R: A Language and Environment for Statistical
+Computing}. Vienna, Austria: R Foundation for Statistical Computing.
+\url{https://www.R-project.org/}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-python}{}}%
@@
 Centrum voor Wiskunde en Informatica (CWI).
+
+\leavevmode\vadjust pre{\hypertarget{ref-pluralize}{}}%
+Rudis, Bob, and Blake Embrey. 2020. \emph{Pluralize: Pluralize and
+'Singularize' Any (English) Word}.
+\url{https://CRAN.R-project.org/package=pluralize}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sas1985sas}{}}%
+SAS Institute. 1985. \emph{SAS User's Guide: Statistics}. Vol. 2. Sas
+Inst.
+
+\leavevmode\vadjust pre{\hypertarget{ref-tidytext}{}}%
+Silge, Julia, and David Robinson. 2016. {``Tidytext: Text Mining and
+Analysis Using Tidy Data Principles in r.''} \emph{JOSS} 1 (3).
+\url{https://doi.org/10.21105/joss.00037}.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Wasserman1975-xr}{}}%
+Wasserman, Anthony I. 1975. {``Issues in Programming Language Design -
+an Overview.''} In.
+
 \leavevmode\vadjust pre{\hypertarget{ref-AlgDesign}{}}%
@@
 \url{https://CRAN.R-project.org/package=AlgDesign}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-cycdesignn}{}}%
+Whittaker, D., E. R. Williams, and J. A John. 2022. \emph{CycDesign: A
+Package for the Computer Generation of Experimental Designs,} (version
+7.0). \url{https://vsni.co.uk/software/cycdesign}.
 
@@
 \url{https://ggplot2.tidyverse.org}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wickham2019-mj}{}}%
+Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
+D'agostino McGowan, Romain Francois, Garrett Grolemund, et al. 2019.
+{``Welcome to the Tidyverse.''} \emph{Journal of Open Source Software} 4
+(43): 1686.
 
@@
 \url{https://bookdown.org/yihui/rmarkdown}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ctv}{}}%
+Zeileis, Achim. 2005. {``{CRAN} Task Views.''} \emph{R News} 5 (1):
+39--40. \url{https://CRAN.R-project.org/doc/Rnews/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ineq}{}}%
+---------. 2014. \emph{Ineq: Measuring Inequality, Concentration, and
+Poverty}. \url{https://CRAN.R-project.org/package=ineq}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-colorspace}{}}%
+Zeileis, Achim, Jason C. Fisher, Kurt Hornik, Ross Ihaka, Claire D.
+McWhite, Paul Murrell, Reto Stauffer, and Claus O. Wilke. 2020.
+{``{colorspace}: A Toolbox for Manipulating and Assessing Colors and
+Palettes.''} \emph{Journal of Statistical Software} 96 (1): 1--49.
+\url{https://doi.org/10.18637/jss.v096.i01}.
 

### 2022-06-11 paper.Rmd
fix links and cross references in pdf version

@@
 ctv <- function(view) {
-  ifelse(knitr::is_html_output(), glue("[{view}](https://cran.r-project.org/web/views/{view}.html)"), glue("ctv{[view]}", .open = "[", .close = "]"))
+  ifelse(knitr::is_html_output(), glue("[{view}](https://cran.r-project.org/web/views/{view}.html)"), glue("\\ctv{[view]}", .open = "[", .close = "]"))
 }
@@
 The paper is organised as follows.
-Section \@ref(data) briefly describes the data source used for the analysis in Section \@ref(eda); Section \@ref(eda) presents some insights into the state of the current DoE packages by exploratory data analysis of package download data, text descriptions and comparisons with other CRAN task views; Section \@ref(design) discuss the interface designs of DoE packages, and we conclude with a discussion in Section \@ref(discussion) of future prospects.
+Section `r ref("data")` briefly describes the data source used for the analysis in Section `r ref("eda")`; Section `r ref("eda")` presents some insights into the state of the current DoE packages by exploratory data analysis of package download data, text descriptions and comparisons with other CRAN task views; Section `r ref("design")` discuss the interface designs of DoE packages, and we conclude with a discussion in Section `r ref("discussion")` of future prospects.
 
@@
 All CRAN packages have a title, description, package connections (suggests, dependency and imports) and other meta-information in the DESCRIPTION file.
-We use the text data from the title and description (as accessed in `r as.Date(file.info("_targets/objects/ngram_download_data")$mtime)`) in Section \@ref(topics).
+We use the text data from the title and description (as accessed in `r as.Date(file.info("_targets/objects/ngram_download_data")$mtime)`) in Section `r ref("topics")`.
 
@@
 
-The list of packages in each CRAN task views (as of `r as.Date(file.info("_targets/objects/ctv_pkgs_data")$mtime)`) are used to contrast the characteristics of DoE packages in Section \@ref(silo).
+The list of packages in each CRAN task views (as of `r as.Date(file.info("_targets/objects/ctv_pkgs_data")$mtime)`) are used to contrast the characteristics of DoE packages in Section `r ref("silo")`.
 
@@
 There have been at least 50 DoE packages since 2013 but most of the downloads are concentrated in just a handful of packages.
-For example, Figure \@ref(fig:plot-lorenz) shows a Lorenz curve [@Lorenz1905-tc] for the total package downloads in 2021 of `r tar_read(gini_yearly) %>% filter(year==2021) %>% pull(n)` DoE packages (first released prior to 2021); we can see from Figure \@ref(fig:plot-lorenz) that bottom 90\% of DoE packages (in terms of total download count in 2021) only share about 30\% of total downloads across all DoE packages -- in another words, 70\% of the total downloads are due to 11 packages (10% of the DoE packages).
+For example, Figure `r ref("fig:plot-lorenz")` shows a Lorenz curve [@Lorenz1905-tc] for the total package downloads in 2021 of `r tar_read(gini_yearly) %>% filter(year==2021) %>% pull(n)` DoE packages (first released prior to 2021); we can see from Figure `r ref("fig:plot-lorenz")` that bottom 90\% of DoE packages (in terms of total download count in 2021) only share about 30\% of total downloads across all DoE packages -- in another words, 70\% of the total downloads are due to 11 packages (10% of the DoE packages).
 
@@
  If we consider package downloads as a measure of "wealth", then we can consider using the Gini index [@Gini1921-mf] as a measure of download inequality across packages.
-The ratio of the red region over the total colored regions in Figure \@ref(fig:plot-lorenz) corresponds to the Gini index for 2021.
+The ratio of the red region over the total colored regions in Figure `r ref("fig:plot-lorenz")` corresponds to the Gini index for 2021.
 A Gini index of 0\% indicates equality in downloads across packages while a value of 100\% indicates maximal inequality (all downloads are due to one package).
-In Figure \@ref(fig:download-share), we see that the distributions of the downloads each year have a heavy right tail with the Gini index ranging from `r scales::percent(min(gini_yearly$gini), 0.1)` to `r scales::percent(max(gini_yearly$gini), 0.1)` across the years 2013 to 2021, indicating that there is a high level of inequality of downloads across packages, particularly with more pronounced inequality in the last 5 years.
+In Figure `r ref("fig:download-share")`, we see that the distributions of the downloads each year have a heavy right tail with the Gini index ranging from `r scales::percent(min(gini_yearly$gini), 0.1)` to `r scales::percent(max(gini_yearly$gini), 0.1)` across the years 2013 to 2021, indicating that there is a high level of inequality of downloads across packages, particularly with more pronounced inequality in the last 5 years.
 
@@
 
-While in absolute terms the Gini index is high for `r ctv("ExperimentalDesign")`, the inequality is not as severe as other CRAN task views as shown in Figure \@ref(fig:fig-gini-all-ctvs).
-We can see in Figure \@ref(fig:fig-gini-all-ctvs) that the Gini index is generally increasing for DoE packages (as is generally the case for other CRAN task views as shown in Figure S1 in the Supplementary Materials).
+While in absolute terms the Gini index is high for `r ctv("ExperimentalDesign")`, the inequality is not as severe as other CRAN task views as shown in Figure `r ref("fig:fig-gini-all-ctvs")`.
+We can see in Figure `r ref("fig:fig-gini-all-ctvs")` that the Gini index is generally increasing for DoE packages (as is generally the case for other CRAN task views as shown in Figure S1 in the Supplementary Materials).
 
@@
 
-From Figures \@ref(fig:download-share) and \@ref(fig:fig-gini-all-ctvs), we could draw some indicative states of the field of experimental designs with possible counterfactual interpretations:
+From Figures `r ref("fig:download-share")` and `r ref("fig:fig-gini-all-ctvs")`, we could draw some indicative states of the field of experimental designs with possible counterfactual interpretations:
 
@@
 
-We can see in Figure \@ref(fig:rank-over-time) that most of the top 10 ranking packages have been in the top 10 for the last 9 years with `lhs` steadily climbing up the ranks in the last few years.
+We can see in Figure `r ref("fig:rank-over-time")` that most of the top 10 ranking packages have been in the top 10 for the last 9 years with `lhs` steadily climbing up the ranks in the last few years.
 
@@
 
-Figure \@ref(fig:established-packages) shows a moderate negative correlation between the first release date and the (log of) total download counts of DoE packages in any given year from 2013 to 2021.
+Figure `r ref("fig:established-packages")` shows a moderate negative correlation between the first release date and the (log of) total download counts of DoE packages in any given year from 2013 to 2021.
 This suggest that in general, a package released earlier are more likely to be used today (possibly for legacy reasons or the general inertia for adoption of new packages).
-We can also see in Figure \@ref(fig:established-packages) the most downloaded packages were generally released in 2004 to 2010.
+We can also see in Figure `r ref("fig:established-packages")` the most downloaded packages were generally released in 2004 to 2010.
 
@@
 
-Figure \@ref(fig:wordcloud-over-time) shows some common purposes of DoE packages, based on bigrams in the package title and description.
+Figure `r ref("fig:wordcloud-over-time")` shows some common purposes of DoE packages, based on bigrams in the package title and description.
 A bigram is only shown as the frequency of single word was not insightful and there were not many trigrams common across packages.
 Not surprisingly, the bigram "experimental design" was the most common.
-More interestingly, "optimal design" appeared across different packages (indicated by the size of the word in Figure \@ref(fig:wordcloud-over-time)), and the bigrams "latin hypercube" and "computer experiment" used across a few packages that are downloaded frequently (indicated by the color of the word in Figure \@ref(fig:wordcloud-over-time)).
-Although there exists a separate `r ctv("ClinicalTrials")` task view, the DoE packages clearly include some packages that are of interest to clinical trials as shown by the size of the bigram "clinical trial" in Figure \@ref(fig:wordcloud-over-time).
+More interestingly, "optimal design" appeared across different packages (indicated by the size of the word in Figure `r ref("fig:wordcloud-over-time")`), and the bigrams "latin hypercube" and "computer experiment" used across a few packages that are downloaded frequently (indicated by the color of the word in Figure `r ref("fig:wordcloud-over-time")`).
+Although there exists a separate `r ctv("ClinicalTrials")` task view, the DoE packages clearly include some packages that are of interest to clinical trials as shown by the size of the bigram "clinical trial" in Figure `r ref("fig:wordcloud-over-time")`.
 
@@
 
-Figure \@ref(fig:ctv-summ-plot) shows that the `r ctv("ExperimentalDesign")` have the lowest average number of contributors among all CRAN task views.
-In addition, we can also see in Figure \@ref(fig:ctv-summ-plot)  that the `r ctv("ExperimentalDesign")` task view has one of the least intra-connectivity (the percentage of packages that make use of other packages within the same task view).
-The full connection of DoE packages with other DoE packages is shown in Figure \@ref(fig:plot-doe-network).
+Figure `r ref("fig:ctv-summ-plot")` shows that the `r ctv("ExperimentalDesign")` have the lowest average number of contributors among all CRAN task views.
+In addition, we can also see in Figure `r ref("fig:ctv-summ-plot")`  that the `r ctv("ExperimentalDesign")` task view has one of the least intra-connectivity (the percentage of packages that make use of other packages within the same task view).
+The full connection of DoE packages with other DoE packages is shown in Figure `r ref("fig:plot-doe-network")`.
 These observations suggest that the field of experimental design is one of the least collaborative field and package development generally occur in silos.
@@
 
-In this section, we discuss these interface designs broadly, with examples illustrated from the top downloaded packages (shown in Figure \@ref(fig:rank-over-time)).
+In this section, we discuss these interface designs broadly, with examples illustrated from the top downloaded packages (shown in Figure `r ref("fig:rank-over-time")`).
 We exclude `ez` and `DiceKriging` from the discussion as the former  is predominately visualisation of experimental data and latter is about the analysis of computer experiments in addition to belonging to the same suite of packages as `DiceDesign`.
@@
 
-Computer experiments, as discussed in Section \@ref(topics), which generally involve space filling designs, have been on the rise.
+Computer experiments, as discussed in Section `r ref("topics")`, which generally involve space filling designs, have been on the rise.
 The exemplar for this type of designs are `lhs` and `DiceDesign`.
@@
 
-The `agricolae` package [@agricolae] was the most downloaded DoE packages in 2013 to 2018 (Figure \@ref(fig:rank-over-time)) and possibly prior to 2013 (data unavailable).
+The `agricolae` package [@agricolae] was the most downloaded DoE packages in 2013 to 2018 (Figure `r ref("fig:rank-over-time")`) and possibly prior to 2013 (data unavailable).
 This package is the prime example of constructing designs based on menu functions, e.g.
@@
 
-Through the exploratory data analysis of three data sources (package download logs, package metadata and CRAN task views) outlined in Section \@ref(data), we have observed that the the total download of DoE packages are concentrated only on a handful of R-packages although these represent a diverse set in comparison with other CRAN task views.
+Through the exploratory data analysis of three data sources (package download logs, package metadata and CRAN task views) outlined in Section `r ref("data")`, we have observed that the the total download of DoE packages are concentrated only on a handful of R-packages although these represent a diverse set in comparison with other CRAN task views.
 Furthermore, the data suggests that the field of experimental design is the least collaborative field; out of all CRAN task views, it has one of the least average number of authors and the lowest intra-connectivity (any dependency on other packages within its own task view).
@@
 
-The interface design (discussed in Section \@ref(design)) reveals that the most widely used DoE packages generally have functions that is 1) a menu format (i.e.
+The interface design (discussed in Section `r ref("design")`) reveals that the most widely used DoE packages generally have functions that is 1) a menu format (i.e.
 the name of the function correspond to a particular recipe of the design or optimal search algorithm) and 2) context is often a second-thought (e.g.

### 2022-06-11 paper/arxiv/arxiv.Rmd
fix links and cross references in pdf version

@@
   - name: Emi Tanaka
-    thanks: emitanaka.org
+    thanks: Corresponding author
     department: Department of Econometrics and Business Statistics
@@
   \usepackage{colortbl}
+  \usepackage{bera}
   \usepackage{xcolor}
@@
   \def\tightlist{}
+  \newcommand{\ctv}[1]{\href{http://CRAN.R-project.org/view=#1}{\emph{#1}}}
 bibliography: ../../paper.bib

### 2022-06-11 paper/arxiv/arxiv.tex
fix links and cross references in pdf version

@@
     Emi Tanaka
-    \thanks{emitanaka.org}
+    \thanks{Corresponding author}
    \\
@@
 \usepackage{colortbl}
+\usepackage{bera}
 \usepackage{xcolor}
@@
 \def\tightlist{}
+\newcommand{\ctv}[1]{\href{http://CRAN.R-project.org/view=#1}{\emph{#1}}}
 
@@
 topic) where R-packages in experimental design are in
-ctv\{ExperimentalDesign\} task view and R-packages in survey designs are
-in ctv\{OfficialStatistics\} task view. The full list of available
-topics can be see in Table S1 in the Supplementary Materials and also
-accessed via the \texttt{ctv} package (Zeileis 2005). A subset of
-experimental designs are segregated into ctv\{ClinicalTrials\} task
-view, where the focus is on clinical trials with primary interest in
-sample size calculations. This paper focuses on the packages in
-ctv\{ExperimentalDesign\} task view, henceforth referred to as ``DoE
-packages''.
+\href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}}
+task view and R-packages in survey designs are in
+\href{http://CRAN.R-project.org/view=OfficialStatistics}{\emph{OfficialStatistics}}
+task view. The full list of available topics can be see in Table S1 in
+the Supplementary Materials and also accessed via the \texttt{ctv}
+package (Zeileis 2005). A subset of experimental designs are segregated
+into
+\href{http://CRAN.R-project.org/view=ClinicalTrials}{\emph{ClinicalTrials}}
+task view, where the focus is on clinical trials with primary interest
+in sample size calculations. This paper focuses on the packages in
+\href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}}
+task view, henceforth referred to as ``DoE packages''.
 
-In ctv\{ExperimentalDesign\} task view, there are 114 R packages for
-experimental design and analysis of data from experiments. The sheer
-quantity and variation of experimental designs in the R-packages are
-arguably unmatched with any other programming languages, e.g.~in Python
-(Rossum 1995), only a handful of libraries that generate design of
-experiment exist (namely \texttt{pyDOE}, \texttt{pyDOE2},
-\texttt{dexpy}, \texttt{experimenter} and \texttt{GPdoemd}) with limited
-type of designs. Thus, the study of DoE packages, based on quantitative
-and qualitative data, can give us an objective view to the state of the
-current experimental designs in practice.
+In
+\href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}}
+task view, there are 114 R packages for experimental design and analysis
+of data from experiments. The sheer quantity and variation of
+experimental designs in the R-packages are arguably unmatched with any
+other programming languages, e.g.~in Python (Rossum 1995), only a
+handful of libraries that generate design of experiment exist (namely
+\texttt{pyDOE}, \texttt{pyDOE2}, \texttt{dexpy}, \texttt{experimenter}
+and \texttt{GPdoemd}) with limited type of designs. Thus, the study of
+DoE packages, based on quantitative and qualitative data, can give us an
+objective view to the state of the current experimental designs in
+practice.
 
@@
 
-The paper is organised as follows. Section @ref(data) briefly describes
-the data source used for the analysis in Section @ref(eda); Section
-@ref(eda) presents some insights into the state of the current DoE
+The paper is organised as follows. Section \ref{data} briefly describes
+the data source used for the analysis in Section \ref{eda}; Section
+\ref{eda} presents some insights into the state of the current DoE
 packages by exploratory data analysis of package download data, text
 descriptions and comparisons with other CRAN task views; Section
-@ref(design) discuss the interface designs of DoE packages, and we
-conclude with a discussion in Section @ref(discussion) of future
+\ref{design} discuss the interface designs of DoE packages, and we
+conclude with a discussion in Section \ref{discussion} of future
 prospects.
@@
 DESCRIPTION file. We use the text data from the title and description
-(as accessed in 2022-05-23) in Section @ref(topics).
+(as accessed in 2022-05-23) in Section \ref{topics}.
 
@@
 The list of packages in each CRAN task views (as of 2022-05-25) are used
-to contrast the characteristics of DoE packages in Section @ref(silo).
+to contrast the characteristics of DoE packages in Section \ref{silo}.
 
@@
 downloads are concentrated in just a handful of packages. For example,
-Figure @ref(fig:plot-lorenz) shows a Lorenz curve (Lorenz 1905) for the
+Figure \ref{fig:plot-lorenz} shows a Lorenz curve (Lorenz 1905) for the
 total package downloads in 2021 of 113 DoE packages (first released
-prior to 2021); we can see from Figure @ref(fig:plot-lorenz) that bottom
+prior to 2021); we can see from Figure \ref{fig:plot-lorenz} that bottom
 90\% of DoE packages (in terms of total download count in 2021) only
@@
 inequality across packages. The ratio of the red region over the total
-colored regions in Figure @ref(fig:plot-lorenz) corresponds to the Gini
+colored regions in Figure \ref{fig:plot-lorenz} corresponds to the Gini
 index for 2021. A Gini index of 0\% indicates equality in downloads
 across packages while a value of 100\% indicates maximal inequality (all
-downloads are due to one package). In Figure @ref(fig:download-share),
+downloads are due to one package). In Figure \ref{fig:download-share},
 we see that the distributions of the downloads each year have a heavy
@@
 While in absolute terms the Gini index is high for
-ctv\{ExperimentalDesign\}, the inequality is not as severe as other CRAN
-task views as shown in Figure @ref(fig:fig-gini-all-ctvs). We can see in
-Figure @ref(fig:fig-gini-all-ctvs) that the Gini index is generally
-increasing for DoE packages (as is generally the case for other CRAN
-task views as shown in Figure S1 in the Supplementary Materials).
+\href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}},
+the inequality is not as severe as other CRAN task views as shown in
+Figure \ref{fig:fig-gini-all-ctvs}. We can see in Figure
+\ref{fig:fig-gini-all-ctvs} that the Gini index is generally increasing
+for DoE packages (as is generally the case for other CRAN task views as
+shown in Figure S1 in the Supplementary Materials).
 
@@
 
-From Figures @ref(fig:download-share) and @ref(fig:fig-gini-all-ctvs),
+From Figures \ref{fig:download-share} and \ref{fig:fig-gini-all-ctvs},
 we could draw some indicative states of the field of experimental
@@
 
-We can see in Figure @ref(fig:rank-over-time) that most of the top 10
+We can see in Figure \ref{fig:rank-over-time} that most of the top 10
 ranking packages have been in the top 10 for the last 9 years with
@@
 
-Figure @ref(fig:established-packages) shows a moderate negative
+Figure \ref{fig:established-packages} shows a moderate negative
 correlation between the first release date and the (log of) total
@@
 adoption of new packages). We can also see in Figure
-@ref(fig:established-packages) the most downloaded packages were
+\ref{fig:established-packages} the most downloaded packages were
 generally released in 2004 to 2010.
@@
 
-Figure @ref(fig:wordcloud-over-time) shows some common purposes of DoE
+Figure \ref{fig:wordcloud-over-time} shows some common purposes of DoE
 packages, based on bigrams in the package title and description. A
@@
 packages (indicated by the size of the word in Figure
-@ref(fig:wordcloud-over-time)), and the bigrams ``latin hypercube'' and
+\ref{fig:wordcloud-over-time}), and the bigrams ``latin hypercube'' and
 ``computer experiment'' used across a few packages that are downloaded
 frequently (indicated by the color of the word in Figure
-@ref(fig:wordcloud-over-time)). Although there exists a separate
-ctv\{ClinicalTrials\} task view, the DoE packages clearly include some
-packages that are of interest to clinical trials as shown by the size of
-the bigram ``clinical trial'' in Figure @ref(fig:wordcloud-over-time).
+\ref{fig:wordcloud-over-time}). Although there exists a separate
+\href{http://CRAN.R-project.org/view=ClinicalTrials}{\emph{ClinicalTrials}}
+task view, the DoE packages clearly include some packages that are of
+interest to clinical trials as shown by the size of the bigram
+``clinical trial'' in Figure \ref{fig:wordcloud-over-time}.
 
@@
 
-Figure @ref(fig:ctv-summ-plot) shows that the ctv\{ExperimentalDesign\}
+Figure \ref{fig:ctv-summ-plot} shows that the
+\href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}}
 have the lowest average number of contributors among all CRAN task
-views. In addition, we can also see in Figure @ref(fig:ctv-summ-plot)
-that the ctv\{ExperimentalDesign\} task view has one of the least
-intra-connectivity (the percentage of packages that make use of other
-packages within the same task view). The full connection of DoE packages
-with other DoE packages is shown in Figure @ref(fig:plot-doe-network).
-These observations suggest that the field of experimental design is one
-of the least collaborative field and package development generally occur
-in silos.
+views. In addition, we can also see in Figure \ref{fig:ctv-summ-plot}
+that the
+\href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}}
+task view has one of the least intra-connectivity (the percentage of
+packages that make use of other packages within the same task view). The
+full connection of DoE packages with other DoE packages is shown in
+Figure \ref{fig:plot-doe-network}. These observations suggest that the
+field of experimental design is one of the least collaborative field and
+package development generally occur in silos.
 
@@
 examples illustrated from the top downloaded packages (shown in Figure
-@ref(fig:rank-over-time)). We exclude \texttt{ez} and
+\ref{fig:rank-over-time}). We exclude \texttt{ez} and
 \texttt{DiceKriging} from the discussion as the former is predominately
@@
 
-Computer experiments, as discussed in Section @ref(topics), which
+Computer experiments, as discussed in Section \ref{topics}, which
 generally involve space filling designs, have been on the rise. The
@@
 downloaded DoE packages in 2013 to 2018 (Figure
-@ref(fig:rank-over-time)) and possibly prior to 2013 (data unavailable).
+\ref{fig:rank-over-time}) and possibly prior to 2013 (data unavailable).
 This package is the prime example of constructing designs based on menu
@@
 download logs, package metadata and CRAN task views) outlined in Section
-@ref(data), we have observed that the the total download of DoE packages
+\ref{data}, we have observed that the the total download of DoE packages
 are concentrated only on a handful of R-packages although these
@@
 
-The interface design (discussed in Section @ref(design)) reveals that
+The interface design (discussed in Section \ref{design}) reveals that
 the most widely used DoE packages generally have functions that is 1) a

### 2022-06-11 paper.Rmd
fix html template

@@
   # below is to have a JSS paper style template
-  pagedown::jss_paged:
+  pagedown::html_paged:
     template: assets/html-template.html
@@
     self_contained: false
+    css: ["assets/html-template.css"]
 ---

### 2022-06-11 _targets.R
update targets to update papers as well

@@ -111,6 +111,7 @@ list(
   
   
   
-  #tar_render(report, "paper.Rmd"),
+  tar_render(report, "paper.Rmd"),
+  tar_render(arxiv, "paper/arxiv/arxiv.Rmd"),
   NULL
 )

### 2022-06-11 README.Rmd
add instruction to README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = ""
)
library(tidyverse)
paper <- here::here("paper.Rmd")
meta <- rmarkdown::yaml_front_matter(paper)
```

# `r meta$title`

This is the repo for the paper titled "`r meta$title`". You can find the arxiv version of the paper in `paper/arxiv/arxiv.pdf`. To reproduce the results of the paper, you will need to run the analysis by calling

```r
targets::tar_make()
```

The paper uses the developmental version of the [`cranscrub` package in the branch `paper-doe-review`](https://github.com/numbats/cranscrub/tree/paper-doe-review) which can be installed by

```r
remotes::install_github("numbats/cranscrub", ref = "paper-doe-review")
```

## Authors

```{r, echo = FALSE}
map_chr(meta$author, ~{
  paste0("- ", .x$name, ", ", .x$affiliation)
}) %>% 
  cli::cli_li() 
```

## Abstract

`r meta$abstract`

### 2022-06-12 paper.Rmd
fixed supp, added renv, added about import for algdesign per dewi's comments

@@
 The sheer quantity and variation of experimental designs in the R-packages are arguably unmatched with any other programming languages, e.g.
-in Python [@python], only a handful of libraries that generate design of experiment exist (namely `pyDOE`, `pyDOE2`, `dexpy`, `experimenter` and `GPdoemd`) with limited type of designs.
+in Python [@python], only a handful of packages that generate design of experiment exist (namely `pyDOE`, `pyDOE2`, `dexpy`, `experimenter` and `GPdoemd`) with limited type of designs.
 Thus, the study of DoE packages, based on quantitative and qualitative data, can give us an objective view to the state of the current experimental designs in practice.
@@
 We can see in Figure `r ref("fig:rank-over-time")` that most of the top 10 ranking packages have been in the top 10 for the last 9 years with `lhs` steadily climbing up the ranks in the last few years.
+It should be noted that the download of one package can prompt a download of another package; the most notable dependency network is `AlgDesign` and `agricolae`, where the former is an import for the latter.
+The full dependency network within the DoE packages is shown later in Figure \@ref(fig:plot-doe-network).
 
@@
 
-This paper uses the `targets` framework [@targets] for reproducibility, `knitr` [@knitr] and `rmarkdown` [@rmarkdown] for creating reproducible documents, `ggplot2` [@ggplot2], `ggraph` [@ggraph], `ggwordcloud` [@ggwordcloud] and `colorspace` [@colorspace] for visualisation, `kableExtra` [@kableExtra] for customising the table in the Supplementary Material and `tidyverse` [@Wickham2019-mj], `tidytext` [@tidytext], `pluralize` [@pluralize], and `ineq` [@ineq] for data processing and manipulation, and `cranlogs` [@cranlogs] and `ctv` [@ctv] for extracting data.
+This paper uses the `targets` framework [@targets] and `renv` [@renv] for reproducibility, `knitr` [@knitr] and `rmarkdown` [@rmarkdown] for creating reproducible documents, `ggplot2` [@ggplot2], `ggraph` [@ggraph], `ggwordcloud` [@ggwordcloud] and `colorspace` [@colorspace] for visualisation, `kableExtra` [@kableExtra] for customising the table in the Supplementary Material and `tidyverse` [@Wickham2019-mj], `tidytext` [@tidytext], `pluralize` [@pluralize], and `ineq` [@ineq] for data processing and manipulation, and `cranlogs` [@cranlogs] and `ctv` [@ctv] for extracting data.
 All code to reproduce this paper is found at `r url("https://github.com/emitanaka/paper-DoE-review")`.

### 2022-06-12 paper/arxiv/arxiv.tex
fixed supp, added renv, added about import for algdesign per dewi's comments

@@
 other programming languages, e.g.~in Python (Rossum 1995), only a
-handful of libraries that generate design of experiment exist (namely
+handful of packages that generate design of experiment exist (namely
 \texttt{pyDOE}, \texttt{pyDOE2}, \texttt{dexpy}, \texttt{experimenter}
@@
 ranking packages have been in the top 10 for the last 9 years with
-\texttt{lhs} steadily climbing up the ranks in the last few years.
+\texttt{lhs} steadily climbing up the ranks in the last few years. It
+should be noted that the download of one package can prompt a download
+of another package; the most notable dependency network is
+\texttt{AlgDesign} and \texttt{agricolae}, where the former is an import
+for the latter. The full dependency network within the DoE packages is
+shown later in Figure @ref(fig:plot-doe-network).
 

### 2022-06-12 supp.Rmd
fixed supp, added renv, added about import for algdesign per dewi's comments

@@
   - Dewi Amaliah 
-bibliography: paper.bib
+bibliography: pkgs.bib
 keywords: ["experimental design", "CRAN task view", "user interface"]
 output: 
-  bookdown::pdf_document2
-header-includes:
-  - \renewcommand{\figurename}{Figure S}
-  - \makeatletter
-  - \def\fnum@figure{\figurename\thefigure}
-  - \makeatother
+  bookdown::pdf_document2:
+    toc: false
+    keep_tex: true
+    citation_package: biblatex
+header-includes: |
+  \usepackage{colortbl}
+  \usepackage{bera}
+  \usepackage{xcolor}
+  \usepackage{hyperref}
+  \usepackage{booktabs}
+  \usepackage[utf8]{inputenc}
+  \usepackage{caption}
+  \renewcommand{\figurename}{Figure S}
+  \makeatletter
+  \def\fnum@figure{\figurename\thefigure}
+  \makeatother
+  \renewcommand{\tablename}{Table S}
+  \makeatletter
+  \def\fnum@table{\tablename\thetable}
+  \makeatother
+
+  \usepackage{bera}
 ---
-
-```{css, echo = FALSE}
-h1, h2, h3, div > * {
- font-family: "Helvetica Neue",Helvetica,Arial,sans-serif;
-}
-
-```
 
@@
 tar_load(gini_yearly_ctv)
+
+#renvLock <- jsonlite::read_json("renv.lock")
+#sink("pkgs.bib")
+#knitr::write_bib(names(renvLock$Packages))
+#sink()
 ```
@@
                digits = c(0, 0, 0, 1, 2, 0),
-               booktabs = TRUE ) %>%
- kableExtra::kable_styling(latex_options = c("striped", "scale_down")) %>% 
+               booktabs = TRUE,
+               longtable = TRUE) %>%
+ kableExtra::kable_styling(latex_options = c("striped", "scale_down", "repeat_header"),
+                           repeat_header_continued = "\\textit{(Continued on next page...)}") %>% 
  kableExtra::column_spec(2, width = "12em") %>% 
@@
 ```
+
+# Session information {-}
+
+```{r session-info, echo = FALSE}
+renv_lock <- jsonlite::read_json("renv.lock")
+pander::pander(sessionInfo(package = names(renv_lock$Packages)), compact = FALSE)
+```

### 2022-06-12 supp.tex
fixed supp, added renv, added about import for algdesign per dewi's comments

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Supplmentary material for ``Current status and prospects of R-packages for the design of experiments''},
  pdfauthor={Emi Tanaka; Dewi Amaliah},
  pdfkeywords={experimental design, CRAN task view, user interface},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{colortbl}
\usepackage{bera}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage[utf8]{inputenc}
\usepackage{caption}
\renewcommand{\figurename}{Figure S}
\makeatletter
\def\fnum@figure{\figurename\thefigure}
\makeatother
\renewcommand{\tablename}{Table S}
\makeatletter
\def\fnum@table{\tablename\thetable}
\makeatother

\usepackage{bera}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{biblatex}
\addbibresource{pkgs.bib}

\title{Supplmentary material for ``Current status and prospects of R-packages for the design of experiments''}
\author{Emi Tanaka \and Dewi Amaliah}
\date{}

\begin{document}
\maketitle

\begin{figure}[htbp]

{\centering \includegraphics{figures/fig-gini-all-ctvs-1}

}

\caption{The points show the Gini index of the download counts by year facetted by CRAN task view with the color showing the number of packages. The grey line shows the distribution of the Gini index across years for all other CRAN task views. The facets are ordered by increasing value of the Gini index in 2021.}\label{fig:fig-gini-all-ctvs}
\end{figure}

\begin{longtable}[t]{l>{\raggedright\arraybackslash}p{12em}>{\raggedleft\arraybackslash}p{5em}>{\raggedleft\arraybackslash}p{5em}>{\raggedleft\arraybackslash}p{5em}>{\raggedleft\arraybackslash}p{5em}}
\caption{\label{tab:ctv-summ-table}A summary table for the CRAN task view that shows in order: the name of the task view, the full topic name, the total of packages, the total number of contributors, the average number of contributors, and the intra-connectivity. The intra-connectivity measures the percentage of packages that depends, suggest or imports at least one other package within the same task view. A low intra-connectivity suggests that development within the topic mostly occur in silos whilst high  intra-connectivity suggests that there are more interactions within the topic. The row is ordered by the average number of contributors.}\\
\toprule
Name & Topic & \# of packages & Total \# of contributors & Average \# of contributors & Intra-connectivity (\%)\\
\midrule
\endfirsthead
\caption[]{\label{tab:ctv-summ-table}A summary table for the CRAN task view that shows in order: the name of the task view, the full topic name, the total of packages, the total number of contributors, the average number of contributors, and the intra-connectivity. The intra-connectivity measures the percentage of packages that depends, suggest or imports at least one other package within the same task view. A low intra-connectivity suggests that development within the topic mostly occur in silos whilst high  intra-connectivity suggests that there are more interactions within the topic. The row is ordered by the average number of contributors. \textit{(continued)}}\\
\toprule
Name & Topic & \# of packages & Total \# of contributors & Average \# of contributors & Intra-connectivity (\%)\\
\midrule
\endhead
\midrule
\multicolumn{6}{r@{}}{\textit{(Continued on next page...)}}\
\endfoot
\bottomrule
\endlastfoot
\cellcolor{gray!6}{ExperimentalDesign} & \cellcolor{gray!6}{Design of Experiments (DoE) \& Analysis of Experimental Data} & \cellcolor{gray!6}{112} & \cellcolor{gray!6}{211} & \cellcolor{gray!6}{2.27} & \cellcolor{gray!6}{29}\\
SportsAnalytics & Sports Analytics & 78 & 149 & 2.32 & 17\\
\cellcolor{gray!6}{MedicalImaging} & \cellcolor{gray!6}{Medical Image Analysis} & \cellcolor{gray!6}{32} & \cellcolor{gray!6}{53} & \cellcolor{gray!6}{2.44} & \cellcolor{gray!6}{44}\\
MetaAnalysis & Meta-Analysis & 157 & 351 & 2.63 & 43\\
\cellcolor{gray!6}{ChemPhys} & \cellcolor{gray!6}{Chemometrics and Computational Physics} & \cellcolor{gray!6}{75} & \cellcolor{gray!6}{162} & \cellcolor{gray!6}{2.65} & \cellcolor{gray!6}{28}\\
\addlinespace
ClinicalTrials & Clinical Trial Design, Monitoring, and Analysis & 59 & 141 & 2.76 & 31\\
\cellcolor{gray!6}{Distributions} & \cellcolor{gray!6}{Probability Distributions} & \cellcolor{gray!6}{257} & \cellcolor{gray!6}{597} & \cellcolor{gray!6}{2.88} & \cellcolor{gray!6}{45}\\
Survival & Survival Analysis & 239 & 558 & 2.92 & 73\\
\cellcolor{gray!6}{ExtremeValue} & \cellcolor{gray!6}{Extreme Value Analysis} & \cellcolor{gray!6}{37} & \cellcolor{gray!6}{89} & \cellcolor{gray!6}{2.95} & \cellcolor{gray!6}{41}\\
Optimization & Optimization and Mathematical Programming & 136 & 328 & 3.04 & 29\\
\addlinespace
\cellcolor{gray!6}{TimeSeries} & \cellcolor{gray!6}{Time Series Analysis} & \cellcolor{gray!6}{339} & \cellcolor{gray!6}{793} & \cellcolor{gray!6}{3.04} & \cellcolor{gray!6}{58}\\
OfficialStatistics & Official Statistics \& Survey Statistics & 131 & 325 & 3.06 & 40\\
\cellcolor{gray!6}{Hydrology} & \cellcolor{gray!6}{Hydrological Data and Modeling} & \cellcolor{gray!6}{100} & \cellcolor{gray!6}{252} & \cellcolor{gray!6}{3.09} & \cellcolor{gray!6}{24}\\
NumericalMathematics & Numerical Mathematics & 115 & 271 & 3.14 & 63\\
\cellcolor{gray!6}{Databases} & \cellcolor{gray!6}{Databases with R} & \cellcolor{gray!6}{43} & \cellcolor{gray!6}{95} & \cellcolor{gray!6}{3.23} & \cellcolor{gray!6}{77}\\
\addlinespace
WebTechnologies & Web Technologies and Services & 201 & 428 & 3.25 & 90\\
\cellcolor{gray!6}{NaturalLanguageProcessing} & \cellcolor{gray!6}{Natural Language Processing} & \cellcolor{gray!6}{56} & \cellcolor{gray!6}{130} & \cellcolor{gray!6}{3.30} & \cellcolor{gray!6}{62}\\
Bayesian & Bayesian Inference & 213 & 621 & 3.35 & 49\\
\cellcolor{gray!6}{FunctionalData} & \cellcolor{gray!6}{Functional Data Analysis} & \cellcolor{gray!6}{40} & \cellcolor{gray!6}{109} & \cellcolor{gray!6}{3.38} & \cellcolor{gray!6}{60}\\
Robust & Robust Statistical Methods & 59 & 136 & 3.41 & 75\\
\addlinespace
\cellcolor{gray!6}{Psychometrics} & \cellcolor{gray!6}{Psychometric Models and Methods} & \cellcolor{gray!6}{230} & \cellcolor{gray!6}{567} & \cellcolor{gray!6}{3.41} & \cellcolor{gray!6}{69}\\
Tracking & Processing and Analysis of Tracking Data & 46 & 149 & 3.46 & 48\\
\cellcolor{gray!6}{Cluster} & \cellcolor{gray!6}{Cluster Analysis \& Finite Mixture Models} & \cellcolor{gray!6}{108} & \cellcolor{gray!6}{305} & \cellcolor{gray!6}{3.47} & \cellcolor{gray!6}{39}\\
Econometrics & Econometrics & 152 & 363 & 3.50 & 81\\
\cellcolor{gray!6}{Finance} & \cellcolor{gray!6}{Empirical Finance} & \cellcolor{gray!6}{158} & \cellcolor{gray!6}{426} & \cellcolor{gray!6}{3.61} & \cellcolor{gray!6}{57}\\
\addlinespace
MissingData & Missing Data & 210 & 740 & 3.99 & 42\\
\cellcolor{gray!6}{SpatioTemporal} & \cellcolor{gray!6}{Handling and Analyzing Spatio-Temporal Data} & \cellcolor{gray!6}{81} & \cellcolor{gray!6}{269} & \cellcolor{gray!6}{4.04} & \cellcolor{gray!6}{70}\\
Spatial & Analysis of Spatial Data & 197 & 618 & 4.26 & 83\\
\cellcolor{gray!6}{GraphicalModels} & \cellcolor{gray!6}{Graphical Models} & \cellcolor{gray!6}{32} & \cellcolor{gray!6}{109} & \cellcolor{gray!6}{4.38} & \cellcolor{gray!6}{78}\\
Pharmacokinetics & Analysis of Pharmacokinetic Data & 29 & 109 & 4.55 & 21\\
\addlinespace
\cellcolor{gray!6}{HighPerformanceComputing} & \cellcolor{gray!6}{High-Performance and Parallel Computing with R} & \cellcolor{gray!6}{83} & \cellcolor{gray!6}{315} & \cellcolor{gray!6}{4.75} & \cellcolor{gray!6}{63}\\
DifferentialEquations & Differential Equations & 27 & 114 & 4.96 & 56\\
\cellcolor{gray!6}{Environmetrics} & \cellcolor{gray!6}{Analysis of Ecological and Environmental Data} & \cellcolor{gray!6}{93} & \cellcolor{gray!6}{383} & \cellcolor{gray!6}{5.02} & \cellcolor{gray!6}{74}\\
MachineLearning & Machine Learning \& Statistical Learning & 102 & 488 & 5.63 & 50\\
\cellcolor{gray!6}{TeachingStatistics} & \cellcolor{gray!6}{Teaching Statistics} & \cellcolor{gray!6}{46} & \cellcolor{gray!6}{236} & \cellcolor{gray!6}{6.35} & \cellcolor{gray!6}{57}\\
\addlinespace
ReproducibleResearch & Reproducible Research & 102 & 524 & 6.49 & 76\\
\cellcolor{gray!6}{ModelDeployment} & \cellcolor{gray!6}{Model Deployment with R} & \cellcolor{gray!6}{31} & \cellcolor{gray!6}{146} & \cellcolor{gray!6}{6.55} & \cellcolor{gray!6}{74}\\*
\end{longtable}

!6\}\{ModelDeployment\} \& \cellcolor{gray!6}{Model Deployment with R} \& \cellcolor{gray!6}{31} \& \cellcolor{gray!6}{146} \& \cellcolor{gray!6}{6.55} \& \cellcolor{gray!6}{74}\textbackslash*
\textbackslash end\{longtable\}

\hypertarget{session-information}{%
\section*{Session information}\label{session-information}}
\addcontentsline{toc}{section}{Session information}

\textbf{R version 4.1.2 (2021-11-01)}

\textbf{Platform:} x86\_64-apple-darwin17.0 (64-bit)

\textbf{locale:}
en\_AU.UTF-8\textbar\textbar en\_AU.UTF-8\textbar\textbar en\_AU.UTF-8\textbar\textbar C\textbar\textbar en\_AU.UTF-8\textbar\textbar en\_AU.UTF-8

\textbf{attached base packages:}

\textbf{other attached packages:}

\begin{itemize}
\tightlist
\item
  AsioHeaders(v.1.16.1-1)
\item
  BH(v.1.75.0-0)
\item
  DBI(v.1.1.1)
\item
  DT(v.0.23)
\item
  MASS(v.7.3-57)
\item
  Matrix(v.1.4-1)
\item
  R6(v.2.5.1)
\item
  RColorBrewer(v.1.1-3)
\item
  Rcpp(v.1.0.8.3)
\item
  RcppArmadillo(v.0.10.8.1.0)
\item
  RcppEigen(v.0.3.3.9.1)
\item
  SnowballC(v.0.7.0)
\item
  V8(v.4.2.0)
\item
  anytime(v.0.3.9)
\item
  askpass(v.1.1)
\item
  assertthat(v.0.2.1)
\item
  backports(v.1.3.0)
\item
  base64enc(v.0.1-3)
\item
  base64url(v.1.4)
\item
  beeswarm(v.0.4.0)
\item
  bit(v.4.0.4)
\item
  bit64(v.4.0.5)
\item
  blob(v.1.2.2)
\item
  bookdown(v.0.24)
\item
  broom(v.0.8.0)
\item
  bslib(v.0.3.1)
\item
  cachem(v.1.0.6)
\item
  callr(v.3.7.0)
\item
  cellranger(v.1.1.0)
\item
  cli(v.3.3.0)
\item
  clipr(v.0.7.1)
\item
  codetools(v.0.2-18)
\item
  colorspace(v.2.0-3)
\item
  commonmark(v.1.7)
\item
  cpp11(v.0.4.2)
\item
  cranlogs(v.2.1.1)
\item
  cranscrub(v.0.0.0.9000)
\item
  crayon(v.1.5.1)
\item
  crosstalk(v.1.2.0)
\item
  ctv(v.0.9-2)
\item
  curl(v.4.3.2)
\item
  data.table(v.1.14.2)
\item
  dbplyr(v.2.1.1)
\item
  digest(v.0.6.29)
\item
  distributional(v.0.2.2)
\item
  dplyr(v.1.0.9)
\item
  dtplyr(v.1.1.0)
\item
  ellipsis(v.0.3.2)
\item
  evaluate(v.0.14)
\item
  fable(v.0.3.1)
\item
  fabletools(v.0.3.1)
\item
  fansi(v.1.0.3)
\item
  farver(v.2.1.0)
\item
  fastmap(v.1.1.0)
\item
  feasts(v.0.2.2)
\item
  fontawesome(v.0.2.2)
\item
  forcats(v.0.5.1)
\item
  fs(v.1.5.2)
\item
  gargle(v.1.2.0)
\item
  generics(v.0.1.2)
\item
  ggbeeswarm(v.0.6.0)
\item
  ggforce(v.0.3.3)
\item
  ggnetwork(v.0.5.10)
\item
  ggnewscale(v.0.4.5)
\item
  ggplot2(v.3.3.6)
\item
  ggraph(v.2.0.5)
\item
  ggrepel(v.0.9.1)
\item
  ggwordcloud(v.0.5.0)
\item
  glue(v.1.6.2)
\item
  googledrive(v.2.0.0)
\item
  googlesheets4(v.1.0.0)
\item
  graphlayouts(v.0.8.0)
\item
  gridExtra(v.2.3)
\item
  gtable(v.0.3.0)
\item
  haven(v.2.4.3)
\item
  here(v.1.0.1)
\item
  highr(v.0.9)
\item
  hms(v.1.1.1)
\item
  htmltools(v.0.5.2)
\item
  htmlwidgets(v.1.5.4)
\item
  httpuv(v.1.6.3)
\item
  httr(v.1.4.3)
\item
  hunspell(v.3.0.1)
\item
  ids(v.1.0.1)
\item
  igraph(v.1.3.1)
\item
  ineq(v.0.2-13)
\item
  isoband(v.0.2.5)
\item
  janeaustenr(v.0.1.5)
\item
  jquerylib(v.0.1.4)
\item
  jsonlite(v.1.8.0)
\item
  kableExtra(v.1.3.4)
\item
  knitr(v.1.37)
\item
  labeling(v.0.4.2)
\item
  later(v.1.3.0)
\item
  lattice(v.0.20-45)
\item
  lazyeval(v.0.2.2)
\item
  lifecycle(v.1.0.1)
\item
  lubridate(v.1.8.0)
\item
  magrittr(v.2.0.3)
\item
  mgcv(v.1.8-40)
\item
  mime(v.0.12)
\item
  modelr(v.0.1.8)
\item
  munsell(v.0.5.0)
\item
  network(v.1.17.2)
\item
  networkD3(v.0.4)
\item
  nlme(v.3.1-157)
\item
  numDeriv(v.2016.8-1.1)
\item
  openssl(v.2.0.2)
\item
  pacman(v.0.5.1)
\item
  pagedown(v.0.16)
\item
  parsedate(v.1.3.0)
\item
  patchwork(v.1.1.1)
\item
  pillar(v.1.7.0)
\item
  pkgconfig(v.2.0.3)
\item
  pkgsearch(v.3.1.0)
\item
  plotly(v.4.10.0)
\item
  pluralize(v.0.2.0)
\item
  plyr(v.1.8.6)
\item
  png(v.0.1-7)
\item
  polyclip(v.1.10-0)
\item
  prettyunits(v.1.1.1)
\item
  processx(v.3.5.2)
\item
  progress(v.1.2.2)
\item
  progressr(v.0.9.0)
\item
  promises(v.1.2.0.1)
\item
  ps(v.1.6.0)
\item
  purrr(v.0.3.4)
\item
  qdapRegex(v.0.7.5)
\item
  rappdirs(v.0.3.3)
\item
  readr(v.2.1.2)
\item
  readxl(v.1.3.1)
\item
  rematch(v.1.0.1)
\item
  rematch2(v.2.1.2)
\item
  renv(v.0.15.5)
\item
  reprex(v.2.0.1)
\item
  reshape(v.0.8.8)
\item
  rlang(v.1.0.2)
\item
  rmarkdown(v.2.11)
\item
  rprojroot(v.2.0.2)
\item
  rstudioapi(v.0.13)
\item
  rticles(v.0.22)
\item
  rvest(v.1.0.2)
\item
  sass(v.0.4.0)
\item
  scales(v.1.2.0)
\item
  selectr(v.0.4-2)
\item
  servr(v.0.24)
\item
  shiny(v.1.7.1)
\item
  slider(v.0.2.2)
\item
  sna(v.2.7)
\item
  sourcetools(v.0.1.7)
\item
  splitstackshape(v.1.4.8)
\item
  statnet.common(v.4.6.0)
\item
  stringi(v.1.7.6)
\item
  stringr(v.1.4.0)
\item
  svglite(v.2.0.0)
\item
  sys(v.3.4)
\item
  systemfonts(v.1.0.3)
\item
  tarchetypes(v.0.6.0)
\item
  targets(v.0.12.0)
\item
  tibble(v.3.1.7)
\item
  tidygraph(v.1.2.0)
\item
  tidyr(v.1.2.0)
\item
  tidyselect(v.1.1.2)
\item
  tidytext(v.0.3.2)
\item
  tidyverse(v.1.3.1)
\item
  tinytex(v.0.36)
\item
  tokenizers(v.0.2.1)
\item
  tsibble(v.1.1.0)
\item
  tweenr(v.1.0.2)
\item
  tzdb(v.0.2.0)
\item
  utf8(v.1.2.2)
\item
  uuid(v.1.0-3)
\item
  vctrs(v.0.4.1)
\item
  vipor(v.0.4.5)
\item
  viridis(v.0.6.2)
\item
  viridisLite(v.0.4.0)
\item
  visNetwork(v.2.1.0)
\item
  vroom(v.1.5.7)
\item
  warp(v.0.2.0)
\item
  webshot(v.0.5.2)
\item
  websocket(v.1.4.1)
\item
  withr(v.2.5.0)
\item
  xfun(v.0.29)
\item
  xml2(v.1.3.2)
\item
  xtable(v.1.8-4)
\item
  yaml(v.2.3.5)
\end{itemize}

\textbf{loaded via a namespace (and not attached):}

\begin{itemize}
\tightlist
\item
  utils(v.4.1.2)
\item
  DiceDesign(v.1.9)
\item
  tools(v.4.1.2)
\item
  compiler(v.4.1.2)
\item
  datasets(v.4.1.2)
\item
  base(v.4.1.2)
\item
  grDevices(v.4.1.2)
\item
  grid(v.4.1.2)
\item
  methods(v.4.1.2)
\item
  pander(v.0.6.5)
\item
  graphics(v.4.1.2)
\item
  BsMD(v.2020.4.30)
\item
  stats(v.4.1.2)
\item
  rsm(v.2.10.3)
\end{itemize}

\printbibliography

\end{document}

### 2022-06-12 _targets.R

@@ -112,6 +112,7 @@ list(
   
   
   tar_render(report, "paper.Rmd"),
+  tar_render(supp, "supp.Rmd"),
   tar_render(arxiv, "paper/arxiv/arxiv.Rmd"),
   NULL
 )

### 2022-06-13 paper.Rmd
rechecked up to Section 3.1 (inclusive)

@@
 We observe that experimental designs in practice appear to be sufficiently manufactured by a small number of packages and the development of experimental designs occur in silos.
-We discuss also the interface designs of widely utilised R packages in the field of experimental design and discuss its future prospects for the field.
+We discuss also the interface designs of widely utilised R packages in the field of experimental design and discuss its future prospects to advance the field in practice.
 bibliography: paper.bib
 knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'index.html')) })
-keywords: ["experimental design", "CRAN task view", "user interface", "application programming interface", "software design"]
+keywords: ["experimental design", "CRAN task view", "interface design"]
 output: 
-  #bookdown::html_document2: 
-  #  theme: paper
-  # below is to have a JSS paper style template
   pagedown::html_paged:
@@
 
-```{css, echo = FALSE}
-h1, h2, h3, div > * {
- font-family: "Helvetica Neue",Helvetica,Arial,sans-serif;
-}
-```
-
 ```{r setup, cache = FALSE, include = FALSE}
@@
 library(glue)
+library(ggwordcloud)
 library(tidyverse)
@@
 tar_load(npkgs)
-tar_load(gini_yearly)
+tar_load(gini_yearly_doe)
 tar_load(gini_yearly_ctv)
@@
 A carefully crafted data collection scheme is therefore critical to optimise the information from data.
-Of all the data collection schemes, the field of experimental designs is specifically devoted to planning an experiment largely based on the founding principles by @Fisher1935-qc or an optimisation framework like those described in @Pukelsheim2006-sv.
-These experimental designs are often constructed with the aid of a statistical software, such as R [@R], so software usage can aid in understanding the practice of experimental design.
-Given the large gain that can be made from collecting higher quality experimental data, we benefit from understanding the current state of the experimental designs in practice.
+The field of experimental designs is specifically devoted to planning the collection of experimental data largely based on the founding principles by @Fisher1935-qc or an optimisation framework like those described in @Pukelsheim2006-sv.
+These experimental designs are often constructed with the aid of a statistical software, such as R [@R], so the usage of experimental design software can inform us about some aspects of experimental designs in practice.
 
@@
 Methods for data collection can be dichotomised by the type of data collected -- namely, experimental or observational -- or alternatively, categorised as experimental design (including quasi-experimental design) or survey design.
-This dichotomisation, to a great extent, is seen in the [Comprehensive R Archive Network (CRAN) task views](https://cran.r-project.org/web/views/) (a volunteer maintained list of R-packages by topic) where R-packages in experimental design are in `r ctv("ExperimentalDesign")` task view and R-packages in survey designs are in `r ctv("OfficialStatistics")` task view.
-The full list of available topics can be see in Table S1 in the Supplementary Materials and also accessed via the `ctv` package [@ctv].
+This dichotomisation, to a great extent, is seen in the [Comprehensive R Archive Network (CRAN) task views](https://cran.r-project.org/web/views/) (a volunteer maintained list of R-packages by topic) where R-packages for experimental design are in `r ctv("ExperimentalDesign")` task view and R-packages for survey designs are in `r ctv("OfficialStatistics")` task view.
+The full list of available topics can be see in Table S1 in the Supplementary Materials.
 A subset of experimental designs are segregated into `r ctv("ClinicalTrials")` task view, where the focus is on clinical trials with primary interest in sample size calculations.
@@
 The immense popularity of `tidyverse` [a collection of R-packages for various stages of data analysis that place enormous emphasis on the interface design by @Wickham2019-mj] is a testament to the impact an interface design can have in practice.
+The practice of experimental design could be advanced by adopting similar interface design principles across DoE packages.
 
 The paper is organised as follows.
-Section `r ref("data")` briefly describes the data source used for the analysis in Section `r ref("eda")`; Section `r ref("eda")` presents some insights into the state of the current DoE packages by exploratory data analysis of package download data, text descriptions and comparisons with other CRAN task views; Section `r ref("design")` discuss the interface designs of DoE packages, and we conclude with a discussion in Section `r ref("discussion")` of future prospects.
+Section `r ref("data")` briefly describes the data source used for the analysis in Section `r ref("eda")`; Section `r ref("eda")` presents some insights into the state of the current DoE packages by exploratory data analysis of package download data, text descriptions and comparisons with other CRAN task views; Section `r ref("design")` discuss the interface designs of widely used DoE packages, and we conclude with a discussion in Section `r ref("discussion")` of future prospects in the software development of experimental designs.
 
@@
 
-To study the DoE packages, we analyse data using three sources of data as described next.
+To study the DoE packages, we analyse data using three sources of data as described below.
 
@@
 The summary data can be easily accessed with the `cranlogs` package [@cranlogs].
-This paper uses the data from the beginning of 2013 to end of 2021 (a total of 9 years) for the DoE packages.
+This paper uses the data from the beginning of 2013 to end of 2021 (a total of 9 years) for the packages in the CRAN task views.
 
@@
 
+CRAN task views are volunteer maintained list of R-packages on CRAN relevant to the corresponding topic.
+There are a total `r n_distinct(gini_yearly_ctv$ctv)` CRAN task views.
+Table S1 in the Supplementary Materials list the full available topics from the `ctv` package [@ctv].
 The list of packages in each CRAN task views (as of `r as.Date(file.info("_targets/objects/ctv_pkgs_data")$mtime)`) are used to contrast the characteristics of DoE packages in Section `r ref("silo")`.
@@
 
-All results presented are derived from exploratory data analysis of observational data; consequently, all interpretations are somewhat speculative and 
-may not be indicative of the true state of the field of experimental design.
+In this section, we derive some conjecture based on the analysis of the data described in Section `r ref("data")`.
+All results presented are from the exploratory data analysis of observational data, consequently, all interpretations are somewhat speculative and may not be indicative of the true state of the field of experimental design.
 In particular, any analysis over time are confounded by the fact that the nature of users and package management have also changed over the years.
@@
 A subset of DoE packages is not primarily about design of experiments but about the analysis of experimental data.
-A complete delineation of these packages is difficult as many have some functions that can aid decisions or constructions of experimental designs (and any categorisation is prone to our subjective bias) so we opted not to remove any DoE packages in the analysis.
+A complete delineation of these packages is difficult as there is almost always at least one function that can aid decisions or constructions of experimental designs (and any categorisation is prone to our subjective bias) so we opted not to remove any DoE packages in the analysis.
 
-## Small, but diverse, set of packages are sufficient for most designs {#popular}
+## Small, but diverse, set of packages are sufficient for most experimental designs in practice {#popular}
 
@@
 
-```{r plot-lorenz, fig.height = 4, fig.width = 4, fig.cap = "Lorenz curve of the total download count for DoE packages in 2021. The red line corresponds to the line of perfect equality."}
+```{r plot-lorenz, fig.height = 4, fig.width = 4, fig.cap = "Lorenz curve of the total download count for DoE packages in 2021.
+The red line corresponds to the line of perfect equality.
+The green region shows the area under the Lorenz curve and the red region shows the area of gap in equality."}
 tar_read(plot_lorenz2021) +
@@
 A Gini index of 0\% indicates equality in downloads across packages while a value of 100\% indicates maximal inequality (all downloads are due to one package).
-In Figure `r ref("fig:download-share")`, we see that the distributions of the downloads each year have a heavy right tail with the Gini index ranging from `r scales::percent(min(gini_yearly$gini), 0.1)` to `r scales::percent(max(gini_yearly$gini), 0.1)` across the years 2013 to 2021, indicating that there is a high level of inequality of downloads across packages, particularly with more pronounced inequality in the last 5 years.
+In Figure `r ref("fig:download-share")`, we see that the distributions of the package downloads each year have a heavy right tail with the Gini index ranging from `r scales::percent(min(gini_yearly_doe$gini), 0.1)` to `r scales::percent(max(gini_yearly_doe$gini), 0.1)` across the years 2013 to 2021, indicating that there is a high level of inequality of package downloads, particularly with more pronounced inequality in the last 6 years.
 
@@
 The label on the bottom of the plot shows the Gini index for downloads and the number of packages with full year of download count in the corresponding year.
-In the last 5 years, the Gini index is consistent above 60\\% each year indicating that most downloads are due to a relatively small number of packages."}
-tar_read(plot_download_dist)
+In the last 6 years, the Gini index is consistent above 60\\% each year indicating that most downloads are due to a relatively small number of packages.
+"}
+tar_read(plot_download_dist) +
+  labs(y = "Number of downloads")
 ```
 
-While in absolute terms the Gini index is high for `r ctv("ExperimentalDesign")`, the inequality is not as severe as other CRAN task views as shown in Figure `r ref("fig:fig-gini-all-ctvs")`.
-We can see in Figure `r ref("fig:fig-gini-all-ctvs")` that the Gini index is generally increasing for DoE packages (as is generally the case for other CRAN task views as shown in Figure S1 in the Supplementary Materials).
+The increase in the number of packages that are not highly downloaded may mean that **_there are more packages to construct niche experimental designs_**.
+Some examples of these packages include `qtlDesign`, `PwrGSD` and `Crossover` made for QTL experiments, group sequential designs and crossover trials, respectively.
+These packages would naturally have a smaller number of potential users.
+Counterfactual to this, the increase could be due to other external factors, such as an increase in the number of skilled developers (and thus more pacakge contributions), a change in CRAN policy or management to add packages (either to CRAN and/or task view), and/or that new packages are still yet to amass users.
+While there is an argument that low download counts are due to the low utility and/or quality of the packages, packages in CRAN task views are selected by expert maintainers, thus we can reasonably assume that any package listed in CRAN task view are of a decent utility and quality.
 
-```{r fig-gini-all-ctvs, fig.height = 4, fig.width = 4, fig.cap = "The points show the Gini index of the download counts by year for ExperimentalDesign CRAN task view with the color showing the number of packages.
+If the downloads are reflective of the experimental designs used in practice, **_small set of packages appear to be sufficient for most to construct the full set of designs of experiments needed in practice_**.
+Packages of course evolve and the top downloaded packages have had regular updates that may have broadened its scope from its previous releases.
+
+While in absolute terms the Gini index is high for `r ctv("ExperimentalDesign")` task view (`r scales::percent(min(gini_yearly_doe$gini), 0.1)` to `r scales::percent(max(gini_yearly_doe$gini), 0.1)`), the inequality is not as severe as other CRAN task views as shown in Figure `r ref("fig:fig-gini-all-ctvs")`.
+We can see in Figure `r ref("fig:fig-gini-all-ctvs")` that the Gini index is generally increasing for DoE packages (as is generally the case for other CRAN task views as shown in Figure S1 in the Supplementary Materials) but most other CRAN task views have a Gini index of over 75%.
+This suggests that other CRAN task views may have dominant standards and in comparison to other topics, there are more **_diverse approaches to designing experiments_** and thus no single DoE package is dominant.
+This observation however doesn't take into account other approaches to generating experimental designs, like the proprietary software, CycDesignN [@cycdesignn] and SAS [@sas1985sas], which may be widely used.
+
+```{r fig-gini-all-ctvs, fig.height = 4, fig.width = 4, fig.cap = "The points show the Gini index of the download counts by year for ExperimentalDesign task view with the color showing the number of packages.
 The grey line shows the distribution of the Gini index across years for all other CRAN task views.
-See Figure S2 in the Supplementary Material for the line graph for the Gini index across years by each CRAN task view."}
+See Figure S1 in the Supplementary Material for the line graph for the Gini index across years by each CRAN task view."}
 gini_yearly_ctv %>% 
@@
 
-From Figures `r ref("fig:download-share")` and `r ref("fig:fig-gini-all-ctvs")`, we could draw some indicative states of the field of experimental designs with possible counterfactual interpretations:
-
-* The increase in the number of packages that are not highly downloaded may mean that **_there are more packages to construct niche experimental designs_**.
-Some examples of these packages include `qtlDesign`, `PwrGSD` and `Crossover` made for QTL experiments, group sequential designs and crossover trials, respectively.
-These packages would naturally have a smaller number of potential users.
-Counterfactual to this, the increase could be due to other external factors, such as an increase in a number of skilled contributors, a change in CRAN policy or management to add packages (either to CRAN and/or task view), and/or that new packages are still yet to amass users.
-While there is an argument that low download counts are due to the low utility and/or quality of the packages, packages in CRAN task views are selected by expert maintainers; thus we can reasonably assume that any package listed in CRAN task view are of a decent utility and quality.
-* If the downloads are reflective of the experimental designs used in practice, **_small set of packages appear to be sufficient for most to construct the full set of designs of experiments needed in practice_**.
-Packages of course evolve and the top downloaded packages have had regular updates that may have broadened its scope.
-* While small subset of DoE packages are most frequently used, none of these packages are dominant in the field (judging from comparisons with other CRAN task views).
-This suggests that there are **_diverse approaches to designing experiments_**.
-This observation doesn't take into account other approaches to generating experimental designs, like the proprietary software, CycDesignN [@cycdesignn] and SAS [@sas1985sas], which may be widely used.
-
-## There is a lack of adaptation of new or innovative designs  {#ranking}
+## There is a lack of adaptation of new or innovative experimental designs  {#ranking}
 
@@
 
-```{r rank-over-time, fig.width = 5.5, fig.height = 4.5, fig.cap = "The plot shows the rank of top 10 downloaded packages by year. "}
-tar_read(plot_rank_dist)
+```{r rank-over-time, fig.width = 5.5, fig.height = 4.5, fig.cap = "The plot shows the rank of top 10 downloaded packages by year.
+Packages that do not appear in top 10 for at least two periods are omitted from the plot.
+Most packages are consistently in the top 10 for the period shown."}
+tar_read(plot_rank_dist) + labs(color = "Package")
 ```
@@
 The blue line corresponds to the least squares fit of a simple linear regression model.
-The label in the right-hand upper corner shows the sample correlation coefficient between the first release date and the log, with base 10, of the total download count.
+The label in the right-hand upper corner shows the sample correlation coefficient between the first release date and the log (with base 10) of the total download count.
 The high-leverage point on the far left belongs to `conf.design`, authored by one of the earlier contributors to R."}
@@
 
-```{r wordcloud-over-time, fig.width = 5.5, fig.height = 5, fig.cap = "The above figure shows the word cloud of bigrams from the title and descriptions of the CRAN packages.
+```{r wordcloud-over-time, fig.width = 5.5, fig.height = 5, fig.cap = "The above figure shows the word cloud of bigrams from the title and descriptions of the DoE packages.
 The size shows how often the bigram appears across the DoE packages and the color are relative to the total download count in 2021 of the packages that contain the bigram."}
@@
 In this section, we discuss these interface designs broadly, with examples illustrated from the top downloaded packages (shown in Figure `r ref("fig:rank-over-time")`).
-We exclude `ez` and `DiceKriging` from the discussion as the former  is predominately visualisation of experimental data and latter is about the analysis of computer experiments in addition to belonging to the same suite of packages as `DiceDesign`.
-
-`DoE.wrapper` [@DoE.wrapper]
+We exclude `ez` and `DiceKriging` from the discussion as the former  is predominately visualisation of experimental data and latter is about the analysis of computer experiments in addition to belonging to the same suite of packages as `DiceDesign` [@DiceDesign].
 
@@
 Computer experiments, as discussed in Section `r ref("topics")`, which generally involve space filling designs, have been on the rise.
-The exemplar for this type of designs are `lhs` and `DiceDesign`.
+The exemplar for this type of designs are `lhs` [@lhs] and `DiceDesign`.
 For the `lhs` package, functions like `randomLHS()`, `optimumLHS()`, and `maximinLHS()`, require users to specify the sample size ($n$) and the number of variables ($p$) then it generates a Latin hypercube sample, based on different optimisation schemes (in this case, random, S optimal and maxmin criteria, respectively; see package documentation for more details).
@@
 The construction of some of these designs are catalogue-based, e.g.
-`oa.design()` in `DoE.base` and the entire package `FrF2.catlg128` [@FrF2.catlg128] is a list of fractional 2-level factorial designs.
+`oa.design()` in `DoE.base`.
 
@@
 
+`r if(knitr::is_latex_output()) '<!--'`
+
+# Supplemental materials
+
+The supplementary materials can be found [here](supp.pdf).
+
+`r if(knitr::is_latex_output()) '-->'`
+
 # Acknowledgement
 
-This paper uses the `targets` framework [@targets] and `renv` [@renv] for reproducibility, `knitr` [@knitr] and `rmarkdown` [@rmarkdown] for creating reproducible documents, `ggplot2` [@ggplot2], `ggraph` [@ggraph], `ggwordcloud` [@ggwordcloud] and `colorspace` [@colorspace] for visualisation, `kableExtra` [@kableExtra] for customising the table in the Supplementary Material and `tidyverse` [@Wickham2019-mj], `tidytext` [@tidytext], `pluralize` [@pluralize], and `ineq` [@ineq] for data processing and manipulation, and `cranlogs` [@cranlogs] and `ctv` [@ctv] for extracting data.
+This paper uses `targets` [@targets] and `renv` [@renv] for reproducibility, `knitr` [@knitr] and `rmarkdown` [@rmarkdown] for creating reproducible documents, `ggplot2` [@ggplot2], `ggraph` [@ggraph], `ggwordcloud` [@ggwordcloud] and `colorspace` [@colorspace] for visualisation, `kableExtra` [@kableExtra] for customising the table in the Supplementary Material and `tidyverse` [@Wickham2019-mj], `tidytext` [@tidytext], `pluralize` [@pluralize], and `ineq` [@ineq] for data processing and manipulation, and `cranlogs` [@cranlogs] and `ctv` [@ctv] for extracting data.
 All code to reproduce this paper is found at `r url("https://github.com/emitanaka/paper-DoE-review")`.

### 2022-06-13 R/plot.R

@@ -103,20 +103,26 @@ plot_rank_distribution <- function(data, nrank = 10, stat) {
   dat <- data %>% 
     mutate(year = as.factor(year)) %>% 
     filter(rank <= nrank)
-  g <- ggplot(dat, aes(x = year, y = {{stat}}, color = package, group = package)) + 
+  pkgs_repeat_toprank <- dat %>% 
+    count(package) %>% 
+    filter(n > 1) %>% 
+    pull(package)
+  g <- dat %>% 
+    filter(package %in% pkgs_repeat_toprank) %>% 
+    ggplot(aes(x = year, y = {{stat}}, color = package, group = package)) + 
     geom_point() + 
     geom_line() +
     labs(x = "Year", y = "Download rank by year") +
     scale_color_manual(values = c("#E69F00", "#56B4E9", "#009E73", 
                                   "#F0E442", "#0072B2", "#D55E00", 
                                   "#CC79A7", "#000000", "#888888",
-                                  "#88CCEE", "#CC6677", "#DDCC77", 
-                                  "#117733", "#DDAA33", "#BB5566"),
-                       breaks = dat %>% 
-                         filter(year == 2021) %>% 
+                                  "#88CCEE", "#CC6677", "seagreen4", 
+                                  "violetred4"),
+                       breaks = data %>% 
+                         filter(package %in% pkgs_repeat_toprank,
+                                year == 2021) %>% 
                          arrange(desc(total)) %>% 
-                         pull(package) %>% 
-                         union(dat$package))
+                         pull(package))
   if(rlang::as_string(rlang::ensym(stat))=="rank") {
     g + scale_y_reverse(breaks = 1:nrank)
   } else {

### 2022-06-13 _targets.R

@@ -63,17 +63,18 @@ list(
   tar_target(top2yrs, cran_rank2$package[1:ntop]),
   tar_target(top5yrs, cran_rank5$package[1:ntop]),
   tar_target(updates_duration_5, pkg_updates_duration(top2yrs, duration = 5)),
-  tar_target(gini_yearly, gini_coef(cran_rank_yearly)), 
   tar_target(gini_yearly_ctv, gini_coef_by_ctv(cran_rank_yearly_all)), 
+  tar_target(gini_yearly_doe, gini_yearly_ctv %>% 
+               filter(ctv == "ExperimentalDesign")),
   tar_target(plot_lorenz2021, plot_lorenz_curve(cran_rank_yearly, 2021)), 
   tar_target(plot_download_dist_all, plot_download_distribution(cran_rank_yearly_all, gini_yearly_ctv) + 
                plot_setup() + facet_wrap(~ctv, ncol = 4)),
-  tar_target(plot_download_dist, plot_download_distribution(cran_rank_yearly, gini_yearly) + plot_setup()),
+  tar_target(plot_download_dist, plot_download_distribution(cran_rank_yearly, gini_yearly_doe) + plot_setup()),
   tar_target(plot_rank_dist, plot_rank_distribution(cran_rank_yearly, stat = rank) + plot_setup()),
   tar_target(plot_rank_download_dist, plot_rank_distribution(cran_rank_yearly, stat = total) + plot_setup()),
   tar_target(plot_download_trend_5, plot_download_trend(cran_download_data5, updates_duration_5) + plot_setup()),
   ## scrub data
-  tar_target(plot_scrub_download_dist, plot_download_distribution(cran_scrub_rank_yearly, gini_yearly) + plot_setup()),
+  tar_target(plot_scrub_download_dist, plot_download_distribution(cran_scrub_rank_yearly, gini_yearly_doe) + plot_setup()),
   tar_target(plot_scrub_rank_dist, plot_rank_distribution(cran_scrub_rank_yearly, stat = rank) + plot_setup()),
   tar_target(plot_scrub_trend, plot_download_trend(cran_scrub_data5, updates_duration_5) + plot_setup()),
   #tar_target(plot_scrub_trend_middle, plot_download_trend(cran_scrub_data5, updates_duration_middle) + plot_setup()),

### 2022-06-14 paper.Rmd
Section 3.3 redrafted

@@
 
-## There is a lack of adaptation of new or innovative experimental designs  {#ranking}
+## The field of experimental design is slow-changing  {#ranking}
 
@@
 
-Figure `r ref("fig:established-packages")` shows a moderate negative correlation between the first release date and the (log of) total download counts of DoE packages in any given year from 2013 to 2021.
-This suggest that in general, a package released earlier are more likely to be used today (possibly for legacy reasons or the general inertia for adoption of new packages).
-We can also see in Figure `r ref("fig:established-packages")` the most downloaded packages were generally released in 2004 to 2010.
+Figure `r ref("fig:release-date-vs-download")` shows a moderate negative correlation between the first release date and the (log of) total download counts of DoE packages in any given year from 2013 to 2021.
+This suggest that in general, a package released earlier are more likely to be used today (possibly for legacy reasons or the general inertia to adopt new packages).
+We can also see in Figure `r ref("fig:release-date-vs-download")` the most downloaded packages were generally released in 2004 to 2010.
 
-```{r established-packages, fig.height = 5, fig.cap = "The above figure shows the total download (in log scale) of a package in the corresponding year against the first release date of the package.
+```{r release-date-vs-download, fig.height = 5, fig.cap = "The above figure shows the total download (in log scale) of a package in the corresponding year against the first release date of the package.
 The blue line corresponds to the least squares fit of a simple linear regression model.
@@
 
-The lack of change in the top ranking packages are indicative that there has been generally little need for new or innovative experimental design needs for the mass.
-We do however also see that the the top downloaded packages generally have more updates so it is possible that the packages have improved or broadened the scope of its usage.
+The consistency in the top 10 ranking packages (Figure `r ref("fig:rank-over-time")`) and the fact that most downloaded DoE packages are first released more than 10 years ago (Figure `r ref("fig:release-date-vs-download")`) indicates that either the existing packages filled the needs of mass in practice or no new packages were compelling for many to switch their practice.
+We do however also see that the the top downloaded packages generally have more updates (see Figure `r ref("fig:release-date-vs-download")`) so it is possible that the packages have improved or broadened the scope of its usage.
 
-## Computer experiments and optimal designs are of interest {#topics}
+## Optimal designs are of interest {#topics}
 
 Figure `r ref("fig:wordcloud-over-time")` shows some common purposes of DoE packages, based on bigrams in the package title and description.
-A bigram is only shown as the frequency of single word was not insightful and there were not many trigrams common across packages.
+We show only the bigrams as unigram was not insightful and there were not many trigrams common across packages.
+In counting the bigrams, we processed the text data as follows:
+
+1. We standardised the words to lower case and removed pluralization. 
+2. Multiple mentions of the same bigram within a package was counted as one (e.g. `AlgDesign` mentions "experimental design" four times in the title and the description but this is counted as one). 
+3.
+Bigrams that consist the stop words (provided in `tidytext::stop_words` in addition to other words we deemed irrelevant, e.g.
+"provide", "e.g.", "calculate" and so on; full list is shown in the code provided in the link under Section `r ref("pkgs")`) are removed.
+
 Not surprisingly, the bigram "experimental design" was the most common.
-More interestingly, "optimal design" appeared across different packages (indicated by the size of the word in Figure `r ref("fig:wordcloud-over-time")`), and the bigrams "latin hypercube" and "computer experiment" used across a few packages that are downloaded frequently (indicated by the color of the word in Figure `r ref("fig:wordcloud-over-time")`).
-Although there exists a separate `r ctv("ClinicalTrials")` task view, the DoE packages clearly include some packages that are of interest to clinical trials as shown by the size of the bigram "clinical trial" in Figure `r ref("fig:wordcloud-over-time")`.
+More interestingly, "optimal design" and "sequential design" appeared across different packages (indicated by the size of the word in Figure `r ref("fig:wordcloud-over-time")`), and the bigrams "latin hypercube" and "computer experiment" are used across a few packages that are downloaded frequently (indicated by the color of the word in Figure `r ref("fig:wordcloud-over-time")`).
+Sequential design, Latin hypercube sampling and computer experiments (which generally include space filling designs like Latin hypercube sampling) are designs that generally operate by optimising a user-selected criterion and can be classified as optimal designs.
 
@@
 ```
+
+Although there exists a separate `r ctv("ClinicalTrials")` task view, the DoE packages clearly include some packages that are of interest to clinical trials as shown by the size of the bigram "clinical trial" (and related bigrams like "dose finding" and "phase ii") in Figure `r ref("fig:wordcloud-over-time")`.
 
@@
 
-# Acknowledgement
+# Acknowledgement {#pkgs}
 

### 2022-06-14 paper.Rmd
section 3 redraft done

@@
 
-Figure `r ref("fig:ctv-summ-plot")` shows that the `r ctv("ExperimentalDesign")` have the lowest average number of contributors among all CRAN task views.
+Figure `r ref("fig:ctv-summ-plot")` shows that the `r ctv("ExperimentalDesign")` task view has the lowest average number of contributors among all `r n_distinct(gini_yearly_ctv$ctv)` CRAN task views.
 In addition, we can also see in Figure `r ref("fig:ctv-summ-plot")`  that the `r ctv("ExperimentalDesign")` task view has one of the least intra-connectivity (the percentage of packages that make use of other packages within the same task view).
 The full connection of DoE packages with other DoE packages is shown in Figure `r ref("fig:plot-doe-network")`.
-These observations suggest that the field of experimental design is one of the least collaborative field and package development generally occur in silos.
+These observations suggest that the **_field of experimental design is one of the least collaborative field_** and package development generally occur in silos.
 
@@
 ```{r plot-doe-network, fig.width = 6, fig.height = 6, fig.cap  = "Package connections (depends, suggests and imports) within the DoE packages.
+The direction of the arrow shows the connection of packags where the package on the tail of the arrow is a depedency, suggestion or import for the package on the head of the arrow.
 DoE packages that do not depend, suggest or import another DoE package is not shown."}
@@
 In this section, we discuss these interface designs broadly, with examples illustrated from the top downloaded packages (shown in Figure `r ref("fig:rank-over-time")`).
-We exclude `ez` and `DiceKriging` from the discussion as the former  is predominately visualisation of experimental data and latter is about the analysis of computer experiments in addition to belonging to the same suite of packages as `DiceDesign` [@DiceDesign].
+We exclude `ez` and `DiceKriging` from the discussion as the former  is predominately visualisation of experimental data and the latter is about the analysis of computer experiments in addition to belonging to the same suite of packages as `DiceDesign` [@DiceDesign].
 

### 2022-06-15 paper.Rmd
done

@@
     Re-running an experiment is generally costly and in some cases impossible due to limited resources, so the design of an experiment plays a critical role in increasing the quality of experimental data.
-In this article we describe the current state of the R-packages for the design of experiments through an exploratory data analysis of package downloads, package metadata, and the comparison of characteristics with other topics.
-We observe that experimental designs in practice appear to be sufficiently manufactured by a small number of packages and the development of experimental designs occur in silos.
+In this paper we describe the current state of the R-packages for the design of experiments through an exploratory data analysis of package downloads, package metadata, and the comparison of characteristics with other topics.
+We observe that experimental designs in practice appear to be sufficiently manufactured by a small number of packages and the development of experimental designs often occur in silos.
 We discuss also the interface designs of widely utilised R packages in the field of experimental design and discuss its future prospects to advance the field in practice.
@@
 tar_load(gini_yearly_ctv)
+tar_load(doe_lorenz2021)
+
+lorenz_value <- function(p) {
+  doe_lorenz2021 %>% 
+    mutate(pdist = abs(p - .env$p)) %>% 
+    filter(pdist == min(pdist)) %>% 
+    pull(L) %>% 
+    magrittr::multiply_by(100) %>% 
+    round()
+}
 ```
@@
 A carefully crafted data collection scheme is therefore critical to optimise the information from data.
-The field of experimental designs is specifically devoted to planning the collection of experimental data largely based on the founding principles by @Fisher1935-qc or an optimisation framework like those described in @Pukelsheim2006-sv.
-These experimental designs are often constructed with the aid of a statistical software, such as R [@R], so the usage of experimental design software can inform us about some aspects of experimental designs in practice.
+The field of experimental designs is specifically devoted to planning the collection of experimental data, largely based on the founding principles by @Fisher1935-qc or an optimisation framework like those described in @Pukelsheim2006-sv.
+These experimental designs are often constructed with the aid of a statistical software, such as R [@R], Python [@python], SAS [@sas1985sas], and so on, so the usage of experimental design software can inform us about some aspects of experimental designs in practice.
 
@@
 The sheer quantity and variation of experimental designs in the R-packages are arguably unmatched with any other programming languages, e.g.
-in Python [@python], only a handful of packages that generate design of experiment exist (namely `pyDOE`, `pyDOE2`, `dexpy`, `experimenter` and `GPdoemd`) with limited type of designs.
+in Python, only a handful of packages that generate design of experiment exist (namely `pyDOE`, `pyDOE2`, `dexpy`, `experimenter` and `GPdoemd`) with limited type of designs.
 Thus, the study of DoE packages, based on quantitative and qualitative data, can give us an objective view to the state of the current experimental designs in practice.
@@
 The paper is organised as follows.
-Section `r ref("data")` briefly describes the data source used for the analysis in Section `r ref("eda")`; Section `r ref("eda")` presents some insights into the state of the current DoE packages by exploratory data analysis of package download data, text descriptions and comparisons with other CRAN task views; Section `r ref("design")` discuss the interface designs of widely used DoE packages, and we conclude with a discussion in Section `r ref("discussion")` of future prospects in the software development of experimental designs.
+Section `r ref("data")` briefly describes the data source used for the analysis in Section `r ref("eda")`; Section `r ref("eda")` presents some insights into the state of the current DoE packages by the exploratory data analysis of package download data, text descriptions and comparisons with other CRAN task views; Section `r ref("design")` discusses the interface designs of widely used DoE packages, and we conclude with a discussion in Section `r ref("discussion")` of future prospects in the software development of experimental designs.
 
@@
 
-All CRAN packages have a title, description, package connections (suggests, dependency and imports) and other meta-information in the DESCRIPTION file.
+All CRAN packages have a title, description, package connections (suggests, depends and imports of other packages) and other meta-information in the DESCRIPTION file.
 We use the text data from the title and description (as accessed in `r as.Date(file.info("_targets/objects/ngram_download_data")$mtime)`) in Section `r ref("topics")`.
@@
 Table S1 in the Supplementary Materials list the full available topics from the `ctv` package [@ctv].
-The list of packages in each CRAN task views (as of `r as.Date(file.info("_targets/objects/ctv_pkgs_data")$mtime)`) are used to contrast the characteristics of DoE packages in Section `r ref("silo")`.
+The list of packages in each CRAN task views (as of `r as.Date(file.info("_targets/objects/ctv_pkgs_data")$mtime)`) are used to contrast the characteristics of DoE packages in Sections `r ref("popular")` and `r ref("silo")`.
 
@@
 
-There have been at least 50 DoE packages since 2013 but most of the downloads are concentrated in just a handful of packages.
-For example, Figure `r ref("fig:plot-lorenz")` shows a Lorenz curve [@Lorenz1905-tc] for the total package downloads in 2021 of `r tar_read(gini_yearly) %>% filter(year==2021) %>% pull(n)` DoE packages (first released prior to 2021); we can see from Figure `r ref("fig:plot-lorenz")` that bottom 90\% of DoE packages (in terms of total download count in 2021) only share about 30\% of total downloads across all DoE packages -- in another words, 70\% of the total downloads are due to 11 packages (10% of the DoE packages).
+There are at least 50 DoE packages since 2013 but most of the downloads are concentrated in just a handful of packages.
+For example, Figure `r ref("fig:plot-lorenz")` shows a Lorenz curve [@Lorenz1905-tc] for the total package downloads in 2021 of `r tar_read(gini_yearly) %>% filter(year==2021) %>% pull(n)` DoE packages (first released prior to 2021); we can see from Figure `r ref("fig:plot-lorenz")` that bottom 90\% of DoE packages (in terms of total download count in 2021) only share about `r lorenz_value(0.9)`\% of total downloads across all DoE packages -- in another words, `r 100 - lorenz_value(0.9)`\% of the total downloads are due to `r round(npkgs * 0.1)` packages (10% of the DoE packages).
 
@@
 The red line corresponds to the line of perfect equality.
-The green region shows the area under the Lorenz curve and the red region shows the area of gap in equality."}
+The yellow region shows the area under the Lorenz curve and the red region shows the area of gap in equality."}
 tar_read(plot_lorenz2021) +
@@
 
-```{r download-share, fig.width = 5.5, fig.height = 4, fig.cap = "The distribution of the number of downloads for each DoE package by year.
+```{r download-share, fig.width = 5.5, fig.height = 4, fig.cap = "The distribution of the number of downloads for DoE packages by year.
 Packages were removed in any year if it was released in that year or later so that each download count is for the full year.
 The label on the bottom of the plot shows the Gini index for downloads and the number of packages with full year of download count in the corresponding year.
-In the last 6 years, the Gini index is consistent above 60\\% each year indicating that most downloads are due to a relatively small number of packages.
+In the last 6 years, the Gini index is consistent above 60\\% indicating that most downloads are due to a relatively small number of packages.
 "}
@@
 These packages would naturally have a smaller number of potential users.
-Counterfactual to this, the increase could be due to other external factors, such as an increase in the number of skilled developers (and thus more pacakge contributions), a change in CRAN policy or management to add packages (either to CRAN and/or task view), and/or that new packages are still yet to amass users.
+Counterfactual to this, the increase could be due to other external factors, such as an increase in the number of skilled developers (and thus more package contributions), a change in CRAN policy or management to add packages (either to CRAN and/or task view), and/or that new packages are still yet to amass users.
 While there is an argument that low download counts are due to the low utility and/or quality of the packages, packages in CRAN task views are selected by expert maintainers, thus we can reasonably assume that any package listed in CRAN task view are of a decent utility and quality.
 
-If the downloads are reflective of the experimental designs used in practice, **_small set of packages appear to be sufficient for most to construct the full set of designs of experiments needed in practice_**.
+If the downloads are reflective of the experimental designs used in practice, **_small set of packages appear to be sufficient for most to construct the full set of designs of experiments they need in practice_**.
 Packages of course evolve and the top downloaded packages have had regular updates that may have broadened its scope from its previous releases.
@@
 This suggests that other CRAN task views may have dominant standards and in comparison to other topics, there are more **_diverse approaches to designing experiments_** and thus no single DoE package is dominant.
-This observation however doesn't take into account other approaches to generating experimental designs, like the proprietary software, CycDesignN [@cycdesignn] and SAS [@sas1985sas], which may be widely used.
+This observation however doesn't take into account other approaches to generate experimental designs, like say the proprietary software, CycDesignN [@cycdesignn], which may be widely used.
 
 ```{r fig-gini-all-ctvs, fig.height = 4, fig.width = 4, fig.cap = "The points show the Gini index of the download counts by year for ExperimentalDesign task view with the color showing the number of packages.
-The grey line shows the distribution of the Gini index across years for all other CRAN task views.
+The grey lines show the line plots of the Gini index across years for all other CRAN task views.
 See Figure S1 in the Supplementary Material for the line graph for the Gini index across years by each CRAN task view."}
@@
 We can see in Figure `r ref("fig:rank-over-time")` that most of the top 10 ranking packages have been in the top 10 for the last 9 years with `lhs` steadily climbing up the ranks in the last few years.
-It should be noted that the download of one package can prompt a download of another package; the most notable dependency network is `AlgDesign` and `agricolae`, where the former is an import for the latter.
-The full dependency network within the DoE packages is shown later in Figure \@ref(fig:plot-doe-network).
+It should be noted that the download of one package can prompt a download of another package; the most notable package connection is `AlgDesign` and `agricolae`, where the former is an import for the latter.
+The full network of packages connections within the DoE packages is shown later in Figure `r ref("fig:plot-doe-network")`.
 
@@
 
-The consistency in the top 10 ranking packages (Figure `r ref("fig:rank-over-time")`) and the fact that most downloaded DoE packages are first released more than 10 years ago (Figure `r ref("fig:release-date-vs-download")`) indicates that either the existing packages filled the needs of mass in practice or no new packages were compelling for many to switch their practice.
+The consistency in the top 10 ranking packages (Figure `r ref("fig:rank-over-time")`) and the fact that most downloaded DoE packages are first released more than 10 years ago (Figure `r ref("fig:release-date-vs-download")`) indicates that either the existing packages fulfilled the needs of mass in practice or no new packages were compelling for many to switch their practice.
 We do however also see that the the top downloaded packages generally have more updates (see Figure `r ref("fig:release-date-vs-download")`) so it is possible that the packages have improved or broadened the scope of its usage.
@@
 Figure `r ref("fig:wordcloud-over-time")` shows some common purposes of DoE packages, based on bigrams in the package title and description.
-We show only the bigrams as unigram was not insightful and there were not many trigrams common across packages.
+We only show the bigrams as unigrams were not insightful and there were not many trigrams common across packages.
 In counting the bigrams, we processed the text data as follows:
@@
 3.
-Bigrams that consist the stop words (provided in `tidytext::stop_words` in addition to other words we deemed irrelevant, e.g.
-"provide", "e.g.", "calculate" and so on; full list is shown in the code provided in the link under Section `r ref("pkgs")`) are removed.
+Bigrams that consist the stop words are removed.
+The stop words are sourced from the lexicons in `tidytext::stop_words` in addition to other words we deemed irrelevant, e.g.
+"provide", "e.g.", "calculate" and so on -- full list is shown in the code provided in the link under Section `r ref("pkgs")`.
 
@@
 The UI design is an abstraction to specifying the desired experimental design and its choices enable how a user expresses the specification of the experimental design.
+The API design aids in other developers to leverage existing systems.
 
-In this section, we discuss these interface designs broadly, with examples illustrated from the top downloaded packages (shown in Figure `r ref("fig:rank-over-time")`).
-We exclude `ez` and `DiceKriging` from the discussion as the former  is predominately visualisation of experimental data and the latter is about the analysis of computer experiments in addition to belonging to the same suite of packages as `DiceDesign` [@DiceDesign].
-
-## The case of space-filling and optimal designs
-
-Computer experiments, as discussed in Section `r ref("topics")`, which generally involve space filling designs, have been on the rise.
-The exemplar for this type of designs are `lhs` [@lhs] and `DiceDesign`.
-For the `lhs` package, functions like `randomLHS()`, `optimumLHS()`, and `maximinLHS()`, require users to specify the sample size ($n$) and the number of variables ($p$) then it generates a Latin hypercube sample, based on different optimisation schemes (in this case, random, S optimal and maxmin criteria, respectively; see package documentation for more details).
-Similarly for the `DiceDesign`, there is a comprehensive list of space-filling designs such as `dmaxDesign()`, `lhsDesign()`, `straussDesign()` and `wspDesign()` with input of $n$ and $p$ as before (among extra parameters for some) that implement algorithms that either maximise the entropy, produce a random Latin hypercube sample, based on Strauss process and WSP algorithm, respectively.
-These designs all output a $n \times p$ matrix with values between 0 and 1.
-
-The `AlgDesign` [@AlgDesign] package offers three primary functions for generating optimal designs: `optBlock()`, `optFederov()` and `optMonteCarlo()`.
-These functions in general require a formula in terms of the supplied terms with the choice of criterion (D, A and I for the latter two).
-The difference lies in the search algorithm for the optimal design.
+In this section, we discuss the interface designs of functions that output an experimental design based on three broad areas: factorial designs, recipe designs and augmenting designs.
+The discussion is exclusive to the top downloaded packages (shown in Figure `r ref("fig:rank-over-time")`), with the exception of `ez` and `DiceKriging`, as the former is predominately visualisation of experimental data and the latter is about the analysis of computer experiments in addition to belonging to the same suite of packages as `DiceDesign` [@DiceDesign].
 
@@
 
-Factorial experiments offer a challenge in the construction and allocation of the treatment factors.
-This is reflected in a number of packages that specifically addresses this challenge.
-These include `DoE.base` [@DoE.base] for full factorial and (regular and irregular) orthogonal array designs via `fac.design()` and `oa.design()`, respectively, `FrF2` [@FrF2] for fractional 2-level factorial designs using `FrF2()`, `FrF2Large()` or `pb()` Plankett-Burnman designs [@Plackett1946-ly], `conf.design` [@conf.design] for symmetric confounded factorial designs via `conf.design()`, and `BHH2` [@BHH2] to generate full or fractional 2-level factorial design matrix via `ffDesMatrix()`.
-All these designs generally requires the user to input the number of factors and if the design is allowed to vary in the number of levels, the input of the number of levels for each factor.
-The construction of some of these designs are catalogue-based, e.g.
-`oa.design()` in `DoE.base`.
+Factorial experiments offers a challenge in allocating the treatment factors to experimental units where the full set of factorial treatments cannot be administered (and replicated) and/or the experimental units have some grouping structure.
+The effort to address this challenge is reflected by the number of packages that focus on the construction of factorial designs as described next.
 
-Response surface design, as illustrated by `rsm` [@rsm], is a factorial experiment but the focus of it is to analyse using response surface modelling.
-The two well known response surface designs are Box-Behnken design [@Box1960-wr] and
-central-composite designs [@Box1951-ji] via `bbd()` and `ccd()` for central-composite designs providing formula.
+The `DoE.base` package [@DoE.base] can construct full factorial and (regular and irregular) orthogonal array designs via `fac.design()` and `oa.design()`, respectively.
+The `FrF2` package [@FrF2] constructs regular fractional 2-level factorial designs using `FrF2()` and `FrF2Large()` with the latter for large designs.
+The `conf.design` package  [@conf.design] constructs symmetric confounded factorial designs via `conf.design()`, and `BHH2` package [@BHH2] generates full or fractional 2-level factorial design matrix via `ffDesMatrix()`.
 
-## The case of sequential designs
+While the argument names and underlying algorithms of these functions to generate factorial designs differ, it generally requires the users to input the number of:
 
-Another specialised design is sequential design (also called adaptive sampling), which is best represented by `tgp` [@tgp] -- these require prior information, which are used to inform the next experimental design using `tgp.design()` and `dopt.gp()` which requires users to supply candidate samples to sub-sample from and a model or design from a prior design.
-Follow-up experiments, which can also be classified as a sequential design, is implemented by `BsMD` [@BsMD].
-In `BsMD`, the follow-up design is determined by a model discriminant approach using `MD()`.
+* treatment factors, 
+* experimental runs (or replications), and 
+* levels for each factor if the design is allowed to vary in the number of levels.
 
-## The case of menu functions
+The output of the design is either a special class of list (e.g.
+the `design` class for `DoE.base` and `FrF2` or a `data.frame` for `conf.design`) or a matrix for `BHH2` such that an element or column corresponds to a treatment factor with each value corresponding to one experimental run.
+The treatment factor are generally assigned pseudo names (e.g.
+letters of the alphabet) or an argument exists for users to input treatment names as a character vector.
 
-The `agricolae` package [@agricolae] was the most downloaded DoE packages in 2013 to 2018 (Figure `r ref("fig:rank-over-time")`) and possibly prior to 2013 (data unavailable).
-This package is the prime example of constructing designs based on menu functions, e.g.
+Some forms of factorial designs are known by other names, e.g.
+*response surface designs* are factorial designs where the treatment factors are discrete levels of continuous variables.
+Two types of response surface designs can be constructed using the `rsm` package [@rsm]: Box-Behnken design [@Box1960-wr] and central-composite designs [@Box1951-ji] via functions `bbd()` and `ccd()`, respectively, where the minimum required input is the number of factors.
+Another form of factorial designs is *saturated designs* where typically higher-order interaction effects of treatment factors are confounded with the main effects.
+Plankett-Burnman designs [@Plackett1946-ly] is one type of saturated design which can be generated by the function `pb()` in the `FrF2` package, where the user provides the number of experimental runs and the number of treatment factors.
+
+## The case of recipe designs
+
+The `agricolae` package [@agricolae] is the prime example of constructing designs based on a set of so-called "recipe functions", where each function corresponds to a single class of experimental design.
+E.g.
 `design.crd()`, `design.rcbd()` and `design.split()` construct a completely randomised design, randomised complete block design and split-plot design, respectively.
-Users typically supply the treatment labels (or number of treatments in the case of `design.split()`) and the number of replications as argument to these functions.
+Users typically supply the treatment labels (or number of treatments in the case of `design.split()`) and the number of replications as argument to these functions and the output is a list with one element corresponding to a `data.frame` that contains the design in a table such that the row corresponds to the experimental run and the columns corresponds to the experimental variables (we refer to this format simply as "table format" henceforth).
+
+The use of recipe functions is not limited to classical experimental designs; the `AlgDesign` [@AlgDesign] package offers three primary functions for generating optimal designs: `optBlock()`, `optFederov()` and `optMonteCarlo()`.
+These functions in general require a data and formula in terms of the supplied data with the choice of criterion (e.g.
+D-criterion) with the output as a list with one element corresponding to the design in a table format.
+The difference between these functions lies in the underlying search strategy for the optimal design and the name of the function is a surrogate for the search algorithm.
+
+Computer experiments, which generally involve space filling designs, are implemented in packages like `lhs` [@lhs] and `DiceDesign` [@DiceDesign].
+For the `lhs` package, functions like `randomLHS()`, `optimumLHS()`, and `maximinLHS()`, require users to specify the sample size ($n$) and the number of variables ($p$) then it generates a Latin hypercube sample [@McKay1979-aw], based on different optimisation schemes (in this case, random, S optimal and maxmin criteria, respectively; see package documentation for more details).
+Similarly for the `DiceDesign`, there is a comprehensive list of space-filling designs such as `dmaxDesign()`, `lhsDesign()`, and `wspDesign()` with input of $n$ and $p$ as before (among extra parameters for some) that implement algorithms that either maximise the entropy [@Shewry1987-jn], produce a random Latin hypercube sample, and use the WSP algorithm [@Santiago2012-ui], respectively.
+These designs all output a $n \times p$ matrix with values between 0 and 1.
+Again, these functions employ a recipe style where each functions have a name that correspond to a certain search strategy in generating the experimental design.
+
+## The case of augmenting designs
+
+Some functions in DoE packages require input of existing experimental designs to produce new designs.
+For example, the `DoE.base` package contain some experimental functions, `cross.design()` and `param.design()`, to combine designs with the former taking a Cartesian product of the input designs while the latter using a Taguchi style [@Taguchi1986-ut] to aggregate the designs with the inner and outer arrays.
+The `lhs` package contain the function that `augmentLHS()` to add additional samples to existing Latin hypercube sample.
+
+Another class of augmenting design is *sequential design* (also called *adaptive sampling*), which is best represented by `tgp` [@tgp] -- these require prior information, which are used to inform the next experimental design using `tgp.design()` and `dopt.gp()`.
+The user is required to supply candidate samples to sub-sample from and a model or a prior experimental design.
+Follow-up experiments, which can also be classified as a sequential design, is implemented by `BsMD` [@BsMD] using a model discriminant approach using `MD()`.
 
@@
 
-Through the exploratory data analysis of three data sources (package download logs, package metadata and CRAN task views) outlined in Section `r ref("data")`, we have observed that the the total download of DoE packages are concentrated only on a handful of R-packages although these represent a diverse set in comparison with other CRAN task views.
-Furthermore, the data suggests that the field of experimental design is the least collaborative field; out of all CRAN task views, it has one of the least average number of authors and the lowest intra-connectivity (any dependency on other packages within its own task view).
+Through the exploratory data analysis of three data sources (package download logs, package metadata and CRAN task views) outlined in Section `r ref("data")`, we have observed in Section `r ref("eda")` that the the total download of DoE packages are concentrated only on a handful of R-packages although these represent a diverse set in comparison with other CRAN task views.
+Furthermore, the data suggests that the field of experimental design is the least collaborative field.
+
 There are a number of limitations and short coming in our exploratory data analysis.
@@
 Second, we only use the RStudio CRAN mirror download, which may bias our observations.
-Third, our analysis are limited to R-packages alone, although the sheer quantity and variation of experimental design is unmatched with any other programming languages.
+Third, our analysis are limited to R-packages alone and many practioners may indeed be using other methods to construct experimental designs.
 Finally, all our statements should be treated as speculative rather than conclusive; the data are all observational so no conclusive, generalisable statement are possible.
@@
 
-The interface design (discussed in Section `r ref("design")`) reveals that the most widely used DoE packages generally have functions that is 1) a menu format (i.e.
-the name of the function correspond to a particular recipe of the design or optimal search algorithm) and 2) context is often a second-thought (e.g.
-only the number of factors is needed and the package will assign pseudo factor names or an argument exists for optional input of character vector that corresponds to the factor names).
+The interface design (discussed in Section `r ref("design")`) reveals that the most widely used DoE packages generally have functions that 1) focus on certain aspects of experimental design (e.g.
+factorial structure or augmenting design), 2) is a recipe format (i.e.
+the name of the function is a surrogate to a single class of the design or optimal search algorithm) and 3) context is often a second-thought -- many inputs a single integer corresponding to the number of factors, levels, or experimental runs (or sample size).
+The function will often assign pseudo factor names or there is an optional argument to input a character vector that corresponds to the factor names.
+These interface designs require users to have processed the experiment in statistical terms (often stripping the experimental context away) and simultaneously, users must make a choice of the generation mechanism by selecting an appropriate function.
+Arguably, the current dominant interface designs are not aligned with the way practitioners cognitively design experiments.
+Often the critical part of designing an experiment is understanding the experimental structure and the experimental context can govern or guide the choice of the algorithm to allocate treatments.
+In addition, the nature of recipe functions can obscure the understanding and relation of the designs (e.g.
+how do you go from an unstructured factorial design to split plot design?).
+Each new method to generating an experimental design appear to correspond to a totally new function and any intermediate results are often not easily accessible.
+These may be contributing factors to why developers often work in silos in the field of experimental design.
+
+A consistent cognitive interface design that leverage existing developments and can be easily extensible for new methods will conceivably be of great help to practitioners.
+Some efforts to this end is seen in `DoE.wrapper` [@DoE.wrapper], which contain wrapper recipe functions to other DoE packages such as `lhs`, `AlgDesign` and `FrF2` (see Figure `r ref("fig:plot-doe-network")`) and is also the subject of the developmental package `edibble` [@edibble].
+Undoubtedly, no single developer or package can cater for all experimental design needs so any unifying interface should consider how other developers can contribute or add their methods.
+Future research could benefit from further exploratory data analysis, expanding the study to beyond R-packages, and discussing other aspects of interface designs.
 

### 2022-06-15 paper/arxiv/arxiv.tex
done

@@
 plays a critical role in increasing the quality of experimental data. In
-this article we describe the current state of the R-packages for the
+this paper we describe the current state of the R-packages for the
 design of experiments through an exploratory data analysis of package
@@
 be sufficiently manufactured by a small number of packages and the
-development of experimental designs occur in silos. We discuss also the
-interface designs of widely utilised R packages in the field of
-experimental design and discuss its future prospects for the field.
+development of experimental designs often occur in silos. We discuss
+also the interface designs of widely utilised R packages in the field of
+experimental design and discuss its future prospects to advance the
+field in practice.
 \end{abstract}
@@
 out of it. A carefully crafted data collection scheme is therefore
-critical to optimise the information from data. Of all the data
-collection schemes, the field of experimental designs is specifically
-devoted to planning an experiment largely based on the founding
-principles by Fisher (1935) or an optimisation framework like those
-described in Pukelsheim (2006). These experimental designs are often
-constructed with the aid of a statistical software, such as R (R Core
-Team 2021), so software usage can aid in understanding the practice of
-experimental design. Given the large gain that can be made from
-collecting higher quality experimental data, we benefit from
-understanding the current state of the experimental designs in practice.
+critical to optimise the information from data. The field of
+experimental designs is specifically devoted to planning the collection
+of experimental data, largely based on the founding principles by Fisher
+(1935) or an optimisation framework like those described in Pukelsheim
+(2006). These experimental designs are often constructed with the aid of
+a statistical software, such as R (R Core Team 2021), Python (Rossum
+1995), SAS (SAS Institute 1985), and so on, so the usage of experimental
+design software can inform us about some aspects of experimental designs
+in practice.
 
@@
 Network (CRAN) task views} (a volunteer maintained list of R-packages by
-topic) where R-packages in experimental design are in
+topic) where R-packages for experimental design are in
 \href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}}
-task view and R-packages in survey designs are in
+task view and R-packages for survey designs are in
 \href{http://CRAN.R-project.org/view=OfficialStatistics}{\emph{OfficialStatistics}}
 task view. The full list of available topics can be see in Table S1 in
-the Supplementary Materials and also accessed via the \texttt{ctv}
-package (Zeileis 2005). A subset of experimental designs are segregated
-into
+the Supplementary Materials. A subset of experimental designs are
+segregated into
 \href{http://CRAN.R-project.org/view=ClinicalTrials}{\emph{ClinicalTrials}}
@@
 experimental designs in the R-packages are arguably unmatched with any
-other programming languages, e.g.~in Python (Rossum 1995), only a
-handful of packages that generate design of experiment exist (namely
-\texttt{pyDOE}, \texttt{pyDOE2}, \texttt{dexpy}, \texttt{experimenter}
-and \texttt{GPdoemd}) with limited type of designs. Thus, the study of
-DoE packages, based on quantitative and qualitative data, can give us an
+other programming languages, e.g.~in Python, only a handful of packages
+that generate design of experiment exist (namely \texttt{pyDOE},
+\texttt{pyDOE2}, \texttt{dexpy}, \texttt{experimenter} and
+\texttt{GPdoemd}) with limited type of designs. Thus, the study of DoE
+packages, based on quantitative and qualitative data, can give us an
 objective view to the state of the current experimental designs in
@@
 Wickham et al. 2019) is a testament to the impact an interface design
-can have in practice.
+can have in practice. The practice of experimental design could be
+advanced by adopting similar interface design principles across DoE
+packages.
 
@@
 \ref{eda} presents some insights into the state of the current DoE
-packages by exploratory data analysis of package download data, text
+packages by the exploratory data analysis of package download data, text
 descriptions and comparisons with other CRAN task views; Section
-\ref{design} discuss the interface designs of DoE packages, and we
-conclude with a discussion in Section \ref{discussion} of future
-prospects.
+\ref{design} discusses the interface designs of widely used DoE
+packages, and we conclude with a discussion in Section \ref{discussion}
+of future prospects in the software development of experimental designs.
 
@@
 To study the DoE packages, we analyse data using three sources of data
-as described next.
+as described below.
 
@@
 uses the data from the beginning of 2013 to end of 2021 (a total of 9
-years) for the DoE packages.
+years) for the packages in the CRAN task views.
 
@@
 All CRAN packages have a title, description, package connections
-(suggests, dependency and imports) and other meta-information in the
+(suggests, depends and imports) and other meta-information in the
 DESCRIPTION file. We use the text data from the title and description
@@
 
-The list of packages in each CRAN task views (as of 2022-05-25) are used
-to contrast the characteristics of DoE packages in Section \ref{silo}.
+CRAN task views are volunteer maintained list of R-packages on CRAN
+relevant to the corresponding topic. There are a total 37 CRAN task
+views. Table S1 in the Supplementary Materials list the full available
+topics from the \texttt{ctv} package (Zeileis 2005). The list of
+packages in each CRAN task views (as of 2022-05-25) are used to contrast
+the characteristics of DoE packages in Sections \ref{popular} and
+\ref{silo}.
 
@@
 
-All results presented are derived from exploratory data analysis of
-observational data; consequently, all interpretations are somewhat
-speculative and may not be indicative of the true state of the field of
-experimental design. In particular, any analysis over time are
-confounded by the fact that the nature of users and package management
-have also changed over the years. It should be noted that some DoE
-packages may have been archived or removed from the task view over the
-years so any cross-sectional analysis presented may not reflect the set
-of all DoE packages at that particular time period (although we assume
-such incidences are low).
+In this section, we derive some conjecture based on the analysis of the
+data described in Section \ref{data}. All results presented are from the
+exploratory data analysis of observational data, consequently, all
+interpretations are somewhat speculative and may not be indicative of
+the true state of the field of experimental design. In particular, any
+analysis over time are confounded by the fact that the nature of users
+and package management have also changed over the years. It should be
+noted that some DoE packages may have been archived or removed from the
+task view over the years so any cross-sectional analysis presented may
+not reflect the set of all DoE packages at that particular time period
+(although we assume such incidences are low).
 
@@
 but about the analysis of experimental data. A complete delineation of
-these packages is difficult as many have some functions that can aid
-decisions or constructions of experimental designs (and any
-categorisation is prone to our subjective bias) so we opted not to
-remove any DoE packages in the analysis.
+these packages is difficult as there is almost always at least one
+function that can aid decisions or constructions of experimental designs
+(and any categorisation is prone to our subjective bias) so we opted not
+to remove any DoE packages in the analysis.
 
@@
 \subsection{Small, but diverse, set of packages are sufficient for most
-designs}\label{popular}}
+experimental designs in practice}\label{popular}}
 
@@
 
-\caption{Lorenz curve of the total download count for DoE packages in 2021. The red line corresponds to the line of perfect equality.}\label{fig:plot-lorenz}
+\caption{Lorenz curve of the total download count for DoE packages in 2021.
+The red line corresponds to the line of perfect equality.
+The green region shows the area under the Lorenz curve and the red region shows the area of gap in equality.}\label{fig:plot-lorenz}
 \end{figure}
@@
 downloads are due to one package). In Figure \ref{fig:download-share},
-we see that the distributions of the downloads each year have a heavy
-right tail with the Gini index ranging from 32.5\% to 68.3\% across the
-years 2013 to 2021, indicating that there is a high level of inequality
-of downloads across packages, particularly with more pronounced
-inequality in the last 5 years.
+we see that the distributions of the package downloads each year have a
+heavy right tail with the Gini index ranging from 42.6\% to 74.1\%
+across the years 2013 to 2021, indicating that there is a high level of
+inequality of package downloads, particularly with more pronounced
+inequality in the last 6 years.
 
@@
 The label on the bottom of the plot shows the Gini index for downloads and the number of packages with full year of download count in the corresponding year.
-In the last 5 years, the Gini index is consistent above 60\% each year indicating that most downloads are due to a relatively small number of packages.}\label{fig:download-share}
+In the last 6 years, the Gini index is consistent above 60\% each year indicating that most downloads are due to a relatively small number of packages.
+}\label{fig:download-share}
 \end{figure}
 
+The increase in the number of packages that are not highly downloaded
+may mean that \textbf{\emph{there are more packages to construct niche
+experimental designs}}. Some examples of these packages include
+\texttt{qtlDesign}, \texttt{PwrGSD} and \texttt{Crossover} made for QTL
+experiments, group sequential designs and crossover trials,
+respectively. These packages would naturally have a smaller number of
+potential users. Counterfactual to this, the increase could be due to
+other external factors, such as an increase in the number of skilled
+developers (and thus more pacakge contributions), a change in CRAN
+policy or management to add packages (either to CRAN and/or task view),
+and/or that new packages are still yet to amass users. While there is an
+argument that low download counts are due to the low utility and/or
+quality of the packages, packages in CRAN task views are selected by
+expert maintainers, thus we can reasonably assume that any package
+listed in CRAN task view are of a decent utility and quality.
+
+If the downloads are reflective of the experimental designs used in
+practice, \textbf{\emph{small set of packages appear to be sufficient
+for most to construct the full set of designs of experiments needed in
+practice}}. Packages of course evolve and the top downloaded packages
+have had regular updates that may have broadened its scope from its
+previous releases.
+
 While in absolute terms the Gini index is high for
-\href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}},
-the inequality is not as severe as other CRAN task views as shown in
-Figure \ref{fig:fig-gini-all-ctvs}. We can see in Figure
-\ref{fig:fig-gini-all-ctvs} that the Gini index is generally increasing
-for DoE packages (as is generally the case for other CRAN task views as
-shown in Figure S1 in the Supplementary Materials).
+\href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}}
+task view (42.6\% to 74.1\%), the inequality is not as severe as other
+CRAN task views as shown in Figure \ref{fig:fig-gini-all-ctvs}. We can
+see in Figure \ref{fig:fig-gini-all-ctvs} that the Gini index is
+generally increasing for DoE packages (as is generally the case for
+other CRAN task views as shown in Figure S1 in the Supplementary
+Materials) but most other CRAN task views have a Gini index of over
+75\%. This suggests that other CRAN task views may have dominant
+standards and in comparison to other topics, there are more
+\textbf{\emph{diverse approaches to designing experiments}} and thus no
+single DoE package is dominant. This observation however doesn't take
+into account other approaches to generating experimental designs, like
+the proprietary software, CycDesignN (Whittaker, Williams, and John
+2022) and SAS (SAS Institute 1985), which may be widely used.
 
@@
 
-\caption{The points show the Gini index of the download counts by year for ExperimentalDesign CRAN task view with the color showing the number of packages.
+\caption{The points show the Gini index of the download counts by year for ExperimentalDesign task view with the color showing the number of packages.
 The grey line shows the distribution of the Gini index across years for all other CRAN task views.
-See Figure S2 in the Supplementary Material for the line graph for the Gini index across years by each CRAN task view.}\label{fig:fig-gini-all-ctvs}
+See Figure S1 in the Supplementary Material for the line graph for the Gini index across years by each CRAN task view.}\label{fig:fig-gini-all-ctvs}
 \end{figure}
 
-From Figures \ref{fig:download-share} and \ref{fig:fig-gini-all-ctvs},
-we could draw some indicative states of the field of experimental
-designs with possible counterfactual interpretations:
-
-\begin{itemize}
-\tightlist
-\item
-  The increase in the number of packages that are not highly downloaded
-  may mean that \textbf{\emph{there are more packages to construct niche
-  experimental designs}}. Some examples of these packages include
-  \texttt{qtlDesign}, \texttt{PwrGSD} and \texttt{Crossover} made for
-  QTL experiments, group sequential designs and crossover trials,
-  respectively. These packages would naturally have a smaller number of
-  potential users. Counterfactual to this, the increase could be due to
-  other external factors, such as an increase in a number of skilled
-  contributors, a change in CRAN policy or management to add packages
-  (either to CRAN and/or task view), and/or that new packages are still
-  yet to amass users. While there is an argument that low download
-  counts are due to the low utility and/or quality of the packages,
-  packages in CRAN task views are selected by expert maintainers; thus
-  we can reasonably assume that any package listed in CRAN task view are
-  of a decent utility and quality.
-\item
-  If the downloads are reflective of the experimental designs used in
-  practice, \textbf{\emph{small set of packages appear to be sufficient
-  for most to construct the full set of designs of experiments needed in
-  practice}}. Packages of course evolve and the top downloaded packages
-  have had regular updates that may have broadened its scope.
-\item
-  While small subset of DoE packages are most frequently used, none of
-  these packages are dominant in the field (judging from comparisons
-  with other CRAN task views). This suggests that there are
-  \textbf{\emph{diverse approaches to designing experiments}}. This
-  observation doesn't take into account other approaches to generating
-  experimental designs, like the proprietary software, CycDesignN
-  (Whittaker, Williams, and John 2022) and SAS (SAS Institute 1985),
-  which may be widely used.
-\end{itemize}
-
 \hypertarget{ranking}{%
-\subsection{There is a lack of adaptation of new or innovative
-designs}\label{ranking}}
+\subsection{The field of experimental design is
+slow-changing}\label{ranking}}
 
@@
 
-\caption{The plot shows the rank of top 10 downloaded packages by year. }\label{fig:rank-over-time}
+\caption{The plot shows the rank of top 10 downloaded packages by year.
+Packages that do not appear in top 10 for at least two periods are omitted from the plot.
+Most packages are consistently in the top 10 for the period shown.}\label{fig:rank-over-time}
 \end{figure}
 
-Figure \ref{fig:established-packages} shows a moderate negative
+Figure \ref{fig:release-date-vs-download} shows a moderate negative
 correlation between the first release date and the (log of) total
@@
 This suggest that in general, a package released earlier are more likely
-to be used today (possibly for legacy reasons or the general inertia for
-adoption of new packages). We can also see in Figure
-\ref{fig:established-packages} the most downloaded packages were
+to be used today (possibly for legacy reasons or the general inertia to
+adopt new packages). We can also see in Figure
+\ref{fig:release-date-vs-download} the most downloaded packages were
 generally released in 2004 to 2010.
@@
 
-{\centering \includegraphics{figures/established-packages-1}
+{\centering \includegraphics{figures/release-date-vs-download-1}
 
@@
 The blue line corresponds to the least squares fit of a simple linear regression model.
-The label in the right-hand upper corner shows the sample correlation coefficient between the first release date and the log, with base 10, of the total download count.
-The high-leverage point on the far left belongs to `conf.design`, authored by one of the earlier contributors to R.}\label{fig:established-packages}
+The label in the right-hand upper corner shows the sample correlation coefficient between the first release date and the log (with base 10) of the total download count.
+The high-leverage point on the far left belongs to `conf.design`, authored by one of the earlier contributors to R.}\label{fig:release-date-vs-download}
 \end{figure}
 
-The lack of change in the top ranking packages are indicative that there
-has been generally little need for new or innovative experimental design
-needs for the mass. We do however also see that the the top downloaded
-packages generally have more updates so it is possible that the packages
-have improved or broadened the scope of its usage.
+The consistency in the top 10 ranking packages (Figure
+\ref{fig:rank-over-time}) and the fact that most downloaded DoE packages
+are first released more than 10 years ago (Figure
+\ref{fig:release-date-vs-download}) indicates that either the existing
+packages filled the needs of mass in practice or no new packages were
+compelling for many to switch their practice. We do however also see
+that the the top downloaded packages generally have more updates (see
+Figure \ref{fig:release-date-vs-download}) so it is possible that the
+packages have improved or broadened the scope of its usage.
 
 \hypertarget{topics}{%
-\subsection{Computer experiments and optimal designs are of
-interest}\label{topics}}
+\subsection{Optimal designs are of interest}\label{topics}}
 
 Figure \ref{fig:wordcloud-over-time} shows some common purposes of DoE
-packages, based on bigrams in the package title and description. A
-bigram is only shown as the frequency of single word was not insightful
-and there were not many trigrams common across packages. Not
-surprisingly, the bigram ``experimental design'' was the most common.
-More interestingly, ``optimal design'' appeared across different
-packages (indicated by the size of the word in Figure
-\ref{fig:wordcloud-over-time}), and the bigrams ``latin hypercube'' and
-``computer experiment'' used across a few packages that are downloaded
-frequently (indicated by the color of the word in Figure
-\ref{fig:wordcloud-over-time}). Although there exists a separate
-\href{http://CRAN.R-project.org/view=ClinicalTrials}{\emph{ClinicalTrials}}
-task view, the DoE packages clearly include some packages that are of
-interest to clinical trials as shown by the size of the bigram
-``clinical trial'' in Figure \ref{fig:wordcloud-over-time}.
+packages, based on bigrams in the package title and description. We show
+only the bigrams as unigram was not insightful and there were not many
+trigrams common across packages. In counting the bigrams, we processed
+the text data as follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  We standardised the words to lower case and removed pluralization.
+\item
+  Multiple mentions of the same bigram within a package was counted as
+  one (e.g.~\texttt{AlgDesign} mentions ``experimental design'' four
+  times in the title and the description but this is counted as one).
+\item
+  Bigrams that consist the stop words (provided in
+  \texttt{tidytext::stop\_words} in addition to other words we deemed
+  irrelevant, e.g.~``provide'', ``e.g.'', ``calculate'' and so on; full
+  list is shown in the code provided in the link under Section
+  \ref{pkgs}) are removed.
+\end{enumerate}
+
+Not surprisingly, the bigram ``experimental design'' was the most
+common. More interestingly, ``optimal design'' and ``sequential design''
+appeared across different packages (indicated by the size of the word in
+Figure \ref{fig:wordcloud-over-time}), and the bigrams ``latin
+hypercube'' and ``computer experiment'' are used across a few packages
+that are downloaded frequently (indicated by the color of the word in
+Figure \ref{fig:wordcloud-over-time}). Sequential design, Latin
+hypercube sampling and computer experiments (which generally include
+space filling designs like Latin hypercube sampling) are designs that
+generally operate by optimising a user-selected criterion and can be
+classified as optimal designs.
 
@@
 
-\caption{The above figure shows the word cloud of bigrams from the title and descriptions of the CRAN packages.
+\caption{The above figure shows the word cloud of bigrams from the title and descriptions of the DoE packages.
 The size shows how often the bigram appears across the DoE packages and the color are relative to the total download count in 2021 of the packages that contain the bigram.}\label{fig:wordcloud-over-time}
 \end{figure}
+
+Although there exists a separate
+\href{http://CRAN.R-project.org/view=ClinicalTrials}{\emph{ClinicalTrials}}
+task view, the DoE packages clearly include some packages that are of
+interest to clinical trials as shown by the size of the bigram
+``clinical trial'' (and related bigrams like ``dose finding'' and
+``phase ii'') in Figure \ref{fig:wordcloud-over-time}.
 
@@
 \href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}}
-have the lowest average number of contributors among all CRAN task
-views. In addition, we can also see in Figure \ref{fig:ctv-summ-plot}
-that the
+task view has the lowest average number of contributors among all 37
+CRAN task views. In addition, we can also see in Figure
+\ref{fig:ctv-summ-plot} that the
 \href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}}
@@
 Figure \ref{fig:plot-doe-network}. These observations suggest that the
-field of experimental design is one of the least collaborative field and
-package development generally occur in silos.
+\textbf{\emph{field of experimental design is one of the least
+collaborative field}} and package development generally occur in silos.
 
@@
 
-\caption{Package connections (depends, suggests and imports) within the DoE packages. DoE packages that do not depend, suggest or import another DoE package is not shown.}\label{fig:plot-doe-network}
+\caption{Package connections (depends, suggests and imports) within the DoE packages.
+The direction of the arrow shows the connection of packags where the package on the tail of the arrow is a depedency, suggestion or import for the package on the head of the arrow.
+DoE packages that do not depend, suggest or import another DoE package is not shown.}\label{fig:plot-doe-network}
 \end{figure}
@@
 choices enable how a user expresses the specification of the
-experimental design.
+experimental design. The API design aids in other developers to leverage
+existing systems.
 
@@
 \texttt{DiceKriging} from the discussion as the former is predominately
-visualisation of experimental data and latter is about the analysis of
-computer experiments in addition to belonging to the same suite of
-packages as \texttt{DiceDesign}.
+visualisation of experimental data and the latter is about the analysis
+of computer experiments in addition to belonging to the same suite of
+packages as \texttt{DiceDesign} (Dupuy, Helbert, and Franco 2015).
 
-\texttt{DoE.wrapper} (Grömping 2020)
+\hypertarget{the-case-of-factorial-designs}{%
+\subsection{The case of factorial
+designs}\label{the-case-of-factorial-designs}}
 
-\hypertarget{the-case-of-space-filling-and-optimal-designs}{%
-\subsection{The case of space-filling and optimal
-designs}\label{the-case-of-space-filling-and-optimal-designs}}
+Factorial experiments offer a challenge in the construction and
+allocation of the treatment factors. This is reflected in a number of
+packages that specifically addresses this challenge. These include
+\texttt{DoE.base} (Grömping 2018) for full factorial and (regular and
+irregular) orthogonal array designs via \texttt{fac.design()} and
+\texttt{oa.design()}, respectively, \texttt{FrF2} (Grömping 2014) for
+fractional 2-level factorial designs using \texttt{FrF2()},
+\texttt{FrF2Large()} or \texttt{pb()} Plankett-Burnman designs (Plackett
+and Burman 1946), \texttt{conf.design} (Venables 2013) for symmetric
+confounded factorial designs via \texttt{conf.design()}, and
+\texttt{BHH2} (Barrios 2016) to generate full or fractional 2-level
+factorial design matrix via \texttt{ffDesMatrix()}. All these designs
+generally requires the user to input the number of factors and if the
+design is allowed to vary in the number of levels, the input of the
+number of levels for each factor. The construction of some of these
+designs are catalogue-based, e.g.~\texttt{oa.design()} in
+\texttt{DoE.base}.
+
+Response surface design, as illustrated by \texttt{rsm} (Lenth 2009), is
+a factorial experiment but the focus of it is to analyse using response
+surface modelling. The two well known response surface designs are
+Box-Behnken design (Box and Behnken 1960) and central-composite designs
+(Box and Wilson 1951) via \texttt{bbd()} and \texttt{ccd()} for
+central-composite designs providing formula.
+
+Saturated designs
+
+\hypertarget{the-case-of-menu-designs}{%
+\subsection{The case of menu designs}\label{the-case-of-menu-designs}}
 
@@
 generally involve space filling designs, have been on the rise. The
-exemplar for this type of designs are \texttt{lhs} and
+exemplar for this type of designs are \texttt{lhs} (Carnell 2022) and
 \texttt{DiceDesign}. For the \texttt{lhs} package, functions like
@@
 
-\hypertarget{the-case-of-factorial-designs}{%
-\subsection{The case of factorial
-designs}\label{the-case-of-factorial-designs}}
-
-Factorial experiments offer a challenge in the construction and
-allocation of the treatment factors. This is reflected in a number of
-packages that specifically addresses this challenge. These include
-\texttt{DoE.base} (Grömping 2018) for full factorial and (regular and
-irregular) orthogonal array designs via \texttt{fac.design()} and
-\texttt{oa.design()}, respectively, \texttt{FrF2} (Grömping 2014) for
-fractional 2-level factorial designs using \texttt{FrF2()},
-\texttt{FrF2Large()} or \texttt{pb()} Plankett-Burnman designs (Plackett
-and Burman 1946), \texttt{conf.design} (Venables 2013) for symmetric
-confounded factorial designs via \texttt{conf.design()}, and
-\texttt{BHH2} (Barrios 2016) to generate full or fractional 2-level
-factorial design matrix via \texttt{ffDesMatrix()}. All these designs
-generally requires the user to input the number of factors and if the
-design is allowed to vary in the number of levels, the input of the
-number of levels for each factor. The construction of some of these
-designs are catalogue-based, e.g.~\texttt{oa.design()} in
-\texttt{DoE.base} and the entire package \texttt{FrF2.catlg128}
-(Grömping 2022) is a list of fractional 2-level factorial designs.
-
-Response surface design, as illustrated by \texttt{rsm} (Lenth 2009), is
-a factorial experiment but the focus of it is to analyse using response
-surface modelling. The two well known response surface designs are
-Box-Behnken design (Box and Behnken 1960) and central-composite designs
-(Box and Wilson 1951) via \texttt{bbd()} and \texttt{ccd()} for
-central-composite designs providing formula.
+The \texttt{agricolae} package (de Mendiburu 2021) was the most
+downloaded DoE packages in 2013 to 2018 (Figure
+\ref{fig:rank-over-time}) and possibly prior to 2013 (data unavailable).
+This package is the prime example of constructing designs based on menu
+functions, e.g.~\texttt{design.crd()}, \texttt{design.rcbd()} and
+\texttt{design.split()} construct a completely randomised design,
+randomised complete block design and split-plot design, respectively.
+Users typically supply the treatment labels (or number of treatments in
+the case of \texttt{design.split()}) and the number of replications as
+argument to these functions.
 
@@
 \texttt{MD()}.
-
-\hypertarget{the-case-of-menu-functions}{%
-\subsection{The case of menu
-functions}\label{the-case-of-menu-functions}}
-
-The \texttt{agricolae} package (de Mendiburu 2021) was the most
-downloaded DoE packages in 2013 to 2018 (Figure
-\ref{fig:rank-over-time}) and possibly prior to 2013 (data unavailable).
-This package is the prime example of constructing designs based on menu
-functions, e.g.~\texttt{design.crd()}, \texttt{design.rcbd()} and
-\texttt{design.split()} construct a completely randomised design,
-randomised complete block design and split-plot design, respectively.
-Users typically supply the treatment labels (or number of treatments in
-the case of \texttt{design.split()}) and the number of replications as
-argument to these functions.
 
@@
 
-\hypertarget{acknowledgement}{%
-\section{Acknowledgement}\label{acknowledgement}}
+\hypertarget{pkgs}{%
+\section{Acknowledgement}\label{pkgs}}
 
-This paper uses the \texttt{targets} framework (Landau 2021) for
-reproducibility, \texttt{knitr} (Xie 2015) and \texttt{rmarkdown} (Xie,
-Allaire, and Grolemund 2018) for creating reproducible documents,
-\texttt{ggplot2} (Wickham 2016), \texttt{ggraph} (Pedersen 2021),
-\texttt{ggwordcloud} (Le Pennec and Slowikowski 2019) and
-\texttt{colorspace} (Zeileis et al. 2020) for visualisation,
+This paper uses \texttt{targets} (Landau 2021) and \texttt{renv} (Ushey
+2022) for reproducibility, \texttt{knitr} (Xie 2015) and
+\texttt{rmarkdown} (Xie, Allaire, and Grolemund 2018) for creating
+reproducible documents, \texttt{ggplot2} (Wickham 2016), \texttt{ggraph}
+(Pedersen 2021), \texttt{ggwordcloud} (Le Pennec and Slowikowski 2019)
+and \texttt{colorspace} (Zeileis et al. 2020) for visualisation,
 \texttt{kableExtra} (Zhu 2021) for customising the table in the
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-lhs}{}}%
+Carnell, Rob. 2022. \emph{Lhs: Latin Hypercube Samples}.
+\url{https://CRAN.R-project.org/package=lhs}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-cranlogs}{}}%
@@
 \url{https://CRAN.R-project.org/package=agricolae}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-DiceDesign}{}}%
+Dupuy, Delphine, Céline Helbert, and Jessica Franco. 2015.
+{``{DiceDesign} and {DiceEval}: Two {R} Packages for Design and Analysis
+of Computer Experiments.''} \emph{Journal of Statistical Software} 65
+(11): 1--38. \url{https://www.jstatsoft.org/v65/i11/}.
 
@@
 \url{https://doi.org/10.18637/jss.v085.i05}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-DoE.wrapper}{}}%
----------. 2020. \emph{DoE.wrapper: Wrapper Package for Design of
-Experiments Functionality}.
-\url{https://CRAN.R-project.org/package=DoE.wrapper}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-FrF2.catlg128}{}}%
----------. 2022. \emph{FrF2.catlg128: Catalogues of Resolution IV 128
-Run 2-Level Fractional Factorials up to 33 Factors That Do Have 5-Letter
-Words}. \url{https://CRAN.R-project.org/package=FrF2.catlg128}.
 
@@
 \url{https://doi.org/10.21105/joss.00037}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-renv}{}}%
+Ushey, Kevin. 2022. \emph{Renv: Project Environments}.
+\url{https://CRAN.R-project.org/package=renv}.
 

### 2022-06-15 R/functions.R

@@ -6,6 +6,10 @@ ctv_list <- function(ctv) {
                                     repos = "http://cran.rstudio.com/")[[1]]  
 }
 
+lorenz_data <- function(data) {
+  data.frame(p = ineq::Lc(data$total)$p, L = ineq::Lc(data$total)$L) 
+}
+
 pkg_first_release <- function(data) {
   data %>% 
     group_by(package) %>% 

### 2022-06-15 R/plot.R

@@ -42,38 +42,18 @@ plot_download_distribution <- function(data, ginidata) {
                                            "\nn=", n)))
 }
 
-plot_lorenz_curve <- function(data, year) {
+plot_lorenz_curve <- function(data) {
   
-  dat <- data %>% 
-    filter(year == .env$year)
-  
-  data <- data.frame(p = ineq::Lc(dat$total)$p, L = ineq::Lc(dat$total)$L) 
-  anndata <- data %>% 
-    arrange(desc(L)) %>% 
-    filter(L < 0.5) %>% 
-    slice(1)
   ggplot(data, aes(p * 100, L * 100)) +
     geom_abline(color = "red") +
+    geom_ribbon(aes(ymin = 0, ymax = L * 100),
+                fill = "#FFC107", alpha = 0.5) +
+    geom_ribbon(aes(ymin = L * 100, ymax = p * 100),
+                 fill = "#D81B60", alpha = 0.5) +
     geom_line() + 
     geom_point() +
-    geom_ribbon(aes(ymin = 0, ymax = L * 100),
-                fill = "#009E73", alpha = 0.5) +
-    geom_ribbon(aes(ymin = L * 100, ymax = p * 100),
-                 fill = "#D55E00", alpha = 0.5) +
     annotate("text", x = 75, y = 80, label = "Line of perfect equality",
-             angle = 45)# +
-    # geom_segment(data = anndata,
-    #              aes(x = p * 100, xend = p * 100,
-    #                  y = 0, yend = L * 100),
-    #              linetype = "dashed") +
-    # geom_segment(data = anndata,
-    #              aes(x = 0, xend = p * 100,
-    #                  y = L * 100, yend = L * 100),
-    #              linetype = "dashed") +
-    # scale_x_continuous(breaks = c(0, 25, 50, 75, round(anndata$p * 100), 100), 
-    #                    expand = c(0, 0)) +
-    # scale_y_continuous(breaks = c(0, 25, round(anndata$L * 100), 50, 75, 100),
-    #                    expand = c(0, 0))
+             angle = 45)
   
 }
 

### 2022-06-15 _targets.R

@@ -9,7 +9,7 @@ tar_option_set(packages = c("tidyverse", "rvest", "lubridate",
                             "igraph", "ggraph", "ggrepel", "patchwork",
                             "ggwordcloud", "fable", "feasts", "lubridate",
                             "ineq", "cranscrub", "tidygraph", "ggraph", 
-                            "rlang", "igraph"))
+                            "rlang"))
 
 
 list(
@@ -66,7 +66,10 @@ list(
   tar_target(gini_yearly_ctv, gini_coef_by_ctv(cran_rank_yearly_all)), 
   tar_target(gini_yearly_doe, gini_yearly_ctv %>% 
                filter(ctv == "ExperimentalDesign")),
-  tar_target(plot_lorenz2021, plot_lorenz_curve(cran_rank_yearly, 2021)), 
+  tar_target(doe_lorenz2021, cran_rank_yearly_all %>% 
+               filter(ctv == "ExperimentalDesign", year == 2021) %>% 
+               lorenz_data()), 
+  tar_target(plot_lorenz2021, plot_lorenz_curve(doe_lorenz2021)),
   tar_target(plot_download_dist_all, plot_download_distribution(cran_rank_yearly_all, gini_yearly_ctv) + 
                plot_setup() + facet_wrap(~ctv, ncol = 4)),
   tar_target(plot_download_dist, plot_download_distribution(cran_rank_yearly, gini_yearly_doe) + plot_setup()),
@@ -110,9 +113,8 @@ list(
   tar_target(plot_cluster_network, plot_pkg_network_by_ctv("Cluster")),
   tar_target(plot_ts_network, plot_pkg_network_by_ctv("TimeSeries")),
   
-  
-  
   tar_render(report, "paper.Rmd"),
+  tar_render(index, rmarkdown::render("paper.Rmd", output_file = 'index.html')),
   tar_render(supp, "supp.Rmd"),
   tar_render(arxiv, "paper/arxiv/arxiv.Rmd"),
   NULL

### 2022-06-15 paper/arxiv/arxiv.tex
arxiv tex updated + old supp.html deleted

@@
 All CRAN packages have a title, description, package connections
-(suggests, depends and imports) and other meta-information in the
-DESCRIPTION file. We use the text data from the title and description
-(as accessed in 2022-05-23) in Section \ref{topics}.
+(suggests, depends and imports of other packages) and other
+meta-information in the DESCRIPTION file. We use the text data from the
+title and description (as accessed in 2022-05-23) in Section
+\ref{topics}.
 
@@
 
-There have been at least 50 DoE packages since 2013 but most of the
-downloads are concentrated in just a handful of packages. For example,
-Figure \ref{fig:plot-lorenz} shows a Lorenz curve (Lorenz 1905) for the
-total package downloads in 2021 of 113 DoE packages (first released
-prior to 2021); we can see from Figure \ref{fig:plot-lorenz} that bottom
-90\% of DoE packages (in terms of total download count in 2021) only
-share about 30\% of total downloads across all DoE packages -- in
-another words, 70\% of the total downloads are due to 11 packages (10\%
-of the DoE packages).
+There are at least 50 DoE packages since 2013 but most of the downloads
+are concentrated in just a handful of packages. For example, Figure
+\ref{fig:plot-lorenz} shows a Lorenz curve (Lorenz 1905) for the total
+package downloads in 2021 of 113 DoE packages (first released prior to
+2021); we can see from Figure \ref{fig:plot-lorenz} that bottom 90\% of
+DoE packages (in terms of total download count in 2021) only share about
+27\% of total downloads across all DoE packages -- in another words,
+73\% of the total downloads are due to 11 packages (10\% of the DoE
+packages).
 
@@
 The red line corresponds to the line of perfect equality.
-The green region shows the area under the Lorenz curve and the red region shows the area of gap in equality.}\label{fig:plot-lorenz}
+The yellow region shows the area under the Lorenz curve and the red region shows the area of gap in equality.}\label{fig:plot-lorenz}
 \end{figure}
@@
 
-\caption{The distribution of the number of downloads for each DoE package by year.
+\caption{The distribution of the number of downloads for DoE packages by year.
 Packages were removed in any year if it was released in that year or later so that each download count is for the full year.
 The label on the bottom of the plot shows the Gini index for downloads and the number of packages with full year of download count in the corresponding year.
-In the last 6 years, the Gini index is consistent above 60\% each year indicating that most downloads are due to a relatively small number of packages.
+In the last 6 years, the Gini index is consistent above 60\% indicating that most downloads are due to a relatively small number of packages.
 }\label{fig:download-share}
@@
 other external factors, such as an increase in the number of skilled
-developers (and thus more pacakge contributions), a change in CRAN
+developers (and thus more package contributions), a change in CRAN
 policy or management to add packages (either to CRAN and/or task view),
@@
 practice, \textbf{\emph{small set of packages appear to be sufficient
-for most to construct the full set of designs of experiments needed in
-practice}}. Packages of course evolve and the top downloaded packages
+for most to construct the full set of designs of experiments they need
+in practice}}. Packages of course evolve and the top downloaded packages
 have had regular updates that may have broadened its scope from its
@@
 single DoE package is dominant. This observation however doesn't take
-into account other approaches to generating experimental designs, like
+into account other approaches to generate experimental designs, like say
 the proprietary software, CycDesignN (Whittaker, Williams, and John
-2022) and SAS (SAS Institute 1985), which may be widely used.
+2022), which may be widely used.
 
@@
 \caption{The points show the Gini index of the download counts by year for ExperimentalDesign task view with the color showing the number of packages.
-The grey line shows the distribution of the Gini index across years for all other CRAN task views.
+The grey lines show the line plots of the Gini index across years for all other CRAN task views.
 See Figure S1 in the Supplementary Material for the line graph for the Gini index across years by each CRAN task view.}\label{fig:fig-gini-all-ctvs}
@@
 should be noted that the download of one package can prompt a download
-of another package; the most notable dependency network is
+of another package; the most notable package connection is
 \texttt{AlgDesign} and \texttt{agricolae}, where the former is an import
-for the latter. The full dependency network within the DoE packages is
-shown later in Figure @ref(fig:plot-doe-network).
+for the latter. The full network of packages connections within the DoE
+packages is shown later in Figure \ref{fig:plot-doe-network}.
 
@@
 \ref{fig:release-date-vs-download}) indicates that either the existing
-packages filled the needs of mass in practice or no new packages were
+packages fulfilled the needs of mass in practice or no new packages were
 compelling for many to switch their practice. We do however also see
@@
 Figure \ref{fig:wordcloud-over-time} shows some common purposes of DoE
-packages, based on bigrams in the package title and description. We show
-only the bigrams as unigram was not insightful and there were not many
+packages, based on bigrams in the package title and description. We only
+show the bigrams as unigrams were not insightful and there were not many
 trigrams common across packages. In counting the bigrams, we processed
@@
 \item
-  Bigrams that consist the stop words (provided in
-  \texttt{tidytext::stop\_words} in addition to other words we deemed
-  irrelevant, e.g.~``provide'', ``e.g.'', ``calculate'' and so on; full
-  list is shown in the code provided in the link under Section
-  \ref{pkgs}) are removed.
+  Bigrams that consist the stop words are removed. The stop words are
+  sourced from the lexicons in \texttt{tidytext::stop\_words} in
+  addition to other words we deemed irrelevant, e.g.~``provide'',
+  ``e.g.'', ``calculate'' and so on -- full list is shown in the code
+  provided in the link under Section \ref{pkgs}.
 \end{enumerate}
@@
 
-In this section, we discuss these interface designs broadly, with
-examples illustrated from the top downloaded packages (shown in Figure
-\ref{fig:rank-over-time}). We exclude \texttt{ez} and
-\texttt{DiceKriging} from the discussion as the former is predominately
-visualisation of experimental data and the latter is about the analysis
-of computer experiments in addition to belonging to the same suite of
-packages as \texttt{DiceDesign} (Dupuy, Helbert, and Franco 2015).
+In this section, we discuss the interface designs of functions that
+output an experimental design based on three broad areas: factorial
+designs, recipe designs and augmenting designs. The discussion is
+exclusive to the top downloaded packages (shown in Figure
+\ref{fig:rank-over-time}), with the exception of \texttt{ez} and
+\texttt{DiceKriging}, as the former is predominately visualisation of
+experimental data and the latter is about the analysis of computer
+experiments in addition to belonging to the same suite of packages as
+\texttt{DiceDesign} (Dupuy, Helbert, and Franco 2015).
 
@@
 
-Factorial experiments offer a challenge in the construction and
-allocation of the treatment factors. This is reflected in a number of
-packages that specifically addresses this challenge. These include
-\texttt{DoE.base} (Grömping 2018) for full factorial and (regular and
-irregular) orthogonal array designs via \texttt{fac.design()} and
-\texttt{oa.design()}, respectively, \texttt{FrF2} (Grömping 2014) for
-fractional 2-level factorial designs using \texttt{FrF2()},
-\texttt{FrF2Large()} or \texttt{pb()} Plankett-Burnman designs (Plackett
-and Burman 1946), \texttt{conf.design} (Venables 2013) for symmetric
-confounded factorial designs via \texttt{conf.design()}, and
-\texttt{BHH2} (Barrios 2016) to generate full or fractional 2-level
-factorial design matrix via \texttt{ffDesMatrix()}. All these designs
-generally requires the user to input the number of factors and if the
-design is allowed to vary in the number of levels, the input of the
-number of levels for each factor. The construction of some of these
-designs are catalogue-based, e.g.~\texttt{oa.design()} in
-\texttt{DoE.base}.
+Factorial experiments offers a challenge in allocating the treatment
+factors to experimental units where the full set of factorial treatments
+cannot be administered (and replicated) and/or the experimental units
+have some grouping structure. The effort to address this challenge is
+reflected by the number of packages that focus on the construction of
+factorial designs as described next.
 
-Response surface design, as illustrated by \texttt{rsm} (Lenth 2009), is
-a factorial experiment but the focus of it is to analyse using response
-surface modelling. The two well known response surface designs are
-Box-Behnken design (Box and Behnken 1960) and central-composite designs
-(Box and Wilson 1951) via \texttt{bbd()} and \texttt{ccd()} for
-central-composite designs providing formula.
+The \texttt{DoE.base} package (Grömping 2018) can construct full
+factorial and (regular and irregular) orthogonal array designs via
+\texttt{fac.design()} and \texttt{oa.design()}, respectively. The
+\texttt{FrF2} package (Grömping 2014) constructs regular fractional
+2-level factorial designs using \texttt{FrF2()} and \texttt{FrF2Large()}
+with the latter for large designs. The \texttt{conf.design} package
+(Venables 2013) constructs symmetric confounded factorial designs via
+\texttt{conf.design()}, and \texttt{BHH2} package (Barrios 2016)
+generates full or fractional 2-level factorial design matrix via
+\texttt{ffDesMatrix()}.
 
-Saturated designs
+While the argument names and underlying algorithms of these functions to
+generate factorial designs differ, it generally requires the users to
+input the number of:
 
-\hypertarget{the-case-of-menu-designs}{%
-\subsection{The case of menu designs}\label{the-case-of-menu-designs}}
+\begin{itemize}
+\tightlist
+\item
+  treatment factors,
+\item
+  experimental runs (or replications), and
+\item
+  levels for each factor if the design is allowed to vary in the number
+  of levels.
+\end{itemize}
 
-Computer experiments, as discussed in Section \ref{topics}, which
-generally involve space filling designs, have been on the rise. The
-exemplar for this type of designs are \texttt{lhs} (Carnell 2022) and
-\texttt{DiceDesign}. For the \texttt{lhs} package, functions like
-\texttt{randomLHS()}, \texttt{optimumLHS()}, and \texttt{maximinLHS()},
-require users to specify the sample size (\(n\)) and the number of
-variables (\(p\)) then it generates a Latin hypercube sample, based on
-different optimisation schemes (in this case, random, S optimal and
-maxmin criteria, respectively; see package documentation for more
-details). Similarly for the \texttt{DiceDesign}, there is a
-comprehensive list of space-filling designs such as
-\texttt{dmaxDesign()}, \texttt{lhsDesign()}, \texttt{straussDesign()}
-and \texttt{wspDesign()} with input of \(n\) and \(p\) as before (among
-extra parameters for some) that implement algorithms that either
-maximise the entropy, produce a random Latin hypercube sample, based on
-Strauss process and WSP algorithm, respectively. These designs all
-output a \(n \times p\) matrix with values between 0 and 1.
+The output of the design is either a special class of list (e.g.~the
+\texttt{design} class for \texttt{DoE.base} and \texttt{FrF2} or a
+\texttt{data.frame} for \texttt{conf.design}) or a matrix for
+\texttt{BHH2} such that an element or column corresponds to a treatment
+factor with each value corresponding to one experimental run. The
+treatment factor are generally assigned pseudo names (e.g.~letters of
+the alphabet) or an argument exists for users to input treatment names
+as a character vector.
 
-The \texttt{AlgDesign} (Wheeler 2022) package offers three primary
-functions for generating optimal designs: \texttt{optBlock()},
-\texttt{optFederov()} and \texttt{optMonteCarlo()}. These functions in
-general require a formula in terms of the supplied terms with the choice
-of criterion (D, A and I for the latter two). The difference lies in the
-search algorithm for the optimal design.
+Some forms of factorial designs are known by other names,
+e.g.~\emph{response surface designs} are factorial designs where the
+treatment factors are discrete levels of continuous variables. Two types
+of response surface designs can be constructed using the \texttt{rsm}
+package (Lenth 2009): Box-Behnken design (Box and Behnken 1960) and
+central-composite designs (Box and Wilson 1951) via functions
+\texttt{bbd()} and \texttt{ccd()}, respectively, where the minimum
+required input is the number of factors. Another form of factorial
+designs is \emph{saturated designs} where typically higher-order
+interaction effects of treatment factors are confounded with the main
+effects. Plankett-Burnman designs (Plackett and Burman 1946) is one type
+of saturated design which can be generated by the function \texttt{pb()}
+in the \texttt{FrF2} package, where the user provides the number of
+experimental runs and the number of treatment factors.
 
-The \texttt{agricolae} package (de Mendiburu 2021) was the most
-downloaded DoE packages in 2013 to 2018 (Figure
-\ref{fig:rank-over-time}) and possibly prior to 2013 (data unavailable).
-This package is the prime example of constructing designs based on menu
-functions, e.g.~\texttt{design.crd()}, \texttt{design.rcbd()} and
-\texttt{design.split()} construct a completely randomised design,
+\hypertarget{the-case-of-recipe-designs}{%
+\subsection{The case of recipe
+designs}\label{the-case-of-recipe-designs}}
+
+The \texttt{agricolae} package (de Mendiburu 2021) is the prime example
+of constructing designs based on a set of so-called ``recipe
+functions'', where each function corresponds to a single class of
+experimental design. E.g. \texttt{design.crd()}, \texttt{design.rcbd()}
+and \texttt{design.split()} construct a completely randomised design,
 randomised complete block design and split-plot design, respectively.
@@
 the case of \texttt{design.split()}) and the number of replications as
-argument to these functions.
+argument to these functions and the output is a list with one element
+corresponding to a \texttt{data.frame} that contains the design in a
+table such that the row corresponds to the experimental run and the
+columns corresponds to the experimental variables (we refer to this
+format simply as ``table format'' henceforth).
 
-\hypertarget{the-case-of-sequential-designs}{%
-\subsection{The case of sequential
-designs}\label{the-case-of-sequential-designs}}
+The use of recipe functions is not limited to classical experimental
+designs; the \texttt{AlgDesign} (Wheeler 2022) package offers three
+primary functions for generating optimal designs: \texttt{optBlock()},
+\texttt{optFederov()} and \texttt{optMonteCarlo()}. These functions in
+general require a data and formula in terms of the supplied data with
+the choice of criterion (e.g.~D-criterion) with the output as a list
+with one element corresponding to the design in a table format. The
+difference between these functions lies in the underlying search
+strategy for the optimal design and the name of the function is a
+surrogate for the search algorithm.
 
-Another specialised design is sequential design (also called adaptive
-sampling), which is best represented by \texttt{tgp} (Gramacy and Taddy
-2010) -- these require prior information, which are used to inform the
-next experimental design using \texttt{tgp.design()} and
-\texttt{dopt.gp()} which requires users to supply candidate samples to
-sub-sample from and a model or design from a prior design. Follow-up
-experiments, which can also be classified as a sequential design, is
-implemented by \texttt{BsMD} (Barrios 2020). In \texttt{BsMD}, the
-follow-up design is determined by a model discriminant approach using
-\texttt{MD()}.
+Computer experiments, which generally involve space filling designs, are
+implemented in packages like \texttt{lhs} (Carnell 2022) and
+\texttt{DiceDesign} (Dupuy, Helbert, and Franco 2015). For the
+\texttt{lhs} package, functions like \texttt{randomLHS()},
+\texttt{optimumLHS()}, and \texttt{maximinLHS()}, require users to
+specify the sample size (\(n\)) and the number of variables (\(p\)) then
+it generates a Latin hypercube sample (McKay, Beckman, and Conover
+1979), based on different optimisation schemes (in this case, random, S
+optimal and maxmin criteria, respectively; see package documentation for
+more details). Similarly for the \texttt{DiceDesign}, there is a
+comprehensive list of space-filling designs such as
+\texttt{dmaxDesign()}, \texttt{lhsDesign()}, and \texttt{wspDesign()}
+with input of \(n\) and \(p\) as before (among extra parameters for
+some) that implement algorithms that either maximise the entropy (Shewry
+and Wynn 1987), produce a random Latin hypercube sample, and use the WSP
+algorithm (Santiago, Claeys-Bruno, and Sergent 2012), respectively.
+These designs all output a \(n \times p\) matrix with values between 0
+and 1. Again, these functions employ a recipe style where each functions
+have a name that correspond to a certain search strategy in generating
+the experimental design.
+
+\hypertarget{the-case-of-augmenting-designs}{%
+\subsection{The case of augmenting
+designs}\label{the-case-of-augmenting-designs}}
+
+Some functions in DoE packages require input of existing experimental
+designs to produce new designs. For example, the \texttt{DoE.base}
+package contain some experimental functions, \texttt{cross.design()} and
+\texttt{param.design()}, to combine designs with the former taking a
+Cartesian product of the input designs while the latter using a Taguchi
+style (Taguchi 1986) to aggregate the designs with the inner and outer
+arrays. The \texttt{lhs} package contain the function that
+\texttt{augmentLHS()} to add additional samples to existing Latin
+hypercube sample.
+
+Another class of augmenting design is \emph{sequential design} (also
+called \emph{adaptive sampling}), which is best represented by
+\texttt{tgp} (Gramacy and Taddy 2010) -- these require prior
+information, which are used to inform the next experimental design using
+\texttt{tgp.design()} and \texttt{dopt.gp()}. The user is required to
+supply candidate samples to sub-sample from and a model or a prior
+experimental design. Follow-up experiments, which can also be classified
+as a sequential design, is implemented by \texttt{BsMD} (Barrios 2020)
+using a model discriminant approach using \texttt{MD()}.
 
@@
 download logs, package metadata and CRAN task views) outlined in Section
-\ref{data}, we have observed that the the total download of DoE packages
-are concentrated only on a handful of R-packages although these
-represent a diverse set in comparison with other CRAN task views.
-Furthermore, the data suggests that the field of experimental design is
-the least collaborative field; out of all CRAN task views, it has one of
-the least average number of authors and the lowest intra-connectivity
-(any dependency on other packages within its own task view). There are a
-number of limitations and short coming in our exploratory data analysis.
-First, CRAN task views are volunteer maintained so some experimental
-design packages may not be included in the DoE packages. Second, we only
-use the RStudio CRAN mirror download, which may bias our observations.
-Third, our analysis are limited to R-packages alone, although the sheer
-quantity and variation of experimental design is unmatched with any
-other programming languages. Finally, all our statements should be
-treated as speculative rather than conclusive; the data are all
-observational so no conclusive, generalisable statement are possible.
-Regardless, the data driven nature of our analysis give an objective
-insight into the field of experimental design.
+\ref{data}, we have observed in Section \ref{eda} that the the total
+download of DoE packages are concentrated only on a handful of
+R-packages although these represent a diverse set in comparison with
+other CRAN task views. Furthermore, the data suggests that the field of
+experimental design is the least collaborative field.
+
+There are a number of limitations and short coming in our exploratory
+data analysis. First, CRAN task views are volunteer maintained so some
+experimental design packages may not be included in the DoE packages.
+Second, we only use the RStudio CRAN mirror download, which may bias our
+observations. Third, our analysis are limited to R-packages alone and
+many practioners may indeed be using other methods to construct
+experimental designs. Finally, all our statements should be treated as
+speculative rather than conclusive; the data are all observational so no
+conclusive, generalisable statement are possible. Regardless, the data
+driven nature of our analysis give an objective insight into the field
+of experimental design.
 
 The interface design (discussed in Section \ref{design}) reveals that
-the most widely used DoE packages generally have functions that is 1) a
-menu format (i.e.~the name of the function correspond to a particular
-recipe of the design or optimal search algorithm) and 2) context is
-often a second-thought (e.g.~only the number of factors is needed and
-the package will assign pseudo factor names or an argument exists for
-optional input of character vector that corresponds to the factor
-names).
+the most widely used DoE packages generally have functions that 1) focus
+on certain aspects of experimental design (e.g.~factorial structure or
+augmenting design), 2) is a recipe format (i.e.~the name of the function
+is a surrogate to a single class of the design or optimal search
+algorithm) and 3) context is often a second-thought -- many inputs a
+single integer corresponding to the number of factors, levels, or
+experimental runs (or sample size). The function will often assign
+pseudo factor names or there is an optional argument to input a
+character vector that corresponds to the factor names. These interface
+designs require users to have processed the experiment in statistical
+terms (often stripping the experimental context away) and
+simultaneously, users must make a choice of the generation mechanism by
+selecting an appropriate function. Arguably, the current dominant
+interface designs are not aligned with the way practitioners cognitively
+design experiments. Often the critical part of designing an experiment
+is understanding the experimental structure and the experimental context
+can govern or guide the choice of the algorithm to allocate treatments.
+In addition, the nature of recipe functions can obscure the
+understanding and relation of the designs (e.g.~how do you go from an
+unstructured factorial design to split plot design?). Each new method to
+generating an experimental design appear to correspond to a totally new
+function and any intermediate results are often not easily accessible.
+These may be contributing factors to why developers often work in silos
+in the field of experimental design.
+
+A consistent cognitive interface design that leverage existing
+developments and can be easily extensible for new methods will
+conceivably be of great help to practitioners. Some efforts to this end
+is seen in \texttt{DoE.wrapper} (Grömping 2020), which contain wrapper
+recipe functions to other DoE packages such as \texttt{lhs},
+\texttt{AlgDesign} and \texttt{FrF2} (see Figure
+\ref{fig:plot-doe-network}) and is also the subject of the developmental
+package \texttt{edibble} (Tanaka 2021). Undoubtedly, no single developer
+or package can cater for all experimental design needs so any unifying
+interface should consider how other developers can contribute or add
+their methods. Future research could benefit from further exploratory
+data analysis, expanding the study to beyond R-packages, and discussing
+other aspects of interface designs.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-DoE.wrapper}{}}%
+---------. 2020. \emph{DoE.wrapper: Wrapper Package for Design of
+Experiments Functionality}.
+\url{https://CRAN.R-project.org/package=DoE.wrapper}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-targets}{}}%
@@
 (70): 209--19. \url{https://doi.org/10.2307/2276207}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-McKay1979-aw}{}}%
+McKay, M D, R J Beckman, and W J Conover. 1979. {``Comparison of Three
+Methods for Selecting Values of Input Variables in the Analysis of
+Output from a Computer Code.''} \emph{Technometrics: A Journal of
+Statistics for the Physical, Chemical, and Engineering Sciences} 21 (2):
+239--45. \url{https://doi.org/10.1080/00401706.1979.10489755}.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Santiago2012-ui}{}}%
+Santiago, J, M Claeys-Bruno, and M Sergent. 2012. {``Construction of
+Space-Filling Designs Using {WSP} Algorithm for High Dimensional
+Spaces.''} \emph{Chemometrics and Intelligent Laboratory Systems} 113
+(April): 26--31. \url{https://doi.org/10.1016/j.chemolab.2011.06.003}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-sas1985sas}{}}%
@@
 Inst.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Shewry1987-jn}{}}%
+Shewry, M C, and H P Wynn. 1987. {``Maximum Entropy Sampling.''}
+\emph{Journal of Applied Statistics} 14 (2): 165--70.
+\url{https://doi.org/10.1080/02664768700000020}.
 
@@
 \url{https://doi.org/10.21105/joss.00037}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Taguchi1986-ut}{}}%
+Taguchi, Genichi. 1986. \emph{Introduction to Quality Engineering:
+Designing Quality into Products and Processes}. Quality Resources.
+
+\leavevmode\vadjust pre{\hypertarget{ref-edibble}{}}%
+Tanaka, Emi. 2021. {``{edibble R-package}.''} \emph{GitHub Repository}.
+\url{https://github.com/emitanaka/edibble}; GitHub.
 

### 2022-06-15 supp.Rmd
update renv + supp

@@
 ---
-title: Supplmentary material for "Current status and prospects of R-packages for the design of experiments"
+title: Supplementary material for "Current status and prospects of R-packages for the design of experiments"
 author:
@@
 library(ggraph)
+library(pander)
 source("R/plot.R")
@@
 tar_load(gini_yearly_ctv)
-
-#renvLock <- jsonlite::read_json("renv.lock")
-#sink("pkgs.bib")
-#knitr::write_bib(names(renvLock$Packages))
-#sink()
 ```
@@
 ```{r session-info, echo = FALSE}
+
 renv_lock <- jsonlite::read_json("renv.lock")
-pander::pander(sessionInfo(package = names(renv_lock$Packages)), compact = FALSE)
+pkgs <- intersect(setdiff(names(renv_lock$Packages), "cranscrub"),
+                  rownames(available.packages(repos = "https://cran.rstudio.com/")))
+pander::pander(sessionInfo(package = pkgs), compact = FALSE)
 ```

### 2022-06-15 supp.tex
update renv + supp

@@
 \hypersetup{
-  pdftitle={Supplmentary material for ``Current status and prospects of R-packages for the design of experiments''},
+  pdftitle={Supplementary material for ``Current status and prospects of R-packages for the design of experiments''},
   pdfauthor={Emi Tanaka; Dewi Amaliah},
@@
 
-\title{Supplmentary material for ``Current status and prospects of R-packages for the design of experiments''}
+\title{Supplementary material for ``Current status and prospects of R-packages for the design of experiments''}
 \author{Emi Tanaka \and Dewi Amaliah}
@@
 
-!6\}\{ModelDeployment\} \& \cellcolor{gray!6}{Model Deployment with R} \& \cellcolor{gray!6}{31} \& \cellcolor{gray!6}{146} \& \cellcolor{gray!6}{6.55} \& \cellcolor{gray!6}{74}\textbackslash*
-\textbackslash end\{longtable\}
-
 \hypertarget{session-information}{%
@@
 \item
-  MASS(v.7.3-57)
+  MASS(v.7.3-54)
 \item
-  Matrix(v.1.4-1)
+  Matrix(v.1.3-4)
 \item
@@
 \item
-  RColorBrewer(v.1.1-3)
+  RColorBrewer(v.1.1-2)
 \item
-  Rcpp(v.1.0.8.3)
+  Rcpp(v.1.0.8)
 \item
@@
 \item
+  coda(v.0.19-4)
+\item
   codetools(v.0.2-18)
 \item
-  colorspace(v.2.0-3)
+  colorspace(v.2.0-2)
 \item
@@
 \item
-  cranscrub(v.0.0.0.9000)
-\item
-  crayon(v.1.5.1)
+  crayon(v.1.4.2)
 \item
@@
 \item
-  fansi(v.1.0.3)
+  fansi(v.1.0.2)
 \item
@@
 \item
-  ggplot2(v.3.3.6)
+  ggplot2(v.3.3.5)
 \item
@@
 \item
-  glue(v.1.6.2)
+  glue(v.1.6.1)
 \item
@@
 \item
-  httr(v.1.4.3)
+  httr(v.1.4.2)
 \item
@@
 \item
-  igraph(v.1.3.1)
+  igraph(v.1.2.11)
 \item
@@
 \item
-  jsonlite(v.1.8.0)
+  jsonlite(v.1.7.3)
 \item
@@
 \item
-  magrittr(v.2.0.3)
+  magrittr(v.2.0.2)
 \item
-  mgcv(v.1.8-40)
+  mgcv(v.1.8-38)
 \item
@@
 \item
-  nlme(v.3.1-157)
+  nlme(v.3.1-153)
 \item
@@
   pagedown(v.0.16)
+\item
+  pander(v.0.6.6)
 \item
@@
 \item
+  remotes(v.2.4.2)
+\item
   renv(v.0.15.5)
@@
 \item
-  scales(v.1.2.0)
+  scales(v.1.1.1)
 \item
@@
 \item
-  tibble(v.3.1.7)
+  tibble(v.3.1.6)
 \item
@@
 \item
-  withr(v.2.5.0)
+  withr(v.2.4.3)
 \item
@@
 \item
-  yaml(v.2.3.5)
+  yaml(v.2.2.2)
 \end{itemize}
@@
   utils(v.4.1.2)
-\item
-  DiceDesign(v.1.9)
 \item
@@
 \item
-  pander(v.0.6.5)
-\item
   graphics(v.4.1.2)
 \item
-  BsMD(v.2020.4.30)
-\item
   stats(v.4.1.2)
-\item
-  rsm(v.2.10.3)
 \end{itemize}

### 2022-06-18 README.Rmd
update README
update readme
update README with arxiv link
make it easier for people to navigate the repo

@@
 This is the repo for the paper titled "`r meta$title`".
-You can find the arxiv version of the paper in `paper/arxiv/arxiv.pdf`.
+
+📜 You can find the arxiv version of the paper [here](https://arxiv.org/abs/2206.07532).
+
+The paper uses the developmental version of the [`cranscrub` package in the branch `paper-doe-review`](https://github.com/numbats/cranscrub/tree/paper-doe-review). This can be installed by
+
+```r
+remotes::install_github("numbats/cranscrub", ref = "paper-doe-review")
+```
+
 To reproduce the results of the paper, you will need to run the analysis by calling
@@
 
-The paper uses the developmental version of the [`cranscrub` package in the branch `paper-doe-review`](https://github.com/numbats/cranscrub/tree/paper-doe-review) which can be installed by
+The primary functions for analysis are:
 
-```r
-remotes::install_github("numbats/cranscrub", ref = "paper-doe-review")
-```
+* `R/functions.R`
+* `R/plot.R`
+* `_targets.R`
+
+The contents of the paper can be found in `paper.Rmd` and the supplementary materials in `supp.Rmd`.
 
@@
 `r meta$abstract`
+
+## Supplementary material
+
+The supplementary material can be found [here](https://emitanaka.org/paper-DoE-review/supp.pdf).

### 2022-06-18 other/analysis/analysis-download-logs.Rmd
make it easier for people to navigate the repo

---
title: "Download logs"
author: "Emi Tanaka"
date: "24/03/2021"
output: html_document
---

```{r library, include=FALSE}
library(tidyverse)
library(magrittr)
library(rvest)
library(glue)
library(lubridate)
library(plotly)
library(cranlogs)
library(ggbeeswarm)
knitr::opts_chunk$set(echo = FALSE, 
                      cache = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      cache.path = "cache/")
```

```{r cache, eval = FALSE}
fs::dir_ls("cache") %>% 
  str_subset(".RData$") %>% 
  walk(load)
```

```{r ctv}
doe_pkgs <- ctv:::.get_pkgs_from_ctv_or_repos("ExperimentalDesign", 
                                              repos = "http://cran.rstudio.com/")[[1]]
survey_pkgs <- ctv:::.get_pkgs_from_ctv_or_repos("OfficialStatistics", 
                                              repos = "http://cran.rstudio.com/")[[1]]
npkgs <- length(doe_pkgs)
```

```{r all_pkgs}
url <- "http://cran.rstudio.com/web/packages/packages.rds"
db <- readRDS(url(url)) %>% 
  as.data.frame()%>% 
  mutate(Description = str_replace_all(Description, "\n", " "),
         Description = str_squish(Description),
         Title = str_replace_all(Title, "\n", " "))
```

```{r extra-pkgs}
exclude_pkgs <- c("argonR", "card", "crispRdesignR", "genotypeR", "genepi", "recipes",
                  "agridat", "desplot", "DiceView", "DOS", "DOS2", "EngrExpt")
cdoe_pkgs <- db %>% 
  filter(str_detect(Title, "[dD]esign") | Package %in% doe_pkgs) %>% 
  filter(!Title %in% exclude_pkgs) %>% 
  pull(Package) 
```

```{r}
fdoe_pkgs <- db %>% 
  filter(str_detect(Title, "[dD]esign") | 
           Package %in% doe_pkgs) %>% 
  mutate(taskview_doe = ifelse(Package %in% doe_pkgs, "yes", "no"),
         taskview_survey = ifelse(Package %in% survey_pkgs, "yes", "no"),
         criteria = case_when(str_detect(tolower(paste(Title, Description)),
                                         "(experiment|block design|factorial design|clinical trial|optimal design|response surface design|stepped wedge design|split-plot design|phase i|crossover design|cluster randomi[sz]ed|non-inferiority trial)") ~ "DoE",
                              TRUE ~ "No")) %>% 
  mutate(construct = case_when(str_detect(tolower(paste(Title, Description)), "construct|comput(e|ing)|creat(e|ing)|find(ing)") ~ "yes",
                               TRUE ~ "no"
                                          )) %>% 
  arrange(desc(criteria), construct) %>% 
  select(Package, Title, Description, taskview_doe, taskview_survey, criteria, construct) %T>% 
  write_csv(file = "output/design_classification.csv") %>% 
  filter(criteria=="DoE") %>% 
  pull(Package)
```

```{r}
Rd2list <- function(Rd){
    names(Rd) <- substring(sapply(Rd, attr, "Rd_tag"),2);
    temp_args <- Rd$arguments;

    Rd$arguments <- NULL;
    myrd <- lapply(Rd, unlist);
    myrd <- lapply(myrd, paste, collapse="");

    temp_args <- temp_args[sapply(temp_args , attr, "Rd_tag") == "\\item"];
    temp_args <- lapply(temp_args, lapply, paste, collapse="");
    temp_args <- lapply(temp_args, "names<-", c("arg", "description"));
    myrd$arguments <- temp_args;
    return(myrd);
}

getHelpList <- function(fn, pkg){
    thefile <- do.call("help", list(topic = fn, package = pkg))
    myrd <- utils:::.getHelpFile(thefile);
    Rd2list(myrd)$description %>% 
      str_replace_all("\\n", "") %>% 
      str_squish() %>% 
      str_trim() 
}
pkg <- "agricolae"
library(pkg, character.only = TRUE)
fn <- lsf.str(glue::glue("package:{pkg}"))[2]
getHelpList(fn, pkg)
```

```{r cran-data}
end <- Sys.Date() - 2 # usually 1-2 days data are not available yet
start <- end - years(2) + days(2)
dldat <- cran_downloads(fdoe_pkgs, from = start, to = end)
```

```{r summary}
dldat_sum <- dldat %>% 
  group_by(package) %>% 
  summarise(total = sum(count)) %>% 
  arrange(desc(total)) %>% 
  mutate(rank = 1:n(),
         taskview = package %in% doe_pkgs)
```

```{r graph}
g1 <- ggplot(dldat_sum, aes(x = taskview, y = total/1000, label = package, text = rank, color = taskview)) + 
  geom_beeswarm() +
  scale_y_log10(label = comma) + 
  labs(x = "",
       y = "Download counts in the last 2 year ('000s)") + 
  theme(axis.ticks.length.x = unit(0, "mm"))

ggplotly(g1)
```

```{r updates}
pkg_url <- "https://cran.r-project.org/web/packages/{pkg}/index.html"
pkg_archive <- "https://cran.r-project.org/src/contrib/Archive/{pkg}/"
pkg_updates <- map(fdoe_pkgs, function(pkg) {
    last_update <- read_html(glue(pkg_url)) %>% 
      html_table() %>% 
      .[[1]] %>% 
      filter(X1=="Published:") %>% 
      pull(X2) %>% 
      ymd()

    archive_dates <- tryCatch({ 
        read_html(glue(pkg_archive)) %>% 
          html_table() %>%
          .[[1]] %>% 
          pull(`Last modified`) %>% 
          ymd_hm() %>% 
          na.omit() %>% 
          as.Date()
      }, error = function(e) {
        NULL
      })
    c(archive_dates, last_update)
  })
names(pkg_updates) <- doe_pkgs

updates <- unlist(pkg_updates) %>% 
  enframe("package", "update") %>% 
  # unlist converts date to integers
  mutate(update = as.Date(update, origin = "1970-01-01"),
         # need to get rid of the numbers appended to pkg names
         package = str_extract(package, paste0(doe_pkgs, collapse="|"))) 
```

```{r first}
pkgsum <- updates %>% 
  group_by(package) %>% 
  summarise(first = min(update),
            nupdates = n()) %>% 
  left_join(dldat_sum, by = "package")

g2 <- ggplot(pkgsum, aes(first, total, label = package, text = rank, size = nupdates)) + 
  geom_point(alpha = 0.4) + 
  scale_y_log10() + 
  #geom_smooth() +
  NULL
ggplotly(g2)
```

* Packages released after 2017 September doesn't seem to be listed in CRAN Task View

```{r}
g3 <- ggplot(pkgsum, aes(nupdates, total, label = package, text = rank)) + 
  geom_point(alpha = 0.4) + 
  scale_y_log10() + 
  scale_x_log10() + 
  #geom_smooth() +
  NULL
ggplotly(g3)
```

### 2022-06-18 other/analysis/doe-pkgs-analysis.Rmd
make it easier for people to navigate the repo

--- 2021-10-03 analysis/doe-pkgs-analysis.Rmd
@@
 output: html_document
-runtime: shiny
 ---
@@
 #summarise the package based on its date of published 
-pkg_tidy_long_type %>%
-  filter(relevance != "no") %>%
+pkg_year <- pkg_type %>%
   group_by(year) %>%
   count() %>%
-  ggplot(aes(x = year, y = n)) +
+  ungroup()
+
+ggplot(pkg_year, aes(x = year, y = n, group = 1)) +
   geom_line() +
   geom_point()
-```
-
-```{r}
-pkg_tidy_long_type %>%
-  filter(relevance != "no") %>%
-  group_by(domain) %>%
-  count() %>%
-  arrange(desc(n))
-```
-
-```{r}
-pkg_tidy_long_main <- pkg_tidy %>%
-  unnest(main) %>%
-  separate(main, into = c("funct_name", "funct_type", "funct_usage"), sep = ' - ')
-```
-
-```{r}
-pkg_tidy_long_main %>%
-  filter(relevance != "no") %>%
-  group_by(funct_type) %>%
-  count() %>%
-  arrange(desc(n))
-```
-## Network analysis
-
-### Create the depends/imports/suggest data frame
-
-```{r}
-# make list of doe package 
-list_pkg <- as.list(pkg_tidy)$L2
-
-# create function to extract the depends, imports, and suggests 
-get_pkg <- function(var){
-  # select only the package name and depends/imports/suggests
-  tmp_df <- pkg_tidy %>%
-    select(L2, author, var) %>%
-    # separate every depends/imports/suggests
-    cSplit(var, sep = " ,")
-
-  df <- tmp_df %>%
-    pivot_longer(c(3:ncol(tmp_df)), names_to = "type", values_to = "pkg") %>%
-    filter(!is.na(pkg)) %>%
-    # check if the depends/imports/suggests is also a DoE package (is in list_pkg)
-    mutate(doe_pkg_related = ifelse(pkg %in% list_pkg, "yes", "no")) %>%
-    # filter only package whose depends/imports/suggests is a DoE
-    filter(doe_pkg_related == "yes") %>%
-    # tidy the data to only have the type: depends/imports/suggests
-    cSplit("type", sep = "_") %>%
-    select(-type_2) %>%
-    rename(type = type_1)
-
-  return(df)
-}
-```
-
-### Create network data
-
-```{r}
-# create data frame from the above function
-depend_df <- get_pkg("depends")
-import_df <- get_pkg("imports")
-suggest_df <- get_pkg("suggests")
-
-# bind data frames
-pkg_edges <- rbind(depend_df, import_df, suggest_df) %>%
-  select(-doe_pkg_related) %>%
-  select(pkg, L2, type, author)
-
-# extract unique package that has depends/imports/suggests
-from <- pkg_edges %>%
-  distinct(L2)
-
-# extract unique package that being depends/imports/suggests
-to <- pkg_edges %>%
-  distinct(pkg)
-
-# create nodes from the two previous dataframe
-pkg_nodes_tmp <- full_join(from, to, by = c("L2" = "pkg")) %>%
-  left_join(pkg_tidy, by = "L2") %>%
-  select(L2, author)
-
-#### tidy author
-
-## separate author 
-aut_tidy <- pkg_nodes_tmp %>%
-  # separate each aut/ctb/cph for each package
-  separate_rows(author, sep=", (?=[A-Z])") %>%
-  # filter only role aut
-  filter(!str_detect(author, "ctb") & !str_detect(author, "National") & !str_detect(author, "Columbia"))
-
-# replace character inside angle and square bracket
-aut_tidy$author <- rm_angle(aut_tidy$author, replacement = "")
-aut_tidy$author <- rm_square(aut_tidy$author, replacement = "")
-
-# remove contributors written as sentence, also remove cph 
-aut_tidy <- aut_tidy %>%
-  mutate(author = ifelse(str_detect(author, "Contributor"), "Yves Deville",
-                         ifelse(str_detect(author, "based"), "Ernesto Barrios",
-                                ifelse(str_detect(author, "contributions"), "John Lawson", author))))
-
-# extract author who wrote >1 packages
-aut_group <- aut_tidy %>%
-  group_by(author) %>%
-  count() %>%
-  arrange(desc(n)) %>%
-  filter(n > 1)
-
-## tidy author in pkg_nodes_tmp 
-pkg_nodes_tmp$author <- rm_angle(pkg_nodes_tmp$author, replacement = "")
-pkg_nodes_tmp$author <- rm_square(pkg_nodes_tmp$author, replacement = "")
-
-```
-
-```{r}
-aut_tidy_join <- left_join(aut_tidy, aut_group, by = "author") %>%
-  group_by(L2) %>%
-  arrange(desc(n), .by_group = TRUE)
-
-pkg_nodes <- aut_tidy_join[!duplicated(aut_tidy_join$L2), ] %>%
-  select(L2, aut_gr)
-
-# nodes for all of the author in DoE packages
-pkg_nodes_full <- pkg_tidy%>%
-  select(L2)
-
-# create the network package 
-pkg_net <- pkg_edges %>%
-  full_join(pkg_nodes, by = "L2")
-```
-
-> Note
-
-Some packages are authored by the same persons. For Example, Dice
-
-## Create network using igraph
-
-```{r}
-routes_igraph <- graph_from_data_frame(d = pkg_edges, vertices = pkg_nodes, directed = TRUE)
-
-# set.seed(1234567)
-# plot(routes_igraph, edge.arrow.size = 0.2, layout = layout_with_graphopt)
-# Notes: The label overlapped, it is hard to see the network
-```
-
-## Create network using tidygraph and ggraph
-
-```{r}
-# create the input based on the igraph object
-routes_tidy <- routes_igraph %>% as_tbl_graph() %>%
-  activate(edges)
-
-routes_tidy
-```
-
-```{r}
-ggraph(routes_tidy, layout = "stress") + 
-  geom_edge_link(aes(color = type),
-                 arrow = arrow(length = unit(1, 'mm')), 
-                 end_cap = circle(1, 'mm')) + 
-  geom_node_point(aes(color = aut_gr)) + 
-  geom_node_text(aes(label = name), repel = TRUE, size = 2) +
-  theme_graph()
-```
-
-```{r}
-aut_tidy_1st_aut <- aut_tidy_join[!duplicated(aut_tidy_join$L2), ] %>%
-  select(L2, author)
-
-aut_gr <- aut_tidy_1st_aut %>%
-  group_by(author) %>%
-  count() %>%
-  ungroup() %>%
-  arrange(desc(n)) %>%
-  mutate(aut_gr = as.factor(ifelse(n > 1, author, "other")))
-
-pkg_nodes_1aut <- left_join(aut_tidy_1st_aut, aut_gr, by = "author") %>%
-  select(L2, aut_gr) %>%
-  mutate(first_aut = factor(aut_gr, levels = c("Ulrike Groemping",
-                                           "John Lawson",
-                                           "Robert B. Gramacy",
-                                           "Delphine Dupuy",
-                                           "Ernesto Barrios",
-                                           "Kevin Wright",
-                                           "Peng Liu",
-                                           "Shwetank Lall",
-                                           "other"))) %>%
-  select(L2, first_aut)
-
-routes_igraph_1aut <- graph_from_data_frame(d = pkg_edges, vertices = pkg_nodes_1aut, directed = TRUE)
-
-# create the input based on the igraph object
-routes_tidy_1aut <- routes_igraph_1aut %>% as_tbl_graph() %>%
-  activate(edges)
-
-ggraph(routes_tidy_1aut, layout = "stress") + 
-  geom_edge_link(aes(color = type),
-                 arrow = arrow(length = unit(1, 'mm')), 
-                 end_cap = circle(1, 'mm')) + 
-  geom_node_point(aes(color = first_aut)) + 
-  geom_node_text(aes(label = name), repel = TRUE, size = 2) +
-  theme_graph()
-```
-
-# Network graph with all the packages in CRAN task view
-
-```{r}
-routes_igraph_full <- graph_from_data_frame(d = pkg_edges, vertices = pkg_nodes_full, directed = TRUE)
-
-routes_tidy_full <- routes_igraph_full %>% as_tbl_graph() %>%
-  activate(edges)
-
-routes_tidy_full
-```
-
-```{r}
-ggraph(routes_tidy_full, layout = "stress") + 
-  geom_edge_link(aes(color = type),
-                 arrow = arrow(length = unit(1, 'mm')), 
-                 end_cap = circle(1, 'mm')) + 
-  geom_node_point() + 
-  geom_node_text(aes(label = name), repel = TRUE, size = 2) +
-  theme_graph()
-```
-
-Note: Hard to read. Hence, with all of the package in CRAN task view, I will make a heatmap instead.
-
-```{r}
-pkg.mat = get.adjacency(routes_igraph_full, sparse = FALSE)
-heatmap(pkg.mat[,109:1],Rowv = NA, Colv = NA,scale="none")
-```
-
-Next:
-1. Color the nodes based on author (Binary variable whether the pkg is authored by Ulrike or not)
-2. Using interactive graph with D3
-3. Write on the paper
-
-# Interactive Plot with visnetwork
-
-```{r}
-nodes_d3 <- pkg_nodes_1aut %>%
-  rowid_to_column("id") %>%
-  mutate(id = id - 1)
-
-edges_d3 <-routes_tidy_1aut %>%
-  activate(edges) %>%
-  data.frame() %>%
-  mutate(from = from - 1,
-         to = to - 1)
-
-forceNetwork(Links = edges_d3, Nodes = nodes_d3, Source = "from", Target = "to", 
-             NodeID = "L2", Group = "first_aut", 
-             opacity = 1, fontSize = 16, zoom = TRUE)
-```
-
-# Comparison to other CRAN TASK VIEW
-
-## Survey
-
-```{r}
-list_pkg_survey <- ctv:::.get_pkgs_from_ctv_or_repos("OfficialStatistics", 
-                                                repos = "http://cran.rstudio.com/")[[1]]
-url <- "http://cran.rstudio.com/web/packages/packages.rds"
-db <- readRDS(url(url)) %>% 
-  as.data.frame()
-survey_pkg <- db %>%
-  filter(Package %in% list_pkg_survey) %>%
-  select(Package, Depends, Imports, Suggests)
-
-get_pkg_survey <- function(var){
-  # select only the package name and depends/imports/suggests
-  tmp_df <- survey_pkg %>%
-    select(Package, var) %>%
-    # separate every depends/imports/suggests
-    cSplit(var, sep = " ,")
-
-  df <- tmp_df %>%
-    pivot_longer(c(2:ncol(tmp_df)), names_to = "type", values_to = "pkg") %>%
-    filter(!is.na(pkg)) %>%
-    # check if the depends/imports/suggests is also a DoE package (is in list_pkg)
-    mutate(survey_pkg_related = ifelse(pkg %in% list_pkg_survey, "yes", "no")) %>%
-    # filter only package whose depends/imports/suggests is a DoE
-    filter(survey_pkg_related == "yes") %>%
-    # tidy the data to only have the type: depends/imports/suggests
-    cSplit("type", sep = "_") %>%
-    select(-type_2) %>%
-    rename(type = type_1)
-
-  return(df)
-}
-
-survey_depends <- get_pkg_survey("Depends")
-survey_imports <- get_pkg_survey("Imports")
-survey_suggests <- get_pkg_survey("Suggests")
-
-pkg_edges_survey <- rbind(survey_depends, survey_imports, survey_suggests)
-
-survey_from <- pkg_edges_survey %>%
-  distinct(Package)
-
-# extract unique package that being depends/imports/suggests
-survey_to <- pkg_edges_survey %>%
-  distinct(pkg)
-
-# create nodes from the two previous dataframe
-pkg_nodes_survey <- full_join(survey_from, survey_to, by = c("Package" = "pkg"))
-```
-
-## Timeseries
-
-```{r}
-list_pkg_ts <- ctv:::.get_pkgs_from_ctv_or_repos("TimeSeries", 
-                                                repos = "http://cran.rstudio.com/")[[1]]
-
-ts_pkg <- db %>%
-  filter(Package %in% list_pkg_ts) %>%
-  select(Package, Depends, Imports, Suggests)
-
-get_pkg_ts <- function(var){
-  # select only the package name and depends/imports/suggests
-  tmp_df <- ts_pkg %>%
-    select(Package, var) %>%
-    # separate every depends/imports/suggests
-    cSplit(var, sep = " ,")
-
-  df <- tmp_df %>%
-    pivot_longer(c(2:ncol(tmp_df)), names_to = "type", values_to = "pkg") %>%
-    filter(!is.na(pkg)) %>%
-    # check if the depends/imports/suggests is also a DoE package (is in list_pkg)
-    mutate(survey_pkg_related = ifelse(pkg %in% list_pkg_ts, "yes", "no")) %>%
-    # filter only package whose depends/imports/suggests is a DoE
-    filter(survey_pkg_related == "yes") %>%
-    # tidy the data to only have the type: depends/imports/suggests
-    cSplit("type", sep = "_") %>%
-    select(-type_2) %>%
-    rename(type = type_1)
-
-  return(df)
-}
-
-ts_depends <- get_pkg_ts("Depends")
-ts_imports <- get_pkg_ts("Imports")
-ts_suggests <- get_pkg_ts("Suggests")
-
-pkg_edges_ts <- rbind(ts_depends, ts_imports, ts_suggests)
-
-ts_from <- pkg_edges_ts %>%
-  distinct(Package)
-
-# extract unique package that being depends/imports/suggests
-ts_to <- pkg_edges_ts %>%
-  distinct(pkg)
-
-# create nodes from the two previous dataframe
-pkg_nodes_ts <- full_join(ts_from, ts_to, by = c("Package" = "pkg"))
 ```
@@
 
+# The domain of package based on year
+
 ```{r}
-sankey_data <- pkg_type %>%
-  select(domain, year) %>%
-  group_by(domain, year) %>%
+pkg_type_year <- pkg_type %>%
+  group_by(year, domain) %>%
   count() %>%
-  ungroup()
+  ungroup() %>%
+  left_join(pkg_year, by="year", suffix = c("_dom", "_year")) %>%
+  mutate(pct = round(n_dom/n_year*100, 0)) %>%
+  mutate(dom = case_when(domain == "general" ~ "Gral",
+                         domain == "clinical trial" ~ "CT",
+                         domain == "genetics" ~ "Gtic",
+                         domain == "computer experiment" ~ "CE",
+                         domain == "agriculture" ~ "Ag",
+                         domain == "industrial experiment" ~ "IE",
+                         domain == "environmental science" ~ "Env.S",
+                         domain == "consumer behaviour" ~ "CB",
+                         domain == "equivalence studies" ~ "Eq.S",
+                         domain == "geospatial study" ~ "GS",
+                         domain == "microbiology" ~ "Mbg",
+                         domain == "sensory studies" ~ "SS"))
 
-domain <- sankey_data %>%
-  distinct(domain) %>%
-  rename(node = domain)
+ggplot(pkg_type_year, aes(x = pct, y = dom)) +
+  geom_col() +
+  facet_wrap(~year, scales = "free_y") +
+  theme_bw()
 
-year <- sankey_data %>%
-  distinct(node = year)
+pkg_type_year %>%
+  mutate(dom = reorder(dom, pct)) %>%
+  group_by(year, dom) %>%
+  arrange(desc(pct)) %>%
+  ungroup() %>%
+  mutate(dom = factor(paste(dom, year, sep = "__"), 
+                       levels = rev(paste(dom, year, sep = "__")))) %>%
+  ggplot(aes(dom, pct)) +
+  geom_col() +
+  geom_text(aes(label=pct), size = 3, color = "black", 
+            hjust = -0.05, vjust=-0.05) +
+  facet_wrap(~year, scales = "free_y") +
+  scale_x_discrete(labels = function(x) gsub("__.+$", "", x)) +
+  coord_flip() +
+  theme_bw() +
+  xlab("Domain") +
+  ylab("Percentage") +
+  theme(panel.grid.major.y = element_blank())
 
-db_sankey <- rbind(domain, year)
-
-routes_sankey <- graph_from_data_frame(d = sankey_data, vertices = db_sankey, directed = TRUE)
-
-routes_sankey_tidy <- routes_sankey %>% as_tbl_graph() %>%
-  activate(edges)
-
-nodes_sankey <- db_sankey %>%
-  rowid_to_column("id") %>%
-  mutate(id = id - 1)
-
-edges_sankey <-routes_sankey_tidy %>%
-  activate(edges) %>%
-  data.frame() %>%
-  mutate(from = from - 1,
-         to = to - 1)
-
-sankeyNetwork(Links = edges_sankey, Nodes = nodes_sankey, Source = 'from', 
-                         Target = 'to', Value = 'n', NodeID = 'node',
-                         units = 'packages')
 ```

### 2022-12-13 paper.Rmd
huge update for index with cleaner pipeline

@@
     csl: assets/html-template.csl
-    self_contained: false
+    self_contained: true
     css: ["assets/html-template.css"]
@@
 ```{r setup, cache = FALSE, include = FALSE}
-library(targets)
 library(glue)
@@
 library(ggraph)
-source("R/plot.R")
+theme_set(
+  theme(
+    panel.background = element_rect(fill = NA),
+    panel.grid = element_line(color = "lightgray"),
+    axis.text = element_text(color = "black"),
+    axis.line = element_line(color = "black", size = 0.7),
+    axis.ticks.length = unit(1.4, "mm"),
+    axis.ticks = element_line(color = "black", size = 0.7),
+    axis.title = element_text(color = "black", face = "bold"),
+    strip.background = element_rect(color = "black",
+                                    fill = "white"),
+    strip.text = element_text(color = "black",
+                              margin = margin(5, 5, 5, 5)),
+    plot.title.position = "plot",
+    plot.title = element_text(color = "black", hjust = 0)
+  )
+)
 knitr::opts_chunk$set(echo = FALSE, 
@@
 ref <- function(x) {
-  ifelse(knitr::is_html_output(), glue("\\@ref({x})"), glue("\\ref{[x]}", .open = "[", .close = "]"))
+  case_when(knitr::is_html_output() ~ glue("\\@ref({x})"),
+            TRUE ~ glue("\\ref{[x]}", .open = "[", .close = "]"))
 }
@@
 }
+rpkg <- function(x) {
+  glue("`{x}`")
+}
 
-tar_load(npkgs)
-tar_load(gini_yearly_doe)
-tar_load(gini_yearly_ctv)
-tar_load(doe_lorenz2021)
+ctv_pkg_data <- readRDS("output/ctv_pkg_data.rds")
+ctv_summary_table <- readRDS("output/ctv_summary_table.rds")
+doe_bigram_all_with_download_data <- readRDS("output/doe_bigram_all_with_download_data.rds")
+doe_pkg_network <- readRDS("output/doe_pkg_network.rds")
+ctv_gini_yearly <- readRDS("output/ctv_gini_yearly.rds")
+ctv_lorenz_yearly <- readRDS("output/ctv_lorenz_yearly.rds")
+ctv_rank_yearly <- readRDS("output/ctv_rank_yearly.rds")
+access_date_pkg_downloads <- readRDS("output/access_date_pkg_downloads.rds")
+
+doe_npkgs <- ctv_summary_table %>% 
+  filter(name=="ExperimentalDesign") %>% 
+  pull(n)
+
+doe_npkgs_2021 <- ctv_gini_yearly %>% 
+  filter(year==2021, name=="ExperimentalDesign") %>% 
+  pull(n)
+
+doe_ginis <- ctv_gini_yearly %>% 
+  filter(name == "ExperimentalDesign") %>% 
+  pull(gini)
 
 lorenz_value <- function(p) {
-  doe_lorenz2021 %>% 
+  ctv_lorenz_yearly %>% 
+    filter(name == "ExperimentalDesign",
+           year == 2021) %>% 
+    unpack(lorenz_data) %>% 
     mutate(pdist = abs(p - .env$p)) %>% 
@@
 
-<!-- Unlike many other data collection schemes, experiments provides the greatest control and consequently, a greater opportunity to statistically optimise the information drawn from the experimental data.
--->
-
 Methods for data collection can be dichotomised by the type of data collected -- namely, experimental or observational -- or alternatively, categorised as experimental design (including quasi-experimental design) or survey design.
@@
 
-In `r ctv("ExperimentalDesign")` task view, there are `r npkgs` R packages for experimental design and analysis of data from experiments.
+In `r ctv("ExperimentalDesign")` task view, there are `r doe_npkgs` R packages for experimental design and analysis of data from experiments.
 The sheer quantity and variation of experimental designs in the R-packages are arguably unmatched with any other programming languages, e.g.
@@
 Certain programming language design can hinder or discourage development of a reliable program [@Wasserman1975-xr].
-The immense popularity of `tidyverse` [a collection of R-packages for various stages of data analysis that place enormous emphasis on the interface design by @Wickham2019-mj] is a testament to the impact an interface design can have in practice.
+The immense popularity of `r rpkg("tidyverse")` [a collection of R-packages for various stages of data analysis that place enormous emphasis on the interface design by @Wickham2019-mj] is a testament to the impact an interface design can have in practice.
 The practice of experimental design could be advanced by adopting similar interface design principles across DoE packages.
@@
 The RStudio mirror is also the only server that provides a comprehensive daily download logs of R and R-packages since October 2012.
-The summary data can be easily accessed with the `cranlogs` package [@cranlogs].
+The summary data can be easily accessed with the `r rpkg("cranlogs")` package [@cranlogs].
 This paper uses the data from the beginning of 2013 to end of 2021 (a total of 9 years) for the packages in the CRAN task views.
@@
 All CRAN packages have a title, description, package connections (suggests, depends and imports of other packages) and other meta-information in the DESCRIPTION file.
-We use the text data from the title and description (as accessed in `r as.Date(file.info("_targets/objects/ngram_download_data")$mtime)`) in Section `r ref("topics")`.
+We use the text data from the title and description (as accessed in `r access_date_pkg_downloads`) in Section `r ref("topics")`.
 
@@
 CRAN task views are volunteer maintained list of R-packages on CRAN relevant to the corresponding topic.
-There are a total `r n_distinct(gini_yearly_ctv$ctv)` CRAN task views.
-Table S1 in the Supplementary Materials list the full available topics from the `ctv` package [@ctv].
-The list of packages in each CRAN task views (as of `r as.Date(file.info("_targets/objects/ctv_pkgs_data")$mtime)`) are used to contrast the characteristics of DoE packages in Sections `r ref("popular")` and `r ref("silo")`.
+There are a total `r n_distinct(ctv_pkg_data$name)` CRAN task views.
+Table S1 in the Supplementary Materials list the full available topics from the `r rpkg("ctv")` package [@ctv].
+The list of packages in each CRAN task views (as of `r access_date_pkg_downloads`) are used to contrast the characteristics of DoE packages in Sections `r ref("popular")` and `r ref("silo")`.
 
@@
 There are at least 50 DoE packages since 2013 but most of the downloads are concentrated in just a handful of packages.
-For example, Figure `r ref("fig:plot-lorenz")` shows a Lorenz curve [@Lorenz1905-tc] for the total package downloads in 2021 of `r tar_read(gini_yearly) %>% filter(year==2021) %>% pull(n)` DoE packages (first released prior to 2021); we can see from Figure `r ref("fig:plot-lorenz")` that bottom 90\% of DoE packages (in terms of total download count in 2021) only share about `r lorenz_value(0.9)`\% of total downloads across all DoE packages -- in another words, `r 100 - lorenz_value(0.9)`\% of the total downloads are due to `r round(npkgs * 0.1)` packages (10% of the DoE packages).
+For example, Figure `r ref("fig:plot-lorenz")` shows a Lorenz curve [@Lorenz1905-tc] for the total package downloads in 2021 of `r doe_npkgs_2021` DoE packages (first released prior to 2021); we can see from Figure `r ref("fig:plot-lorenz")` that bottom 90\% of DoE packages (in terms of total download count in 2021) only share about `r lorenz_value(0.9)`\% of total downloads across all DoE packages -- in another words, `r 100 - lorenz_value(0.9)`\% of the total downloads are due to `r round(doe_npkgs_2021 * 0.1)` packages (10% of the DoE packages).
 
@@
 The yellow region shows the area under the Lorenz curve and the red region shows the area of gap in equality."}
-tar_read(plot_lorenz2021) +
-  plot_setup() +
+ctv_lorenz_yearly %>% 
+  filter(name == "ExperimentalDesign",
+         year == 2021) %>% 
+  ggplot(aes(lorenz_data$p * 100, lorenz_data$L * 100)) +
+  geom_abline(color = "red") +
+  geom_ribbon(aes(ymin = 0, ymax = lorenz_data$L * 100),
+                fill = "#FFC107", alpha = 0.5) +
+  geom_ribbon(aes(ymin = lorenz_data$L * 100, ymax = lorenz_data$p * 100),
+                 fill = "#D81B60", alpha = 0.5) +
+  geom_line() + 
+  geom_point() +
+  annotate("text", x = 75, y = 80, 
+           label = "Line of perfect equality",
+            angle = 45) +
   labs(y = "Cumulative share of downloads (%)",
@@
 A Gini index of 0\% indicates equality in downloads across packages while a value of 100\% indicates maximal inequality (all downloads are due to one package).
-In Figure `r ref("fig:download-share")`, we see that the distributions of the package downloads each year have a heavy right tail with the Gini index ranging from `r scales::percent(min(gini_yearly_doe$gini), 0.1)` to `r scales::percent(max(gini_yearly_doe$gini), 0.1)` across the years 2013 to 2021, indicating that there is a high level of inequality of package downloads, particularly with more pronounced inequality in the last 6 years.
+In Figure `r ref("fig:download-share")`, we see that the distributions of the package downloads each year have a heavy right tail with the Gini index ranging from `r scales::percent(min(doe_ginis), 0.1)` to `r scales::percent(max(doe_ginis), 0.1)` across the years 2013 to 2021, indicating that there is a high level of inequality of package downloads, particularly with more pronounced inequality in the last 6 years.
 
@@
 "}
-tar_read(plot_download_dist) +
-  labs(y = "Number of downloads")
+ctv_rank_yearly %>% 
+    filter(name == "ExperimentalDesign") %>% 
+    mutate(year = as.factor(year)) %>% 
+    ggplot(aes(x = year, y = total)) + 
+    geom_violin() +
+    geom_boxplot(width = 0.1) + 
+    labs(x = "Year", y = "Number of downloads") + 
+    scale_y_log10(label = scales::comma, expand = c(0.1, 0)) +
+    geom_label(data = ctv_gini_yearly %>% 
+                        filter(name == "ExperimentalDesign") %>% 
+                        mutate(year = as.factor(year)), 
+               size = 3,
+               aes(y = 900, 
+                   label = paste0(scales::percent(gini, 0.1),
+                                   "\nn=", n)))
 ```
@@
 The increase in the number of packages that are not highly downloaded may mean that **_there are more packages to construct niche experimental designs_**.
-Some examples of these packages include `qtlDesign`, `PwrGSD` and `Crossover` made for QTL experiments, group sequential designs and crossover trials, respectively.
+Some examples of these packages include `r rpkg("qtlDesign")`, `r rpkg("PwrGSD")` and `r rpkg("Crossover")` made for QTL experiments, group sequential designs and crossover trials, respectively.
 These packages would naturally have a smaller number of potential users.
@@
 
-While in absolute terms the Gini index is high for `r ctv("ExperimentalDesign")` task view (`r scales::percent(min(gini_yearly_doe$gini), 0.1)` to `r scales::percent(max(gini_yearly_doe$gini), 0.1)`), the inequality is not as severe as other CRAN task views as shown in Figure `r ref("fig:fig-gini-all-ctvs")`.
+While in absolute terms the Gini index is high for `r ctv("ExperimentalDesign")` task view (`r scales::percent(min(doe_ginis), 0.1)` to `r scales::percent(max(doe_ginis), 0.1)`), the inequality is not as severe as other CRAN task views as shown in Figure `r ref("fig:fig-gini-all-ctvs")`.
 We can see in Figure `r ref("fig:fig-gini-all-ctvs")` that the Gini index is generally increasing for DoE packages (as is generally the case for other CRAN task views as shown in Figure S1 in the Supplementary Materials) but most other CRAN task views have a Gini index of over 75%.
@@
 See Figure S1 in the Supplementary Material for the line graph for the Gini index across years by each CRAN task view."}
-gini_yearly_ctv %>% 
-  ungroup() %>%  
-  arrange(ctv, year) %>% 
-  mutate(ctv = fct_reorder(ctv, gini, last),
+ctv_gini_yearly %>% 
+  arrange(name, year) %>% 
+  mutate(name = fct_reorder(name, gini, last),
          year = factor(year),
          gini = 100 * gini) %>%  
-  ggplot(aes(year, gini, group = ctv)) + 
+  ggplot(aes(year, gini, group = name)) + 
   geom_line(color = "grey") +
-  geom_line(data = ~filter(.x, ctv=="ExperimentalDesign")) + 
-  geom_point(data = ~filter(.x, ctv=="ExperimentalDesign"),
+  geom_line(data = ~filter(.x, name=="ExperimentalDesign")) + 
+  geom_point(data = ~filter(.x, name=="ExperimentalDesign"),
              aes(color = n)) +
-  plot_setup() +
   labs(x = "Year", y = "Gini index (%)", color = "# of packages",
@@
 
-We can see in Figure `r ref("fig:rank-over-time")` that most of the top 10 ranking packages have been in the top 10 for the last 9 years with `lhs` steadily climbing up the ranks in the last few years.
-It should be noted that the download of one package can prompt a download of another package; the most notable package connection is `AlgDesign` and `agricolae`, where the former is an import for the latter.
+We can see in Figure `r ref("fig:rank-over-time")` that most of the top 10 ranking packages have been in the top 10 for the last 9 years with `r rpkg("lhs")` steadily climbing up the ranks in the last few years.
+It should be noted that the download of one package can prompt a download of another package; the most notable package connection is `r rpkg("AlgDesign")` and `r rpkg("agricolae")`, where the former is an import for the latter.
 The full network of packages connections within the DoE packages is shown later in Figure `r ref("fig:plot-doe-network")`.
@@
 Most packages are consistently in the top 10 for the period shown."}
-tar_read(plot_rank_dist) + labs(color = "Package")
+nrank <- 10
+pkgs_repeat_toprank <- ctv_rank_yearly %>% 
+  filter(name=="ExperimentalDesign") %>% 
+  filter(rank <= nrank) %>% 
+  count(package) %>% 
+  filter(n > 1) %>% 
+  pull(package)
+
+ctv_rank_yearly %>% 
+  filter(name=="ExperimentalDesign",
+         package %in% pkgs_repeat_toprank) %>% 
+  mutate(year = as.factor(year)) %>% 
+  ggplot(aes(year, rank, color = package)) +
+  geom_point() + 
+  geom_line(aes(group = package)) + 
+  scale_y_reverse(breaks = 1:nrank) +
+  geom_text(aes(label = package),
+            data = ~filter(.x, year == 2021),
+            hjust = 0,
+            nudge_x = 0.1) +
+  guides(color = "none") + 
+  scale_x_discrete(expand = expansion(c(0.05, 0.25))) +
+  labs(color = "Package", x = "Year", y = "Download rank") +
+  scale_color_manual(values = c("#E69F00", "#56B4E9", "#009E73", 
+                                  "#F0E442", "#0072B2", "#D55E00", 
+                                  "#CC79A7", "#000000", "#888888",
+                                  "#88CCEE", "#CC6677", "seagreen4", 
+                                  "violetred4"))
 ```
@@
 The high-leverage point on the far left belongs to `conf.design`, authored by one of the earlier contributors to R."}
-tar_read(cran_rank_yearly) %>%
-  mutate(nupdates = map2_int(package, year, ~{
-    tar_read(updates_all) %>% 
-      filter(package == .x,
-             lubridate::year(update) <= .y) %>% 
-      nrow()
-  })) %>% 
+ctv_rank_yearly %>% 
+  filter(name == "ExperimentalDesign") %>% 
   ggplot(aes(first, total/1000)) +
@@
   scale_color_continuous(trans = "log10") + 
-  plot_setup() +
   labs(x = "First release date",
@@
 1. We standardised the words to lower case and removed pluralization. 
-2. Multiple mentions of the same bigram within a package was counted as one (e.g. `AlgDesign` mentions "experimental design" four times in the title and the description but this is counted as one). 
+2.
+Multiple mentions of the same bigram within a package was counted as one (e.g.
+`r rpkg("AlgDesign")` mentions "experimental design" four times in the title and the description but this is counted as one).
 3.
@@
 The size shows how often the bigram appears across the DoE packages and the color are relative to the total download count in 2021 of the packages that contain the bigram."}
-tar_read(plot_wordcloud2) + 
+doe_bigram_all_with_download_data %>%
+  filter(perc >= 1, n > 1, year == 2021) %>%
+  ggplot(aes(
+    label = word,
+    color = perc,
+    size = n
+  )) +
+  geom_text_wordcloud(shape = "square",
+                      show.legend = TRUE) +
+  guides(size = "none") +
+  theme(legend.position = "bottom",
+        text = element_text(size = 18)) +
+  guides(color = guide_colorbar(title.position = "top")) +
   scale_size_area(max_size = 5) +
@@
 
-Figure `r ref("fig:ctv-summ-plot")` shows that the `r ctv("ExperimentalDesign")` task view has the lowest average number of contributors among all `r n_distinct(gini_yearly_ctv$ctv)` CRAN task views.
+Figure `r ref("fig:ctv-summ-plot")` shows that the `r ctv("ExperimentalDesign")` task view has the lowest average number of contributors among all `r n_distinct(ctv_gini_yearly$name)` CRAN task views.
 In addition, we can also see in Figure `r ref("fig:ctv-summ-plot")`  that the `r ctv("ExperimentalDesign")` task view has one of the least intra-connectivity (the percentage of packages that make use of other packages within the same task view).
@@
 The actual numerical values are show in Table S1 in the Supplementary Material."}
-tar_read(ctv_summary_table) %>% 
+ctv_summary_table %>% 
   ggplot(aes(avg_contributors, intra_perc)) +
@@
        color_new = "Total number of packages") +
-  plot_setup() +
   theme(legend.position = "bottom",
@@
 DoE packages that do not depend, suggest or import another DoE package is not shown."}
-tar_read(plot_doe_network) +
+doe_pkg_network %>%
+  ggraph('fr') +
+  geom_edge_link(aes(color = Type),
+                 arrow = arrow(length = unit(3, 'mm')),
+                 end_cap = circle(3, 'mm')) +
+  geom_node_point(size = 4) +
+  geom_node_text(aes(label = name), repel = TRUE, size = 3.5) +
+  theme_graph(base_family = "Helvetica") +
+  labs(color = "") +
   theme(legend.position = "bottom")
@@
 In this section, we discuss the interface designs of functions that output an experimental design based on three broad areas: factorial designs, recipe designs and augmenting designs.
-The discussion is exclusive to the top downloaded packages (shown in Figure `r ref("fig:rank-over-time")`), with the exception of `ez` and `DiceKriging`, as the former is predominately visualisation of experimental data and the latter is about the analysis of computer experiments in addition to belonging to the same suite of packages as `DiceDesign` [@DiceDesign].
+The discussion is exclusive to the top downloaded packages (shown in Figure `r ref("fig:rank-over-time")`), with the exception of `r rpkg("ez")` and `r rpkg("DiceKriging")`, as the former is predominately visualisation of experimental data and the latter is about the analysis of computer experiments in addition to belonging to the same suite of packages as `r rpkg("DiceDesign")` [@DiceDesign].
 
@@
 
-The `DoE.base` package [@DoE.base] can construct full factorial and (regular and irregular) orthogonal array designs via `fac.design()` and `oa.design()`, respectively.
-The `FrF2` package [@FrF2] constructs regular fractional 2-level factorial designs using `FrF2()` and `FrF2Large()` with the latter for large designs.
-The `conf.design` package  [@conf.design] constructs symmetric confounded factorial designs via `conf.design()`, and `BHH2` package [@BHH2] generates full or fractional 2-level factorial design matrix via `ffDesMatrix()`.
+The `r rpkg("DoE.base")` package [@DoE.base] can construct full factorial and (regular and irregular) orthogonal array designs via `fac.design()` and `oa.design()`, respectively.
+The `r rpkg("FrF2")` package [@FrF2] constructs regular fractional 2-level factorial designs using `FrF2()` and `FrF2Large()` with the latter for large designs.
+The `r rpkg("conf.design")` package  [@conf.design] constructs symmetric confounded factorial designs via `conf.design()`, and `r rpkg("BHH2")` package [@BHH2] generates full or fractional 2-level factorial design matrix via `ffDesMatrix()`.
 
@@
 The output of the design is either a special class of list (e.g.
-the `design` class for `DoE.base` and `FrF2` or a `data.frame` for `conf.design`) or a matrix for `BHH2` such that an element or column corresponds to a treatment factor with each value corresponding to one experimental run.
+the `design` class for `r rpkg("DoE.base")` and `r rpkg("FrF2")` or a `data.frame` for `r rpkg("conf.design")`) or a matrix for `BHH2` such that an element or column corresponds to a treatment factor with each value corresponding to one experimental run.
 The treatment factor are generally assigned pseudo names (e.g.
@@
 *response surface designs* are factorial designs where the treatment factors are discrete levels of continuous variables.
-Two types of response surface designs can be constructed using the `rsm` package [@rsm]: Box-Behnken design [@Box1960-wr] and central-composite designs [@Box1951-ji] via functions `bbd()` and `ccd()`, respectively, where the minimum required input is the number of factors.
+Two types of response surface designs can be constructed using the `r rpkg("rsm")` package [@rsm]: Box-Behnken design [@Box1960-wr] and central-composite designs [@Box1951-ji] via functions `bbd()` and `ccd()`, respectively, where the minimum required input is the number of factors.
 Another form of factorial designs is *saturated designs* where typically higher-order interaction effects of treatment factors are confounded with the main effects.
-Plankett-Burnman designs [@Plackett1946-ly] is one type of saturated design which can be generated by the function `pb()` in the `FrF2` package, where the user provides the number of experimental runs and the number of treatment factors.
+Plankett-Burnman designs [@Plackett1946-ly] is one type of saturated design which can be generated by the function `pb()` in the `r rpkg("FrF2")` package, where the user provides the number of experimental runs and the number of treatment factors.
 
@@
 
-The `agricolae` package [@agricolae] is the prime example of constructing designs based on a set of so-called "recipe functions", where each function corresponds to a single class of experimental design.
+The `r rpkg("agricolae")` package [@agricolae] is the prime example of constructing designs based on a set of so-called "recipe functions", where each function corresponds to a single class of experimental design.
 E.g.
@@
 
-The use of recipe functions is not limited to classical experimental designs; the `AlgDesign` [@AlgDesign] package offers three primary functions for generating optimal designs: `optBlock()`, `optFederov()` and `optMonteCarlo()`.
+The use of recipe functions is not limited to classical experimental designs; the `r rpkg("AlgDesign")` [@AlgDesign] package offers three primary functions for generating optimal designs: `optBlock()`, `optFederov()` and `optMonteCarlo()`.
 These functions in general require a data and formula in terms of the supplied data with the choice of criterion (e.g.
@@
 
-Computer experiments, which generally involve space filling designs, are implemented in packages like `lhs` [@lhs] and `DiceDesign` [@DiceDesign].
-For the `lhs` package, functions like `randomLHS()`, `optimumLHS()`, and `maximinLHS()`, require users to specify the sample size ($n$) and the number of variables ($p$) then it generates a Latin hypercube sample [@McKay1979-aw], based on different optimisation schemes (in this case, random, S optimal and maxmin criteria, respectively; see package documentation for more details).
-Similarly for the `DiceDesign`, there is a comprehensive list of space-filling designs such as `dmaxDesign()`, `lhsDesign()`, and `wspDesign()` with input of $n$ and $p$ as before (among extra parameters for some) that implement algorithms that either maximise the entropy [@Shewry1987-jn], produce a random Latin hypercube sample, and use the WSP algorithm [@Santiago2012-ui], respectively.
+Computer experiments, which generally involve space filling designs, are implemented in packages like `r rpkg("lhs")` [@lhs] and `r rpkg("DiceDesign")` [@DiceDesign].
+For the `r rpkg("lhs")` package, functions like `randomLHS()`, `optimumLHS()`, and `maximinLHS()`, require users to specify the sample size ($n$) and the number of variables ($p$) then it generates a Latin hypercube sample [@McKay1979-aw], based on different optimisation schemes (in this case, random, S optimal and maxmin criteria, respectively; see package documentation for more details).
+Similarly for the `r rpkg("DiceDesign")`, there is a comprehensive list of space-filling designs such as `dmaxDesign()`, `lhsDesign()`, and `wspDesign()` with input of $n$ and $p$ as before (among extra parameters for some) that implement algorithms that either maximise the entropy [@Shewry1987-jn], produce a random Latin hypercube sample, and use the WSP algorithm [@Santiago2012-ui], respectively.
 These designs all output a $n \times p$ matrix with values between 0 and 1.
@@
 Some functions in DoE packages require input of existing experimental designs to produce new designs.
-For example, the `DoE.base` package contain some experimental functions, `cross.design()` and `param.design()`, to combine designs with the former taking a Cartesian product of the input designs while the latter using a Taguchi style [@Taguchi1986-ut] to aggregate the designs with the inner and outer arrays.
-The `lhs` package contain the function that `augmentLHS()` to add additional samples to existing Latin hypercube sample.
+For example, the `r rpkg("DoE.base")` package contain some experimental functions, `cross.design()` and `param.design()`, to combine designs with the former taking a Cartesian product of the input designs while the latter using a Taguchi style [@Taguchi1986-ut] to aggregate the designs with the inner and outer arrays.
+The `r rpkg("lhs")` package contain the function that `augmentLHS()` to add additional samples to existing Latin hypercube sample.
 
-Another class of augmenting design is *sequential design* (also called *adaptive sampling*), which is best represented by `tgp` [@tgp] -- these require prior information, which are used to inform the next experimental design using `tgp.design()` and `dopt.gp()`.
+Another class of augmenting design is *sequential design* (also called *adaptive sampling*), which is best represented by `r rpkg("tgp")` [@tgp] -- these require prior information, which are used to inform the next experimental design using `tgp.design()` and `dopt.gp()`.
 The user is required to supply candidate samples to sub-sample from and a model or a prior experimental design.
-Follow-up experiments, which can also be classified as a sequential design, is implemented by `BsMD` [@BsMD] using a model discriminant approach using `MD()`.
+Follow-up experiments, which can also be classified as a sequential design, is implemented by `r rpkg("BsMD")` [@BsMD] using a model discriminant approach using `MD()`.
 
@@
 A consistent cognitive interface design that leverage existing developments and can be easily extensible for new methods will conceivably be of great help to practitioners.
-Some efforts to this end is seen in `DoE.wrapper` [@DoE.wrapper], which contain wrapper recipe functions to other DoE packages such as `lhs`, `AlgDesign` and `FrF2` (see Figure `r ref("fig:plot-doe-network")`) and is also the subject of the developmental package `edibble` [@edibble].
+Some efforts to this end is seen in `r rpkg("DoE.wrapper")` [@DoE.wrapper], which contain wrapper recipe functions to other DoE packages such as `r rpkg("lhs")`, `r rpkg("AlgDesign")` and `r rpkg("FrF2")` (see Figure `r ref("fig:plot-doe-network")`) and is also the subject of the developmental package `r rpkg("edibble")` [@edibble].
 Undoubtedly, no single developer or package can cater for all experimental design needs so any unifying interface should consider how other developers can contribute or add their methods.
@@
 
-This paper uses `targets` [@targets] and `renv` [@renv] for reproducibility, `knitr` [@knitr] and `rmarkdown` [@rmarkdown] for creating reproducible documents, `ggplot2` [@ggplot2], `ggraph` [@ggraph], `ggwordcloud` [@ggwordcloud] and `colorspace` [@colorspace] for visualisation, `kableExtra` [@kableExtra] for customising the table in the Supplementary Material and `tidyverse` [@Wickham2019-mj], `tidytext` [@tidytext], `pluralize` [@pluralize], and `ineq` [@ineq] for data processing and manipulation, and `cranlogs` [@cranlogs] and `ctv` [@ctv] for extracting data.
+This paper uses `r rpkg("targets")` [@targets] and `r rpkg("renv")` [@renv] for reproducibility, `r rpkg("knitr")` [@knitr] and `r rpkg("rmarkdown")` [@rmarkdown] for creating reproducible documents, `r rpkg("ggplot2")` [@ggplot2], `r rpkg("ggraph")` [@ggraph], `r rpkg("ggwordcloud")` [@ggwordcloud] and `colorspace` [@colorspace] for visualisation, `r rpkg("kableExtra")` [@kableExtra] for customising the table in the Supplementary Material and `r rpkg("tidyverse")` [@Wickham2019-mj], `r rpkg("tidytext")` [@tidytext], `r rpkg("pluralize")` [@pluralize], and `r rpkg("ineq")` [@ineq] for data processing and manipulation, and `r rpkg("cranlogs")` [@cranlogs] and `r rpkg("ctv")` [@ctv] for extracting data.
 All code to reproduce this paper is found at `r url("https://github.com/emitanaka/paper-DoE-review")`.
-
-# References

### 2022-12-13 supp.tex
huge update for index with cleaner pipeline

@@
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={Supplementary material for ``Current status and prospects of R-packages for the design of experiments''},
-  pdfauthor={Emi Tanaka; Dewi Amaliah},
-  pdfkeywords={experimental design, CRAN task view, user interface},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
 \usepackage[margin=1in]{geometry}
@@
 \addbibresource{pkgs.bib}
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={Supplementary material for ``Current status and prospects of R-packages for the design of experiments''},
+  pdfauthor={Emi Tanaka; Dewi Amaliah},
+  pdfkeywords={experimental design, CRAN task view, user interface},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
@@
 \endlastfoot
-\cellcolor{gray!6}{ExperimentalDesign} & \cellcolor{gray!6}{Design of Experiments (DoE) \& Analysis of Experimental Data} & \cellcolor{gray!6}{112} & \cellcolor{gray!6}{211} & \cellcolor{gray!6}{2.27} & \cellcolor{gray!6}{29}\\
-SportsAnalytics & Sports Analytics & 78 & 149 & 2.32 & 17\\
-\cellcolor{gray!6}{MedicalImaging} & \cellcolor{gray!6}{Medical Image Analysis} & \cellcolor{gray!6}{32} & \cellcolor{gray!6}{53} & \cellcolor{gray!6}{2.44} & \cellcolor{gray!6}{44}\\
-MetaAnalysis & Meta-Analysis & 157 & 351 & 2.63 & 43\\
-\cellcolor{gray!6}{ChemPhys} & \cellcolor{gray!6}{Chemometrics and Computational Physics} & \cellcolor{gray!6}{75} & \cellcolor{gray!6}{162} & \cellcolor{gray!6}{2.65} & \cellcolor{gray!6}{28}\\
+\cellcolor{gray!6}{ExperimentalDesign} & \cellcolor{gray!6}{Design of Experiments (DoE) \& Analysis of Experimental Data} & \cellcolor{gray!6}{105} & \cellcolor{gray!6}{192} & \cellcolor{gray!6}{2.23} & \cellcolor{gray!6}{30}\\
+SportsAnalytics & Sports Analytics & 87 & 160 & 2.25 & 16\\
+\cellcolor{gray!6}{MedicalImaging} & \cellcolor{gray!6}{Medical Image Analysis} & \cellcolor{gray!6}{31} & \cellcolor{gray!6}{48} & \cellcolor{gray!6}{2.32} & \cellcolor{gray!6}{45}\\
+ChemPhys & Chemometrics and Computational Physics & 73 & 144 & 2.45 & 25\\
+\cellcolor{gray!6}{ClinicalTrials} & \cellcolor{gray!6}{Clinical Trial Design, Monitoring, and Analysis} & \cellcolor{gray!6}{55} & \cellcolor{gray!6}{121} & \cellcolor{gray!6}{2.58} & \cellcolor{gray!6}{33}\\
 \addlinespace
-ClinicalTrials & Clinical Trial Design, Monitoring, and Analysis & 59 & 141 & 2.76 & 31\\
-\cellcolor{gray!6}{Distributions} & \cellcolor{gray!6}{Probability Distributions} & \cellcolor{gray!6}{257} & \cellcolor{gray!6}{597} & \cellcolor{gray!6}{2.88} & \cellcolor{gray!6}{45}\\
-Survival & Survival Analysis & 239 & 558 & 2.92 & 73\\
-\cellcolor{gray!6}{ExtremeValue} & \cellcolor{gray!6}{Extreme Value Analysis} & \cellcolor{gray!6}{37} & \cellcolor{gray!6}{89} & \cellcolor{gray!6}{2.95} & \cellcolor{gray!6}{41}\\
-Optimization & Optimization and Mathematical Programming & 136 & 328 & 3.04 & 29\\
+MetaAnalysis & Meta-Analysis & 153 & 338 & 2.59 & 44\\
+\cellcolor{gray!6}{Distributions} & \cellcolor{gray!6}{Probability Distributions} & \cellcolor{gray!6}{250} & \cellcolor{gray!6}{580} & \cellcolor{gray!6}{2.88} & \cellcolor{gray!6}{46}\\
+Survival & Survival Analysis & 234 & 556 & 2.95 & 71\\
+\cellcolor{gray!6}{ExtremeValue} & \cellcolor{gray!6}{Extreme Value Analysis} & \cellcolor{gray!6}{37} & \cellcolor{gray!6}{89} & \cellcolor{gray!6}{2.97} & \cellcolor{gray!6}{41}\\
+Hydrology & Hydrological Data and Modeling & 96 & 236 & 3.03 & 20\\
 \addlinespace
-\cellcolor{gray!6}{TimeSeries} & \cellcolor{gray!6}{Time Series Analysis} & \cellcolor{gray!6}{339} & \cellcolor{gray!6}{793} & \cellcolor{gray!6}{3.04} & \cellcolor{gray!6}{58}\\
-OfficialStatistics & Official Statistics \& Survey Statistics & 131 & 325 & 3.06 & 40\\
-\cellcolor{gray!6}{Hydrology} & \cellcolor{gray!6}{Hydrological Data and Modeling} & \cellcolor{gray!6}{100} & \cellcolor{gray!6}{252} & \cellcolor{gray!6}{3.09} & \cellcolor{gray!6}{24}\\
-NumericalMathematics & Numerical Mathematics & 115 & 271 & 3.14 & 63\\
-\cellcolor{gray!6}{Databases} & \cellcolor{gray!6}{Databases with R} & \cellcolor{gray!6}{43} & \cellcolor{gray!6}{95} & \cellcolor{gray!6}{3.23} & \cellcolor{gray!6}{77}\\
+\cellcolor{gray!6}{Optimization} & \cellcolor{gray!6}{Optimization and Mathematical Programming} & \cellcolor{gray!6}{136} & \cellcolor{gray!6}{329} & \cellcolor{gray!6}{3.04} & \cellcolor{gray!6}{29}\\
+Pharmacokinetics & Analysis of Pharmacokinetic Data & 27 & 72 & 3.04 & 19\\
+\cellcolor{gray!6}{TimeSeries} & \cellcolor{gray!6}{Time Series Analysis} & \cellcolor{gray!6}{333} & \cellcolor{gray!6}{791} & \cellcolor{gray!6}{3.08} & \cellcolor{gray!6}{57}\\
+OfficialStatistics & Official Statistics \& Survey Statistics & 129 & 328 & 3.11 & 38\\
+\cellcolor{gray!6}{NumericalMathematics} & \cellcolor{gray!6}{Numerical Mathematics} & \cellcolor{gray!6}{115} & \cellcolor{gray!6}{273} & \cellcolor{gray!6}{3.14} & \cellcolor{gray!6}{63}\\
 \addlinespace
-WebTechnologies & Web Technologies and Services & 201 & 428 & 3.25 & 90\\
-\cellcolor{gray!6}{NaturalLanguageProcessing} & \cellcolor{gray!6}{Natural Language Processing} & \cellcolor{gray!6}{56} & \cellcolor{gray!6}{130} & \cellcolor{gray!6}{3.30} & \cellcolor{gray!6}{62}\\
-Bayesian & Bayesian Inference & 213 & 621 & 3.35 & 49\\
-\cellcolor{gray!6}{FunctionalData} & \cellcolor{gray!6}{Functional Data Analysis} & \cellcolor{gray!6}{40} & \cellcolor{gray!6}{109} & \cellcolor{gray!6}{3.38} & \cellcolor{gray!6}{60}\\
-Robust & Robust Statistical Methods & 59 & 136 & 3.41 & 75\\
+DifferentialEquations & Differential Equations & 26 & 71 & 3.27 & 54\\
+\cellcolor{gray!6}{Bayesian} & \cellcolor{gray!6}{Bayesian Inference} & \cellcolor{gray!6}{207} & \cellcolor{gray!6}{589} & \cellcolor{gray!6}{3.29} & \cellcolor{gray!6}{49}\\
+WebTechnologies & Web Technologies and Services & 197 & 429 & 3.30 & 87\\
+\cellcolor{gray!6}{Tracking} & \cellcolor{gray!6}{Processing and Analysis of Tracking Data} & \cellcolor{gray!6}{47} & \cellcolor{gray!6}{146} & \cellcolor{gray!6}{3.34} & \cellcolor{gray!6}{45}\\
+Databases & Databases with R & 41 & 94 & 3.34 & 78\\
 \addlinespace
-\cellcolor{gray!6}{Psychometrics} & \cellcolor{gray!6}{Psychometric Models and Methods} & \cellcolor{gray!6}{230} & \cellcolor{gray!6}{567} & \cellcolor{gray!6}{3.41} & \cellcolor{gray!6}{69}\\
-Tracking & Processing and Analysis of Tracking Data & 46 & 149 & 3.46 & 48\\
-\cellcolor{gray!6}{Cluster} & \cellcolor{gray!6}{Cluster Analysis \& Finite Mixture Models} & \cellcolor{gray!6}{108} & \cellcolor{gray!6}{305} & \cellcolor{gray!6}{3.47} & \cellcolor{gray!6}{39}\\
-Econometrics & Econometrics & 152 & 363 & 3.50 & 81\\
-\cellcolor{gray!6}{Finance} & \cellcolor{gray!6}{Empirical Finance} & \cellcolor{gray!6}{158} & \cellcolor{gray!6}{426} & \cellcolor{gray!6}{3.61} & \cellcolor{gray!6}{57}\\
+\cellcolor{gray!6}{NaturalLanguageProcessing} & \cellcolor{gray!6}{Natural Language Processing} & \cellcolor{gray!6}{55} & \cellcolor{gray!6}{129} & \cellcolor{gray!6}{3.35} & \cellcolor{gray!6}{62}\\
+CausalInference & Causal Inference & 156 & 439 & 3.35 & 24\\
+\cellcolor{gray!6}{Psychometrics} & \cellcolor{gray!6}{Psychometric Models and Methods} & \cellcolor{gray!6}{225} & \cellcolor{gray!6}{561} & \cellcolor{gray!6}{3.46} & \cellcolor{gray!6}{70}\\
+Robust & Robust Statistical Methods & 56 & 129 & 3.46 & 75\\
+\cellcolor{gray!6}{Cluster} & \cellcolor{gray!6}{Cluster Analysis \& Finite Mixture Models} & \cellcolor{gray!6}{106} & \cellcolor{gray!6}{303} & \cellcolor{gray!6}{3.50} & \cellcolor{gray!6}{39}\\
 \addlinespace
-MissingData & Missing Data & 210 & 740 & 3.99 & 42\\
-\cellcolor{gray!6}{SpatioTemporal} & \cellcolor{gray!6}{Handling and Analyzing Spatio-Temporal Data} & \cellcolor{gray!6}{81} & \cellcolor{gray!6}{269} & \cellcolor{gray!6}{4.04} & \cellcolor{gray!6}{70}\\
-Spatial & Analysis of Spatial Data & 197 & 618 & 4.26 & 83\\
-\cellcolor{gray!6}{GraphicalModels} & \cellcolor{gray!6}{Graphical Models} & \cellcolor{gray!6}{32} & \cellcolor{gray!6}{109} & \cellcolor{gray!6}{4.38} & \cellcolor{gray!6}{78}\\
-Pharmacokinetics & Analysis of Pharmacokinetic Data & 29 & 109 & 4.55 & 21\\
+Econometrics & Econometrics & 150 & 387 & 3.67 & 84\\
+\cellcolor{gray!6}{Finance} & \cellcolor{gray!6}{Empirical Finance} & \cellcolor{gray!6}{154} & \cellcolor{gray!6}{438} & \cellcolor{gray!6}{3.85} & \cellcolor{gray!6}{57}\\
+FunctionalData & Functional Data Analysis & 39 & 130 & 3.97 & 59\\
+\cellcolor{gray!6}{MissingData} & \cellcolor{gray!6}{Missing Data} & \cellcolor{gray!6}{207} & \cellcolor{gray!6}{735} & \cellcolor{gray!6}{4.00} & \cellcolor{gray!6}{42}\\
+Epidemiology & Epidemiology & 76 & 264 & 4.24 & 55\\
 \addlinespace
-\cellcolor{gray!6}{HighPerformanceComputing} & \cellcolor{gray!6}{High-Performance and Parallel Computing with R} & \cellcolor{gray!6}{83} & \cellcolor{gray!6}{315} & \cellcolor{gray!6}{4.75} & \cellcolor{gray!6}{63}\\
-DifferentialEquations & Differential Equations & 27 & 114 & 4.96 & 56\\
-\cellcolor{gray!6}{Environmetrics} & \cellcolor{gray!6}{Analysis of Ecological and Environmental Data} & \cellcolor{gray!6}{93} & \cellcolor{gray!6}{383} & \cellcolor{gray!6}{5.02} & \cellcolor{gray!6}{74}\\
-MachineLearning & Machine Learning \& Statistical Learning & 102 & 488 & 5.63 & 50\\
-\cellcolor{gray!6}{TeachingStatistics} & \cellcolor{gray!6}{Teaching Statistics} & \cellcolor{gray!6}{46} & \cellcolor{gray!6}{236} & \cellcolor{gray!6}{6.35} & \cellcolor{gray!6}{57}\\
+\cellcolor{gray!6}{SpatioTemporal} & \cellcolor{gray!6}{Handling and Analyzing Spatio-Temporal Data} & \cellcolor{gray!6}{74} & \cellcolor{gray!6}{258} & \cellcolor{gray!6}{4.27} & \cellcolor{gray!6}{72}\\
+Spatial & Analysis of Spatial Data & 192 & 608 & 4.32 & 82\\
+\cellcolor{gray!6}{GraphicalModels} & \cellcolor{gray!6}{Graphical Models} & \cellcolor{gray!6}{31} & \cellcolor{gray!6}{108} & \cellcolor{gray!6}{4.48} & \cellcolor{gray!6}{81}\\
+HighPerformanceComputing & High-Performance and Parallel Computing with R & 86 & 315 & 4.64 & 62\\
+\cellcolor{gray!6}{Environmetrics} & \cellcolor{gray!6}{Analysis of Ecological and Environmental Data} & \cellcolor{gray!6}{92} & \cellcolor{gray!6}{386} & \cellcolor{gray!6}{5.11} & \cellcolor{gray!6}{74}\\
 \addlinespace
-ReproducibleResearch & Reproducible Research & 102 & 524 & 6.49 & 76\\
-\cellcolor{gray!6}{ModelDeployment} & \cellcolor{gray!6}{Model Deployment with R} & \cellcolor{gray!6}{31} & \cellcolor{gray!6}{146} & \cellcolor{gray!6}{6.55} & \cellcolor{gray!6}{74}\\*
+MachineLearning & Machine Learning \& Statistical Learning & 101 & 494 & 5.73 & 50\\
+\cellcolor{gray!6}{TeachingStatistics} & \cellcolor{gray!6}{Teaching Statistics} & \cellcolor{gray!6}{45} & \cellcolor{gray!6}{237} & \cellcolor{gray!6}{6.51} & \cellcolor{gray!6}{58}\\
+ModelDeployment & Model Deployment with R & 31 & 146 & 6.58 & 74\\
+\cellcolor{gray!6}{ReproducibleResearch} & \cellcolor{gray!6}{Reproducible Research} & \cellcolor{gray!6}{101} & \cellcolor{gray!6}{540} & \cellcolor{gray!6}{6.72} & \cellcolor{gray!6}{76}\\*
 \end{longtable}
+
+!6\}\{ReproducibleResearch\} \& \cellcolor{gray!6}{Reproducible Research} \& \cellcolor{gray!6}{101} \& \cellcolor{gray!6}{540} \& \cellcolor{gray!6}{6.72} \& \cellcolor{gray!6}{76}\textbackslash*
+\textbackslash end\{longtable\}
 

### 2022-12-13 R/functions.R

@@ -1,25 +1,31 @@
 
-
-
-ctv_list <- function(ctv) {
-  ctv:::.get_pkgs_from_ctv_or_repos(ctv, 
-                                    repos = "http://cran.rstudio.com/")[[1]]  
+save_output <- function(obj) {
+  # file name derived from object
+  fn <- rlang::as_string(rlang::enexpr(obj))
+  saveRDS(obj, here::here(glue::glue("output/{fn}.rds")))
 }
 
-lorenz_data <- function(data) {
-  data.frame(p = ineq::Lc(data$total)$p, L = ineq::Lc(data$total)$L) 
+ctv_pkgs <- function(ctv_table, ctv_name) {
+  ctv_table %>% 
+    filter(name == ctv_name) %>% 
+    pull(packages) %>% 
+    .[[1]]
 }
 
-pkg_first_release <- function(data) {
-  data %>% 
-    group_by(package) %>% 
-    summarise(first = min(update)) %>% 
-    ungroup()
+ctv_pkg_db <- function(ctv_table) {
+  unlist(ctv_table$packages) %>% 
+    unique() %>% 
+    cranscrub::pkg_db() %>% 
+    cranscrub::pkg_db_clean_depends() %>% 
+    cranscrub::pkg_db_clean_imports() %>% 
+    cranscrub::pkg_db_clean_suggests() %>% 
+    cranscrub::pkg_db_clean_author()
 }
 
-pkg_update_dates <- function(pkgs, message = FALSE) {
-  res <- lapply(pkgs, function(pkg) {
-    message(paste("Package", pkg))
+
+pkg_update_dates <- function(pkgs, message = TRUE) {
+  res <- map(pkgs, function(pkg) {
+    if(message) message(paste("Package", pkg))
     pkgsearch::cran_package_history(pkg) %>% 
       select(package = Package, update = date) %>% 
       mutate(update = as.Date(update))
@@ -27,80 +33,75 @@ pkg_update_dates <- function(pkgs, message = FALSE) {
   do.call("rbind", res)
 }
 
-
-cran_downloads_duration <- function(pkgs, start_year) {
-  # it can have trouble downloading long periods so break it up by year
-  map_dfr(start_year:2021, ~{
-    start <- as.Date(paste0(.x, "-01-01"))
-    end <- as.Date(paste0(.x, "-12-31"))
-    cran_downloads(pkgs, from = start, to = end)  
+pkg_cran_downloads <- function(pkgs, start, end) {
[…]

### 2022-12-13 R/plot.R

@@ -1,166 +1,0 @@
-
-plot_setup <- function() {
-  theme(panel.background = element_rect(fill = NA),
-        panel.grid = element_line(color = "lightgray"),
-        axis.text = element_text(color = "black"),
-        axis.line = element_line(color = "black", size = 0.7),
-        axis.ticks.length = unit(1.4, "mm"),
-        axis.ticks = element_line(color = "black", size = 0.7),
-        axis.title = element_text(color = "black", face = "bold"),
-        strip.background = element_rect(color = "black",
-                                        fill = "white"),
-        strip.text = element_text(color = "black", 
-                                  margin = margin(5, 5, 5, 5)),
-        plot.title.position = "plot",
-        plot.title = element_text(color = "black", hjust = 0))
-}
-
-plot_stl_model <- function(data, pkg, updates, trend, speriod1, speriod2) {
-  data %>% 
-    filter(package == pkg) %>% 
-    as_tsibble(index = date) %>% 
-    model(STL(count ~ trend(window = !!trend) + season(period = !!speriod1) + season(period = !!speriod2),
-              robust = TRUE)) %>% 
-    components() %>% 
-    autoplot() + 
-    geom_vline(data = filter(updates, package == pkg), 
-               aes(xintercept = update), color = "red")
-  
-}
-
-plot_download_distribution <- function(data, ginidata) {
-  data %>% 
-    mutate(year = as.factor(year)) %>% 
-    ggplot(aes(x = year, y = total)) + 
-    geom_violin() +
-    geom_boxplot(width = 0.1) + 
-    labs(x = "Year", y = "Downloads") + 
-    scale_y_log10(label = comma, expand = c(0.1, 0)) +
-    geom_label(data = mutate(ginidata, year = as.factor(year)), 
-               size = 3,
-               aes(y = 900, label = paste0(scales::percent(gini, 0.1),
-                                           "\nn=", n)))
-}
-
-plot_lorenz_curve <- function(data) {
-  
-  ggplot(data, aes(p * 100, L * 100)) +
-    geom_abline(color = "red") +
-    geom_ribbon(aes(ymin = 0, ymax = L * 100),
-                fill = "#FFC107", alpha = 0.5) +
-    geom_ribbon(aes(ymin = L * 100, ymax = p * 100),
-                 fill = "#D81B60", alpha = 0.5) +
-    geom_line() + 
-    geom_point() +
[…]

### 2022-12-13 _targets.R

@@ -2,117 +2,49 @@ library(targets)
 library(ggplot2)
 library(tarchetypes)
 source("R/functions.R")
-source("R/plot.R")
-tar_option_set(packages = c("tidyverse", "rvest", "lubridate",
-                            "cranlogs", "glue", "scales", "colorspace",
-                            "tidytext", "pluralize", "kableExtra", 
-                            "igraph", "ggraph", "ggrepel", "patchwork",
-                            "ggwordcloud", "fable", "feasts", "lubridate",
-                            "ineq", "cranscrub", "tidygraph", "ggraph", 
-                            "rlang"))
+tar_option_set(packages = c("tidyverse", "lubridate", "cranlogs", "glue", 
+                            "scales", "tidytext", "pluralize", "patchwork",
+                            "ineq", "cranscrub", "rlang", "igraph", "tidygraph"))
 
 
 list(
-  tar_target(doe_pkgs,  cranscrub::ctv_pkgs("ExperimentalDesign")),
-  tar_target(ctv_pkgs_df, cranscrub::ctv_table()),
-  tar_target(npkgs, length(doe_pkgs)),
-  tar_target(doe_db, cranscrub::pkg_db(doe_pkgs)),
-  tar_target(updates_all, pkg_update_dates(doe_pkgs)),
-  tar_target(first_release, pkg_first_release(updates_all)),
+  tar_target(ctv_table, cranscrub::ctv_table()),
+  tar_target(doe_pkgs,  ctv_pkgs(ctv_table, "ExperimentalDesign")),
+  tar_target(doe_npkgs, length(doe_pkgs)),
+  tar_target(pkg_db_clean, ctv_pkg_db(ctv_table)),
+  # this part takes a long time to complete (~6 hours)
+  tar_target(pkg_downloads, pkg_cran_downloads(pkg_db_clean$Package, 
+                                               start = 2013, end = 2021)),
+  tar_target(pkg_updates, pkg_update_dates(pkg_db_clean$Package)),
+  tar_target(ctv_pkg_data, ctv_combine_pkg_data(ctv_table,
+                                                pkg_downloads, 
+                                                pkg_db_clean,
+                                                pkg_updates)),
[…]

### 2022-12-13 paper/arxiv/arxiv.Rmd
update arxiv

@@
   \newcommand{\ctv}[1]{\href{http://CRAN.R-project.org/view=#1}{\emph{#1}}}
-bibliography: ../../paper.bib
+bibliography: paper.bib
 biblio-style: unsrt
@@
 ```{r, include = FALSE}
+file.copy(here::here("paper.bib"), here::here("paper/arxiv/paper.bib"))
+file.copy(here::here("supp.pdf"), here::here("paper/arxiv/anc/supp.pdf"))
+zip("anc", "anc/")
+
 paper <- here::here("paper.Rmd")

### 2022-12-13 paper/arxiv/arxiv.tex
update arxiv

@@
 \href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}}
-task view, there are 114 R packages for experimental design and analysis
+task view, there are 105 R packages for experimental design and analysis
 of data from experiments. The sheer quantity and variation of
@@
 meta-information in the DESCRIPTION file. We use the text data from the
-title and description (as accessed in 2022-05-23) in Section
+title and description (as accessed in 2022-12-12) in Section
 \ref{topics}.
@@
 CRAN task views are volunteer maintained list of R-packages on CRAN
-relevant to the corresponding topic. There are a total 37 CRAN task
+relevant to the corresponding topic. There are a total 39 CRAN task
 views. Table S1 in the Supplementary Materials list the full available
 topics from the \texttt{ctv} package (Zeileis 2005). The list of
-packages in each CRAN task views (as of 2022-05-25) are used to contrast
+packages in each CRAN task views (as of 2022-12-12) are used to contrast
 the characteristics of DoE packages in Sections \ref{popular} and
@@
 \ref{fig:plot-lorenz} shows a Lorenz curve (Lorenz 1905) for the total
-package downloads in 2021 of 113 DoE packages (first released prior to
+package downloads in 2021 of 102 DoE packages (first released prior to
 2021); we can see from Figure \ref{fig:plot-lorenz} that bottom 90\% of
 DoE packages (in terms of total download count in 2021) only share about
-27\% of total downloads across all DoE packages -- in another words,
-73\% of the total downloads are due to 11 packages (10\% of the DoE
+32\% of total downloads across all DoE packages -- in another words,
+68\% of the total downloads are due to 10 packages (10\% of the DoE
 packages).
@@
 we see that the distributions of the package downloads each year have a
-heavy right tail with the Gini index ranging from 42.6\% to 74.1\%
+heavy right tail with the Gini index ranging from 32.7\% to 69.1\%
 across the years 2013 to 2021, indicating that there is a high level of
@@
 \href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}}
-task view (42.6\% to 74.1\%), the inequality is not as severe as other
+task view (32.7\% to 69.1\%), the inequality is not as severe as other
 CRAN task views as shown in Figure \ref{fig:fig-gini-all-ctvs}. We can
@@
 \href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}}
-task view has the lowest average number of contributors among all 37
+task view has the lowest average number of contributors among all 39
 CRAN task views. In addition, we can also see in Figure
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
 \hypertarget{refs}{}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-BHH2}{}}%
-Barrios, Ernesto. 2016. \emph{Bhh2: Useful Functions for Box, Hunter and
+Barrios, Ernesto. 2016. \emph{BHH2: Useful Functions for Box, Hunter and
 Hunter II}. \url{https://CRAN.R-project.org/package=BHH2}.
@@
 \bibliographystyle{unsrt}
-\bibliography{../../paper.bib}
+\bibliography{paper.bib}
 

### 2022-12-13 _targets.R
update index.html

@@ -46,7 +46,7 @@ list(
   tar_target(save_access_date_pkg_downloads, save_output(access_date_pkg_downloads)),
 
   tar_render(report, "paper.Rmd"),
-  tar_render(index, rmarkdown::render("paper.Rmd", output_file = 'index.html')),
+  tar_target(index, rmarkdown::render("paper.Rmd", output_file = 'index.html')),
   tar_render(supp, "supp.Rmd"),
   tar_render(arxiv, "paper/arxiv/arxiv.Rmd"),
   NULL

### 2022-12-14 paper.Rmd
use amercian english and tidy up grammar

@@
 abstract: >
-    Re-running an experiment is generally costly and in some cases impossible due to limited resources, so the design of an experiment plays a critical role in increasing the quality of experimental data.
-In this paper we describe the current state of the R-packages for the design of experiments through an exploratory data analysis of package downloads, package metadata, and the comparison of characteristics with other topics.
-We observe that experimental designs in practice appear to be sufficiently manufactured by a small number of packages and the development of experimental designs often occur in silos.
-We discuss also the interface designs of widely utilised R packages in the field of experimental design and discuss its future prospects to advance the field in practice.
+    Re-running an experiment is generally costly and, in some cases, impossible due to limited resources; therefore, the design of an experiment plays a critical role in increasing the quality of experimental data.
+In this paper, we describe the current state of R-packages for the design of experiments through an exploratory data analysis of package downloads, package metadata, and a comparison of characteristics with other topics.
+We observed that experimental designs in practice appear to be sufficiently manufactured by a small number of packages, and the development of experimental designs often occurs in silos.
+We also discuss the interface designs of widely utilized R packages in the field of experimental design and discuss their future prospects for advancing the field in practice.
 bibliography: paper.bib
@@
 output: 
+  #word_document: default
   pagedown::html_paged:
@@
 
-The critical role of data collection is well captured in the expression "garbage in, garbage out" -- in other words, if the collected data is rubbish then no analysis, however complex it may be, can make something out of it.
-A carefully crafted data collection scheme is therefore critical to optimise the information from data.
-The field of experimental designs is specifically devoted to planning the collection of experimental data, largely based on the founding principles by @Fisher1935-qc or an optimisation framework like those described in @Pukelsheim2006-sv.
-These experimental designs are often constructed with the aid of a statistical software, such as R [@R], Python [@python], SAS [@sas1985sas], and so on, so the usage of experimental design software can inform us about some aspects of experimental designs in practice.
+The critical role of data collection is well captured in the expression “garbage in, garbage out” – in other words, if the collected data are rubbish, then no analysis, however complex it may be, can make something out of it.
+Therefore, a carefully crafted data-collection scheme is critical for optimizing the information from the data.
+The field of experimental design is specifically devoted to planning the collection of experimental data, largely based on the founding principles of @Fisher1935-qc or an optimization framework like those described in  @Pukelsheim2006-sv.
+These experimental designs are often constructed with the aid of statistical software such as R [@R], Python [@python], and SAS [@sas1985sas]; thus the use of experimental design software can inform us about some aspects of experimental designs in practice.
 
-Methods for data collection can be dichotomised by the type of data collected -- namely, experimental or observational -- or alternatively, categorised as experimental design (including quasi-experimental design) or survey design.
-This dichotomisation, to a great extent, is seen in the [Comprehensive R Archive Network (CRAN) task views](https://cran.r-project.org/web/views/) (a volunteer maintained list of R-packages by topic) where R-packages for experimental design are in `r ctv("ExperimentalDesign")` task view and R-packages for survey designs are in `r ctv("OfficialStatistics")` task view.
-The full list of available topics can be see in Table S1 in the Supplementary Materials.
-A subset of experimental designs are segregated into `r ctv("ClinicalTrials")` task view, where the focus is on clinical trials with primary interest in sample size calculations.
-This paper focuses on the packages in `r ctv("ExperimentalDesign")` task view, henceforth referred to as "DoE packages".
+Methods for data collection can be dichotomized by the type of data collected -- namely, experimental or observational -- or alternatively, categorized as experimental design (including quasi-experimental design) or survey design.
+This dichotomization, to a great extent, is seen in the [Comprehensive R Archive Network (CRAN) task views](https://cran.r-project.org/web/views/) (a volunteer maintained list of R-packages by topic) where R-packages for experimental design are in `r ctv("ExperimentalDesign")` task view and R-packages for survey designs are in `r ctv("OfficialStatistics")` task view.
+A full list of available topics is provided in Table S1 in the Supplementary Materials.
+A subset of experimental designs is segregated into the `r ctv("ClinicalTrials")` task view, where the focus is on clinical trials with primary interest in sample size calculations.
+This paper focuses on packages in `r ctv("ExperimentalDesign")` task view, henceforth referred to as "DoE packages".
 
-In `r ctv("ExperimentalDesign")` task view, there are `r doe_npkgs` R packages for experimental design and analysis of data from experiments.
-The sheer quantity and variation of experimental designs in the R-packages are arguably unmatched with any other programming languages, e.g.
-in Python, only a handful of packages that generate design of experiment exist (namely `pyDOE`, `pyDOE2`, `dexpy`, `experimenter` and `GPdoemd`) with limited type of designs.
-Thus, the study of DoE packages, based on quantitative and qualitative data, can give us an objective view to the state of the current experimental designs in practice.
+From the `r ctv("ExperimentalDesign")` task view, there are `r doe_npkgs` R packages for the experimental design and analysis of data from experiments.
+The sheer quantity and variation of experimental designs in the R-packages are arguably unmatched with any other programming languages, for example, in Python, only a handful of packages that generate design of experiment exist (namely `pyDOE`, `pyDOE2`, `dexpy`, `experimenter`, and `GPdoemd`) with a limited type of design.
+Thus, the study of DoE packages, based on quantitative and qualitative data, can provide an objective view of the state of current experimental designs in practice.
 
-A utility of software can also be described by its design to facilitate clear expression and interpretation of the desired experimental design.
-Certain programming language design can hinder or discourage development of a reliable program [@Wasserman1975-xr].
-The immense popularity of `r rpkg("tidyverse")` [a collection of R-packages for various stages of data analysis that place enormous emphasis on the interface design by @Wickham2019-mj] is a testament to the impact an interface design can have in practice.
-The practice of experimental design could be advanced by adopting similar interface design principles across DoE packages.
+The utility of the software can also be described by its design to facilitate the clear expression and interpretation of the desired experimental design.
+Certain programming language designs can hinder or discourage the development of reliable programs [@Wasserman1975-xr].
+The immense popularity of `r rpkg("tidyverse")` [a collection of R-packages for various stages of data analysis that places enormous emphasis on the interface design by @Wickham2019-mj] is a testament to the impact that an interface design can have in practice.
+The practice of experimental design can be advanced by adopting similar interface design principles across the DoE packages.
 
-The paper is organised as follows.
+The remainder of this paper is organized as follows.
 Section `r ref("data")` briefly describes the data source used for the analysis in Section `r ref("eda")`; Section `r ref("eda")` presents some insights into the state of the current DoE packages by the exploratory data analysis of package download data, text descriptions and comparisons with other CRAN task views; Section `r ref("design")` discusses the interface designs of widely used DoE packages, and we conclude with a discussion in Section `r ref("discussion")` of future prospects in the software development of experimental designs.
@@
 
-The Comprehensive R Archive Network (CRAN) is a network of servers located across the world that store mirrored versions of R and R-packages.
+The Comprehensive R Archive Network (CRAN) is a network of servers located across the world that stores mirrored versions of the R and R packages.
 The most popular network is the RStudio mirror (the default server for those that use the RStudio IDE).
-The RStudio mirror is also the only server that provides a comprehensive daily download logs of R and R-packages since October 2012.
-The summary data can be easily accessed with the `r rpkg("cranlogs")` package [@cranlogs].
-This paper uses the data from the beginning of 2013 to end of 2021 (a total of 9 years) for the packages in the CRAN task views.
+The RStudio mirror is also the only server that provides a comprehensive daily download logs of R and R packages since October 2012.
+The summary data can be easily accessed using the `r rpkg("cranlogs")` package [@cranlogs].
+This paper uses the data from the beginning of 2013 to the end of 2021 (a total of nine years) for the packages in the CRAN task views.
 
@@
 
-All CRAN packages have a title, description, package connections (suggests, depends and imports of other packages) and other meta-information in the DESCRIPTION file.
-We use the text data from the title and description (as accessed in `r access_date_pkg_downloads`) in Section `r ref("topics")`.
+All CRAN packages have a title, description, package connections (suggests, depends, and imports of other packages), and other meta-information in the DESCRIPTION file.
+We use text data from the title and description (accessed in `r access_date_pkg_downloads`) in Section `r ref("topics")`.
 
@@
 
-CRAN task views are volunteer maintained list of R-packages on CRAN relevant to the corresponding topic.
-There are a total `r n_distinct(ctv_pkg_data$name)` CRAN task views.
-Table S1 in the Supplementary Materials list the full available topics from the `r rpkg("ctv")` package [@ctv].
-The list of packages in each CRAN task views (as of `r access_date_pkg_downloads`) are used to contrast the characteristics of DoE packages in Sections `r ref("popular")` and `r ref("silo")`.
+CRAN task views are volunteer-maintained lists of R-packages on CRAN relevant to the corresponding topic.
+There were `r n_distinct(ctv_pkg_data$name)` CRAN task views in total.
+Table S1 in the Supplementary Materials list the available topics from the `r rpkg("ctv")` package [@ctv].
+The list of packages in each CRAN task view (as of `r access_date_pkg_downloads`) is used to contrast the characteristics of the DoE packages in Sections `r ref("popular")` and `r ref("silo")`.
 
@@
 
-In this section, we derive some conjecture based on the analysis of the data described in Section `r ref("data")`.
-All results presented are from the exploratory data analysis of observational data, consequently, all interpretations are somewhat speculative and may not be indicative of the true state of the field of experimental design.
-In particular, any analysis over time are confounded by the fact that the nature of users and package management have also changed over the years.
-It should be noted that some DoE packages may have been archived or removed from the task view over the years so any cross-sectional analysis presented may not reflect the set of all DoE packages at that particular time period (although we assume such incidences are low).
+In this section, we derive some conjectures based on an analysis of the data described in Section `r ref("data")`.
+All results presented are from exploratory data analysis of observational data, consequently, all interpretations are somewhat speculative and may not be indicative of the true state of the field of experimental design.
+In particular, any analysis over time is confounded by the fact that the nature of users and package management has changed over the years.
+It should be noted that some DoE packages may have been archived or removed from the task view over the years; therefore, any cross-sectional analysis presented may not reflect the set of all DoE packages at that particular time period (although we assume such incidences are low).
 
-A subset of DoE packages is not primarily about design of experiments but about the analysis of experimental data.
-A complete delineation of these packages is difficult as there is almost always at least one function that can aid decisions or constructions of experimental designs (and any categorisation is prone to our subjective bias) so we opted not to remove any DoE packages in the analysis.
+A subset of DoE packages is not primarily about the design of experiments but about the analysis of experimental data.
+A complete delineation of these packages is difficult, as there is almost always at least one function that can aid decisions or constructions of experimental designs (and any categorization is prone to our subjective bias); therefore we opted not to remove any DoE packages in the analysis.
 
@@
 
-There are at least 50 DoE packages since 2013 but most of the downloads are concentrated in just a handful of packages.
-For example, Figure `r ref("fig:plot-lorenz")` shows a Lorenz curve [@Lorenz1905-tc] for the total package downloads in 2021 of `r doe_npkgs_2021` DoE packages (first released prior to 2021); we can see from Figure `r ref("fig:plot-lorenz")` that bottom 90\% of DoE packages (in terms of total download count in 2021) only share about `r lorenz_value(0.9)`\% of total downloads across all DoE packages -- in another words, `r 100 - lorenz_value(0.9)`\% of the total downloads are due to `r round(doe_npkgs_2021 * 0.1)` packages (10% of the DoE packages).
+There are at least 50 DoE packages since 2013 but most of the downloads are concentrated in only a handful of packages.
+For example, Figure `r ref("fig:plot-lorenz")` shows a Lorenz curve [@Lorenz1905-tc] for the total package downloads in 2021 for `r doe_npkgs_2021` DoE packages (first released prior to 2021).
+We can see from Figure `r ref("fig:plot-lorenz")` that the bottom 90\% of DoE packages (in terms of total download count in 2021) only share approximately `r lorenz_value(0.9)`\% of total downloads across all DoE packages; in other words, `r 100 - lorenz_value(0.9)`\% of the total downloads are due to `r round(doe_npkgs_2021 * 0.1)` packages (10% of the DoE packages).
 
@@
 The red line corresponds to the line of perfect equality.
-The yellow region shows the area under the Lorenz curve and the red region shows the area of gap in equality."}
+The yellow region shows the area under the Lorenz curve and the red region shows the area of the gap in equality."}
 ctv_lorenz_yearly %>% 
@@
  If we consider package downloads as a measure of "wealth", then we can consider using the Gini index [@Gini1921-mf] as a measure of download inequality across packages.
-The ratio of the red region over the total colored regions in Figure `r ref("fig:plot-lorenz")` corresponds to the Gini index for 2021.
-A Gini index of 0\% indicates equality in downloads across packages while a value of 100\% indicates maximal inequality (all downloads are due to one package).
-In Figure `r ref("fig:download-share")`, we see that the distributions of the package downloads each year have a heavy right tail with the Gini index ranging from `r scales::percent(min(doe_ginis), 0.1)` to `r scales::percent(max(doe_ginis), 0.1)` across the years 2013 to 2021, indicating that there is a high level of inequality of package downloads, particularly with more pronounced inequality in the last 6 years.
+The ratio of the red region to the total colored regions in Figure `r ref("fig:plot-lorenz")` corresponds to the Gini index for 2021.
+A Gini index of 0\% indicates equality in downloads across packages whereas a value of 100\% indicates maximal inequality (all downloads are due to one package).
+In Figure `r ref("fig:download-share")`, we see that the distributions of the package downloads each year have a heavy right tail with the Gini index ranging from `r scales::percent(min(doe_ginis), 0.1)` to `r scales::percent(max(doe_ginis), 0.1)` across the years 2013 to 2021, indicating that there is a high level of inequality in  package downloads, particularly with more pronounced inequality in the last six years.
 
-```{r download-share, fig.width = 5.5, fig.height = 4, fig.cap = "The distribution of the number of downloads for DoE packages by year.
-Packages were removed in any year if it was released in that year or later so that each download count is for the full year.
-The label on the bottom of the plot shows the Gini index for downloads and the number of packages with full year of download count in the corresponding year.
-In the last 6 years, the Gini index is consistent above 60\\% indicating that most downloads are due to a relatively small number of packages.
+```{r download-share, fig.width = 5.5, fig.height = 4, fig.cap = "Distribution of number of downloads for DoE packages by year.
+Packages were removed in any year if they were released in that year or later so that each download count was for the full year.
+The label at the bottom of the plot shows the Gini index for downloads and the number of packages with a full download count in the corresponding year.
+In the last six years, the Gini index has consistently exceeded 60\\%, indicating that most downloads are due to a relatively small number of packages.
 "}
@@
 
-The increase in the number of packages that are not highly downloaded may mean that **_there are more packages to construct niche experimental designs_**.
+An increase in the number of packages that are not highly downloaded may mean that **_there are more packages to construct niche experimental designs_**.
 Some examples of these packages include `r rpkg("qtlDesign")`, `r rpkg("PwrGSD")` and `r rpkg("Crossover")` made for QTL experiments, group sequential designs and crossover trials, respectively.
-These packages would naturally have a smaller number of potential users.
-Counterfactual to this, the increase could be due to other external factors, such as an increase in the number of skilled developers (and thus more package contributions), a change in CRAN policy or management to add packages (either to CRAN and/or task view), and/or that new packages are still yet to amass users.
-While there is an argument that low download counts are due to the low utility and/or quality of the packages, packages in CRAN task views are selected by expert maintainers, thus we can reasonably assume that any package listed in CRAN task view are of a decent utility and quality.
+These packages would naturally have fewer potential users.
+Counterfactual to this, the increase could be due to other external factors, such as an increase in the number of skilled developers (and thus more package contributions), a change in CRAN policy or management to add packages (either to CRAN and/or task view), and/or the fact that new packages are still yet to amass users.
+While there is an argument that low download counts are due to the low utility and/or quality of packages, packages in CRAN task views are selected by expert maintainers.
+We can reasonably assume that any package listed in the CRAN task view has an acceptable utility and quality.
 
-If the downloads are reflective of the experimental designs used in practice, **_small set of packages appear to be sufficient for most to construct the full set of designs of experiments they need in practice_**.
-Packages of course evolve and the top downloaded packages have had regular updates that may have broadened its scope from its previous releases.
+If the downloads are reflective of the experimental designs used in practice, **_a small set of packages appears to be sufficient for most users to construct the full set of designs of experiments they need in practice_**.
+Packages of course evolve, and the top downloaded packages have had regular updates that may have broadened their scope from their previous releases.
 
-While in absolute terms the Gini index is high for `r ctv("ExperimentalDesign")` task view (`r scales::percent(min(doe_ginis), 0.1)` to `r scales::percent(max(doe_ginis), 0.1)`), the inequality is not as severe as other CRAN task views as shown in Figure `r ref("fig:fig-gini-all-ctvs")`.
-We can see in Figure `r ref("fig:fig-gini-all-ctvs")` that the Gini index is generally increasing for DoE packages (as is generally the case for other CRAN task views as shown in Figure S1 in the Supplementary Materials) but most other CRAN task views have a Gini index of over 75%.
-This suggests that other CRAN task views may have dominant standards and in comparison to other topics, there are more **_diverse approaches to designing experiments_** and thus no single DoE package is dominant.
-This observation however doesn't take into account other approaches to generate experimental designs, like say the proprietary software, CycDesignN [@cycdesignn], which may be widely used.
+While in absolute terms the Gini index is high for the `r ctv("ExperimentalDesign")` task view (`r scales::percent(min(doe_ginis), 0.1)` to `r scales::percent(max(doe_ginis), 0.1)`), the inequality is not as severe as in other CRAN task views as shown in Figure `r ref("fig:fig-gini-all-ctvs")`.
+We can see in Figure `r ref("fig:fig-gini-all-ctvs")` that the Gini index generally increases over time for DoE packages (as is generally the case for other CRAN task views as shown in Figure S1 in the Supplementary Materials), but most other CRAN task views have a Gini index of over 75%.
+This suggests that other CRAN task views may have dominant standards, and in comparison to other topics, there are more **_diverse approaches to designing experiments_**, and thus, no single DoE package is dominant.
+However, this observation doe not consider other approaches to generate experimental designs, such as the proprietary software, CycDesignN [@cycdesignn], which may be widely used.
 
-```{r fig-gini-all-ctvs, fig.height = 4, fig.width = 4, fig.cap = "The points show the Gini index of the download counts by year for ExperimentalDesign task view with the color showing the number of packages.
-The grey lines show the line plots of the Gini index across years for all other CRAN task views.
-See Figure S1 in the Supplementary Material for the line graph for the Gini index across years by each CRAN task view."}
+```{r fig-gini-all-ctvs, fig.height = 4, fig.width = 4, fig.cap = "The points show the Gini index of the download counts by year for the ExperimentalDesign task view with the color showing the number of packages.
+The gray lines show the line plots of the Gini index across years for all other CRAN task views.
+See Figure S1 in the Supplementary Material for the line graph of the Gini index across years for each CRAN task view."}
 ctv_gini_yearly %>% 
@@
 
-We can see in Figure `r ref("fig:rank-over-time")` that most of the top 10 ranking packages have been in the top 10 for the last 9 years with `r rpkg("lhs")` steadily climbing up the ranks in the last few years.
-It should be noted that the download of one package can prompt a download of another package; the most notable package connection is `r rpkg("AlgDesign")` and `r rpkg("agricolae")`, where the former is an import for the latter.
-The full network of packages connections within the DoE packages is shown later in Figure `r ref("fig:plot-doe-network")`.
+We can see in Figure `r ref("fig:rank-over-time")` that most of the top 10 ranking packages have been in the top 10 for the last nine years with `r rpkg("lhs")` steadily climbing up the ranks in the last few years.
+It should be noted that the download of one package can prompt the download of another package; the most notable package connection is `r rpkg("AlgDesign")` and `r rpkg("agricolae")`, where the former is an import for the latter.
+The full network of package connections within the DoE packages is shown in Figure `r ref("fig:plot-doe-network")`.
 
-```{r rank-over-time, fig.width = 5.5, fig.height = 4.5, fig.cap = "The plot shows the rank of top 10 downloaded packages by year.
-Packages that do not appear in top 10 for at least two periods are omitted from the plot.
+```{r rank-over-time, fig.width = 5.5, fig.height = 4.5, fig.cap = "The plot shows the rank of the top 10 packages downloaded by year.
+Packages that did not appear in the top 10 for at least two periods were omitted from the plot.
 Most packages are consistently in the top 10 for the period shown."}
@@
 
-Figure `r ref("fig:release-date-vs-download")` shows a moderate negative correlation between the first release date and the (log of) total download counts of DoE packages in any given year from 2013 to 2021.
-This suggest that in general, a package released earlier are more likely to be used today (possibly for legacy reasons or the general inertia to adopt new packages).
-We can also see in Figure `r ref("fig:release-date-vs-download")` the most downloaded packages were generally released in 2004 to 2010.
+Figure `r ref("fig:release-date-vs-download")` shows a moderate negative correlation between the first release date and the (log of) total download counts of DoE packages for any given year from 2013 to 2021.
+This suggests that packages released earlier are more likely to be used today (possibly for legacy reasons or the general inertia to adopt new packages).
+We can also see in Figure `r ref("fig:release-date-vs-download")` that most downloaded packages were released from 2004 to 2010.
 
 ```{r release-date-vs-download, fig.height = 5, fig.cap = "The above figure shows the total download (in log scale) of a package in the corresponding year against the first release date of the package.
-The blue line corresponds to the least squares fit of a simple linear regression model.
-The label in the right-hand upper corner shows the sample correlation coefficient between the first release date and the log (with base 10) of the total download count.
-The high-leverage point on the far left belongs to `conf.design`, authored by one of the earlier contributors to R."}
+The blue line corresponds to the least-squares fit of the simple linear regression model.
+The label in the upper right-hand corner shows the sample correlation coefficient between the first release date and log (with base 10) of the total download count.
+The high leverage point on the far left belongs to `conf.design`, authored by one of the earlier contributors to R."}
 ctv_rank_yearly %>% 
@@
 
-The consistency in the top 10 ranking packages (Figure `r ref("fig:rank-over-time")`) and the fact that most downloaded DoE packages are first released more than 10 years ago (Figure `r ref("fig:release-date-vs-download")`) indicates that either the existing packages fulfilled the needs of mass in practice or no new packages were compelling for many to switch their practice.
-We do however also see that the the top downloaded packages generally have more updates (see Figure `r ref("fig:release-date-vs-download")`) so it is possible that the packages have improved or broadened the scope of its usage.
+The consistency in the top 10 ranking packages (Figure `r ref("fig:rank-over-time")`) and the fact that most downloaded DoE packages were first released more than 10 years ago (Figure `r ref("fig:release-date-vs-download")`) indicate that either the existing packages fulfilled the needs of mass in practice or no new packages were compelling for many to switch their practice.
+However, we also see that the top downloaded packages generally have more updates (see Figure `r ref("fig:release-date-vs-download")`); therefore, it is possible that the packages have improved or broadened the scope of their usage.
 
@@
 
-Figure `r ref("fig:wordcloud-over-time")` shows some common purposes of DoE packages, based on bigrams in the package title and description.
-We only show the bigrams as unigrams were not insightful and there were not many trigrams common across packages.
-In counting the bigrams, we processed the text data as follows:
+Figure `r ref("fig:wordcloud-over-time")` shows some of the common purposes of DoE packages, based on bigrams in the package title and description.
+We only show the bigrams as unigrams were not insightful, and there were not many trigrams common across packages.
+To count the bigrams, we processed the text data as follows:
 
-1. We standardised the words to lower case and removed pluralization. 
+1. We standardized the words to the lower case and removed pluralization. 
 2.
-Multiple mentions of the same bigram within a package was counted as one (e.g.
-`r rpkg("AlgDesign")` mentions "experimental design" four times in the title and the description but this is counted as one).
+Multiple mentions of the same bigram within a package were counted as one (for example, `r rpkg("AlgDesign")` mentions "experimental design" four times in the title and the description, but this is counted as one).
 3.
-Bigrams that consist the stop words are removed.
+Bigrams consisting of stop words were removed.
 The stop words are sourced from the lexicons in `tidytext::stop_words` in addition to other words we deemed irrelevant, e.g.
@@
 
-Not surprisingly, the bigram "experimental design" was the most common.
+Unsurprisingly, the bigram "experimental design" was the most common.
 More interestingly, "optimal design" and "sequential design" appeared across different packages (indicated by the size of the word in Figure `r ref("fig:wordcloud-over-time")`), and the bigrams "latin hypercube" and "computer experiment" are used across a few packages that are downloaded frequently (indicated by the color of the word in Figure `r ref("fig:wordcloud-over-time")`).
-Sequential design, Latin hypercube sampling and computer experiments (which generally include space filling designs like Latin hypercube sampling) are designs that generally operate by optimising a user-selected criterion and can be classified as optimal designs.
+Sequential design, Latin hypercube sampling and computer experiments (which generally include space-filling designs such as Latin hypercube sampling) generally operate by optimizing a user-selected criterion and can be classified as optimal designs.
 
 ```{r wordcloud-over-time, fig.width = 5.5, fig.height = 5, fig.cap = "The above figure shows the word cloud of bigrams from the title and descriptions of the DoE packages.
-The size shows how often the bigram appears across the DoE packages and the color are relative to the total download count in 2021 of the packages that contain the bigram."}
+The size shows how often the bigram appears across the DoE packages and the color is relative to the total download count in 2021 for the packages that contain the bigram."}
 doe_bigram_all_with_download_data %>%
@@
 
-Figure `r ref("fig:ctv-summ-plot")` shows that the `r ctv("ExperimentalDesign")` task view has the lowest average number of contributors among all `r n_distinct(ctv_gini_yearly$name)` CRAN task views.
+Figure `r ref("fig:ctv-summ-plot")` shows that the `r ctv("ExperimentalDesign")` task view has the lowest average number of contributors among all the `r n_distinct(ctv_gini_yearly$name)` CRAN task views.
 In addition, we can also see in Figure `r ref("fig:ctv-summ-plot")`  that the `r ctv("ExperimentalDesign")` task view has one of the least intra-connectivity (the percentage of packages that make use of other packages within the same task view).
-The full connection of DoE packages with other DoE packages is shown in Figure `r ref("fig:plot-doe-network")`.
-These observations suggest that the **_field of experimental design is one of the least collaborative field_** and package development generally occur in silos.
+The full connection between DoE packages is shown in Figure `r ref("fig:plot-doe-network")`.
+These observations suggest that **_experimental design is one of the least collaborative fields_** and package development generally occurs in silos.
 
-```{r ctv-summ-plot, fig.width = 5.5, fig.height = 6, fig.cap = "The above figure is a scatterplot of the intra-connectivity (the percentage of packages that depends, suggest or imports at least one other package within the same task view) and the average number of contributors for each CRAN task view.
-A low intra-connectivity suggests that development within the topic mostly occur in silos whilst high  intra-connectivity suggests that there are more interactions within the topic.
-The color shows the number of packages, the size of the point corresponds to the total number of contributors, and the text labels show the CRAN task view name.
-The label of ExperimentalDesign task view is colored red.
+```{r ctv-summ-plot, fig.width = 5.5, fig.height = 6, fig.cap = "The above figure is a scatterplot of intra-connectivity (the percentage of packages that depends, suggests or imports at least one other package within the same task view) and the average number of contributors for each CRAN task view.
+Low intra-connectivity suggests that development within the topic mostly occurs in silos, while high  intra-connectivity suggests that there are more interactions within the topic.
+The color shows the number of packages, the size of the point corresponds to the total number of contributors, and the text labels show the CRAN task view names.
+The label of ExperimentalDesign task view is colored in red.
 The task views in the bottom-left corner are topics that are more indicative of contributors working in silos.
-The actual numerical values are show in Table S1 in the Supplementary Material."}
+The actual numerical values are listed in Table S1 in the Supplementary Material."}
 ctv_summary_table %>% 
@@
 
-```{r plot-doe-network, fig.width = 6, fig.height = 6, fig.cap  = "Package connections (depends, suggests and imports) within the DoE packages.
-The direction of the arrow shows the connection of packags where the package on the tail of the arrow is a depedency, suggestion or import for the package on the head of the arrow.
-DoE packages that do not depend, suggest or import another DoE package is not shown."}
+```{r plot-doe-network, fig.width = 6, fig.height = 6, fig.cap  = "Package connections (depends, suggests and imports) within DoE packages.
+The direction of the arrow shows the connection of packags where the package on the tail of the arrow is a dependency, suggestion or import for the package on the head of the arrow.
+DoE packages that do not depend, suggest or import another DoE package are not shown."}
 doe_pkg_network %>%
@@
 In software design, there are two interface designs to consider: user interface (UI) and application programming interface (API).
-UI is concerned with the interaction of the software by the user, while the API is concerned with how different programs interact and is predominately of the interest to the developer.
-The UI design is an abstraction to specifying the desired experimental design and its choices enable how a user expresses the specification of the experimental design.
-The API design aids in other developers to leverage existing systems.
+The UI is concerned with the interaction of the software by the user, while the API is concerned with how different programs interact and is predominately of interest to the developer.
+The UI design is an abstraction that specifies the desired experimental design, and its choices enable how a user expresses the specification of an experimental design.
+The API design aids other developers in leveraging existing systems.
 
-In this section, we discuss the interface designs of functions that output an experimental design based on three broad areas: factorial designs, recipe designs and augmenting designs.
-The discussion is exclusive to the top downloaded packages (shown in Figure `r ref("fig:rank-over-time")`), with the exception of `r rpkg("ez")` and `r rpkg("DiceKriging")`, as the former is predominately visualisation of experimental data and the latter is about the analysis of computer experiments in addition to belonging to the same suite of packages as `r rpkg("DiceDesign")` [@DiceDesign].
+In this section, we discuss the interface designs of functions that output an experimental design based on three broad areas: factorial, recipe and augmenting designs.
+The discussion is exclusive to the top downloaded packages (shown in Figure `r ref("fig:rank-over-time")`), with the exception of `r rpkg("ez")` and `r rpkg("DiceKriging")`, as the former is predominately visualization of experimental data and the latter is about the analysis of computer experiments in addition to belonging to the same suite of packages as `r rpkg("DiceDesign")` [@DiceDesign].
 
@@
 
-Factorial experiments offers a challenge in allocating the treatment factors to experimental units where the full set of factorial treatments cannot be administered (and replicated) and/or the experimental units have some grouping structure.
-The effort to address this challenge is reflected by the number of packages that focus on the construction of factorial designs as described next.
+Factorial experiments offer a challenge in allocating the treatment factors to experimental units where the full set of factorial treatments cannot be administered (and replicated) and/or the experimental units have a grouping structure.
+The effort to address this challenge is reflected in the number of packages that focus on the construction of factorial designs as described next.
 
 The `r rpkg("DoE.base")` package [@DoE.base] can construct full factorial and (regular and irregular) orthogonal array designs via `fac.design()` and `oa.design()`, respectively.
-The `r rpkg("FrF2")` package [@FrF2] constructs regular fractional 2-level factorial designs using `FrF2()` and `FrF2Large()` with the latter for large designs.
-The `r rpkg("conf.design")` package  [@conf.design] constructs symmetric confounded factorial designs via `conf.design()`, and `r rpkg("BHH2")` package [@BHH2] generates full or fractional 2-level factorial design matrix via `ffDesMatrix()`.
+The `r rpkg("FrF2")` package [@FrF2] constructs regular fractional 2-level factorial designs using `FrF2()` and `FrF2Large()`, with the latter for large designs.
+The `r rpkg("conf.design")` package  [@conf.design] constructs symmetric confounded factorial designs via `conf.design()` and the `r rpkg("BHH2")` package [@BHH2] generates a full or fractional 2-level factorial design matrix via `ffDesMatrix()`.
 
@@
 * experimental runs (or replications), and 
-* levels for each factor if the design is allowed to vary in the number of levels.
+* levels for each factor if the design was allowed to vary in the number of levels.
 
@@
 the `design` class for `r rpkg("DoE.base")` and `r rpkg("FrF2")` or a `data.frame` for `r rpkg("conf.design")`) or a matrix for `BHH2` such that an element or column corresponds to a treatment factor with each value corresponding to one experimental run.
-The treatment factor are generally assigned pseudo names (e.g.
+The treatment factors are generally assigned pseudo names (e.g.
 letters of the alphabet) or an argument exists for users to input treatment names as a character vector.
 
-Some forms of factorial designs are known by other names, e.g.
-*response surface designs* are factorial designs where the treatment factors are discrete levels of continuous variables.
+Some forms of factorial design are known by other names, for example, *response surface designs* are factorial designs where the treatment factors are discrete levels of continuous variables.
 Two types of response surface designs can be constructed using the `r rpkg("rsm")` package [@rsm]: Box-Behnken design [@Box1960-wr] and central-composite designs [@Box1951-ji] via functions `bbd()` and `ccd()`, respectively, where the minimum required input is the number of factors.
-Another form of factorial designs is *saturated designs* where typically higher-order interaction effects of treatment factors are confounded with the main effects.
-Plankett-Burnman designs [@Plackett1946-ly] is one type of saturated design which can be generated by the function `pb()` in the `r rpkg("FrF2")` package, where the user provides the number of experimental runs and the number of treatment factors.
+Another form of factorial design is the *saturated designs* where higher-order interaction effects of treatment factors are typically confounded with the main effects.
+Plankett-Burnman designs [@Plackett1946-ly] are a type of saturated design that can be generated by the function `pb()` in the `r rpkg("FrF2")` package, where the user provides the number of experimental runs and the number of treatment factors.
 
@@
 The `r rpkg("agricolae")` package [@agricolae] is the prime example of constructing designs based on a set of so-called "recipe functions", where each function corresponds to a single class of experimental design.
-E.g.
-`design.crd()`, `design.rcbd()` and `design.split()` construct a completely randomised design, randomised complete block design and split-plot design, respectively.
-Users typically supply the treatment labels (or number of treatments in the case of `design.split()`) and the number of replications as argument to these functions and the output is a list with one element corresponding to a `data.frame` that contains the design in a table such that the row corresponds to the experimental run and the columns corresponds to the experimental variables (we refer to this format simply as "table format" henceforth).
+For example, `design.crd()`, `design.rcbd()` and `design.split()` construct completely randomized, randomized complete block and split-plot designs, respectively.
+Users typically supply treatment labels (or the number of treatments in the case of `design.split()`) and the number of replications as arguments for these functions.
+The output is a list with one element corresponding to a `data.frame` that contains the design in a table such that the row corresponds to the experimental run and the columns correspond to the experimental variables (we refer to this format simply as "table format" henceforth).
 
 The use of recipe functions is not limited to classical experimental designs; the `r rpkg("AlgDesign")` [@AlgDesign] package offers three primary functions for generating optimal designs: `optBlock()`, `optFederov()` and `optMonteCarlo()`.
-These functions in general require a data and formula in terms of the supplied data with the choice of criterion (e.g.
-D-criterion) with the output as a list with one element corresponding to the design in a table format.
-The difference between these functions lies in the underlying search strategy for the optimal design and the name of the function is a surrogate for the search algorithm.
+In general, these functions require data and formulas in terms of the supplied data variables with the choice of the criterion (e.g.
+the D-criterion), with the output as a list with one element corresponding to the design in a table format.
+The difference between these functions lies in the underlying search strategy for optimal designs, and the name of the function is a surrogate for the search algorithm.
 
-Computer experiments, which generally involve space filling designs, are implemented in packages like `r rpkg("lhs")` [@lhs] and `r rpkg("DiceDesign")` [@DiceDesign].
-For the `r rpkg("lhs")` package, functions like `randomLHS()`, `optimumLHS()`, and `maximinLHS()`, require users to specify the sample size ($n$) and the number of variables ($p$) then it generates a Latin hypercube sample [@McKay1979-aw], based on different optimisation schemes (in this case, random, S optimal and maxmin criteria, respectively; see package documentation for more details).
-Similarly for the `r rpkg("DiceDesign")`, there is a comprehensive list of space-filling designs such as `dmaxDesign()`, `lhsDesign()`, and `wspDesign()` with input of $n$ and $p$ as before (among extra parameters for some) that implement algorithms that either maximise the entropy [@Shewry1987-jn], produce a random Latin hypercube sample, and use the WSP algorithm [@Santiago2012-ui], respectively.
-These designs all output a $n \times p$ matrix with values between 0 and 1.
-Again, these functions employ a recipe style where each functions have a name that correspond to a certain search strategy in generating the experimental design.
+Computer experiments, which generally involve space-filling designs, are implemented in packages such as `r rpkg("lhs")` [@lhs] and `r rpkg("DiceDesign")` [@DiceDesign].
+For the `r rpkg("lhs")` package, functions such as `randomLHS()`, `optimumLHS()`, and `maximinLHS()`, require users to specify the sample size ($n$) and the number of variables ($p$), then it generates a Latin hypercube sample [@McKay1979-aw], based on different optimization schemes (in this case, random, S optimal and maxmin criteria, respectively; see package documentation for more details).
+Similarly for `r rpkg("DiceDesign")`, there is a comprehensive list of space-filling designs such as `dmaxDesign()`, `lhsDesign()`, and `wspDesign()` with input of $n$ and $p$ as before (among additional parameters for some) that implement algorithms that either maximize the entropy [@Shewry1987-jn], produce a random Latin hypercube sample, or use the WSP algorithm [@Santiago2012-ui], respectively.
+These designs output an $n \times p$ matrix with values between 0 and 1.
+Again, these functions employ a recipe style, in which each function has a name that corresponds to a certain search strategy to generate the experimental design.
 
@@
 
-Some functions in DoE packages require input of existing experimental designs to produce new designs.
-For example, the `r rpkg("DoE.base")` package contain some experimental functions, `cross.design()` and `param.design()`, to combine designs with the former taking a Cartesian product of the input designs while the latter using a Taguchi style [@Taguchi1986-ut] to aggregate the designs with the inner and outer arrays.
-The `r rpkg("lhs")` package contain the function that `augmentLHS()` to add additional samples to existing Latin hypercube sample.
+Some functions in the DoE packages require the input of existing experimental designs to produce new designs.
+For example, the `r rpkg("DoE.base")` package contains some experimental functions, `cross.design()` and `param.design()`, to combine designs with the former taking a Cartesian product of the input designs while the latter uses a Taguchi style [@Taguchi1986-ut] to aggregate the designs with the inner and outer arrays.
+The `r rpkg("lhs")` package contains a function that `augmentLHS()` to add additional samples to the existing Latin hypercube sample.
 
-Another class of augmenting design is *sequential design* (also called *adaptive sampling*), which is best represented by `r rpkg("tgp")` [@tgp] -- these require prior information, which are used to inform the next experimental design using `tgp.design()` and `dopt.gp()`.
-The user is required to supply candidate samples to sub-sample from and a model or a prior experimental design.
-Follow-up experiments, which can also be classified as a sequential design, is implemented by `r rpkg("BsMD")` [@BsMD] using a model discriminant approach using `MD()`.
+Another class of augmenting design is *sequential design* (also called *adaptive sampling*), which is best represented by `r rpkg("tgp")` [@tgp].
+This requires prior information that is used to inform the next experimental design using `tgp.design()` and `dopt.gp()`.
+The user is required to supply candidate samples to subsample from and a model or a prior experimental design.
+Follow-up experiments, which can also be classified as sequential designs, is implemented by `r rpkg("BsMD")` [@BsMD] using a model-discriminant approach using `MD()`.
 
@@
 
-Through the exploratory data analysis of three data sources (package download logs, package metadata and CRAN task views) outlined in Section `r ref("data")`, we have observed in Section `r ref("eda")` that the the total download of DoE packages are concentrated only on a handful of R-packages although these represent a diverse set in comparison with other CRAN task views.
-Furthermore, the data suggests that the field of experimental design is the least collaborative field.
+Through the exploratory data analysis of the three data sources (package download logs, package metadata and CRAN task views) outlined in Section `r ref("data")`, we observed in Section `r ref("eda")` that the the total download of DoE packages is concentrated only on a handful of R-packages although these represent a diverse set in comparison with other CRAN task views.
+Furthermore, the data suggest that experimental design is the least collaborative field.
 
-There are a number of limitations and short coming in our exploratory data analysis.
+There are a number of limitations and shortcomings to our exploratory data analysis.
 First, CRAN task views are volunteer maintained so some experimental design packages may not be included in the DoE packages.
-Second, we only use the RStudio CRAN mirror download, which may bias our observations.
-Third, our analysis are limited to R-packages alone and many practioners may indeed be using other methods to construct experimental designs.
-Finally, all our statements should be treated as speculative rather than conclusive; the data are all observational so no conclusive, generalisable statement are possible.
-Regardless, the data driven nature of our analysis give an objective insight into the field of experimental design.
+Second, we used only the RStudio CRAN mirror download, which may have biased our observations.
+Third, our analysis was limited to R-packages alone and many practitioners may use other methods to construct experimental designs.
+Finally, all our statements should be treated as speculative rather than conclusive; the data are all observational so no conclusive, generalizable statement is possible.
+Regardless, the data-driven nature of our analysis provides objective insight into the field of experimental design.
 
 The interface design (discussed in Section `r ref("design")`) reveals that the most widely used DoE packages generally have functions that 1) focus on certain aspects of experimental design (e.g.
-factorial structure or augmenting design), 2) is a recipe format (i.e.
-the name of the function is a surrogate to a single class of the design or optimal search algorithm) and 3) context is often a second-thought -- many inputs a single integer corresponding to the number of factors, levels, or experimental runs (or sample size).
-The function will often assign pseudo factor names or there is an optional argument to input a character vector that corresponds to the factor names.
-These interface designs require users to have processed the experiment in statistical terms (often stripping the experimental context away) and simultaneously, users must make a choice of the generation mechanism by selecting an appropriate function.
-Arguably, the current dominant interface designs are not aligned with the way practitioners cognitively design experiments.
-Often the critical part of designing an experiment is understanding the experimental structure and the experimental context can govern or guide the choice of the algorithm to allocate treatments.
-In addition, the nature of recipe functions can obscure the understanding and relation of the designs (e.g.
-how do you go from an unstructured factorial design to split plot design?).
-Each new method to generating an experimental design appear to correspond to a totally new function and any intermediate results are often not easily accessible.
-These may be contributing factors to why developers often work in silos in the field of experimental design.
+factorial structure or augmenting design), 2) are a recipe format (i.e.
+the name of the function is a surrogate to a single class of the design or optimal search algorithm) and 3) context is often a second thought -- many inputs a single integer corresponding to the number of factors, levels, or experimental runs (or sample size).
+The function will often assign pseudo-factor names or there is an optional argument to input a character vector that corresponds to the factor names.
+These interface designs require users to have processed the experiment in statistical terms (often stripping the experimental context away) and simultaneously, users must choose the generation mechanism by selecting an appropriate function.
+Arguably, the current dominant interface designs are not aligned with the way practitioners cognitively design their experiments.
+Often, the critical part of designing an experiment is to understand the experimental structure, and the experimental context can govern or guide the choice of algorithm to allocate treatments.
+In addition, the nature of recipe functions can obscure the understanding and relation of designs (e.g., how do you go from an unstructured factorial design to split plot design?).
+Each new method for generating an experimental design appears to correspond to a completely new function, and intermediate results are often not easily accessible.
+These factors may contribute to why developers often work in silos in the field of experimental design.
 
-A consistent cognitive interface design that leverage existing developments and can be easily extensible for new methods will conceivably be of great help to practitioners.
-Some efforts to this end is seen in `r rpkg("DoE.wrapper")` [@DoE.wrapper], which contain wrapper recipe functions to other DoE packages such as `r rpkg("lhs")`, `r rpkg("AlgDesign")` and `r rpkg("FrF2")` (see Figure `r ref("fig:plot-doe-network")`) and is also the subject of the developmental package `r rpkg("edibble")` [@edibble].
-Undoubtedly, no single developer or package can cater for all experimental design needs so any unifying interface should consider how other developers can contribute or add their methods.
+A consistent cognitive interface design that leverages existing developments and can be easily extended to new methods will conceivably be of great help to practitioners.
+Some efforts to this end are seen in `r rpkg("DoE.wrapper")` [@DoE.wrapper], which contains wrapper recipe functions for other DoE packages such as `r rpkg("lhs")`, `r rpkg("AlgDesign")` and `r rpkg("FrF2")` (see Figure `r ref("fig:plot-doe-network")`), and is also the subject of the developmental package `r rpkg("edibble")` [@edibble].
+Undoubtedly, no single developer or package can cater to all experimental designs; therefore, any unifying interface should consider how other developers can contribute or add their methods.
 Future research could benefit from further exploratory data analysis, expanding the study to beyond R-packages, and discussing other aspects of interface designs.
@@
 
-This paper uses `r rpkg("targets")` [@targets] and `r rpkg("renv")` [@renv] for reproducibility, `r rpkg("knitr")` [@knitr] and `r rpkg("rmarkdown")` [@rmarkdown] for creating reproducible documents, `r rpkg("ggplot2")` [@ggplot2], `r rpkg("ggraph")` [@ggraph], `r rpkg("ggwordcloud")` [@ggwordcloud] and `colorspace` [@colorspace] for visualisation, `r rpkg("kableExtra")` [@kableExtra] for customising the table in the Supplementary Material and `r rpkg("tidyverse")` [@Wickham2019-mj], `r rpkg("tidytext")` [@tidytext], `r rpkg("pluralize")` [@pluralize], and `r rpkg("ineq")` [@ineq] for data processing and manipulation, and `r rpkg("cranlogs")` [@cranlogs] and `r rpkg("ctv")` [@ctv] for extracting data.
-All code to reproduce this paper is found at `r url("https://github.com/emitanaka/paper-DoE-review")`.
+This paper uses `r rpkg("targets")` [@targets] and `r rpkg("renv")` [@renv] for reproducibility, `r rpkg("knitr")` [@knitr] and `r rpkg("rmarkdown")` [@rmarkdown] for creating reproducible documents, `r rpkg("ggplot2")` [@ggplot2], `r rpkg("ggraph")` [@ggraph], `r rpkg("ggwordcloud")` [@ggwordcloud] and `colorspace` [@colorspace] for visualization, `r rpkg("kableExtra")` [@kableExtra] for customizing the table in the Supplementary Material and `r rpkg("tidyverse")` [@Wickham2019-mj], `r rpkg("tidytext")` [@tidytext], `r rpkg("pluralize")` [@pluralize], and `r rpkg("ineq")` [@ineq] for data processing and manipulation, and `r rpkg("cranlogs")` [@cranlogs] and `r rpkg("ctv")` [@ctv] for extracting data.
+All codes used to reproduce this paper are found in `r url("https://github.com/emitanaka/paper-DoE-review")`.

### 2022-12-14 paper.Rmd
reference  setups for specific outlets to accommodate differences in non-numbered/numbered sections and output types (PDF/HTML)

@@
     css: ["assets/html-template.css"]
+params:
+  ref_setup_fn: "ref-setup.R"
 ---
@@
                       warning = FALSE)
+source(params$ref_setup_fn)
+```
 
-ctv <- function(view) {
-  ifelse(knitr::is_html_output(), glue("[{view}](https://cran.r-project.org/web/views/{view}.html)"), glue("\\ctv{[view]}", .open = "[", .close = "]"))
-}
-CRANpkg <- function(pkg) {
-  ifelse(knitr::is_html_output(), glue("[{pkg}](https://cran.r-project.org/web/packages/{pkg}/index.html)"), glue("\\CRANpkg{[pkg]}", .open = "[", .close = "]"))
-}
-ref <- function(x) {
-  case_when(knitr::is_html_output() ~ glue("\\@ref({x})"),
-            TRUE ~ glue("\\ref{[x]}", .open = "[", .close = "]"))
-}
-url <- function(x) {
-  ifelse(knitr::is_html_output(), x, glue("\\url{[x]}", .open = "[", .close = "]"))
-}
-rpkg <- function(x) {
-  glue("`{x}`")
-}
-
+```{r data, include = FALSE}
 ctv_pkg_data <- readRDS("output/ctv_pkg_data.rds")
@@
 Certain programming language designs can hinder or discourage the development of reliable programs [@Wasserman1975-xr].
-The immense popularity of `r rpkg("tidyverse")` [a collection of R-packages for various stages of data analysis that places enormous emphasis on the interface design by @Wickham2019-mj] is a testament to the impact that an interface design can have in practice.
+The immense popularity of `r CRANpkg("tidyverse")` [a collection of R-packages for various stages of data analysis that places enormous emphasis on the interface design by @Wickham2019-mj] is a testament to the impact that an interface design can have in practice.
 The practice of experimental design can be advanced by adopting similar interface design principles across the DoE packages.
@@
 The remainder of this paper is organized as follows.
-Section `r ref("data")` briefly describes the data source used for the analysis in Section `r ref("eda")`; Section `r ref("eda")` presents some insights into the state of the current DoE packages by the exploratory data analysis of package download data, text descriptions and comparisons with other CRAN task views; Section `r ref("design")` discusses the interface designs of widely used DoE packages, and we conclude with a discussion in Section `r ref("discussion")` of future prospects in the software development of experimental designs.
+`r secref("data", "data")` briefly describes the data source used for the analysis; `r secref("eda", "Exploratory data analysis" )` presents some insights into the state of the current DoE packages by the exploratory data analysis of package download data, text descriptions and comparisons with other CRAN task views; `r secref("design", "Interface design")` discusses the interface designs of widely used DoE packages, and we conclude with a discussion in `r secref("discussion", "Discussion")` of future prospects in the software development of experimental designs.
 
@@
 The RStudio mirror is also the only server that provides a comprehensive daily download logs of R and R packages since October 2012.
-The summary data can be easily accessed using the `r rpkg("cranlogs")` package [@cranlogs].
+The summary data can be easily accessed using the `r CRANpkg("cranlogs")` package [@cranlogs].
 This paper uses the data from the beginning of 2013 to the end of 2021 (a total of nine years) for the packages in the CRAN task views.
@@
 All CRAN packages have a title, description, package connections (suggests, depends, and imports of other packages), and other meta-information in the DESCRIPTION file.
-We use text data from the title and description (accessed in `r access_date_pkg_downloads`) in Section `r ref("topics")`.
+We use text data from the title and description (accessed in `r access_date_pkg_downloads`).
 
@@
 There were `r n_distinct(ctv_pkg_data$name)` CRAN task views in total.
-Table S1 in the Supplementary Materials list the available topics from the `r rpkg("ctv")` package [@ctv].
-The list of packages in each CRAN task view (as of `r access_date_pkg_downloads`) is used to contrast the characteristics of the DoE packages in Sections `r ref("popular")` and `r ref("silo")`.
+Table S1 in the Supplementary Materials list the available topics from the `r CRANpkg("ctv")` package [@ctv].
+The list of packages in each CRAN task view (as of `r access_date_pkg_downloads`) is used to contrast the characteristics of the DoE packages.
 
-# Explorative data analysis {#eda}
+# Exploratory data analysis {#eda}
 
-In this section, we derive some conjectures based on an analysis of the data described in Section `r ref("data")`.
+In this section, we derive some conjectures based on an analysis of the data described in `r secref("data", "Data")`.
 All results presented are from exploratory data analysis of observational data, consequently, all interpretations are somewhat speculative and may not be indicative of the true state of the field of experimental design.
@@
 
-## Small, but diverse, set of packages are sufficient for most experimental designs in practice {#popular}
+## Small, but diverse, set of packages are sufficient for most experimental designs in practice
 
@@
 An increase in the number of packages that are not highly downloaded may mean that **_there are more packages to construct niche experimental designs_**.
-Some examples of these packages include `r rpkg("qtlDesign")`, `r rpkg("PwrGSD")` and `r rpkg("Crossover")` made for QTL experiments, group sequential designs and crossover trials, respectively.
+Some examples of these packages include `r CRANpkg("qtlDesign")`, `r CRANpkg("PwrGSD")` and `r CRANpkg("Crossover")` made for QTL experiments, group sequential designs and crossover trials, respectively.
 These packages would naturally have fewer potential users.
@@
 
-## The field of experimental design is slow-changing  {#ranking}
+## The field of experimental design is slow-changing
 
-We can see in Figure `r ref("fig:rank-over-time")` that most of the top 10 ranking packages have been in the top 10 for the last nine years with `r rpkg("lhs")` steadily climbing up the ranks in the last few years.
-It should be noted that the download of one package can prompt the download of another package; the most notable package connection is `r rpkg("AlgDesign")` and `r rpkg("agricolae")`, where the former is an import for the latter.
+We can see in Figure `r ref("fig:rank-over-time")` that most of the top 10 ranking packages have been in the top 10 for the last nine years with `r CRANpkg("lhs")` steadily climbing up the ranks in the last few years.
+It should be noted that the download of one package can prompt the download of another package; the most notable package connection is `r CRANpkg("AlgDesign")` and `r CRANpkg("agricolae")`, where the former is an import for the latter.
 The full network of package connections within the DoE packages is shown in Figure `r ref("fig:plot-doe-network")`.
@@
 
-## Optimal designs are of interest {#topics}
+## Optimal designs are of interest
 
@@
 2.
-Multiple mentions of the same bigram within a package were counted as one (for example, `r rpkg("AlgDesign")` mentions "experimental design" four times in the title and the description, but this is counted as one).
+Multiple mentions of the same bigram within a package were counted as one (for example, `r CRANpkg("AlgDesign")` mentions "experimental design" four times in the title and the description, but this is counted as one).
 3.
@@
 The stop words are sourced from the lexicons in `tidytext::stop_words` in addition to other words we deemed irrelevant, e.g.
-"provide", "e.g.", "calculate" and so on -- full list is shown in the code provided in the link under Section `r ref("pkgs")`.
+"provide", "e.g.", "calculate" and so on -- full list is shown in the code provided in the link under `r secref("pkgs", "Acknowledgement")`.
 
@@
 
-## Development of experimental designs occur in silos {#silo}
+## Development of experimental designs occur in silos
 
@@
 In this section, we discuss the interface designs of functions that output an experimental design based on three broad areas: factorial, recipe and augmenting designs.
-The discussion is exclusive to the top downloaded packages (shown in Figure `r ref("fig:rank-over-time")`), with the exception of `r rpkg("ez")` and `r rpkg("DiceKriging")`, as the former is predominately visualization of experimental data and the latter is about the analysis of computer experiments in addition to belonging to the same suite of packages as `r rpkg("DiceDesign")` [@DiceDesign].
+The discussion is exclusive to the top downloaded packages (shown in Figure `r ref("fig:rank-over-time")`), with the exception of `r CRANpkg("ez")` and `r CRANpkg("DiceKriging")`, as the former is predominately visualization of experimental data and the latter is about the analysis of computer experiments in addition to belonging to the same suite of packages as `r CRANpkg("DiceDesign")` [@DiceDesign].
 
@@
 
-The `r rpkg("DoE.base")` package [@DoE.base] can construct full factorial and (regular and irregular) orthogonal array designs via `fac.design()` and `oa.design()`, respectively.
-The `r rpkg("FrF2")` package [@FrF2] constructs regular fractional 2-level factorial designs using `FrF2()` and `FrF2Large()`, with the latter for large designs.
-The `r rpkg("conf.design")` package  [@conf.design] constructs symmetric confounded factorial designs via `conf.design()` and the `r rpkg("BHH2")` package [@BHH2] generates a full or fractional 2-level factorial design matrix via `ffDesMatrix()`.
+The `r CRANpkg("DoE.base")` package [@DoE.base] can construct full factorial and (regular and irregular) orthogonal array designs via `fac.design()` and `oa.design()`, respectively.
+The `r CRANpkg("FrF2")` package [@FrF2] constructs regular fractional 2-level factorial designs using `FrF2()` and `FrF2Large()`, with the latter for large designs.
+The `r CRANpkg("conf.design")` package  [@conf.design] constructs symmetric confounded factorial designs via `conf.design()` and the `r CRANpkg("BHH2")` package [@BHH2] generates a full or fractional 2-level factorial design matrix via `ffDesMatrix()`.
 
@@
 The output of the design is either a special class of list (e.g.
-the `design` class for `r rpkg("DoE.base")` and `r rpkg("FrF2")` or a `data.frame` for `r rpkg("conf.design")`) or a matrix for `BHH2` such that an element or column corresponds to a treatment factor with each value corresponding to one experimental run.
+the `design` class for `r CRANpkg("DoE.base")` and `r CRANpkg("FrF2")` or a `data.frame` for `r CRANpkg("conf.design")`) or a matrix for `BHH2` such that an element or column corresponds to a treatment factor with each value corresponding to one experimental run.
 The treatment factors are generally assigned pseudo names (e.g.
@@
 Some forms of factorial design are known by other names, for example, *response surface designs* are factorial designs where the treatment factors are discrete levels of continuous variables.
-Two types of response surface designs can be constructed using the `r rpkg("rsm")` package [@rsm]: Box-Behnken design [@Box1960-wr] and central-composite designs [@Box1951-ji] via functions `bbd()` and `ccd()`, respectively, where the minimum required input is the number of factors.
+Two types of response surface designs can be constructed using the `r CRANpkg("rsm")` package [@rsm]: Box-Behnken design [@Box1960-wr] and central-composite designs [@Box1951-ji] via functions `bbd()` and `ccd()`, respectively, where the minimum required input is the number of factors.
 Another form of factorial design is the *saturated designs* where higher-order interaction effects of treatment factors are typically confounded with the main effects.
-Plankett-Burnman designs [@Plackett1946-ly] are a type of saturated design that can be generated by the function `pb()` in the `r rpkg("FrF2")` package, where the user provides the number of experimental runs and the number of treatment factors.
+Plankett-Burnman designs [@Plackett1946-ly] are a type of saturated design that can be generated by the function `pb()` in the `r CRANpkg("FrF2")` package, where the user provides the number of experimental runs and the number of treatment factors.
 
@@
 
-The `r rpkg("agricolae")` package [@agricolae] is the prime example of constructing designs based on a set of so-called "recipe functions", where each function corresponds to a single class of experimental design.
+The `r CRANpkg("agricolae")` package [@agricolae] is the prime example of constructing designs based on a set of so-called "recipe functions", where each function corresponds to a single class of experimental design.
 For example, `design.crd()`, `design.rcbd()` and `design.split()` construct completely randomized, randomized complete block and split-plot designs, respectively.
@@
 
-The use of recipe functions is not limited to classical experimental designs; the `r rpkg("AlgDesign")` [@AlgDesign] package offers three primary functions for generating optimal designs: `optBlock()`, `optFederov()` and `optMonteCarlo()`.
+The use of recipe functions is not limited to classical experimental designs; the `r CRANpkg("AlgDesign")` [@AlgDesign] package offers three primary functions for generating optimal designs: `optBlock()`, `optFederov()` and `optMonteCarlo()`.
 In general, these functions require data and formulas in terms of the supplied data variables with the choice of the criterion (e.g.
@@
 
-Computer experiments, which generally involve space-filling designs, are implemented in packages such as `r rpkg("lhs")` [@lhs] and `r rpkg("DiceDesign")` [@DiceDesign].
-For the `r rpkg("lhs")` package, functions such as `randomLHS()`, `optimumLHS()`, and `maximinLHS()`, require users to specify the sample size ($n$) and the number of variables ($p$), then it generates a Latin hypercube sample [@McKay1979-aw], based on different optimization schemes (in this case, random, S optimal and maxmin criteria, respectively; see package documentation for more details).
-Similarly for `r rpkg("DiceDesign")`, there is a comprehensive list of space-filling designs such as `dmaxDesign()`, `lhsDesign()`, and `wspDesign()` with input of $n$ and $p$ as before (among additional parameters for some) that implement algorithms that either maximize the entropy [@Shewry1987-jn], produce a random Latin hypercube sample, or use the WSP algorithm [@Santiago2012-ui], respectively.
+Computer experiments, which generally involve space-filling designs, are implemented in packages such as `r CRANpkg("lhs")` [@lhs] and `r CRANpkg("DiceDesign")` [@DiceDesign].
+For the `r CRANpkg("lhs")` package, functions such as `randomLHS()`, `optimumLHS()`, and `maximinLHS()`, require users to specify the sample size ($n$) and the number of variables ($p$), then it generates a Latin hypercube sample [@McKay1979-aw], based on different optimization schemes (in this case, random, S optimal and maxmin criteria, respectively; see package documentation for more details).
+Similarly for `r CRANpkg("DiceDesign")`, there is a comprehensive list of space-filling designs such as `dmaxDesign()`, `lhsDesign()`, and `wspDesign()` with input of $n$ and $p$ as before (among additional parameters for some) that implement algorithms that either maximize the entropy [@Shewry1987-jn], produce a random Latin hypercube sample, or use the WSP algorithm [@Santiago2012-ui], respectively.
 These designs output an $n \times p$ matrix with values between 0 and 1.
@@
 Some functions in the DoE packages require the input of existing experimental designs to produce new designs.
-For example, the `r rpkg("DoE.base")` package contains some experimental functions, `cross.design()` and `param.design()`, to combine designs with the former taking a Cartesian product of the input designs while the latter uses a Taguchi style [@Taguchi1986-ut] to aggregate the designs with the inner and outer arrays.
-The `r rpkg("lhs")` package contains a function that `augmentLHS()` to add additional samples to the existing Latin hypercube sample.
+For example, the `r CRANpkg("DoE.base")` package contains some experimental functions, `cross.design()` and `param.design()`, to combine designs with the former taking a Cartesian product of the input designs while the latter uses a Taguchi style [@Taguchi1986-ut] to aggregate the designs with the inner and outer arrays.
+The `r CRANpkg("lhs")` package contains a function that `augmentLHS()` to add additional samples to the existing Latin hypercube sample.
 
-Another class of augmenting design is *sequential design* (also called *adaptive sampling*), which is best represented by `r rpkg("tgp")` [@tgp].
+Another class of augmenting design is *sequential design* (also called *adaptive sampling*), which is best represented by `r CRANpkg("tgp")` [@tgp].
 This requires prior information that is used to inform the next experimental design using `tgp.design()` and `dopt.gp()`.
 The user is required to supply candidate samples to subsample from and a model or a prior experimental design.
-Follow-up experiments, which can also be classified as sequential designs, is implemented by `r rpkg("BsMD")` [@BsMD] using a model-discriminant approach using `MD()`.
+Follow-up experiments, which can also be classified as sequential designs, is implemented by `r CRANpkg("BsMD")` [@BsMD] using a model-discriminant approach using `MD()`.
 
@@
 
-Through the exploratory data analysis of the three data sources (package download logs, package metadata and CRAN task views) outlined in Section `r ref("data")`, we observed in Section `r ref("eda")` that the the total download of DoE packages is concentrated only on a handful of R-packages although these represent a diverse set in comparison with other CRAN task views.
+Through the exploratory data analysis of the three data sources (package download logs, package metadata and CRAN task views) outlined in `r secref("data", "Data")`, we observed in `r secref("eda", "Exploratory data analysis")` that the the total download of DoE packages is concentrated only on a handful of R-packages although these represent a diverse set in comparison with other CRAN task views.
 Furthermore, the data suggest that experimental design is the least collaborative field.
@@
 
-The interface design (discussed in Section `r ref("design")`) reveals that the most widely used DoE packages generally have functions that 1) focus on certain aspects of experimental design (e.g.
+The interface design (discussed in `r secref("design", "Interface design")`) reveals that the most widely used DoE packages generally have functions that 1) focus on certain aspects of experimental design (e.g.
 factorial structure or augmenting design), 2) are a recipe format (i.e.
@@
 A consistent cognitive interface design that leverages existing developments and can be easily extended to new methods will conceivably be of great help to practitioners.
-Some efforts to this end are seen in `r rpkg("DoE.wrapper")` [@DoE.wrapper], which contains wrapper recipe functions for other DoE packages such as `r rpkg("lhs")`, `r rpkg("AlgDesign")` and `r rpkg("FrF2")` (see Figure `r ref("fig:plot-doe-network")`), and is also the subject of the developmental package `r rpkg("edibble")` [@edibble].
+Some efforts to this end are seen in `r CRANpkg("DoE.wrapper")` [@DoE.wrapper], which contains wrapper recipe functions for other DoE packages such as `r CRANpkg("lhs")`, `r CRANpkg("AlgDesign")` and `r CRANpkg("FrF2")` (see Figure `r ref("fig:plot-doe-network")`), and is also the subject of the developmental package `r CRANpkg("edibble")` [@edibble].
 Undoubtedly, no single developer or package can cater to all experimental designs; therefore, any unifying interface should consider how other developers can contribute or add their methods.
@@
 
-This paper uses `r rpkg("targets")` [@targets] and `r rpkg("renv")` [@renv] for reproducibility, `r rpkg("knitr")` [@knitr] and `r rpkg("rmarkdown")` [@rmarkdown] for creating reproducible documents, `r rpkg("ggplot2")` [@ggplot2], `r rpkg("ggraph")` [@ggraph], `r rpkg("ggwordcloud")` [@ggwordcloud] and `colorspace` [@colorspace] for visualization, `r rpkg("kableExtra")` [@kableExtra] for customizing the table in the Supplementary Material and `r rpkg("tidyverse")` [@Wickham2019-mj], `r rpkg("tidytext")` [@tidytext], `r rpkg("pluralize")` [@pluralize], and `r rpkg("ineq")` [@ineq] for data processing and manipulation, and `r rpkg("cranlogs")` [@cranlogs] and `r rpkg("ctv")` [@ctv] for extracting data.
+This paper uses `r CRANpkg("targets")` [@targets] and `r CRANpkg("renv")` [@renv] for reproducibility, `r CRANpkg("knitr")` [@knitr] and `r CRANpkg("rmarkdown")` [@rmarkdown] for creating reproducible documents, `r CRANpkg("ggplot2")` [@ggplot2], `r CRANpkg("ggraph")` [@ggraph], `r CRANpkg("ggwordcloud")` [@ggwordcloud] and `colorspace` [@colorspace] for visualization, `r CRANpkg("kableExtra")` [@kableExtra] for customizing the table in the Supplementary Material and `r CRANpkg("tidyverse")` [@Wickham2019-mj], `r CRANpkg("tidytext")` [@tidytext], `r CRANpkg("pluralize")` [@pluralize], and `r CRANpkg("ineq")` [@ineq] for data processing and manipulation, and `r CRANpkg("cranlogs")` [@cranlogs] and `r CRANpkg("ctv")` [@ctv] for extracting data.
 All codes used to reproduce this paper are found in `r url("https://github.com/emitanaka/paper-DoE-review")`.

### 2022-12-14 paper/arxiv/arxiv.Rmd
reference  setups for specific outlets to accommodate differences in non-numbered/numbered sections and output types (PDF/HTML)

@@
   \newcommand{\ctv}[1]{\href{http://CRAN.R-project.org/view=#1}{\emph{#1}}}
+  \newcommand{\CRANpkg}[1]{\href{https://cran.r-project.org/web/packages/#1/index.html}{\texttt{#1}}}
 bibliography: paper.bib
@@
 
-```{r child=paper}
+```{r, echo= FALSE, results="asis"}
+params <- list()
+params$ref_setup_fn <- "paper/arxiv/ref-setup.R"
+cat(knitr::knit_child(paper, envir = environment(), quiet = TRUE))
 ```

### 2022-12-14 paper/arxiv/arxiv.tex
reference  setups for specific outlets to accommodate differences in non-numbered/numbered sections and output types (PDF/HTML)

--- 2022-06-15 paper/arxiv/arxiv.tex
@@
 \newcommand{\ctv}[1]{\href{http://CRAN.R-project.org/view=#1}{\emph{#1}}}
+\newcommand{\CRANpkg}[1]{\href{https://cran.r-project.org/web/packages/#1/index.html}{\texttt{#1}}}
 
@@
 \begin{abstract}
-Re-running an experiment is generally costly and in some cases
-impossible due to limited resources, so the design of an experiment
-plays a critical role in increasing the quality of experimental data. In
-this paper we describe the current state of the R-packages for the
-design of experiments through an exploratory data analysis of package
-downloads, package metadata, and the comparison of characteristics with
-other topics. We observe that experimental designs in practice appear to
-be sufficiently manufactured by a small number of packages and the
-development of experimental designs often occur in silos. We discuss
-also the interface designs of widely utilised R packages in the field of
-experimental design and discuss its future prospects to advance the
-field in practice.
+Re-running an experiment is generally costly and, in some cases,
+impossible due to limited resources; therefore, the design of an
+experiment plays a critical role in increasing the quality of
+experimental data. In this paper, we describe the current state of
+R-packages for the design of experiments through an exploratory data
+analysis of package downloads, package metadata, and a comparison of
+characteristics with other topics. We observed that experimental designs
+in practice appear to be sufficiently manufactured by a small number of
+packages, and the development of experimental designs often occurs in
+silos. We also discuss the interface designs of widely utilized R
+packages in the field of experimental design and discuss their future
+prospects for advancing the field in practice.
 \end{abstract}
@@
 The critical role of data collection is well captured in the expression
-``garbage in, garbage out'' -- in other words, if the collected data is
-rubbish then no analysis, however complex it may be, can make something
-out of it. A carefully crafted data collection scheme is therefore
-critical to optimise the information from data. The field of
-experimental designs is specifically devoted to planning the collection
-of experimental data, largely based on the founding principles by Fisher
-(1935) or an optimisation framework like those described in Pukelsheim
+``garbage in, garbage out'' -- in other words, if the collected data are
+rubbish, then no analysis, however complex it may be, can make something
+out of it. Therefore, a carefully crafted data-collection scheme is
+critical for optimizing the information from the data. The field of
+experimental design is specifically devoted to planning the collection
+of experimental data, largely based on the founding principles of Fisher
+(1935) or an optimization framework like those described in Pukelsheim
 (2006). These experimental designs are often constructed with the aid of
-a statistical software, such as R (R Core Team 2021), Python (Rossum
-1995), SAS (SAS Institute 1985), and so on, so the usage of experimental
-design software can inform us about some aspects of experimental designs
-in practice.
+statistical software such as R (R Core Team 2021), Python (Rossum 1995),
+and SAS (SAS Institute 1985); thus the use of experimental design
+software can inform us about some aspects of experimental designs in
+practice.
 
-Methods for data collection can be dichotomised by the type of data
+Methods for data collection can be dichotomized by the type of data
 collected -- namely, experimental or observational -- or alternatively,
-categorised as experimental design (including quasi-experimental design)
-or survey design. This dichotomisation, to a great extent, is seen in
+categorized as experimental design (including quasi-experimental design)
+or survey design. This dichotomization, to a great extent, is seen in
 the \href{https://cran.r-project.org/web/views/}{Comprehensive R Archive
@@
 \href{http://CRAN.R-project.org/view=OfficialStatistics}{\emph{OfficialStatistics}}
-task view. The full list of available topics can be see in Table S1 in
-the Supplementary Materials. A subset of experimental designs are
-segregated into
+task view. A full list of available topics is provided in Table S1 in
+the Supplementary Materials. A subset of experimental designs is
+segregated into the
 \href{http://CRAN.R-project.org/view=ClinicalTrials}{\emph{ClinicalTrials}}
 task view, where the focus is on clinical trials with primary interest
-in sample size calculations. This paper focuses on the packages in
+in sample size calculations. This paper focuses on packages in
 \href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}}
@@
 
-In
+From the
 \href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}}
-task view, there are 114 R packages for experimental design and analysis
-of data from experiments. The sheer quantity and variation of
+task view, there are 105 R packages for the experimental design and
+analysis of data from experiments. The sheer quantity and variation of
 experimental designs in the R-packages are arguably unmatched with any
-other programming languages, e.g.~in Python, only a handful of packages
-that generate design of experiment exist (namely \texttt{pyDOE},
-\texttt{pyDOE2}, \texttt{dexpy}, \texttt{experimenter} and
-\texttt{GPdoemd}) with limited type of designs. Thus, the study of DoE
-packages, based on quantitative and qualitative data, can give us an
-objective view to the state of the current experimental designs in
-practice.
+other programming languages, for example, in Python, only a handful of
+packages that generate design of experiment exist (namely
+\texttt{pyDOE}, \texttt{pyDOE2}, \texttt{dexpy}, \texttt{experimenter},
+and \texttt{GPdoemd}) with a limited type of design. Thus, the study of
+DoE packages, based on quantitative and qualitative data, can provide an
+objective view of the state of current experimental designs in practice.
 
-A utility of software can also be described by its design to facilitate
-clear expression and interpretation of the desired experimental design.
-Certain programming language design can hinder or discourage development
-of a reliable program (Wasserman 1975). The immense popularity of
-\texttt{tidyverse} (a collection of R-packages for various stages of
-data analysis that place enormous emphasis on the interface design by
-Wickham et al. 2019) is a testament to the impact an interface design
-can have in practice. The practice of experimental design could be
-advanced by adopting similar interface design principles across DoE
-packages.
+The utility of the software can also be described by its design to
+facilitate the clear expression and interpretation of the desired
+experimental design. Certain programming language designs can hinder or
+discourage the development of reliable programs (Wasserman 1975). The
+immense popularity of
+\href{https://cran.r-project.org/web/packages/tidyverse/index.html}{\texttt{tidyverse}}
+(a collection of R-packages for various stages of data analysis that
+places enormous emphasis on the interface design by Wickham et al. 2019)
+is a testament to the impact that an interface design can have in
+practice. The practice of experimental design can be advanced by
+adopting similar interface design principles across the DoE packages.
 
-The paper is organised as follows. Section \ref{data} briefly describes
-the data source used for the analysis in Section \ref{eda}; Section
+The remainder of this paper is organized as follows. Section \ref{data}
+briefly describes the data source used for the analysis; Section
 \ref{eda} presents some insights into the state of the current DoE
@@
 The Comprehensive R Archive Network (CRAN) is a network of servers
-located across the world that store mirrored versions of R and
-R-packages. The most popular network is the RStudio mirror (the default
+located across the world that stores mirrored versions of the R and R
+packages. The most popular network is the RStudio mirror (the default
 server for those that use the RStudio IDE). The RStudio mirror is also
 the only server that provides a comprehensive daily download logs of R
-and R-packages since October 2012. The summary data can be easily
-accessed with the \texttt{cranlogs} package (Csárdi 2019). This paper
-uses the data from the beginning of 2013 to end of 2021 (a total of 9
-years) for the packages in the CRAN task views.
+and R packages since October 2012. The summary data can be easily
+accessed using the
+\href{https://cran.r-project.org/web/packages/cranlogs/index.html}{\texttt{cranlogs}}
+package (Csárdi 2019). This paper uses the data from the beginning of
+2013 to the end of 2021 (a total of nine years) for the packages in the
+CRAN task views.
 
@@
 All CRAN packages have a title, description, package connections
-(suggests, depends and imports) and other meta-information in the
-DESCRIPTION file. We use the text data from the title and description
-(as accessed in 2022-05-23) in Section \ref{topics}.
+(suggests, depends, and imports of other packages), and other
+meta-information in the DESCRIPTION file. We use text data from the
+title and description (accessed in 2022-12-12).
 
@@
 
-CRAN task views are volunteer maintained list of R-packages on CRAN
-relevant to the corresponding topic. There are a total 37 CRAN task
-views. Table S1 in the Supplementary Materials list the full available
-topics from the \texttt{ctv} package (Zeileis 2005). The list of
-packages in each CRAN task views (as of 2022-05-25) are used to contrast
-the characteristics of DoE packages in Sections \ref{popular} and
-\ref{silo}.
+CRAN task views are volunteer-maintained lists of R-packages on CRAN
+relevant to the corresponding topic. There were 39 CRAN task views in
+total. Table S1 in the Supplementary Materials list the available topics
+from the
+\href{https://cran.r-project.org/web/packages/ctv/index.html}{\texttt{ctv}}
+package (Zeileis 2005). The list of packages in each CRAN task view (as
+of 2022-12-12) is used to contrast the characteristics of the DoE
+packages.
 
 \hypertarget{eda}{%
-\section{Explorative data analysis}\label{eda}}
+\section{Exploratory data analysis}\label{eda}}
 
-In this section, we derive some conjecture based on the analysis of the
-data described in Section \ref{data}. All results presented are from the
+In this section, we derive some conjectures based on an analysis of the
+data described in Section \ref{data}. All results presented are from
 exploratory data analysis of observational data, consequently, all
@@
 the true state of the field of experimental design. In particular, any
-analysis over time are confounded by the fact that the nature of users
-and package management have also changed over the years. It should be
-noted that some DoE packages may have been archived or removed from the
-task view over the years so any cross-sectional analysis presented may
-not reflect the set of all DoE packages at that particular time period
-(although we assume such incidences are low).
+analysis over time is confounded by the fact that the nature of users
+and package management has changed over the years. It should be noted
+that some DoE packages may have been archived or removed from the task
+view over the years; therefore, any cross-sectional analysis presented
+may not reflect the set of all DoE packages at that particular time
+period (although we assume such incidences are low).
 
-A subset of DoE packages is not primarily about design of experiments
-but about the analysis of experimental data. A complete delineation of
-these packages is difficult as there is almost always at least one
-function that can aid decisions or constructions of experimental designs
-(and any categorisation is prone to our subjective bias) so we opted not
-to remove any DoE packages in the analysis.
+A subset of DoE packages is not primarily about the design of
+experiments but about the analysis of experimental data. A complete
+delineation of these packages is difficult, as there is almost always at
+least one function that can aid decisions or constructions of
+experimental designs (and any categorization is prone to our subjective
+bias); therefore we opted not to remove any DoE packages in the
+analysis.
 
-\hypertarget{popular}{%
+\hypertarget{small-but-diverse-set-of-packages-are-sufficient-for-most-experimental-designs-in-practice}{%
 \subsection{Small, but diverse, set of packages are sufficient for most
-experimental designs in practice}\label{popular}}
+experimental designs in
+practice}\label{small-but-diverse-set-of-packages-are-sufficient-for-most-experimental-designs-in-practice}}
 
-There have been at least 50 DoE packages since 2013 but most of the
-downloads are concentrated in just a handful of packages. For example,
-Figure \ref{fig:plot-lorenz} shows a Lorenz curve (Lorenz 1905) for the
-total package downloads in 2021 of 113 DoE packages (first released
-prior to 2021); we can see from Figure \ref{fig:plot-lorenz} that bottom
-90\% of DoE packages (in terms of total download count in 2021) only
-share about 30\% of total downloads across all DoE packages -- in
-another words, 70\% of the total downloads are due to 11 packages (10\%
-of the DoE packages).
+There are at least 50 DoE packages since 2013 but most of the downloads
+are concentrated in only a handful of packages. For example, Figure
+\ref{fig:plot-lorenz} shows a Lorenz curve (Lorenz 1905) for the total
+package downloads in 2021 for 102 DoE packages (first released prior to
+2021). We can see from Figure \ref{fig:plot-lorenz} that the bottom 90\%
+of DoE packages (in terms of total download count in 2021) only share
+approximately 32\% of total downloads across all DoE packages; in other
+words, 68\% of the total downloads are due to 10 packages (10\% of the
+DoE packages).
 
@@
 The red line corresponds to the line of perfect equality.
-The green region shows the area under the Lorenz curve and the red region shows the area of gap in equality.}\label{fig:plot-lorenz}
+The yellow region shows the area under the Lorenz curve and the red region shows the area of the gap in equality.}\label{fig:plot-lorenz}
 \end{figure}
@@
 consider using the Gini index (Gini 1921) as a measure of download
-inequality across packages. The ratio of the red region over the total
+inequality across packages. The ratio of the red region to the total
 colored regions in Figure \ref{fig:plot-lorenz} corresponds to the Gini
 index for 2021. A Gini index of 0\% indicates equality in downloads
-across packages while a value of 100\% indicates maximal inequality (all
-downloads are due to one package). In Figure \ref{fig:download-share},
-we see that the distributions of the package downloads each year have a
-heavy right tail with the Gini index ranging from 42.6\% to 74.1\%
-across the years 2013 to 2021, indicating that there is a high level of
-inequality of package downloads, particularly with more pronounced
-inequality in the last 6 years.
+across packages whereas a value of 100\% indicates maximal inequality
+(all downloads are due to one package). In Figure
+\ref{fig:download-share}, we see that the distributions of the package
+downloads each year have a heavy right tail with the Gini index ranging
+from 32.7\% to 69.1\% across the years 2013 to 2021, indicating that
+there is a high level of inequality in package downloads, particularly
+with more pronounced inequality in the last six years.
 
@@
 
-\caption{The distribution of the number of downloads for each DoE package by year.
-Packages were removed in any year if it was released in that year or later so that each download count is for the full year.
-The label on the bottom of the plot shows the Gini index for downloads and the number of packages with full year of download count in the corresponding year.
-In the last 6 years, the Gini index is consistent above 60\% each year indicating that most downloads are due to a relatively small number of packages.
+\caption{Distribution of number of downloads for DoE packages by year.
+Packages were removed in any year if they were released in that year or later so that each download count was for the full year.
+The label at the bottom of the plot shows the Gini index for downloads and the number of packages with a full download count in the corresponding year.
+In the last six years, the Gini index has consistently exceeded 60\%, indicating that most downloads are due to a relatively small number of packages.
 }\label{fig:download-share}
@@
 
-The increase in the number of packages that are not highly downloaded
-may mean that \textbf{\emph{there are more packages to construct niche
+An increase in the number of packages that are not highly downloaded may
+mean that \textbf{\emph{there are more packages to construct niche
 experimental designs}}. Some examples of these packages include
-\texttt{qtlDesign}, \texttt{PwrGSD} and \texttt{Crossover} made for QTL
-experiments, group sequential designs and crossover trials,
-respectively. These packages would naturally have a smaller number of
-potential users. Counterfactual to this, the increase could be due to
-other external factors, such as an increase in the number of skilled
-developers (and thus more pacakge contributions), a change in CRAN
-policy or management to add packages (either to CRAN and/or task view),
-and/or that new packages are still yet to amass users. While there is an
-argument that low download counts are due to the low utility and/or
-quality of the packages, packages in CRAN task views are selected by
-expert maintainers, thus we can reasonably assume that any package
-listed in CRAN task view are of a decent utility and quality.
+\href{https://cran.r-project.org/web/packages/qtlDesign/index.html}{\texttt{qtlDesign}},
+\href{https://cran.r-project.org/web/packages/PwrGSD/index.html}{\texttt{PwrGSD}}
+and
+\href{https://cran.r-project.org/web/packages/Crossover/index.html}{\texttt{Crossover}}
+made for QTL experiments, group sequential designs and crossover trials,
+respectively. These packages would naturally have fewer potential users.
+Counterfactual to this, the increase could be due to other external
+factors, such as an increase in the number of skilled developers (and
+thus more package contributions), a change in CRAN policy or management
+to add packages (either to CRAN and/or task view), and/or the fact that
+new packages are still yet to amass users. While there is an argument
+that low download counts are due to the low utility and/or quality of
+packages, packages in CRAN task views are selected by expert
+maintainers. We can reasonably assume that any package listed in the
+CRAN task view has an acceptable utility and quality.
 
 If the downloads are reflective of the experimental designs used in
-practice, \textbf{\emph{small set of packages appear to be sufficient
-for most to construct the full set of designs of experiments needed in
-practice}}. Packages of course evolve and the top downloaded packages
-have had regular updates that may have broadened its scope from its
-previous releases.
+practice, \textbf{\emph{a small set of packages appears to be sufficient
+for most users to construct the full set of designs of experiments they
+need in practice}}. Packages of course evolve, and the top downloaded
+packages have had regular updates that may have broadened their scope
+from their previous releases.
 
-While in absolute terms the Gini index is high for
+While in absolute terms the Gini index is high for the
 \href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}}
-task view (42.6\% to 74.1\%), the inequality is not as severe as other
-CRAN task views as shown in Figure \ref{fig:fig-gini-all-ctvs}. We can
-see in Figure \ref{fig:fig-gini-all-ctvs} that the Gini index is
-generally increasing for DoE packages (as is generally the case for
-other CRAN task views as shown in Figure S1 in the Supplementary
-Materials) but most other CRAN task views have a Gini index of over
+task view (32.7\% to 69.1\%), the inequality is not as severe as in
+other CRAN task views as shown in Figure \ref{fig:fig-gini-all-ctvs}. We
+can see in Figure \ref{fig:fig-gini-all-ctvs} that the Gini index
+generally increases over time for DoE packages (as is generally the case
+for other CRAN task views as shown in Figure S1 in the Supplementary
+Materials), but most other CRAN task views have a Gini index of over
 75\%. This suggests that other CRAN task views may have dominant
-standards and in comparison to other topics, there are more
-\textbf{\emph{diverse approaches to designing experiments}} and thus no
-single DoE package is dominant. This observation however doesn't take
-into account other approaches to generating experimental designs, like
-the proprietary software, CycDesignN (Whittaker, Williams, and John
-2022) and SAS (SAS Institute 1985), which may be widely used.
+standards, and in comparison to other topics, there are more
+\textbf{\emph{diverse approaches to designing experiments}}, and thus,
+no single DoE package is dominant. However, this observation doe not
+consider other approaches to generate experimental designs, such as the
+proprietary software, CycDesignN (Whittaker, Williams, and John 2022),
+which may be widely used.
 
@@
 
-\caption{The points show the Gini index of the download counts by year for ExperimentalDesign task view with the color showing the number of packages.
-The grey line shows the distribution of the Gini index across years for all other CRAN task views.
-See Figure S1 in the Supplementary Material for the line graph for the Gini index across years by each CRAN task view.}\label{fig:fig-gini-all-ctvs}
+\caption{The points show the Gini index of the download counts by year for the ExperimentalDesign task view with the color showing the number of packages.
+The gray lines show the line plots of the Gini index across years for all other CRAN task views.
+See Figure S1 in the Supplementary Material for the line graph of the Gini index across years for each CRAN task view.}\label{fig:fig-gini-all-ctvs}
 \end{figure}
 
-\hypertarget{ranking}{%
+\hypertarget{the-field-of-experimental-design-is-slow-changing}{%
 \subsection{The field of experimental design is
-slow-changing}\label{ranking}}
+slow-changing}\label{the-field-of-experimental-design-is-slow-changing}}
 
 We can see in Figure \ref{fig:rank-over-time} that most of the top 10
-ranking packages have been in the top 10 for the last 9 years with
-\texttt{lhs} steadily climbing up the ranks in the last few years. It
-should be noted that the download of one package can prompt a download
-of another package; the most notable dependency network is
-\texttt{AlgDesign} and \texttt{agricolae}, where the former is an import
-for the latter. The full dependency network within the DoE packages is
-shown later in Figure @ref(fig:plot-doe-network).
+ranking packages have been in the top 10 for the last nine years with
+\href{https://cran.r-project.org/web/packages/lhs/index.html}{\texttt{lhs}}
+steadily climbing up the ranks in the last few years. It should be noted
+that the download of one package can prompt the download of another
+package; the most notable package connection is
+\href{https://cran.r-project.org/web/packages/AlgDesign/index.html}{\texttt{AlgDesign}}
+and
+\href{https://cran.r-project.org/web/packages/agricolae/index.html}{\texttt{agricolae}},
+where the former is an import for the latter. The full network of
+package connections within the DoE packages is shown in Figure
+\ref{fig:plot-doe-network}.
 
@@
 
-\caption{The plot shows the rank of top 10 downloaded packages by year.
-Packages that do not appear in top 10 for at least two periods are omitted from the plot.
+\caption{The plot shows the rank of the top 10 packages downloaded by year.
+Packages that did not appear in the top 10 for at least two periods were omitted from the plot.
 Most packages are consistently in the top 10 for the period shown.}\label{fig:rank-over-time}
@@
 correlation between the first release date and the (log of) total
-download counts of DoE packages in any given year from 2013 to 2021.
-This suggest that in general, a package released earlier are more likely
-to be used today (possibly for legacy reasons or the general inertia to
-adopt new packages). We can also see in Figure
-\ref{fig:release-date-vs-download} the most downloaded packages were
-generally released in 2004 to 2010.
+download counts of DoE packages for any given year from 2013 to 2021.
+This suggests that packages released earlier are more likely to be used
+today (possibly for legacy reasons or the general inertia to adopt new
+packages). We can also see in Figure \ref{fig:release-date-vs-download}
+that most downloaded packages were released from 2004 to 2010.
 
@@
 \caption{The above figure shows the total download (in log scale) of a package in the corresponding year against the first release date of the package.
-The blue line corresponds to the least squares fit of a simple linear regression model.
-The label in the right-hand upper corner shows the sample correlation coefficient between the first release date and the log (with base 10) of the total download count.
-The high-leverage point on the far left belongs to `conf.design`, authored by one of the earlier contributors to R.}\label{fig:release-date-vs-download}
+The blue line corresponds to the least-squares fit of the simple linear regression model.
+The label in the upper right-hand corner shows the sample correlation coefficient between the first release date and log (with base 10) of the total download count.
+The high leverage point on the far left belongs to `conf.design`, authored by one of the earlier contributors to R.}\label{fig:release-date-vs-download}
 \end{figure}
@@
 \ref{fig:rank-over-time}) and the fact that most downloaded DoE packages
-are first released more than 10 years ago (Figure
-\ref{fig:release-date-vs-download}) indicates that either the existing
-packages filled the needs of mass in practice or no new packages were
-compelling for many to switch their practice. We do however also see
-that the the top downloaded packages generally have more updates (see
-Figure \ref{fig:release-date-vs-download}) so it is possible that the
-packages have improved or broadened the scope of its usage.
+were first released more than 10 years ago (Figure
+\ref{fig:release-date-vs-download}) indicate that either the existing
+packages fulfilled the needs of mass in practice or no new packages were
+compelling for many to switch their practice. However, we also see that
+the top downloaded packages generally have more updates (see Figure
+\ref{fig:release-date-vs-download}); therefore, it is possible that the
+packages have improved or broadened the scope of their usage.
 
-\hypertarget{topics}{%
-\subsection{Optimal designs are of interest}\label{topics}}
+\hypertarget{optimal-designs-are-of-interest}{%
+\subsection{Optimal designs are of
+interest}\label{optimal-designs-are-of-interest}}
 
-Figure \ref{fig:wordcloud-over-time} shows some common purposes of DoE
-packages, based on bigrams in the package title and description. We show
-only the bigrams as unigram was not insightful and there were not many
-trigrams common across packages. In counting the bigrams, we processed
-the text data as follows:
+Figure \ref{fig:wordcloud-over-time} shows some of the common purposes
+of DoE packages, based on bigrams in the package title and description.
+We only show the bigrams as unigrams were not insightful, and there were
+not many trigrams common across packages. To count the bigrams, we
+processed the text data as follows:
 
@@
 \item
-  We standardised the words to lower case and removed pluralization.
+  We standardized the words to the lower case and removed pluralization.
 \item
-  Multiple mentions of the same bigram within a package was counted as
-  one (e.g.~\texttt{AlgDesign} mentions ``experimental design'' four
-  times in the title and the description but this is counted as one).
+  Multiple mentions of the same bigram within a package were counted as
+  one (for example,
+  \href{https://cran.r-project.org/web/packages/AlgDesign/index.html}{\texttt{AlgDesign}}
+  mentions ``experimental design'' four times in the title and the
+  description, but this is counted as one).
 \item
-  Bigrams that consist the stop words (provided in
-  \texttt{tidytext::stop\_words} in addition to other words we deemed
-  irrelevant, e.g.~``provide'', ``e.g.'', ``calculate'' and so on; full
-  list is shown in the code provided in the link under Section
-  \ref{pkgs}) are removed.
+  Bigrams consisting of stop words were removed. The stop words are
+  sourced from the lexicons in \texttt{tidytext::stop\_words} in
+  addition to other words we deemed irrelevant, e.g.~``provide'',
+  ``e.g.'', ``calculate'' and so on -- full list is shown in the code
+  provided in the link under Section \ref{pkgs}.
 \end{enumerate}
 
-Not surprisingly, the bigram ``experimental design'' was the most
-common. More interestingly, ``optimal design'' and ``sequential design''
+Unsurprisingly, the bigram ``experimental design'' was the most common.
+More interestingly, ``optimal design'' and ``sequential design''
 appeared across different packages (indicated by the size of the word in
@@
 hypercube sampling and computer experiments (which generally include
-space filling designs like Latin hypercube sampling) are designs that
-generally operate by optimising a user-selected criterion and can be
-classified as optimal designs.
+space-filling designs such as Latin hypercube sampling) generally
+operate by optimizing a user-selected criterion and can be classified as
+optimal designs.
 
@@
 \caption{The above figure shows the word cloud of bigrams from the title and descriptions of the DoE packages.
-The size shows how often the bigram appears across the DoE packages and the color are relative to the total download count in 2021 of the packages that contain the bigram.}\label{fig:wordcloud-over-time}
+The size shows how often the bigram appears across the DoE packages and the color is relative to the total download count in 2021 for the packages that contain the bigram.}\label{fig:wordcloud-over-time}
 \end{figure}
@@
 
-\hypertarget{silo}{%
+\hypertarget{development-of-experimental-designs-occur-in-silos}{%
 \subsection{Development of experimental designs occur in
-silos}\label{silo}}
+silos}\label{development-of-experimental-designs-occur-in-silos}}
 
@@
 \href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}}
-task view has the lowest average number of contributors among all 37
+task view has the lowest average number of contributors among all the 39
 CRAN task views. In addition, we can also see in Figure
@@
 packages that make use of other packages within the same task view). The
-full connection of DoE packages with other DoE packages is shown in
-Figure \ref{fig:plot-doe-network}. These observations suggest that the
-\textbf{\emph{field of experimental design is one of the least
-collaborative field}} and package development generally occur in silos.
+full connection between DoE packages is shown in Figure
+\ref{fig:plot-doe-network}. These observations suggest that
+\textbf{\emph{experimental design is one of the least collaborative
+fields}} and package development generally occurs in silos.
 
@@
 
-\caption{The above figure is a scatterplot of the intra-connectivity (the percentage of packages that depends, suggest or imports at least one other package within the same task view) and the average number of contributors for each CRAN task view.
-A low intra-connectivity suggests that development within the topic mostly occur in silos whilst high  intra-connectivity suggests that there are more interactions within the topic.
-The color shows the number of packages, the size of the point corresponds to the total number of contributors, and the text labels show the CRAN task view name.
-The label of ExperimentalDesign task view is colored red.
+\caption{The above figure is a scatterplot of intra-connectivity (the percentage of packages that depends, suggests or imports at least one other package within the same task view) and the average number of contributors for each CRAN task view.
+Low intra-connectivity suggests that development within the topic mostly occurs in silos, while high  intra-connectivity suggests that there are more interactions within the topic.
+The color shows the number of packages, the size of the point corresponds to the total number of contributors, and the text labels show the CRAN task view names.
+The label of ExperimentalDesign task view is colored in red.
 The task views in the bottom-left corner are topics that are more indicative of contributors working in silos.
-The actual numerical values are show in Table S1 in the Supplementary Material.}\label{fig:ctv-summ-plot}
+The actual numerical values are listed in Table S1 in the Supplementary Material.}\label{fig:ctv-summ-plot}
 \end{figure}
@@
 
-\caption{Package connections (depends, suggests and imports) within the DoE packages.
-The direction of the arrow shows the connection of packags where the package on the tail of the arrow is a depedency, suggestion or import for the package on the head of the arrow.
-DoE packages that do not depend, suggest or import another DoE package is not shown.}\label{fig:plot-doe-network}
+\caption{Package connections (depends, suggests and imports) within DoE packages.
+The direction of the arrow shows the connection of packags where the package on the tail of the arrow is a dependency, suggestion or import for the package on the head of the arrow.
+DoE packages that do not depend, suggest or import another DoE package are not shown.}\label{fig:plot-doe-network}
 \end{figure}
@@
 In software design, there are two interface designs to consider: user
-interface (UI) and application programming interface (API). UI is
+interface (UI) and application programming interface (API). The UI is
 concerned with the interaction of the software by the user, while the
 API is concerned with how different programs interact and is
-predominately of the interest to the developer. The UI design is an
-abstraction to specifying the desired experimental design and its
-choices enable how a user expresses the specification of the
-experimental design. The API design aids in other developers to leverage
-existing systems.
+predominately of interest to the developer. The UI design is an
+abstraction that specifies the desired experimental design, and its
+choices enable how a user expresses the specification of an experimental
+design. The API design aids other developers in leveraging existing
+systems.
 
-In this section, we discuss these interface designs broadly, with
-examples illustrated from the top downloaded packages (shown in Figure
-\ref{fig:rank-over-time}). We exclude \texttt{ez} and
-\texttt{DiceKriging} from the discussion as the former is predominately
-visualisation of experimental data and the latter is about the analysis
-of computer experiments in addition to belonging to the same suite of
-packages as \texttt{DiceDesign} (Dupuy, Helbert, and Franco 2015).
+In this section, we discuss the interface designs of functions that
+output an experimental design based on three broad areas: factorial,
+recipe and augmenting designs. The discussion is exclusive to the top
+downloaded packages (shown in Figure \ref{fig:rank-over-time}), with the
+exception of
+\href{https://cran.r-project.org/web/packages/ez/index.html}{\texttt{ez}}
+and
+\href{https://cran.r-project.org/web/packages/DiceKriging/index.html}{\texttt{DiceKriging}},
+as the former is predominately visualization of experimental data and
+the latter is about the analysis of computer experiments in addition to
+belonging to the same suite of packages as
+\href{https://cran.r-project.org/web/packages/DiceDesign/index.html}{\texttt{DiceDesign}}
+(Dupuy, Helbert, and Franco 2015).
 
@@
 
-Factorial experiments offer a challenge in the construction and
-allocation of the treatment factors. This is reflected in a number of
-packages that specifically addresses this challenge. These include
-\texttt{DoE.base} (Grömping 2018) for full factorial and (regular and
+Factorial experiments offer a challenge in allocating the treatment
+factors to experimental units where the full set of factorial treatments
+cannot be administered (and replicated) and/or the experimental units
+have a grouping structure. The effort to address this challenge is
+reflected in the number of packages that focus on the construction of
+factorial designs as described next.
+
+The
+\href{https://cran.r-project.org/web/packages/DoE.base/index.html}{\texttt{DoE.base}}
+package (Grömping 2018) can construct full factorial and (regular and
 irregular) orthogonal array designs via \texttt{fac.design()} and
-\texttt{oa.design()}, respectively, \texttt{FrF2} (Grömping 2014) for
-fractional 2-level factorial designs using \texttt{FrF2()},
-\texttt{FrF2Large()} or \texttt{pb()} Plankett-Burnman designs (Plackett
-and Burman 1946), \texttt{conf.design} (Venables 2013) for symmetric
-confounded factorial designs via \texttt{conf.design()}, and
-\texttt{BHH2} (Barrios 2016) to generate full or fractional 2-level
-factorial design matrix via \texttt{ffDesMatrix()}. All these designs
-generally requires the user to input the number of factors and if the
-design is allowed to vary in the number of levels, the input of the
-number of levels for each factor. The construction of some of these
-designs are catalogue-based, e.g.~\texttt{oa.design()} in
-\texttt{DoE.base}.
+\texttt{oa.design()}, respectively. The
+\href{https://cran.r-project.org/web/packages/FrF2/index.html}{\texttt{FrF2}}
+package (Grömping 2014) constructs regular fractional 2-level factorial
+designs using \texttt{FrF2()} and \texttt{FrF2Large()}, with the latter
+for large designs. The
+\href{https://cran.r-project.org/web/packages/conf.design/index.html}{\texttt{conf.design}}
+package (Venables 2013) constructs symmetric confounded factorial
+designs via \texttt{conf.design()} and the
+\href{https://cran.r-project.org/web/packages/BHH2/index.html}{\texttt{BHH2}}
+package (Barrios 2016) generates a full or fractional 2-level factorial
+design matrix via \texttt{ffDesMatrix()}.
 
-Response surface design, as illustrated by \texttt{rsm} (Lenth 2009), is
-a factorial experiment but the focus of it is to analyse using response
-surface modelling. The two well known response surface designs are
-Box-Behnken design (Box and Behnken 1960) and central-composite designs
-(Box and Wilson 1951) via \texttt{bbd()} and \texttt{ccd()} for
-central-composite designs providing formula.
+While the argument names and underlying algorithms of these functions to
+generate factorial designs differ, it generally requires the users to
+input the number of:
 
-Saturated designs
+\begin{itemize}
+\tightlist
+\item
+  treatment factors,
+\item
+  experimental runs (or replications), and
+\item
+  levels for each factor if the design was allowed to vary in the number
+  of levels.
+\end{itemize}
 
-\hypertarget{the-case-of-menu-designs}{%
-\subsection{The case of menu designs}\label{the-case-of-menu-designs}}
+The output of the design is either a special class of list (e.g.~the
+\texttt{design} class for
+\href{https://cran.r-project.org/web/packages/DoE.base/index.html}{\texttt{DoE.base}}
+and
+\href{https://cran.r-project.org/web/packages/FrF2/index.html}{\texttt{FrF2}}
+or a \texttt{data.frame} for
+\href{https://cran.r-project.org/web/packages/conf.design/index.html}{\texttt{conf.design}})
+or a matrix for \texttt{BHH2} such that an element or column corresponds
+to a treatment factor with each value corresponding to one experimental
+run. The treatment factors are generally assigned pseudo names
+(e.g.~letters of the alphabet) or an argument exists for users to input
+treatment names as a character vector.
 
-Computer experiments, as discussed in Section \ref{topics}, which
-generally involve space filling designs, have been on the rise. The
-exemplar for this type of designs are \texttt{lhs} (Carnell 2022) and
-\texttt{DiceDesign}. For the \texttt{lhs} package, functions like
-\texttt{randomLHS()}, \texttt{optimumLHS()}, and \texttt{maximinLHS()},
-require users to specify the sample size (\(n\)) and the number of
-variables (\(p\)) then it generates a Latin hypercube sample, based on
-different optimisation schemes (in this case, random, S optimal and
-maxmin criteria, respectively; see package documentation for more
-details). Similarly for the \texttt{DiceDesign}, there is a
-comprehensive list of space-filling designs such as
-\texttt{dmaxDesign()}, \texttt{lhsDesign()}, \texttt{straussDesign()}
-and \texttt{wspDesign()} with input of \(n\) and \(p\) as before (among
-extra parameters for some) that implement algorithms that either
-maximise the entropy, produce a random Latin hypercube sample, based on
-Strauss process and WSP algorithm, respectively. These designs all
-output a \(n \times p\) matrix with values between 0 and 1.
+Some forms of factorial design are known by other names, for example,
+\emph{response surface designs} are factorial designs where the
+treatment factors are discrete levels of continuous variables. Two types
+of response surface designs can be constructed using the
+\href{https://cran.r-project.org/web/packages/rsm/index.html}{\texttt{rsm}}
+package (Lenth 2009): Box-Behnken design (Box and Behnken 1960) and
+central-composite designs (Box and Wilson 1951) via functions
+\texttt{bbd()} and \texttt{ccd()}, respectively, where the minimum
+required input is the number of factors. Another form of factorial
+design is the \emph{saturated designs} where higher-order interaction
+effects of treatment factors are typically confounded with the main
+effects. Plankett-Burnman designs (Plackett and Burman 1946) are a type
+of saturated design that can be generated by the function \texttt{pb()}
+in the
+\href{https://cran.r-project.org/web/packages/FrF2/index.html}{\texttt{FrF2}}
+package, where the user provides the number of experimental runs and the
+number of treatment factors.
 
-The \texttt{AlgDesign} (Wheeler 2022) package offers three primary
-functions for generating optimal designs: \texttt{optBlock()},
-\texttt{optFederov()} and \texttt{optMonteCarlo()}. These functions in
-general require a formula in terms of the supplied terms with the choice
-of criterion (D, A and I for the latter two). The difference lies in the
-search algorithm for the optimal design.
+\hypertarget{the-case-of-recipe-designs}{%
+\subsection{The case of recipe
+designs}\label{the-case-of-recipe-designs}}
 
-The \texttt{agricolae} package (de Mendiburu 2021) was the most
-downloaded DoE packages in 2013 to 2018 (Figure
-\ref{fig:rank-over-time}) and possibly prior to 2013 (data unavailable).
-This package is the prime example of constructing designs based on menu
-functions, e.g.~\texttt{design.crd()}, \texttt{design.rcbd()} and
-\texttt{design.split()} construct a completely randomised design,
-randomised complete block design and split-plot design, respectively.
-Users typically supply the treatment labels (or number of treatments in
-the case of \texttt{design.split()}) and the number of replications as
-argument to these functions.
+The
+\href{https://cran.r-project.org/web/packages/agricolae/index.html}{\texttt{agricolae}}
+package (de Mendiburu 2021) is the prime example of constructing designs
+based on a set of so-called ``recipe functions'', where each function
+corresponds to a single class of experimental design. For example,
+\texttt{design.crd()}, \texttt{design.rcbd()} and
+\texttt{design.split()} construct completely randomized, randomized
+complete block and split-plot designs, respectively. Users typically
+supply treatment labels (or the number of treatments in the case of
+\texttt{design.split()}) and the number of replications as arguments for
+these functions. The output is a list with one element corresponding to
+a \texttt{data.frame} that contains the design in a table such that the
+row corresponds to the experimental run and the columns correspond to
+the experimental variables (we refer to this format simply as ``table
+format'' henceforth).
 
-\hypertarget{the-case-of-sequential-designs}{%
-\subsection{The case of sequential
-designs}\label{the-case-of-sequential-designs}}
+The use of recipe functions is not limited to classical experimental
+designs; the
+\href{https://cran.r-project.org/web/packages/AlgDesign/index.html}{\texttt{AlgDesign}}
+(Wheeler 2022) package offers three primary functions for generating
+optimal designs: \texttt{optBlock()}, \texttt{optFederov()} and
+\texttt{optMonteCarlo()}. In general, these functions require data and
+formulas in terms of the supplied data variables with the choice of the
+criterion (e.g.~the D-criterion), with the output as a list with one
+element corresponding to the design in a table format. The difference
+between these functions lies in the underlying search strategy for
+optimal designs, and the name of the function is a surrogate for the
+search algorithm.
 
-Another specialised design is sequential design (also called adaptive
-sampling), which is best represented by \texttt{tgp} (Gramacy and Taddy
-2010) -- these require prior information, which are used to inform the
-next experimental design using \texttt{tgp.design()} and
-\texttt{dopt.gp()} which requires users to supply candidate samples to
-sub-sample from and a model or design from a prior design. Follow-up
-experiments, which can also be classified as a sequential design, is
-implemented by \texttt{BsMD} (Barrios 2020). In \texttt{BsMD}, the
-follow-up design is determined by a model discriminant approach using
-\texttt{MD()}.
+Computer experiments, which generally involve space-filling designs, are
+implemented in packages such as
+\href{https://cran.r-project.org/web/packages/lhs/index.html}{\texttt{lhs}}
+(Carnell 2022) and
+\href{https://cran.r-project.org/web/packages/DiceDesign/index.html}{\texttt{DiceDesign}}
+(Dupuy, Helbert, and Franco 2015). For the
+\href{https://cran.r-project.org/web/packages/lhs/index.html}{\texttt{lhs}}
+package, functions such as \texttt{randomLHS()}, \texttt{optimumLHS()},
+and \texttt{maximinLHS()}, require users to specify the sample size
+(\(n\)) and the number of variables (\(p\)), then it generates a Latin
+hypercube sample (McKay, Beckman, and Conover 1979), based on different
+optimization schemes (in this case, random, S optimal and maxmin
+criteria, respectively; see package documentation for more details).
+Similarly for
+\href{https://cran.r-project.org/web/packages/DiceDesign/index.html}{\texttt{DiceDesign}},
+there is a comprehensive list of space-filling designs such as
+\texttt{dmaxDesign()}, \texttt{lhsDesign()}, and \texttt{wspDesign()}
+with input of \(n\) and \(p\) as before (among additional parameters for
+some) that implement algorithms that either maximize the entropy (Shewry
+and Wynn 1987), produce a random Latin hypercube sample, or use the WSP
+algorithm (Santiago, Claeys-Bruno, and Sergent 2012), respectively.
+These designs output an \(n \times p\) matrix with values between 0 and
+1. Again, these functions employ a recipe style, in which each function
+has a name that corresponds to a certain search strategy to generate the
+experimental design.
+
+\hypertarget{the-case-of-augmenting-designs}{%
+\subsection{The case of augmenting
+designs}\label{the-case-of-augmenting-designs}}
+
+Some functions in the DoE packages require the input of existing
+experimental designs to produce new designs. For example, the
+\href{https://cran.r-project.org/web/packages/DoE.base/index.html}{\texttt{DoE.base}}
+package contains some experimental functions, \texttt{cross.design()}
+and \texttt{param.design()}, to combine designs with the former taking a
+Cartesian product of the input designs while the latter uses a Taguchi
+style (Taguchi 1986) to aggregate the designs with the inner and outer
+arrays. The
+\href{https://cran.r-project.org/web/packages/lhs/index.html}{\texttt{lhs}}
+package contains a function that \texttt{augmentLHS()} to add additional
+samples to the existing Latin hypercube sample.
+
+Another class of augmenting design is \emph{sequential design} (also
+called \emph{adaptive sampling}), which is best represented by
+\href{https://cran.r-project.org/web/packages/tgp/index.html}{\texttt{tgp}}
+(Gramacy and Taddy 2010). This requires prior information that is used
+to inform the next experimental design using \texttt{tgp.design()} and
+\texttt{dopt.gp()}. The user is required to supply candidate samples to
+subsample from and a model or a prior experimental design. Follow-up
+experiments, which can also be classified as sequential designs, is
+implemented by
+\href{https://cran.r-project.org/web/packages/BsMD/index.html}{\texttt{BsMD}}
+(Barrios 2020) using a model-discriminant approach using \texttt{MD()}.
 
@@
 
-Through the exploratory data analysis of three data sources (package
+Through the exploratory data analysis of the three data sources (package
 download logs, package metadata and CRAN task views) outlined in Section
-\ref{data}, we have observed that the the total download of DoE packages
-are concentrated only on a handful of R-packages although these
-represent a diverse set in comparison with other CRAN task views.
-Furthermore, the data suggests that the field of experimental design is
-the least collaborative field; out of all CRAN task views, it has one of
-the least average number of authors and the lowest intra-connectivity
-(any dependency on other packages within its own task view). There are a
-number of limitations and short coming in our exploratory data analysis.
-First, CRAN task views are volunteer maintained so some experimental
-design packages may not be included in the DoE packages. Second, we only
-use the RStudio CRAN mirror download, which may bias our observations.
-Third, our analysis are limited to R-packages alone, although the sheer
-quantity and variation of experimental design is unmatched with any
-other programming languages. Finally, all our statements should be
-treated as speculative rather than conclusive; the data are all
-observational so no conclusive, generalisable statement are possible.
-Regardless, the data driven nature of our analysis give an objective
-insight into the field of experimental design.
+\ref{data}, we observed in Section \ref{eda} that the the total download
+of DoE packages is concentrated only on a handful of R-packages although
+these represent a diverse set in comparison with other CRAN task views.
+Furthermore, the data suggest that experimental design is the least
+collaborative field.
+
+There are a number of limitations and shortcomings to our exploratory
+data analysis. First, CRAN task views are volunteer maintained so some
+experimental design packages may not be included in the DoE packages.
+Second, we used only the RStudio CRAN mirror download, which may have
+biased our observations. Third, our analysis was limited to R-packages
+alone and many practitioners may use other methods to construct
+experimental designs. Finally, all our statements should be treated as
+speculative rather than conclusive; the data are all observational so no
+conclusive, generalizable statement is possible. Regardless, the
+data-driven nature of our analysis provides objective insight into the
+field of experimental design.
 
 The interface design (discussed in Section \ref{design}) reveals that
-the most widely used DoE packages generally have functions that is 1) a
-menu format (i.e.~the name of the function correspond to a particular
-recipe of the design or optimal search algorithm) and 2) context is
-often a second-thought (e.g.~only the number of factors is needed and
-the package will assign pseudo factor names or an argument exists for
-optional input of character vector that corresponds to the factor
-names).
+the most widely used DoE packages generally have functions that 1) focus
+on certain aspects of experimental design (e.g.~factorial structure or
+augmenting design), 2) are a recipe format (i.e.~the name of the
+function is a surrogate to a single class of the design or optimal
+search algorithm) and 3) context is often a second thought -- many
+inputs a single integer corresponding to the number of factors, levels,
+or experimental runs (or sample size). The function will often assign
+pseudo-factor names or there is an optional argument to input a
+character vector that corresponds to the factor names. These interface
+designs require users to have processed the experiment in statistical
+terms (often stripping the experimental context away) and
+simultaneously, users must choose the generation mechanism by selecting
+an appropriate function. Arguably, the current dominant interface
+designs are not aligned with the way practitioners cognitively design
+their experiments. Often, the critical part of designing an experiment
+is to understand the experimental structure, and the experimental
+context can govern or guide the choice of algorithm to allocate
+treatments. In addition, the nature of recipe functions can obscure the
+understanding and relation of designs (e.g., how do you go from an
+unstructured factorial design to split plot design?). Each new method
+for generating an experimental design appears to correspond to a
+completely new function, and intermediate results are often not easily
+accessible. These factors may contribute to why developers often work in
+silos in the field of experimental design.
+
+A consistent cognitive interface design that leverages existing
+developments and can be easily extended to new methods will conceivably
+be of great help to practitioners. Some efforts to this end are seen in
+\href{https://cran.r-project.org/web/packages/DoE.wrapper/index.html}{\texttt{DoE.wrapper}}
+(Grömping 2020), which contains wrapper recipe functions for other DoE
+packages such as
+\href{https://cran.r-project.org/web/packages/lhs/index.html}{\texttt{lhs}},
+\href{https://cran.r-project.org/web/packages/AlgDesign/index.html}{\texttt{AlgDesign}}
+and
+\href{https://cran.r-project.org/web/packages/FrF2/index.html}{\texttt{FrF2}}
+(see Figure \ref{fig:plot-doe-network}), and is also the subject of the
+developmental package
+\href{https://cran.r-project.org/web/packages/edibble/index.html}{\texttt{edibble}}
+(Tanaka 2021). Undoubtedly, no single developer or package can cater to
+all experimental designs; therefore, any unifying interface should
+consider how other developers can contribute or add their methods.
+Future research could benefit from further exploratory data analysis,
+expanding the study to beyond R-packages, and discussing other aspects
+of interface designs.
 
@@
 
-This paper uses \texttt{targets} (Landau 2021) and \texttt{renv} (Ushey
-2022) for reproducibility, \texttt{knitr} (Xie 2015) and
-\texttt{rmarkdown} (Xie, Allaire, and Grolemund 2018) for creating
-reproducible documents, \texttt{ggplot2} (Wickham 2016), \texttt{ggraph}
-(Pedersen 2021), \texttt{ggwordcloud} (Le Pennec and Slowikowski 2019)
-and \texttt{colorspace} (Zeileis et al. 2020) for visualisation,
-\texttt{kableExtra} (Zhu 2021) for customising the table in the
-Supplementary Material and \texttt{tidyverse} (Wickham et al. 2019),
-\texttt{tidytext} (Silge and Robinson 2016), \texttt{pluralize} (Rudis
-and Embrey 2020), and \texttt{ineq} (Zeileis 2014) for data processing
-and manipulation, and \texttt{cranlogs} (Csárdi 2019) and \texttt{ctv}
-(Zeileis 2005) for extracting data. All code to reproduce this paper is
-found at \url{https://github.com/emitanaka/paper-DoE-review}.
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
+This paper uses
+\href{https://cran.r-project.org/web/packages/targets/index.html}{\texttt{targets}}
+(Landau 2021) and
+\href{https://cran.r-project.org/web/packages/renv/index.html}{\texttt{renv}}
+(Ushey 2022) for reproducibility,
+\href{https://cran.r-project.org/web/packages/knitr/index.html}{\texttt{knitr}}
+(Xie 2015) and
+\href{https://cran.r-project.org/web/packages/rmarkdown/index.html}{\texttt{rmarkdown}}
+(Xie, Allaire, and Grolemund 2018) for creating reproducible documents,
+\href{https://cran.r-project.org/web/packages/ggplot2/index.html}{\texttt{ggplot2}}
+(Wickham 2016),
+\href{https://cran.r-project.org/web/packages/ggraph/index.html}{\texttt{ggraph}}
+(Pedersen 2021),
+\href{https://cran.r-project.org/web/packages/ggwordcloud/index.html}{\texttt{ggwordcloud}}
+(Le Pennec and Slowikowski 2019) and \texttt{colorspace} (Zeileis et al.
+2020) for visualization,
+\href{https://cran.r-project.org/web/packages/kableExtra/index.html}{\texttt{kableExtra}}
+(Zhu 2021) for customizing the table in the Supplementary Material and
+\href{https://cran.r-project.org/web/packages/tidyverse/index.html}{\texttt{tidyverse}}
+(Wickham et al. 2019),
+\href{https://cran.r-project.org/web/packages/tidytext/index.html}{\texttt{tidytext}}
+(Silge and Robinson 2016),
+\href{https://cran.r-project.org/web/packages/pluralize/index.html}{\texttt{pluralize}}
+(Rudis and Embrey 2020), and
+\href{https://cran.r-project.org/web/packages/ineq/index.html}{\texttt{ineq}}
+(Zeileis 2014) for data processing and manipulation, and
+\href{https://cran.r-project.org/web/packages/cranlogs/index.html}{\texttt{cranlogs}}
+(Csárdi 2019) and
+\href{https://cran.r-project.org/web/packages/ctv/index.html}{\texttt{ctv}}
+(Zeileis 2005) for extracting data. All codes used to reproduce this
+paper are found in \url{https://github.com/emitanaka/paper-DoE-review}.
 

### 2022-12-14 paper/arxiv/ref-setup.R

@@ -1,0 +1,6 @@
+# arxiv
+ctv <- function(view) glue("\\ctv{[view]}", .open = "[", .close = "]")
+CRANpkg <- function(pkg) glue("\\CRANpkg{[pkg]}", .open = "[", .close = "]")
+ref <- function(x) glue("\\ref{[x]}", .open = "[", .close = "]")
+secref <- function(x, name) glue("Section \\ref{[x]}", .open = "[", .close = "]")
+url <- function(x) glue("\\url{[x]}", .open = "[", .close = "]")

### 2022-12-14 ref-setup.R

@@ -1,0 +1,6 @@
+# main HTML
+ctv <- function(view) glue("[{view}](https://cran.r-project.org/web/views/{view}.html)")
+CRANpkg <- function(pkg) glue("[`{pkg}`](https://cran.r-project.org/web/packages/{pkg}/index.html)")
+ref <- function(x) glue("\\@ref({x})")
+secref <- function(x, name) glue("Section \\@ref({x})")
+url <- function(x) glue("[{x}]({x})")

### 2022-12-14 paper.Rmd
fixed reference name

@@
 The remainder of this paper is organized as follows.
-`r secref("data", "data")` briefly describes the data source used for the analysis; `r secref("eda", "Exploratory data analysis" )` presents some insights into the state of the current DoE packages by the exploratory data analysis of package download data, text descriptions and comparisons with other CRAN task views; `r secref("design", "Interface design")` discusses the interface designs of widely used DoE packages, and we conclude with a discussion in `r secref("discussion", "Discussion")` of future prospects in the software development of experimental designs.
+`r secref("data", "Data")` briefly describes the data source used for the analysis; `r secref("eda", "Exploratory data analysis" )` presents some insights into the state of the current DoE packages by the exploratory data analysis of package download data, text descriptions and comparisons with other CRAN task views; `r secref("design", "Interface design")` discusses the interface designs of widely used DoE packages, and we conclude with a discussion in `r secref("discussion", "Discussion")` of future prospects in the software development of experimental designs.
 
